### 2018-05-07 Manuscript.Rmd
Initial commit.

---
title: xxx
preprint: false
author: 
  - name: Tobias Roth
    affiliation: 1
    corresponding: true
    email: t.roth@unibas.ch
  - name: Lukas Kohli
    affiliation: 2
affiliation:
  - code: 1
    address: Zoological Institute, University of Basel, Basel, Switzerland
  - code: 2
    address: Hintermann  Weber AG, Austrasse 2a, 4153 Reinach, Switzerland
abstract: >
  xxx
header-includes: >
  \usepackage{lipsum}
bibliography: Library.bib
output: 
  rticles::peerj_article:
    base_format: bookdown::pdf_document2 
---

```{r settings, echo = FALSE, message=FALSE}
# Libraries
library(knitr)
library(xtable)
library(tidyverse)
library(arm)
library(nlme)
library(simba)

# Load data
load("RData/surv.RData")
load("RData/pl.RData")
load("RData/sites.RData")

# Knitr settings 
opts_chunk$set(echo = FALSE, hide = TRUE, cache = TRUE, warning = FALSE, message = FALSE,
               fig.asp = .6, fig.width = 8, out.width = "100%")

# Theme of ggplot
theme_set(theme_classic())

# Remove comment for xtable
options(xtable.comment = FALSE)
```

# Introduction {-}
xxx

# Materials & Methods {-}

## Monitoring data {-}
- Selection of sample sites based on 1366 K_Standort.csv column "E23_1366".
- Three surveys 2003-2007, 2008-2012 and 2013 - 2017.

## Plant traits {-}

Functional traits:

- SLA: specific leaf area
- CH: canopy height
- SM: Seed mass

Ellenberg indicator values:

- L: light
- N: Nutrient contentent
- T: Temperature
- F: Huminity

Community measures:

- Species richness: number of recorded species per $10m^2$.
- Spatial turnover (beta-diversity): Average turnover between all pair-wise combinations of study plots.
- gamma diversity: Total number of species recorded in all study plots.

## Statistical analyses {-}

Environmental variables were standardized.

```{r dataprep}
surv$yr <- (surv$yearPl - 2010) / 10
surv$Hoehe <- (surv$Hoehe - 1000) / 500
surv$NTOT <- (surv$NTOT2007 - 10) / 10
surv$Neig <- surv$Neig / 10
surv$Preci <- (surv$bio12 - 1000) / 200
# surv$F <- (surv$F - 3) / 0.5
# surv$L <- (surv$L - 3) / 0.5
```

# Results {-}
Different measures of total community structure suggested that plant communities of mountain hay meadow were rather stable between 2003 and 2017: For each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity, average ellenberg values for temperature, nutrients, light, huminity and reaction, and average of species' canopy height, specific leaf area and seed mass did not vary much. For all measures, the average temperal trend per site did not differ significantly from zero (Table \@ref(temporaltrends)). Note that beta- and gamma-diversity are note available for single sites and thus mixed models could not be applied.

```{r temporaltrends, results='asis', }
res <- tibble(Measures = c("Alpha-diversity", "Beta-diversity", "Gamma-Diversity", 
                           "Temperature value", "Huminity value", "Nutrients value", "Light value", 
                           "Canopy height", "Specific leaf area", "Seed mass"))
res$'Period 1' <- ""; res$'Period 2' <- ""; res$'Period 3' <- ""
res$'Temporal-Trend' <- ""; res$'P-value' <- ""

# Alpha diversity
res[1, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$SR, surv$Visit, mean), 2)
tmp <- summary(glmer(SR ~ yr + (yr|aID_STAO), data = surv, family = poisson))$coefficients
res[1, "Temporal-Trend"] <- formatC(tmp["yr", "Estimate"], format = "f", digits = 3)
res[1, "P-value"] <- formatC(tmp["yr", "Pr(>|z|)"], format = "f", digits = 3)

# Beta diversity
getbeta <- function(x) {
  tmp <- simba::sim(pl %>% filter(Visit == x) %>% dplyr::select(aID_STAO, aID_SP, Occ), method = "simpson", 
            listin = TRUE, listout = TRUE)
  formatC(tmp$simpson, format = "f", digits = 2)
}
res[2, c("Period 1", "Period 2", "Period 3")] <- lapply(1:3, getbeta)

# Gamma diversity
res[3, c("Period 1", "Period 2", "Period 3")] <- tapply(pl$aID_SP, pl$Visit, function(x) length(unique(x)))

# Mean community index
res[4, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$T, surv$Visit, mean), 2)
tmp <- summary(lme(T ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
res[4, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
res[4, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)

res[5, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$F, surv$Visit, mean), 2)
tmp <- summary(lme(F ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
res[5, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
res[5, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)

res[6, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$N, surv$Visit, mean), 2)
tmp <- summary(lme(N ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
res[6, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
res[6, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)

res[7, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$L, surv$Visit, mean), 2)
tmp <- summary(lme(L ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
res[7, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
res[7, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)

# Average of functional traits
res[ 8, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$CH, surv$Visit, mean), 2)
tmp <- summary(lmer(CH ~ yr + (yr|aID_STAO), data = surv))$coefficients
res[ 8, "Temporal-Trend"] <- formatC(tmp["yr", "Estimate"], format = "f", digits = 3)
tmp <- anova(lmer(CH ~ yr + (yr|aID_STAO), data = surv), lmer(CH ~ 1 + (yr|aID_STAO), data = surv))[2, "Pr(>Chisq)"]
res[ 8, "P-value"] <- formatC(tmp, format = "f", digits = 3)

res[9, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$SLA, surv$Visit, mean), 2)
tmp <- summary(lme(SLA ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
res[9, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
res[9, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)

res[10, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$SM, surv$Visit, mean), 2)
tmp <- summary(lme(SM ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
res[10, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
res[10, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)

print(xtable(res, align = c("l", "l", rep("r", 5)), label = "temporaltrends", caption = "Average measures of community structure for the three survey periods (in each period all sites are surveyed once). The temporal trends and p-values are based on linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution) with site-ID as random effect. Temporal-trends are given per 10 years. Linear mixed models could not be applied for beta- and gamma-diversity because measures are not available for the single sites."), include.rownames=FALSE)
```

The temporal stability suggested by the community measures were, however, in contrast to the temporal turn-over of recorded species (i.e. species exchange ratio sensu @Hillebrand2018). In average $\pm$ SD the proportion of species that differed between two surveys was `r formatC(100 * mean(sites$Turnover), format = "f", digits = 1)` $\pm$ `r formatC(100 * sd(sites$Turnover), format = "f", digits = 1)`.

```{r}
summary(lm(Turnover ~ NTOT2007 + Hoehe + Neig + Expos, data = sites))
```

```{r}
# Colonization probability
# mean(d$Visit2[d$Visit1 == 0])
# mean(d$Visit3[d$Visit2 == 0])
# 
# col1 <- summary(glmer(Visit2 ~ NTOT2000 + Hoehe + T + N + L + F + (1|aID_SP) + (1|aID_STAO), 
#               data = d %>% filter(Visit1 == 0), family = binomial))
# col2 <- summary(glmer(Visit3 ~ T + N + L + F + (1|aID_SP) + (1|aID_STAO), 
#               data = d %>% filter(Visit2 == 0), family = binomial))
# col3 <- summary(glmer(Visit2 ~ log(CH) + log(SLA) + log(SM) + (1|aID_SP) + (1|aID_STAO), 
#               data = d %>% filter(Visit1 == 0), family = binomial))
# col4 <- summary(glmer(Visit3 ~ log(CH) + log(SLA) + log(SM) + (1|aID_SP) + (1|aID_STAO), 
#               data = d %>% filter(Visit2 == 0), family = binomial))
# 
# # Local survival
# mean(d$Visit2[d$Visit1 == 1])
# mean(d$Visit3[d$Visit2 == 1])
# phi1 <- summary(glmer(Visit2 ~ NTOT2000 + Hoehe + T + N + L + F + (1|aID_SP) + (1|aID_STAO), 
#               data = d %>% filter(Visit1 == 1), family = binomial))
# phi2 <- summary(glmer(Visit3 ~ NTOT2000 + Hoehe + T + N + L + F + (1|aID_SP) + (1|aID_STAO), 
#               data = d %>% filter(Visit2 == 1), family = binomial))
# phi3 <- summary(glmer(Visit2 ~ log(CH) + log(SLA) + log(SM) + (1|aID_SP) + (1|aID_STAO), 
#               data = d %>% filter(Visit1 == 1), family = binomial))
# phi4 <- summary(glmer(Visit3 ~ log(CH) + log(SLA) + log(SM) + (1|aID_SP) + (1|aID_STAO), 
#               data = d %>% filter(Visit2 == 1), family = binomial))
```

# Discussion {-}
xxx

# Conclusions {-}
xxx

# Acknowledgements {-}
xxx

# References {-}

### 2018-05-07 Manuscript.tex
Initial commit.

%% Submissions for peer-review must enable line-numbering
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file, and
%% the corresponding author info requires v1.2

\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions

% ZNK -- Adding headers for pandoc

\setlength{\emergencystretch}{3em}
\providecommand{\tightlist}{
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{lipsum}
\usepackage[unicode=true]{hyperref}
\usepackage{longtable}

\usepackage{lipsum}

\title{xxx}

\author[1]{Tobias Roth}

\corrauthor[1]{Tobias Roth}{\href{mailto:t.roth@unibas.ch}{\nolinkurl{t.roth@unibas.ch}}}
\author[2]{Lukas Kohli}

\affil[1]{Zoological Institute, University of Basel, Basel, Switzerland}
\affil[2]{Hintermann Weber AG, Austrasse 2a, 4153 Reinach, Switzerland}

%
% \author[1]{First Author}
% \author[2]{Second Author}
% \affil[1]{Address of first author}
% \affil[2]{Address of second author}
% \corrauthor[1]{First Author}{f.author@email.com}

%

\begin{abstract}
xxx
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}\label{introduction}
\addcontentsline{toc}{section}{Introduction}

xxx

\section*{Materials \& Methods}\label{materials-methods}
\addcontentsline{toc}{section}{Materials \& Methods}

\subsection*{Monitoring data}\label{monitoring-data}
\addcontentsline{toc}{subsection}{Monitoring data}

\begin{itemize}
\tightlist
\item
  Selection of sample sites based on 1366 K\_Standort.csv column
  ``E23\_1366''.
\item
  Three surveys 2003-2007, 2008-2012 and 2013 - 2017.
\end{itemize}

\subsection*{Plant traits}\label{plant-traits}
\addcontentsline{toc}{subsection}{Plant traits}

Functional traits:

\begin{itemize}
\tightlist
\item
  SLA: specific leaf area
\item
  CH: canopy height
\item
  SM: Seed mass
\end{itemize}

Ellenberg indicator values:

\begin{itemize}
\tightlist
\item
  L: light
\item
  N: Nutrient contentent
\item
  T: Temperature
\item
  F: Huminity
\end{itemize}

Community measures:

\begin{itemize}
\tightlist
\item
  Species richness: number of recorded species per \(10m^2\).
\item
  Spatial turnover (beta-diversity): Average turnover between all
  pair-wise combinations of study plots.
\item
  gamma diversity: Total number of species recorded in all study plots.
\end{itemize}

\subsection*{Statistical analyses}\label{statistical-analyses}
\addcontentsline{toc}{subsection}{Statistical analyses}

Environmental variables were standardized.

\section*{Results}\label{results}
\addcontentsline{toc}{section}{Results}

Different measures of total community structure suggested that plant
communities of mountain hay meadow were rather stable between 2003 and
2017: For each of the three 5-year survey periods the averages of
alpha-, beta- and gamma-diversity, average ellenberg values for
temperature, nutrients, light, huminity and reaction, and average of
species' canopy height, specific leaf area and seed mass did not vary
much. For all measures, the average temperal trend per site did not
differ significantly from zero (Table \ref{temporaltrends}). Note that
beta- and gamma-diversity are note available for single sites and thus
mixed models could not be applied.

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrr}
  \hline
Measures & Period 1 & Period 2 & Period 3 & Temporal-Trend & P-value \\ 
  \hline
Alpha-diversity & 46.36 & 46.72 & 46.45 & 0.002 & 0.896 \\ 
  Beta-diversity & 0.68 & 0.65 & 0.65 &  &  \\ 
  Gamma-Diversity & 517 & 529 & 517 &  &  \\ 
  Temperature value & 3.12 & 3.14 & 3.14 & 0.013 & 0.060 \\ 
  Huminity value & 2.99 & 2.98 & 2.99 & 0.006 & 0.405 \\ 
  Nutrients value & 3.22 & 3.22 & 3.22 & -0.004 & 0.698 \\ 
  Light value & 3.57 & 3.56 & 3.56 & -0.010 & 0.196 \\ 
  Canopy height & -1.24 & -1.22 & -1.23 & 0.013 & 0.307 \\ 
  Specific leaf area & 8.21 & 8.27 & 8.24 & 0.030 & 0.621 \\ 
  Seed mass & -0.34 & -0.32 & -0.33 & 0.010 & 0.596 \\ 
   \hline
\end{tabular}
\caption{Average measures of community structure for the three survey periods (in each period all sites are surveyed once). The temporal trends and p-values are based on linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution) with site-ID as random effect. Temporal-trends are given per 10 years. Linear mixed models could not be applied for beta- and gamma-diversity because measures are not available for the single sites.} 
\label{temporaltrends}
\end{table}

The temporal stability suggested by the community measures were,
however, in contrast to the temporal turn-over of recorded species
(i.e.~species exchange ratio sensu Hillebrand et al. (2018)). In average
\(\pm\) SD the proportion of species that differed between two surveys
was 18.7 \(\pm\) 6.8.

\begin{verbatim}
## 
## Call:
## lm(formula = Turnover ~ NTOT2007 + Hoehe + Neig + Expos, data = sites)
## 
## Residuals:
##       Min        1Q    Median        3Q       Max 
## -0.148019 -0.040275 -0.007297  0.042223  0.205389 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)  3.468e-01  5.191e-02   6.682 6.98e-10 ***
## NTOT2007    -4.629e-03  1.498e-03  -3.091  0.00246 ** 
## Hoehe       -7.279e-05  2.263e-05  -3.216  0.00165 ** 
## Neig        -1.816e-04  7.590e-04  -0.239  0.81125    
## Expos        2.433e-05  2.401e-05   1.014  0.31278    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.06522 on 125 degrees of freedom
## Multiple R-squared:  0.1085, Adjusted R-squared:  0.07997 
## F-statistic: 3.803 on 4 and 125 DF,  p-value: 0.005939
\end{verbatim}

\section*{Discussion}\label{discussion}
\addcontentsline{toc}{section}{Discussion}

xxx

\section*{Conclusions}\label{conclusions}
\addcontentsline{toc}{section}{Conclusions}

xxx

\section*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

xxx

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Hillebrand2018}{}
Hillebrand, Helmut, Bernd Blasius, Elizabeth T. Borer, Jonathan M.
Chase, John A. Downing, Britas Klemens Eriksson, Christopher T.
Filstrup, et al. 2018. ``Biodiversity Change Is Uncoupled from Species
Richness Trends: Consequences for Conservation and Monitoring.''
\emph{Journal of Applied Ecology} 55 (1): 169--84.
doi:\href{https://doi.org/10.1111/1365-2664.12959}{10.1111/1365-2664.12959}.

\end{document}

### 2018-05-07 R/1366_Data_processing.R

@@ -1,0 +1,107 @@
+#======================================================================================================
+#
+# Plant community change in mountain hay meadows
+#
+# DATA PREPARATION
+#
+#======================================================================================================
+
+rm(list=ls(all=TRUE))
+
+#------------------------------------------------------------------------------------------------------
+# Settings
+#------------------------------------------------------------------------------------------------------
+# Libraries
+library(tidyverse)
+library(simba)
+
+# Connection to data base
+db <- src_sqlite(path = "~/Documents/Dropbox/DB_BDM.db", create = FALSE)
+
+#------------------------------------------------------------------------------------------------------
+# Selection of sites
+#------------------------------------------------------------------------------------------------------
+# Import list with sites
+sel <- read_delim("Data/1366 K_Standort.csv", delim = ";") %>% 
+  filter(E23_1366 == "ja") %>% 
+  transmute(aID_STAO = aIdStaoZ9)
+
+# Select only sites where three surveys are available since 2003
+tmp <- tbl(db, "KD_Z9") %>% 
+  as.tibble() %>% 
+  filter(!is.na(match(aID_STAO, sel$aID_STAO)) & !is.na(yearPl) & yearP >= 2003 & yearP <= 2017) %>% 
+  dplyr::select(aID_KD, aID_STAO, yearPl) %>% 
+  left_join(tbl(db, "PL") %>% group_by(aID_KD) %>% summarise(SR = n()), copy = TRUE) %>% 
+  filter(SR >= 0)
+tmp <- table(tmp$aID_STAO)
+sum(tmp != 3)
+sel <- names(tmp[tmp == 3])
+
+# Select surveys and add site data
+surv <- tbl(db, "KD_Z9") %>% 
+  as.tibble() %>% 
+  filter(!is.na(match(aID_STAO, sel)) & !is.na(yearPl) & yearP >= 2003 & yearP <= 2017) %>% 
+  transmute(aID_KD = aID_KD, aID_STAO = aID_STAO, yearP = yearP, yearPl = yearPl, Visit = as.integer(floor((yearP - 1998) / 5))) %>% 
+  left_join(tbl(db, "RAUMDATEN_Z9") %>% 
+              dplyr::select(aID_STAO, Hoehe, Neig, Expos, AS09_72, AS09_17, CACO3, Vernass, BGR_6, 
[…]

### 2018-05-07 R/1366_gains_losses.R

@@ -1,0 +1,57 @@
+#======================================================================================================
+#
+# Plant community change in mountain hay meadows
+#
+# Compare gains and losses
+#
+#======================================================================================================
+
+rm(list=ls(all=TRUE))
+
+#------------------------------------------------------------------------------------------------------
+# Settings and load data
+#------------------------------------------------------------------------------------------------------
+# Libraries
+library(tidyverse)
+library(arm)
+
+# Load data
+load("RData/pl.RData")
+load("RData/sites.RData")
+
+#------------------------------------------------------------------------------------------------------
+# Compile absence data for all species x site combinations
+#------------------------------------------------------------------------------------------------------
+d <- expand.grid(aID_STAO = sites$aID_STAO, aID_SP = unique(pl$aID_SP)) %>% as.tibble() %>% 
+  left_join(pl %>% group_by(aID_SP) %>% summarise(T = min(T), F = min(F), N = min(N), L = min(L))) %>% 
+  left_join(pl %>% filter(yearP >= 2003 & yearP<= 2007) %>% 
+              transmute(aID_STAO = aID_STAO, aID_SP = aID_SP, Occ1 = Occ)) %>% 
+  left_join(pl %>% filter(yearP >= 2008 & yearP <= 2012) %>% 
+              transmute(aID_STAO = aID_STAO, aID_SP = aID_SP, Occ2 = Occ)) %>% 
+  left_join(pl %>% filter(yearP >= 2013 & yearP <= 2017) %>% 
+              transmute(aID_STAO = aID_STAO, aID_SP = aID_SP, Occ3 = Occ)) %>% 
+  replace_na(replace = list(Occ1 = 0, Occ2 = 0, Occ3 = 0)) %>% 
+  left_join(sites) 
+
+#------------------------------------------------------------------------------------------------------
+# Analyse colonization
+#------------------------------------------------------------------------------------------------------
+summary(glmer(Occ2 ~ T + F + N + L + (1|aID_SP) + (1|aID_STAO), data = d %>% filter(Occ1 == 0), family = binomial))
+summary(glmer(Occ3 ~ T + F + N + L + (1|aID_SP) + (1|aID_STAO), data = d %>% filter(Occ2 == 0), family = binomial))
+
+#------------------------------------------------------------------------------------------------------
[…]

### 2018-05-09 Manuscript.Rmd
Auswertung Turnover angepasst und erweitert.

@@
 # Introduction {-}
-xxx
 
@@
 ```{r}
-summary(lm(Turnover ~ NTOT2007 + Hoehe + Neig + Expos, data = sites))
+tt <- sites %>% gather(Period, Turnover, -aID_STAO, -Elevation,  -Inclination, -Temperature, -Precipitation, -NDEP2010)
+
+summary(lmer(Turnover ~ Period * (Temperature + Precipitation + NDEP2010 + Inclination) +
+               (1|aID_STAO), data = tt)) 
+summary(lmer(Turnover ~ Period + Temperature + Precipitation + NDEP2010 + Inclination +
+               (1|aID_STAO), data = tt))
+# Neigung als Mass für Land-use intensity
 ```

### 2018-05-09 R/1366_Data_processing.R

@@ -66,7 +66,8 @@ surv <- surv %>% left_join(
               T = mean(T, na.rm = TRUE), 
               F = mean(F, na.rm = TRUE),
               N = mean(N, na.rm = TRUE),
-              L = mean(L, na.rm = TRUE)), copy = TRUE)
+              L = mean(L, na.rm = TRUE),
+              MV = mean(MV, na.rm = TRUE)), copy = TRUE)
 
 #------------------------------------------------------------------------------------------------------
 # Selection plants data
@@ -88,14 +89,16 @@ pl <- tbl(db, "PL") %>%
 sites <- data.frame(aID_STAO = as.integer(sel)) %>% 
   as.tibble() %>% 
   left_join(tbl(db, "RAUMDATEN_Z9") %>% 
-              dplyr::select(aID_STAO, Hoehe, Neig, Expos, AS09_72, AS09_17, CACO3, Vernass, BGR_6, 
-                            bio1, bio12, bio13, bio14), copy = TRUE) %>% 
+              transmute(aID_STAO = aID_STAO, Elevation = Hoehe, Inclination = Neig, 
+                            Temperature = bio1, Precipitation = bio12), copy = TRUE) %>% 
   left_join(tbl(db, "RAUMDATEN_Z9_NDEP") %>% 
-              dplyr::select(aID_STAO, NTOT2007), copy = TRUE)
+              transmute(aID_STAO = aID_STAO, NDEP2010 = NTOT2010b), copy = TRUE)
 
 for(i in 1:nrow(sites)) {
-  sites[i, "Turnover"] <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i]) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "simpson", 
-      listin = TRUE, listout = TRUE)$simpson %>% mean
+  sites[i, "Turnover1"] <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit <= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", 
+      listin = TRUE, listout = TRUE)$cocogaston %>% mean
+  sites[i, "Turnover2"] <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit >= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", 
+                               listin = TRUE, listout = TRUE)$cocogaston %>% mean
 }
   
 #------------------------------------------------------------------------------------------------------

### 2018-05-13 Manuscript.Rmd
First version of Results chapter.

--- 2018-05-07 Manuscript.Rmd
@@
   - name: Tobias Roth
-    affiliation: 1
+    affiliation: 1, 2
     corresponding: true
@@
 
-```{r settings, echo = FALSE, message=FALSE}
-# Libraries
-library(knitr)
-library(xtable)
-library(tidyverse)
-library(arm)
-library(nlme)
-library(simba)
-
-# Load data
-load("RData/surv.RData")
-load("RData/pl.RData")
-load("RData/sites.RData")
-
-# Knitr settings 
-opts_chunk$set(echo = FALSE, hide = TRUE, cache = TRUE, warning = FALSE, message = FALSE,
-               fig.asp = .6, fig.width = 8, out.width = "100%")
-
-# Theme of ggplot
-theme_set(theme_classic())
-
-# Remove comment for xtable
-options(xtable.comment = FALSE)
+```{r settings, echo = FALSE, message=FALSE, cache=FALSE}
+source("R/1366_Settings.R")
 ```
@@
 ## Plant traits {-}
-
 Functional traits:
@@
 
+### Community measures
+
+The temporal turnover (i.e. species exchange ratio sensu @Hillebrand2018) is the proportion of species that differ between two time points calculated as
+
+$\text{Spatial turnover} = \frac{\text{Species gained} + \text{Species lost}}{\text{Total species observed in both timepoints}}$.
+
 ## Statistical analyses {-}
-
 Environmental variables were standardized.
 
-```{r dataprep}
-surv$yr <- (surv$yearPl - 2010) / 10
-surv$Hoehe <- (surv$Hoehe - 1000) / 500
-surv$NTOT <- (surv$NTOT2007 - 10) / 10
-surv$Neig <- surv$Neig / 10
-surv$Preci <- (surv$bio12 - 1000) / 200
-# surv$F <- (surv$F - 3) / 0.5
-# surv$L <- (surv$L - 3) / 0.5
+```{r helpfunctions}
+# Help functions to get posterior samples
+getposterior <- function(x) {
+  list( 
+    `EL = 2` = plogis(simdat[x,1] + Xsd * simdat[x,2] - simdat[x,3] -  Xsd * simdat[x,4]),
+    `EL = 4` = plogis(simdat[x,1] + Xsd * simdat[x,2] + simdat[x,3] +  Xsd * simdat[x,4]))
+}
+
+compileres <- function(EL) {
+  tt <- map(1:length(X), function(s) map(post, paste(EL)) %>% map_dbl(s))
+  tibble(mean = map_dbl(tt, mean),
+         low = map_dbl(tt, quantile, probs = 0.025),
+         high = map_dbl(tt, quantile, probs = 0.975),
+         X = X,
+         value = paste(EL)
+  )
+}
 ```
@@
 # Results {-}
-Different measures of total community structure suggested that plant communities of mountain hay meadow were rather stable between 2003 and 2017: For each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity, average ellenberg values for temperature, nutrients, light, huminity and reaction, and average of species' canopy height, specific leaf area and seed mass did not vary much.
-For all measures, the average temperal trend per site did not differ significantly from zero (Table \@ref(temporaltrends)).
-Note that beta- and gamma-diversity are note available for single sites and thus mixed models could not be applied.
-
 ```{r temporaltrends, results='asis', }
@@
 
-The temporal stability suggested by the community measures were, however, in contrast to the temporal turn-over of recorded species (i.e.
-species exchange ratio sensu @Hillebrand2018).
-In average $\pm$ SD the proportion of species that differed between two surveys was `r formatC(100 * mean(sites$Turnover), format = "f", digits = 1)` $\pm$ `r formatC(100 * sd(sites$Turnover), format = "f", digits = 1)`.
+The different measures of total community structure suggested that plant communities of mountain hay meadow were stable between 2003 and 2017 (Table \@ref(temporaltrends)): for each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity, average Ellenberg values for temperature, nutrients, light and huminity, and average of species' canopy height, specific leaf area and seed mass did not vary much among the three sampling periods.
+For all measures, the average temperal trend per site did not differ significantly from a constant trend.
+Note that beta- and gamma-diversity are note available for single sites and thus mixed models to calculate the p-value could not be applied.
 
-```{r}
-summary(lm(Turnover ~ NTOT2007 + Hoehe + Neig + Expos, data = sites))
+```{r modturnover}
+tt <- rbind(
+  sites %>% transmute(
+    change = change1, 
+    nochange = nochange1,
+    period = 1),
+  sites %>% transmute(
+    change = change2, 
+    nochange = nochange2,
+    period = 2))
+mod <- glm(cbind(change, nochange) ~ period, data = tt, family = binomial)
 ```
 
-```{r}
-# Colonization probability
-# mean(d$Visit2[d$Visit1 == 0])
-# mean(d$Visit3[d$Visit2 == 0])
-# 
-# col1 <- summary(glmer(Visit2 ~ NTOT2000 + Hoehe + T + N + L + F + (1|aID_SP) + (1|aID_STAO), 
-#               data = d %>% filter(Visit1 == 0), family = binomial))
-# col2 <- summary(glmer(Visit3 ~ T + N + L + F + (1|aID_SP) + (1|aID_STAO), 
-#               data = d %>% filter(Visit2 == 0), family = binomial))
-# col3 <- summary(glmer(Visit2 ~ log(CH) + log(SLA) + log(SM) + (1|aID_SP) + (1|aID_STAO), 
-#               data = d %>% filter(Visit1 == 0), family = binomial))
-# col4 <- summary(glmer(Visit3 ~ log(CH) + log(SLA) + log(SM) + (1|aID_SP) + (1|aID_STAO), 
-#               data = d %>% filter(Visit2 == 0), family = binomial))
-# 
-# # Local survival
-# mean(d$Visit2[d$Visit1 == 1])
-# mean(d$Visit3[d$Visit2 == 1])
-# phi1 <- summary(glmer(Visit2 ~ NTOT2000 + Hoehe + T + N + L + F + (1|aID_SP) + (1|aID_STAO), 
-#               data = d %>% filter(Visit1 == 1), family = binomial))
-# phi2 <- summary(glmer(Visit3 ~ NTOT2000 + Hoehe + T + N + L + F + (1|aID_SP) + (1|aID_STAO), 
-#               data = d %>% filter(Visit2 == 1), family = binomial))
-# phi3 <- summary(glmer(Visit2 ~ log(CH) + log(SLA) + log(SM) + (1|aID_SP) + (1|aID_STAO), 
-#               data = d %>% filter(Visit1 == 1), family = binomial))
-# phi4 <- summary(glmer(Visit3 ~ log(CH) + log(SLA) + log(SM) + (1|aID_SP) + (1|aID_STAO), 
-#               data = d %>% filter(Visit2 == 1), family = binomial))
+The temporal stability as inferred from the community measures were, however, in contrast to the large  temporal turnover of recorded species.
+The percentage $\pm$ SD of species that differ between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))` and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`.
+The decrease in the temporal turnover between the first and second survey to the temporal turnover between the second and third survey was statistically significant (Binomial generalized linear model; effect size = `r coef(mod)[2]`; p = `r summary(mod)$coefficients["period", "Pr(>|z|)"]`).
+
+```{r compofvarianceran}
+modcol <- summary(glmer(Occ ~ 1 + (1|aID_STAO) + (1|aID_SP), data = coldat, family = binomial))
+modls <- summary(glmer(Occ ~ 1 + (1|aID_STAO) + (1|aID_SP), data = survdat, family = binomial))
+getratio <- function(x) x[1] / x[2]
 ```
 
+This spatial turnover might be partly explained by species that remained undetected in one of the surveys, but it might also be the result of species that newly colonized sites (species gains) and species that truely disappeared from sites (species losses).
+Spatial turnover will be high if the probability a site is colonized by a species is high (i.e.
+high colonization probability) and/or if the probability that a species survives at a site between two time points is low (i.e.
+low local survival).
+We found that the variation in local survival and colonization, was largely due to differences between species and only to a lesser extend to the differences between locations: while the variance of the species random effect in a binomial linear mixed model for local survival was about `r getratio(unlist(modls$varcor))` times larger than the variance of the sampling site random effect, in the model for colonization the variance among species was even `r getratio(unlist(modcol$varcor))` times bigger than the variance among sites.
+
+```{r compofvariance}
+# Local survival
+ref <- glmer(Occ ~ 1 + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial)
+modT <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), 
+              data = survdat, family = binomial)
+modF <- glmer(Occ ~ Precipitation * F + (1|aID_SP) + (1|aID_STAO), 
+              data = survdat, family = binomial)
+modN <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), 
+              data = survdat, family = binomial)
+modL <- glmer(Occ ~ Inclination * L + (1|aID_SP) + (1|aID_STAO), 
+              data = survdat, family = binomial)
+tt <- data.frame(anova(ref, modT, modF, modN, modL))
+tt$Model <- c("Random change", "Temperature driven", "Precipitation driven", "Nitrogen driven", "Land-use driven")
+tt <- tt[order(tt$AIC),]
+tt$'Delta-AIC' <- tt$AIC - tt$AIC[1]
+resls <- tt[, c("Model", "Df", "AIC", "Delta-AIC")]
+
+# Colonization probability
+ref <- glmer(Occ ~ 1 + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial)
+modT <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), 
+              data = coldat, family = binomial)
+modF <- glmer(Occ ~ Precipitation * F + (1|aID_SP) + (1|aID_STAO), 
+              data = coldat, family = binomial)
+modN <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), 
+              data = coldat, family = binomial)
+modL <- glmer(Occ ~ Inclination * L + (1|aID_SP) + (1|aID_STAO), 
+              data = coldat, family = binomial)
+tt <- data.frame(anova(ref, modT, modF, modN, modL))
+tt$Model <- c("Random change", "Temperature driven", "Precipitation driven", "Nitrogen driven", "Land-use driven")
+tt <- tt[order(tt$AIC),]
+tt$'Delta-AIC' <- tt$AIC - tt$AIC[1]
+rescol <- tt[, c("Model", "Df", "AIC", "Delta-AIC")]
+```
+```{r resdifeffects}
+tt <- rbind(rescol, resls)
+kable(tt, digits = 2, row.names = FALSE, format = "latex", caption = "Comparision of results from different drivers to explain colonization probability (a) and local survvival (b).
+See methods for the tested models.", booktabs = T) %>%
+    kable_styling() %>%
+    group_rows("(a) Colonization probability", 1, 5, bold = F, italic = T) %>%
+    group_rows("(b) Local survival probability", 6, 10, bold = F, italic = T)
+```
+
+We thus inferred if we could explain differences in colonization probability and local survival with Ellengerg species values for temperature, humidity, nutrient or light.
+We found that both colonization probability and local survival probability was best explained by temperature (i.e.
+highest colonization or local survival probability of species with an Ellenberg value for Temperature that corresponds to the annual average temperature of the site) followed by nutrient availability (i.e.
+highest colonization or local survival probability of species with an Ellenberg value for nutrient availability that corresponds to the annual N deposition rate estimated for the site; Table \@ref(tab:resdifeffects)).
+
+```{r Teff, fig.cap="Colonization (a) and local survival (b) of low temperature species (Ellenberg T = 2; red line) and high temperature species (Ellengerg T = 4) species along the temperature gradient.
+Given are means and 95\\%-Credible Intervals from logistic linear mixed models."}
+X <- seq(0, 10, 0.2)
+Xsd <- (X - 5)
+
+# Colonization
+mod <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial)
+simdat <- arm::sim(mod, nsim)@fixef
+post <- map(1:nsim, getposterior)
+res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
+p1 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
+  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
+  geom_line() +
+    labs(title = "(a) Colonization", 
+         y = "Colonization probability", 
+         x = "Temperature gradient")
+
+# Local survival
+mod <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial)
+simdat <- arm::sim(mod, nsim)@fixef
+post <- map(1:nsim, getposterior)
+res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
+p2 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
+  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
+  geom_line() +
+    labs(title = "(b) Local survival", 
+         y = "Local survival probability", 
+         x = "Temperature gradient")
+multiplot(p1, p2, cols = 2)
+```
+
+```{r Neff, fig.cap="Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
+Given are means and 95\\%-Credible Intervals from logistic linear mixed models."}
+X <- 0:50
+Xsd <- (X - 10) / 10
+
+# Colonization
+mod <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial)
+simdat <- arm::sim(mod, nsim)@fixef
+post <- map(1:nsim, getposterior)
+res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
+p1 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
+  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
+  geom_line() +
+    labs(title = "(a) Colonization", 
+         y = "Colonization probability", 
+         x = "Nitrogen deposition")
+
+# Local survival
+mod <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial)
+simdat <- arm::sim(mod, nsim)@fixef
+post <- map(1:nsim, getposterior)
+res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
+p2 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
+  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
+  geom_line() +
+    labs(title = "(b) Local survival", 
+         y = "Local survival probability", 
+         x = "Nitrogen deposition")
+multiplot(p1, p2, cols = 2)
+```
+
+While the colonization and local survival probability of cold living species (Ellenberg T = 2) declindes along the temperature gradient, the colonization and local survival probability of warm living species (Ellenberg T = 4) increases along the temperature gradient (Fig.
+\@ref(fig:Teff)).
+Similarly for nutrient values: while the colonization and local survival probability of oligotrophic species (Ellenberg N = 2) declindes along the N deposition gradient, the colonization and local survival probability of eutrophic species (N = 4) increases along the N deposition gradient (Fig.
+\@ref(fig:Neff)).
+
+```{r periodeffect}
+# Temperature driven model
+modTcol <- glmer(Occ ~ factor(Period) * Temperature * T + (1|aID_SP) + (1|aID_STAO),
+               data = coldat, family = binomial)
+modTls <- glmer(Occ ~ factor(Period) * Temperature * T + (1|aID_SP) + (1|aID_STAO),
+               data = survdat, family = binomial)
+
+# Nitrogen driven model
+modNcol <- glmer(Occ ~ factor(Period) * NTOT * N + (1|aID_SP) + (1|aID_STAO),
+               data = coldat, family = binomial)
+modNls <- glmer(Occ ~ factor(Period) * NTOT * N + (1|aID_SP) + (1|aID_STAO),
+               data = survdat, family = binomial)
+```
+
+For the temperature driven and nitrogen driven models in table \@ref(tab:resdifeffects) we also tested whether the strength of the effects differed from the first/second to the second/third survey period.
+For the temperatur driven models we found that the differences how species with different Ellenberg T values responded to the average annual temperature of a site tended to decrease from the first/second study period to the second/third study period (colonization probability: period x Ellenberg T x annual mean temperature of site = `r fixef(modTcol)["factor(Period)2:Temperature:T"]`, p = `r summary(modTcol)$coefficients["factor(Period)2:Temperature:T", "Pr(>|z|)"]`; local survival: period x Ellenberg T x annual mean temperature of site = `r fixef(modTls)["factor(Period)2:Temperature:T"]`, p = `r summary(modTls)$coefficients["factor(Period)2:Temperature:T", "Pr(>|z|)"]`).
+In contrast, however, the differences how species with different Ellenberg N values responded to the average N deposition of a site tended to increase from the first/second study period to the second/third study period (colonization probability: period x Ellenberg N x N deposition at site = `r fixef(modNcol)["factor(Period)2:NTOT:N"]`, p = `r summary(modNcol)$coefficients["factor(Period)2:NTOT:N", "Pr(>|z|)"]`; local survival: period x Ellenberg N x N deposition at site = `r fixef(modNls)["factor(Period)2:NTOT:N"]`, p = `r summary(modNls)$coefficients["factor(Period)2:NTOT:N", "Pr(>|z|)"]`).
+
+```{r figconsequences, fig.cap="Beschreibung einfügen."}
+# Change of correlation across study period
+getcor <- function(year) {
+  list(
+    corT = cor(surv %>% filter(yearP >= year & yearP <= (year + 4)) %>% dplyr::select(Temperature, T))[1,2],
+    corN = cor(surv %>% filter(yearP >= year & yearP <= (year + 4)) %>% dplyr::select(NTOT, N))[1,2])
+}
+tt <- map(2003:2013, getcor)
+res <- rbind(
+  tibble(cor = map_dbl(tt, "corN"),
+              driver = "NTOT",
+              year = 2003:2013),
+  tibble(cor = map_dbl(tt, "corT"),
+              driver = "Temperature",
+              year = 2003:2013))
+p1 <- ggplot(res, aes(x = year, y = cor, col = driver)) +
+  geom_point() +
+  geom_line() +
+  ylim(0.5, 1)
+res$yr <- res$year - 2010 
+cormod <- lm(cor ~ yr * driver, data = res)
+
+# Effect on N deposition model
+makemod <- function(year) {
+  mod <- glm(SR ~ Temperature + Precipitation + NTOT + Inclination, family = poisson,
+      data = surv %>% filter(yearP >= year & yearP <= (year + 4)))
+  arm::sim(mod, nsim)@coef
+}
+
+getestim <- function(x) {
+  list(meanN = mean(x[, "NTOT"]),
+       lowN = quantile(x[, "NTOT"], probs = 0.25),
+       upN = quantile(x[, "NTOT"], probs = 0.75),
+       meanT = mean(x[, "Temperature"]),
+       lowT = quantile(x[, "Temperature"], probs = 0.25),
+       upT = quantile(x[, "Temperature"], probs = 0.75))
+}
+tt <- map(2003:2013, makemod) %>% map(getestim)
+res <- rbind(
+  tibble(mean = map_dbl(tt, "meanN"),
+              low = map_dbl(tt, "lowN"),
+              up = map_dbl(tt, "upN"),
+              factor = "NTOT",
+              period = 2003:2013),
+    tibble(mean = map_dbl(tt, "meanT"),
+              low = map_dbl(tt, "lowT"),
+              up = map_dbl(tt, "upT"),
+              factor = "Temperature",
+              period = 2003:2013))
+
+p2 <- ggplot(res, aes(x = period, y = mean, col = factor)) +
+  geom_errorbar(aes(ymin = low, ymax = up), width = 0.2) +
+  geom_point() +
+  geom_line() +
+  geom_abline(intercept = 0, slope = 0) +
+  ylim(-0.1, 0.015)
+multiplot(p1, p2, cols = 2)
+```
+
+It thus seems that temperature is the main driver of species turnover in the first part of the study period and its importance is slightly decreasing over time and nitrogen deposition becomes relatively more important as a driver that shapes the species communities.
+We thus expect to see that the correlation of the mean temperature value of a community with the yearly annual temperature of the site (temperature correlation) is decreasing over the years while the correlation of the mean nutrient value of a community with the N deposition of the site (nitrogen correlation) is increasing over the years.
+Our results tent to be in line with this expectation (linear model: temporal trend of temperature correlation x temporal trend of nitrogen correlation = `r coef(cormod)["yr:driverTemperature"]`, p = `r  summary(cormod)$coefficients["yr:driverTemperature", "Pr(>|t|)"]`), however the change in the correlation over the study period remained week and the temperature correlation remained stronger than the nitrogen correlation over the entire study period (Left panel of Fig.
+\@ref(fig:figconsequences)).
+More important, however, was the effect on our model that we used to infer whether and how strong the spatial variation in N deposition is correlated with Nitrogen deposition.
+If we apply that model the estimated effect of N depositon on total species richness is increasing over the study period (right panel of \@ref(fig:figconsequences)).
+
 # Discussion {-}
-xxx
+- Any empty space that could be caused by any disturbance that let to the local disappearance of species is likely to be fiellied by eutrophic species.
+Disturbance depends on the site, while colonization depends on species characteristics.
+
+- Altough N deposition considerabely declinded between 2005 and 2015, we could not detect major shifts in plant community structure during the same time period.
+
+- Eutrophic species have rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at high N deposition.
+This suggests that it takes much more time to replace eutrophic by oligotrophic species than replacing oligotrophic by eutrophic species.
+
+- Our data on colonization and local survival (i.e. temporal variation) confirm the empirical critical loads that we infered from anlysing spatial co-variation of N deposition and species richness.
+
+- Local survival is higher for low temperature plants --> This could explain the decrease in cumminity change along elevation.
+This could also explain the differences at mount summets were space was empty in the beginning.
+
+- Climatic effects are more likely to be reversed thant effects due to fertilization.
 

### 2018-05-13 Manuscript.tex
First version of Results chapter.

@@
 
-\author[1]{Tobias Roth}
+\author[1, 2]{Tobias Roth}
 
-\corrauthor[1]{Tobias Roth}{\href{mailto:t.roth@unibas.ch}{\nolinkurl{t.roth@unibas.ch}}}
+\corrauthor[1, 2]{Tobias Roth}{\href{mailto:t.roth@unibas.ch}{\nolinkurl{t.roth@unibas.ch}}}
 \author[2]{Lukas Kohli}
@@
 
+\begin{verbatim}
+## Warning: package 'kableExtra' was built under R version 3.4.4
+\end{verbatim}
+
 \section*{Introduction}\label{introduction}
 \addcontentsline{toc}{section}{Introduction}
-
-xxx
 
@@
 
+\subsubsection{Community measures}\label{community-measures}
+
+The temporal turnover (i.e.~species exchange ratio sensu Hillebrand et
+al. (2018)) is the proportion of species that differ between two time
+points calculated as
+
+\(\text{Spatial turnover} = \frac{\text{Species gained} + \text{Species lost}}{\text{Total species observed in both timepoints}}\).
+
 \subsection*{Statistical analyses}\label{statistical-analyses}
@@
 \addcontentsline{toc}{section}{Results}
-
-Different measures of total community structure suggested that plant
-communities of mountain hay meadow were rather stable between 2003 and
-2017: For each of the three 5-year survey periods the averages of
-alpha-, beta- and gamma-diversity, average ellenberg values for
-temperature, nutrients, light, huminity and reaction, and average of
-species' canopy height, specific leaf area and seed mass did not vary
-much. For all measures, the average temperal trend per site did not
-differ significantly from zero (Table \ref{temporaltrends}). Note that
-beta- and gamma-diversity are note available for single sites and thus
-mixed models could not be applied.
 
@@
 
-The temporal stability suggested by the community measures were,
-however, in contrast to the temporal turn-over of recorded species
-(i.e.~species exchange ratio sensu Hillebrand et al. (2018)). In average
-\(\pm\) SD the proportion of species that differed between two surveys
-was 18.7 \(\pm\) 6.8.
+The different measures of total community structure suggested that plant
+communities of mountain hay meadow were stable between 2003 and 2017
+(Table \ref{temporaltrends}): for each of the three 5-year survey
+periods the averages of alpha-, beta- and gamma-diversity, average
+Ellenberg values for temperature, nutrients, light and huminity, and
+average of species' canopy height, specific leaf area and seed mass did
+not vary much among the three sampling periods. For all measures, the
+average temperal trend per site did not differ significantly from a
+constant trend. Note that beta- and gamma-diversity are note available
+for single sites and thus mixed models to calculate the p-value could
+not be applied.
 
-\begin{verbatim}
-## 
-## Call:
-## lm(formula = Turnover ~ NTOT2007 + Hoehe + Neig + Expos, data = sites)
-## 
-## Residuals:
-##       Min        1Q    Median        3Q       Max 
-## -0.148019 -0.040275 -0.007297  0.042223  0.205389 
-## 
-## Coefficients:
-##               Estimate Std. Error t value Pr(>|t|)    
-## (Intercept)  3.468e-01  5.191e-02   6.682 6.98e-10 ***
-## NTOT2007    -4.629e-03  1.498e-03  -3.091  0.00246 ** 
-## Hoehe       -7.279e-05  2.263e-05  -3.216  0.00165 ** 
-## Neig        -1.816e-04  7.590e-04  -0.239  0.81125    
-## Expos        2.433e-05  2.401e-05   1.014  0.31278    
-## ---
-## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
-## 
-## Residual standard error: 0.06522 on 125 degrees of freedom
-## Multiple R-squared:  0.1085, Adjusted R-squared:  0.07997 
-## F-statistic: 3.803 on 4 and 125 DF,  p-value: 0.005939
-\end{verbatim}
+The temporal stability as inferred from the community measures were,
+however, in contrast to the large temporal turnover of recorded species.
+The percentage \(\pm\) SD of species that differ between the first and
+second survey at a site was 37.55 \(\pm\) 10.46 and the percentage of
+species that differ between the second and third survey was 35.63
+\(\pm\) 10.33. The decrease in the temporal turnover between the first
+and second survey to the temporal turnover between the second and third
+survey was statistically significant (Binomial generalized linear model;
+effect size = -0.078; p = 0.027).
+
+This spatial turnover might be partly explained by species that remained
+undetected in one of the surveys, but it might also be the result of
+species that newly colonized sites (species gains) and species that
+truely disappeared from sites (species losses). Spatial turnover will be
+high if the probability a site is colonized by a species is high
+(i.e.~high colonization probability) and/or if the probability that a
+species survives at a site between two time points is low (i.e.~low
+local survival). We found that the variation in local survival and
+colonization, was largely due to differences between species and only to
+a lesser extend to the differences between locations: while the variance
+of the species random effect in a binomial linear mixed model for local
+survival was about 4.46 times larger than the variance of the sampling
+site random effect, in the model for colonization the variance among
+species was even 9.15 times bigger than the variance among sites.
+
+\begin{table}
+
+\caption{\label{tab:resdifeffects}Comparision of results from different drivers to explain colonization probability (a) and local survvival (b). See methods for the tested models.}
+\centering
+\begin{tabular}[t]{lrrr}
+\toprule
+Model & Df & AIC & Delta-AIC\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textit{(a) Colonization probability}}\\
+\hspace{1em}Temperature driven & 6 & 19321.85 & 0.00\\
+\hspace{1em}Nitrogen driven & 6 & 19691.81 & 369.96\\
+\hspace{1em}Precipitation driven & 6 & 20032.57 & 710.73\\
+\hspace{1em}Random change & 3 & 20038.36 & 716.52\\
+\hspace{1em}Land-use driven & 6 & 20041.22 & 719.38\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textit{(b) Local survival probability}}\\
+\hspace{1em}Temperature driven & 6 & 9811.08 & 0.00\\
+\hspace{1em}Nitrogen driven & 6 & 9827.78 & 16.70\\
+\hspace{1em}Land-use driven & 6 & 9834.31 & 23.23\\
+\hspace{1em}Random change & 3 & 9837.76 & 26.68\\
+\hspace{1em}Precipitation driven & 6 & 9838.55 & 27.47\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+We thus inferred if we could explain differences in colonization
+probability and local survival with Ellengerg species values for
+temperature, humidity, nutrient or light. We found that both
+colonization probability and local survival probability was best
+explained by temperature (i.e.~highest colonization or local survival
+probability of species with an Ellenberg value for Temperature that
+corresponds to the annual average temperature of the site) followed by
+nutrient availability (i.e.~highest colonization or local survival
+probability of species with an Ellenberg value for nutrient availability
+that corresponds to the annual N deposition rate estimated for the site;
+Table \ref{tab:resdifeffects}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Manuscript_files/figure-latex/Teff-1} \caption{Colonization (a) and local survival (b) of low temperature species (Ellenberg T = 2; red line) and high temperature species (Ellengerg T = 4) species along the temperature gradient.
+Given are means and 95\%-Credible Intervals from logistic linear mixed models.}\label{fig:Teff}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Manuscript_files/figure-latex/Neff-1} \caption{Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
+Given are means and 95\%-Credible Intervals from logistic linear mixed models.}\label{fig:Neff}
+\end{figure}
+
+While the colonization and local survival probability of cold living
+species (Ellenberg T = 2) declindes along the temperature gradient, the
+colonization and local survival probability of warm living species
+(Ellenberg T = 4) increases along the temperature gradient (Fig.
+\ref{fig:Teff}). Similarly for nutrient values: while the colonization
+and local survival probability of oligotrophic species (Ellenberg N = 2)
+declindes along the N deposition gradient, the colonization and local
+survival probability of eutrophic species (N = 4) increases along the N
+deposition gradient (Fig. \ref{fig:Neff}).
+
+For the temperature driven and nitrogen driven models in table
+\ref{tab:resdifeffects} we also tested whether the strength of the
+effects differed from the first/second to the second/third survey
+period. For the temperatur driven models we found that the differences
+how species with different Ellenberg T values responded to the average
+annual temperature of a site tended to decrease from the first/second
+study period to the second/third study period (colonization probability:
+period x Ellenberg T x annual mean temperature of site = -0.0098, p =
+0.70; local survival: period x Ellenberg T x annual mean temperature of
+site = -0.072, p = 0.051). In contrast, however, the differences how
+species with different Ellenberg N values responded to the average N
+deposition of a site tended to increase from the first/second study
+period to the second/third study period (colonization probability:
+period x Ellenberg N x N deposition at site = 0.025, p = 0.74; local
+survival: period x Ellenberg N x N deposition at site = 0.075, p =
+0.50).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Manuscript_files/figure-latex/figconsequences-1} \caption{Beschreibung einfügen.}\label{fig:figconsequences}
+\end{figure}
+
+It thus seems that temperature is the main driver of species turnover in
+the first part of the study period and its importance is slightly
+decreasing over time and nitrogen deposition becomes relatively more
+important as a driver that shapes the species communities. We thus
+expect to see that the correlation of the mean temperature value of a
+community with the yearly annual temperature of the site (temperature
+correlation) is decreasing over the years while the correlation of the
+mean nutrient value of a community with the N deposition of the site
+(nitrogen correlation) is increasing over the years. Our results tent to
+be in line with this expectation (linear model: temporal trend of
+temperature correlation x temporal trend of nitrogen correlation =
+-0.0015, p = 0.067), however the change in the correlation over the
+study period remained week and the temperature correlation remained
+stronger than the nitrogen correlation over the entire study period
+(Left panel of Fig. \ref{fig:figconsequences}). More important, however,
+was the effect on our model that we used to infer whether and how strong
+the spatial variation in N deposition is correlated with Nitrogen
+deposition. If we apply that model the estimated effect of N depositon
+on total species richness is increasing over the study period (right
+panel of \ref{fig:figconsequences}).
 
@@
 
-xxx
+\begin{itemize}
+\item
+  Any empty space that could be caused by any disturbance that let to
+  the local disappearance of species is likely to be fiellied by
+  eutrophic species. Disturbance depends on the site, while colonization
+  depends on species characteristics.
+\item
+  Altough N deposition considerabely declinded between 2005 and 2015, we
+  could not detect major shifts in plant community structure during the
+  same time period.
+\item
+  Eutrophic species have rather high local survival across the entire
+  deposition gradient, while oligotrophic species have much reduced
+  local survival at high N deposition. This suggests that it takes much
+  more time to replace eutrophic by oligotrophic species than replacing
+  oligotrophic by eutrophic species.
+\item
+  Our data on colonization and local survival (i.e.~temporal variation)
+  confirm the empirical critical loads that we infered from anlysing
+  spatial co-variation of N deposition and species richness.
+\item
+  Local survival is higher for low temperature plants --\textgreater{}
+  This could explain the decrease in cumminity change along elevation.
+  This could also explain the differences at mount summets were space
+  was empty in the beginning.
+\item
+  Climatic effects are more likely to be reversed thant effects due to
+  fertilization.
+\end{itemize}
 

### 2018-05-13 R/1366_Data_processing.R

@@ -43,10 +43,10 @@ surv <- tbl(db, "KD_Z9") %>%
   filter(!is.na(match(aID_STAO, sel)) & !is.na(yearPl) & yearP >= 2003 & yearP <= 2017) %>% 
   transmute(aID_KD = aID_KD, aID_STAO = aID_STAO, yearP = yearP, yearPl = yearPl, Visit = as.integer(floor((yearP - 1998) / 5))) %>% 
   left_join(tbl(db, "RAUMDATEN_Z9") %>% 
-              dplyr::select(aID_STAO, Hoehe, Neig, Expos, AS09_72, AS09_17, CACO3, Vernass, BGR_6, 
-                            bio1, bio12, bio13, bio14), copy = TRUE) %>% 
+              transmute(aID_STAO = aID_STAO, Elevation = Hoehe, Inclination = Neig, 
+                       Temperature = bio1, Precipitation = bio12), copy = TRUE) %>% 
   left_join(tbl(db, "RAUMDATEN_Z9_NDEP") %>% 
-              dplyr::select(aID_STAO, NTOT2007), copy = TRUE)
+              transmute(aID_STAO = aID_STAO, NTOT = NTOT2010b), copy = TRUE)
   
 # Add community measures to 'surv'
 surv <- surv %>% left_join(
@@ -92,19 +92,95 @@ sites <- data.frame(aID_STAO = as.integer(sel)) %>%
               transmute(aID_STAO = aID_STAO, Elevation = Hoehe, Inclination = Neig, 
                             Temperature = bio1, Precipitation = bio12), copy = TRUE) %>% 
   left_join(tbl(db, "RAUMDATEN_Z9_NDEP") %>% 
-              transmute(aID_STAO = aID_STAO, NDEP2010 = NTOT2010b), copy = TRUE)
+              transmute(aID_STAO = aID_STAO, NTOT = NTOT2010b), copy = TRUE)
+
+# for(i in 1:nrow(sites)) {
+#   sites[i, "Turnover1"] <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit <= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", 
+#       listin = TRUE, listout = TRUE)$cocogaston %>% mean
[…]

### 2018-05-13 R/1366_Settings.R

@@ -1,0 +1,101 @@
+#======================================================================================================
+#
+# Plant community change in mountain hay meadows
+#
+# Settings for manuscript
+#
+#======================================================================================================
+
+rm(list=ls(all=TRUE))
+
+
+#------------------------------------------------------------------------------------------------------
+# Libraries
+#------------------------------------------------------------------------------------------------------
+library(knitr)
+library(kableExtra)
+library(xtable)
+library(tidyverse)
+library(arm)
+library(nlme)
+library(simba)
+
+#------------------------------------------------------------------------------------------------------
+# Load data
+#------------------------------------------------------------------------------------------------------
+load("RData/surv.RData")
+load("RData/pl.RData")
+load("RData/sites.RData")
+load("RData/coldat.RData")
+load("RData/survdat.RData")
+
+#------------------------------------------------------------------------------------------------------
+# Settings 
+#------------------------------------------------------------------------------------------------------
+# Knitr-Settings
+opts_chunk$set(echo = FALSE, hide = TRUE, cache = TRUE, warning = FALSE, message = FALSE,
+               fig.asp = .4, fig.width = 8, out.width = "100%")
+
+# Darstellung von R-Berechnungen innerhalb von Text
+inline_hook <- function(x) {
+  if (is.numeric(x)) {
+    x <- format(x, nsmall = 2, digits = 2)
+  }
+  x
+}
+knit_hooks$set(inline = inline_hook)
+
+# Plot settings
+theme_set(theme_classic() + 
+            theme(legend.position="bottom") + 
+            theme(legend.title = element_blank()))
+pd <- position_dodge(0.2) 
+tcol <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", 
+          "#D55E00", "#CC79A7")
+
+# Remove comment for xtable
+options(xtable.comment = FALSE)
+
+# Number of posterior samples
+nsim <- 100
+
+#------------------------------------------------------------------------------------------------------
+# Hilfsfunktionen
+#------------------------------------------------------------------------------------------------------
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+  
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+  
+  numPlots = length(plots)
+  
[…]

### 2018-05-13 R/1366_gains_losses.R

@@ -1,57 +1,0 @@
-#======================================================================================================
-#
-# Plant community change in mountain hay meadows
-#
-# Compare gains and losses
-#
-#======================================================================================================
-
-rm(list=ls(all=TRUE))
-
-#------------------------------------------------------------------------------------------------------
-# Settings and load data
-#------------------------------------------------------------------------------------------------------
-# Libraries
-library(tidyverse)
-library(arm)
-
-# Load data
-load("RData/pl.RData")
-load("RData/sites.RData")
-
-#------------------------------------------------------------------------------------------------------
-# Compile absence data for all species x site combinations
-#------------------------------------------------------------------------------------------------------
-d <- expand.grid(aID_STAO = sites$aID_STAO, aID_SP = unique(pl$aID_SP)) %>% as.tibble() %>% 
-  left_join(pl %>% group_by(aID_SP) %>% summarise(T = min(T), F = min(F), N = min(N), L = min(L))) %>% 
-  left_join(pl %>% filter(yearP >= 2003 & yearP<= 2007) %>% 
-              transmute(aID_STAO = aID_STAO, aID_SP = aID_SP, Occ1 = Occ)) %>% 
-  left_join(pl %>% filter(yearP >= 2008 & yearP <= 2012) %>% 
-              transmute(aID_STAO = aID_STAO, aID_SP = aID_SP, Occ2 = Occ)) %>% 
-  left_join(pl %>% filter(yearP >= 2013 & yearP <= 2017) %>% 
-              transmute(aID_STAO = aID_STAO, aID_SP = aID_SP, Occ3 = Occ)) %>% 
-  replace_na(replace = list(Occ1 = 0, Occ2 = 0, Occ3 = 0)) %>% 
-  left_join(sites) 
-
-#------------------------------------------------------------------------------------------------------
-# Analyse colonization
-#------------------------------------------------------------------------------------------------------
-summary(glmer(Occ2 ~ T + F + N + L + (1|aID_SP) + (1|aID_STAO), data = d %>% filter(Occ1 == 0), family = binomial))
-summary(glmer(Occ3 ~ T + F + N + L + (1|aID_SP) + (1|aID_STAO), data = d %>% filter(Occ2 == 0), family = binomial))
-
-#------------------------------------------------------------------------------------------------------
[…]

### 2018-05-20 R/1366_Data_processing.R
Hide coordination in coordID.

@@ -22,7 +22,7 @@ db <- src_sqlite(path = "~/Documents/Dropbox/DB_BDM.db", create = FALSE)
 # Selection of sites
 #------------------------------------------------------------------------------------------------------
 # Import list with sites
-sel <- read_delim("Data/1366 K_Standort.csv", delim = ";") %>% 
+sel <- read_delim("Data_protected/1366 K_Standort.csv", delim = ";") %>% 
   filter(E23_1366 == "ja") %>% 
   transmute(aID_STAO = aIdStaoZ9)
 
@@ -93,13 +93,6 @@ sites <- data.frame(aID_STAO = as.integer(sel)) %>%
                             Temperature = bio1, Precipitation = bio12), copy = TRUE) %>% 
   left_join(tbl(db, "RAUMDATEN_Z9_NDEP") %>% 
               transmute(aID_STAO = aID_STAO, NTOT = NTOT2010b), copy = TRUE)
-
-# for(i in 1:nrow(sites)) {
-#   sites[i, "Turnover1"] <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit <= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", 
-#       listin = TRUE, listout = TRUE)$cocogaston %>% mean
-#   sites[i, "Turnover2"] <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit >= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", 
-#                                listin = TRUE, listout = TRUE)$cocogaston %>% mean
-# }
 
 for(i in 1:nrow(sites)) {
   tt <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit <= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", listin = TRUE, listout = TRUE)
@@ -176,11 +169,24 @@ survdat <- rbind(
   filter(!is.na(T) & !is.na(F) & !is.na(N) & !is.na(L))
 
 #------------------------------------------------------------------------------------------------------
-# Save data
+# Change siteID and save data
 #------------------------------------------------------------------------------------------------------
+siteIDs <- data.frame(
+  aID_STAO = sites$aID_STAO,
[…]

### 2018-05-20 R/Settings.R
Start documenting repository.

@@ -8,7 +8,6 @@
 
 rm(list=ls(all=TRUE))
 
-
 #------------------------------------------------------------------------------------------------------
 # Libraries
 #------------------------------------------------------------------------------------------------------

### 2018-05-21 Manuscript.Rmd
Hide coordination in coordID.
Version with a lot of experimenting junk in the Manuscript. For the next commit I will need to tidy up.

@@
 ---
-title: xxx
+title: No sign of recovery 
 preprint: false
@@
 abstract: >
-  xxx
+  Nitrogen (N) deposition is a major threat to biodiversity of many habitats.
+The recent intro- duction of cleaner technologies in Switzerland has let to reductions in the emissions of nitro- gen oxides, with affiliated decrease in N-deposition.
+We infered different drivers of community change (i.e.
+Nitrogen deposition, climate warming, land-use change) in Swiss mountain hay meadows.
+The data were obtained from the Swiss biodiversity monitoring.
+  Species gain and losses were best explainde by differences in Ellenberg value for temperature suggesting that climate warming was the most important driver of community change in Mountain hay meadows.
+
 header-includes: >
   \usepackage{lipsum}
-bibliography: Library.bib
+bibliography: Bibliography.bib
 output: 
@@
 ```{r settings, echo = FALSE, message=FALSE, cache=FALSE}
-source("R/1366_Settings.R")
+source("R/Settings.R")
 ```
@@
 # Introduction {-}
+Nitrogen (N) deposition is a major threat to biodiversity.
+The recent introduction of cleaner technologies has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in N-deposition in many parts of Europe.
+However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovey of extant plant communities.
+
+One useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both immigration and extinction, often in a closed range of values (Hillebrand_et_al-2018.pdf).
+
+Here we infered mountain hay meadows in Switzerland. Explain why mountain hay meadows are important. Also explain other threats to mountain hay meadows (climate change, land-use change).
 
@@
 
-### Community measures
+### Community measures {-}
 
@@
 
-The different measures of total community structure suggested that plant communities of mountain hay meadow were stable between 2003 and 2017 (Table \@ref(temporaltrends)): for each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity, average Ellenberg values for temperature, nutrients, light and huminity, and average of species' canopy height, specific leaf area and seed mass did not vary much among the three sampling periods.
-For all measures, the average temperal trend per site did not differ significantly from a constant trend.
+The different measures of total community structure suggested that plant communities in mountain hay meadows were stable between 2003 and 2017 (Table \@ref(temporaltrends)): for each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity, average Ellenberg values for temperature, nutrients, light and huminity, and average of species' canopy height, specific leaf area and seed mass did not vary much among the three sampling periods.
+For all measures, all p-values of the average temperal trends per site were $>0.05$ and only in the case of the average temperature value of the species the p-value was $<0.1$.
 Note that beta- and gamma-diversity are note available for single sites and thus mixed models to calculate the p-value could not be applied.
@@
 
-The temporal stability as inferred from the community measures were, however, in contrast to the large  temporal turnover of recorded species.
-The percentage $\pm$ SD of species that differ between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))` and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`.
-The decrease in the temporal turnover between the first and second survey to the temporal turnover between the second and third survey was statistically significant (Binomial generalized linear model; effect size = `r coef(mod)[2]`; p = `r summary(mod)$coefficients["period", "Pr(>|z|)"]`).
+The temporal stability as inferred from the community measures were, however, in contrast to the large  observed temporal turnover of recorded species.
+The average percentage $\pm$ SD of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
+There is some evidence that this slight decrease in the temporal turnover from the first and second survey to the temporal turnover of the second and third survey is systematic and was not entirly caused by change (Binomial generalized linear model; effect size = `r coef(mod)[2]`; p = `r summary(mod)$coefficients["period", "Pr(>|z|)"]`).
 
@@
 ```
-
-This spatial turnover might be partly explained by species that remained undetected in one of the surveys, but it might also be the result of species that newly colonized sites (species gains) and species that truely disappeared from sites (species losses).
-Spatial turnover will be high if the probability a site is colonized by a species is high (i.e.
-high colonization probability) and/or if the probability that a species survives at a site between two time points is low (i.e.
-low local survival).
-We found that the variation in local survival and colonization, was largely due to differences between species and only to a lesser extend to the differences between locations: while the variance of the species random effect in a binomial linear mixed model for local survival was about `r getratio(unlist(modls$varcor))` times larger than the variance of the sampling site random effect, in the model for colonization the variance among species was even `r getratio(unlist(modcol$varcor))` times bigger than the variance among sites.
 
@@
 ```
+
 ```{r resdifeffects}
@@
 
-We thus inferred if we could explain differences in colonization probability and local survival with Ellengerg species values for temperature, humidity, nutrient or light.
-We found that both colonization probability and local survival probability was best explained by temperature (i.e.
-highest colonization or local survival probability of species with an Ellenberg value for Temperature that corresponds to the annual average temperature of the site) followed by nutrient availability (i.e.
-highest colonization or local survival probability of species with an Ellenberg value for nutrient availability that corresponds to the annual N deposition rate estimated for the site; Table \@ref(tab:resdifeffects)).
+ Spatial turnover will be high if the probability a site is colonized by a species is high (i.e.
+high colonization probability) and/or if the probability that species survives at a site between two time points is low (i.e.
+low local survival).
+We found that the variation in local survival and colonization probability was largely due to differences between species and only to a lesser extend to the differences between locations: while the variance of the species random effect in a binomial linear mixed model for local survival was about `r getratio(unlist(modls$varcor))` times larger than the variance of the sampling site random effect, in the model for colonization the variance among species was even `r getratio(unlist(modcol$varcor))` times bigger than the variance among sites.
+
+ We inferred if we could explain differences in colonization and local survival with Ellenberg species values for temperature, humidity, nutrient or light (Table \@ref(tab:resdifeffects)).
+We found that both colonization probability and local survival probability was best explained by a model containing an interaction between species temperature and the average annual temperature of a site:  highest colonization probability and local survival were found  for the species with an Ellenberg value for Temperature that corresponded to the annual average temperature of the site (Fig.
+\@ref(fig:Teff))..
+While the colonization and local survival probability of cold living species (Ellenberg T = 2) declindes along the temperature gradient, the colonization and local survival probability of warm living species (Ellenberg T = 4) increases along the temperature gradient (Fig.
+\@ref(fig:Teff)).
+
+The second best model contained an interaction between species Ellenberg value for nutrient availability and the average annual N deposition rate of a site (Table \@ref(tab:resdifeffects)): the highest colonization probability and local survival were performed by species with an Ellenberg value for nutrient availability that corresponded to the annual N deposition rate of the site.
+While the colonization and local survival probability of oligotrophic species (Ellenberg N = 2) declindes along the N deposition gradient, the colonization and local survival probability of eutrophic species (N = 4) increases along the N deposition gradient (Fig.
+\@ref(fig:Neff)).
 
@@
 
-While the colonization and local survival probability of cold living species (Ellenberg T = 2) declindes along the temperature gradient, the colonization and local survival probability of warm living species (Ellenberg T = 4) increases along the temperature gradient (Fig.
-\@ref(fig:Teff)).
-Similarly for nutrient values: while the colonization and local survival probability of oligotrophic species (Ellenberg N = 2) declindes along the N deposition gradient, the colonization and local survival probability of eutrophic species (N = 4) increases along the N deposition gradient (Fig.
-\@ref(fig:Neff)).
-
 ```{r periodeffect}
@@
 
-For the temperature driven and nitrogen driven models in table \@ref(tab:resdifeffects) we also tested whether the strength of the effects differed from the first/second to the second/third survey period.
+For the temperature driven and nitrogen driven models in Table \@ref(tab:resdifeffects) we additionally tested whether the strength of the effects on colonization probability and local survival differed from the first/second to the second/third survey period.
 For the temperatur driven models we found that the differences how species with different Ellenberg T values responded to the average annual temperature of a site tended to decrease from the first/second study period to the second/third study period (colonization probability: period x Ellenberg T x annual mean temperature of site = `r fixef(modTcol)["factor(Period)2:Temperature:T"]`, p = `r summary(modTcol)$coefficients["factor(Period)2:Temperature:T", "Pr(>|z|)"]`; local survival: period x Ellenberg T x annual mean temperature of site = `r fixef(modTls)["factor(Period)2:Temperature:T"]`, p = `r summary(modTls)$coefficients["factor(Period)2:Temperature:T", "Pr(>|z|)"]`).
@@
 res <- rbind(
-  tibble(cor = map_dbl(tt, "corN"),
+  tibble(cor = map_dbl(tt, "corN") / tt[[1]]$corN,
               driver = "NTOT",
               year = 2003:2013),
-  tibble(cor = map_dbl(tt, "corT"),
+  tibble(cor = map_dbl(tt, "corT") / tt[[1]]$corT,
               driver = "Temperature",
               year = 2003:2013))
+
+tt <- rbind(surv %>% transmute(CI = T, yr = yearPl - 2010, Type = "T", aID_STAO = aID_STAO),
+            surv %>% transmute(CI = N, yr = yearPl - 2010, Type = "N", aID_STAO = aID_STAO))
+m1 <- lmer(CI ~ Type * yr + (yr|aID_STAO), data = tt)
+m2 <- lmer(CI ~ Type + yr + (yr|aID_STAO), data = tt)
+anova(m1, m2)
+
+m3 <- lmer(CI ~ Type + yr + (yr|aID_STAO), data = tt)
+
 p1 <- ggplot(res, aes(x = year, y = cor, col = driver)) +
@@
   geom_line() +
-  ylim(0.5, 1)
+  ylim(0.95, 1.05)
 res$yr <- res$year - 2010 
@@
 # Effect on N deposition model
+surv$yr <- surv$yearPl - 2010
+summary(glmer(SR ~ yr * NTOT * Temperature + (1|aID_STAO), data = surv, family = poisson))
+
 makemod <- function(year) {
@@
 }
-
 getestim <- function(x) {
   list(meanN = mean(x[, "NTOT"]),
-       lowN = quantile(x[, "NTOT"], probs = 0.25),
-       upN = quantile(x[, "NTOT"], probs = 0.75),
+       lowN = quantile(x[, "NTOT"], probs = 0.025),
+       upN = quantile(x[, "NTOT"], probs = 0.975),
        meanT = mean(x[, "Temperature"]),
@@
               period = 2003:2013))
-
+pd <- position_dodge(0.5)
 p2 <- ggplot(res, aes(x = period, y = mean, col = factor)) +
-  geom_errorbar(aes(ymin = low, ymax = up), width = 0.2) +
-  geom_point() +
-  geom_line() +
+  geom_errorbar(aes(ymin = low, ymax = up), width = 0.5, position=pd) +
+  geom_point(position=pd) +
+  geom_line(position=pd) +
   geom_abline(intercept = 0, slope = 0) +
-  ylim(-0.1, 0.015)
+  ylim(-0.2, 0.05)
 multiplot(p1, p2, cols = 2)
+ # summary(lm(mean ~ factor * period, data = res))
 ```
@@
 
+```{r}
+# Calculate trend for all species
+gettrend <- function(s) {
+  tt <- surv %>% left_join(pl %>% filter(aID_SP == s) %>% dplyr::select(aID_KD, Occ)) %>% 
+    replace_na(list(Occ = 0))
+  trend <- NA
+  try(trend <- glmer(Occ ~ yr + (1|aID_STAO), data = tt, family = binomial) %>% 
+    tidy() %>% 
+    filter(term == "yr") %>% 
+    pull(estimate))
+trend
+}
+res <- pl %>% group_by(aID_SP, T, F, N, L) %>% 
+  summarise(Nsites = n_distinct(aID_STAO),
+            Nrecords = n())
+res$Trend <- map_dbl(res$aID_SP, gettrend)
+
+# Is trend of species related to Ellenberg values
+gls(Trend ~ N, data = res) %>% 
+  tidy
+
+library("nlme")
+vf <- varExp(form=~Nrecords)
+gls(Trend ~ T, data = res %>% filter(!is.na(T)), weights=vf) %>% summary
+gls(Trend ~ F, data = res %>% filter(!is.na(F)), weights=vf) %>% summary
+gls(Trend ~ N, data = res %>% filter(!is.na(N)), weights=vf) %>% summary
+gls(Trend ~ L, data = res %>% filter(!is.na(L)), weights=vf) %>% summary
+
+cor.test(res$T, res$N)
+```
+
+```{r}
+# Borrowoing ideas from community assembly
+library(geometry)
+
+getFric <- function(x) {
+  tt <- pl %>% 
+  filter(!is.na(T) & !is.na(F) & !is.na(N) & !is.na(L)) %>% 
+  filter(aID_KD == x) %>% dplyr::select(T, F, N, L) 
+  geometry::convhulln(tt, "FA")$vol
+}
+
+surv$FRic <- map_dbl(surv$aID_KD, getFric)
+
+summary(lmer(FRic ~ yr + (yr|aID_STAO), data = surv))
+tapply(surv$FRic, surv$Visit, mean)
+
+ex1 <- dbFD(dummy$trait, dummy$abun)
+ex1
+```
+
+--> Es gibt keine Einheitlichen Entwicklungen der Communities. Das heisst aber nicht, dass es keine systematischen Veränderungen gibt!
+
+```{r}
+x <- 28
+
+getsamples <- function(s, nspec, d) {
+  sample(d$T, nspec, replace = FALSE) %>% mean(na.rm = TRUE)
+}
+
+getsim <- function(x) {
+  tt <- pl %>% filter(aID_STAO == x) %>% 
+    group_by(aID_SP, T, F, N, L) %>% 
+    summarise(Occ1 = sum(Visit == 1),
+              Occ2 = sum(Visit == 2),
+              Occ3 = sum(Visit == 3))
+  gain <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
+  lost <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
+  tmpgain <- map_dbl(1:nsim, getsamples, nspec = sum(gain), d = tt)
+  tmplost <- map_dbl(1:nsim, getsamples, nspec = sum(lost), d = tt)
+  list(
+    gain = mean(tt$T[gain], na.rm = TRUE),
+    gainran = mean(tmpgain, na.rm = TRUE),
+    gainransd = sd(tmpgain, na.rm = TRUE),
+    lost = mean(tt$T[lost], na.rm = TRUE),
+    lostran = mean(tmplost, na.rm = TRUE),
+    lostransd = sd(tmplost, na.rm = TRUE))
+}
+
+tt <- map(sites$aID_STAO, getsim)
+
+sites$ngain <-  tt %>% map_dbl(function(x) (x$gain - x$gainran))
+sites$nlost <-  tt %>% map_dbl(function(x) (x$lost - x$lostran))
+
+t.test(sites$ngain)
+t.test(sites$nlost)
+
+lm(ngain ~ Inclination + Temperature + Precipitation  + NTOT, data = sites) %>% 
+  tidy
+
+plot(sites$ngain, sites$nlost)
+
+tapply()
+
+```
+
 # Discussion {-}
@@
 Disturbance depends on the site, while colonization depends on species characteristics.
+
+- The rather large spatial turnover might be partly explained by species that remained undetected in one of the surveys, but it might also be the result of species that newly colonized sites (species gains) and species that truely disappeared from sites (species losses).
 
@@
 
+### Space for time substitution
+Often observational studies infer the change of plant diversity along a gradient of N deposi- tion.
+Thus, they infer how the spatial variation in species richness is related to N deposition and assume that this spatial variation in species richness arose because over time some areas lost more species than others because they chronically experienced higher N deposition.
+Alt- hough there is evidence supporting the use of such a 'space for time substitution' for detecting the effects of N deposition on plant diversity (Stevens et al.
+2010), they can not replace stud- ies that relate temporal patterns in species with N deposition (De Schrijver et al.
+2011).
+While recovery of acidified surface waters has been well investigated (De Vries et al.
+2015), there are only a limited number of studies inferring temporal trends of plant species diversity relat- ed to varying amounts of N-deposition.
+Storkey et al.
+(2015) demonstrated a positive response of biodiversity to reducing N addition from either atmospheric pollution or fertilizers in the Park Grass Experiment: «The proportion of legumes, species richness and diversity increased across the experiment between 1991 and 2012 as N-deposition declined».
+For forest floor vegetation in permanent plots across Europe the exceedance of critical loads of N over a peri- od from 9 to 42 years had negative effects on the cover of oligotrophic plant species, i.e spe- cies that prefer nutrient-poor soils, although species richness remained constant (Dirnböck et al.
+2014).
+Another example of recovery in eutrophicated habitats gives the recovery of species richness in previously fertilized plots (Clark and Tilman 2008).
+In this study, the recorded recovery in species richness within one or two decade was likely due to the species rich vege- tation surrounding the experimental plots, from where immigration was easily feasible.
+
+> Hier sagen, dass der CRL identisch wie der CRL von räumlichen Zusammenhängen ist.
+
 # Conclusions {-}
@@
 # Acknowledgements {-}
-xxx
+We thank the dedicated and qualified botanists who conducted fieldwork.
+The Swiss Federal Office for the Environment (FOEN) kindly provided biodiversity monitoring data and topographic data.
+This work was supported by the FOEN, the Swiss National Science Foundation (grant no.
+31003A_156294), the Swiss Association Pro Petite Camargue Alsacienne, the Fondation de bienfaisance Jeanne Lovioz, and the MAVA Foundation.
 
 # References {-}
+
+```{r}
+# The MS was build with the following R sesstion:
+writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
+```

### 2018-05-21 Manuscript.tex
Hide coordination in coordID.
Version with a lot of experimenting junk in the Manuscript. For the next commit I will need to tidy up.

@@
 
-\title{xxx}
+\title{Plant community changes in mountain hay meadows from 2003 to 2017}
 
@@
 \begin{abstract}
-xxx
+Nitrogen (N) deposition is a major threat to biodiversity of many
+habitats. The recent intro- duction of cleaner technologies in
+Switzerland has let to reductions in the emissions of nitro- gen oxides,
+with affiliated decrease in N-deposition. We infered different drivers
+of community change (i.e.~Nitrogen deposition, climate warming, land-use
+change) in Swiss mountain hay meadows. The data were obtained from the
+Swiss biodiversity monitoring.
 % Dummy abstract text.
@@
 
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[section]
+\theoremstyle{definition}
+\newtheorem{exercise}{Exercise}[section]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+\newtheorem*{solution}{Solution}
 \begin{document}
@@
 
-\begin{verbatim}
-## Warning: package 'kableExtra' was built under R version 3.4.4
-\end{verbatim}
-
 \section*{Introduction}\label{introduction}
 \addcontentsline{toc}{section}{Introduction}
+
+It is an open question whether and how fast the reduction in N
+deposition rates will lead to the recovey of plant communities.
+
+Here we infered mountain hay meadows in Switzerland. Explain why
+mountain hay meadows are important.
 
@@
 
-\subsubsection{Community measures}\label{community-measures}
+\subsubsection*{Community measures}\label{community-measures}
+\addcontentsline{toc}{subsubsection}{Community measures}
 
@@
 Alpha-diversity & 46.36 & 46.72 & 46.45 & 0.002 & 0.896 \\ 
-  Beta-diversity & 0.68 & 0.65 & 0.65 &  &  \\ 
+  Beta-diversity & 0.74 & 0.72 & 0.55 &  &  \\ 
   Gamma-Diversity & 517 & 529 & 517 &  &  \\ 
@@
 The different measures of total community structure suggested that plant
-communities of mountain hay meadow were stable between 2003 and 2017
+communities in mountain hay meadows were stable between 2003 and 2017
 (Table \ref{temporaltrends}): for each of the three 5-year survey
@@
 average of species' canopy height, specific leaf area and seed mass did
-not vary much among the three sampling periods. For all measures, the
-average temperal trend per site did not differ significantly from a
-constant trend. Note that beta- and gamma-diversity are note available
-for single sites and thus mixed models to calculate the p-value could
-not be applied.
+not vary much among the three sampling periods. For all measures, all
+p-values of the average temperal trends per site were \(>0.05\) and only
+in the case of the average temperature value of the species the p-value
+was \(<0.1\). Note that beta- and gamma-diversity are note available for
+single sites and thus mixed models to calculate the p-value could not be
+applied.
 
 The temporal stability as inferred from the community measures were,
-however, in contrast to the large temporal turnover of recorded species.
-The percentage \(\pm\) SD of species that differ between the first and
-second survey at a site was 37.55 \(\pm\) 10.46 and the percentage of
-species that differ between the second and third survey was 35.63
-\(\pm\) 10.33. The decrease in the temporal turnover between the first
-and second survey to the temporal turnover between the second and third
-survey was statistically significant (Binomial generalized linear model;
+however, in contrast to the large observed temporal turnover of recorded
+species. The average percentage \(\pm\) SD of species that differed
+between the first and second survey at a site was NaN \(\pm\) NA\% and
+the percentage of species that differ between the second and third
+survey was NaN \(\pm\) NA\%. There is some evidence that this slight
+decrease in the temporal turnover from the first and second survey to
+the temporal turnover of the second and third survey is systematic and
+was not entirly caused by change (Binomial generalized linear model;
 effect size = -0.078; p = 0.027).
-
-This spatial turnover might be partly explained by species that remained
-undetected in one of the surveys, but it might also be the result of
-species that newly colonized sites (species gains) and species that
-truely disappeared from sites (species losses). Spatial turnover will be
-high if the probability a site is colonized by a species is high
-(i.e.~high colonization probability) and/or if the probability that a
-species survives at a site between two time points is low (i.e.~low
-local survival). We found that the variation in local survival and
-colonization, was largely due to differences between species and only to
-a lesser extend to the differences between locations: while the variance
-of the species random effect in a binomial linear mixed model for local
-survival was about 4.46 times larger than the variance of the sampling
-site random effect, in the model for colonization the variance among
-species was even 9.15 times bigger than the variance among sites.
 
@@
 
-We thus inferred if we could explain differences in colonization
-probability and local survival with Ellengerg species values for
-temperature, humidity, nutrient or light. We found that both
+Spatial turnover will be high if the probability a site is colonized by
+a species is high (i.e.~high colonization probability) and/or if the
+probability that species survives at a site between two time points is
+low (i.e.~low local survival). We found that the variation in local
+survival and colonization probability was largely due to differences
+between species and only to a lesser extend to the differences between
+locations: while the variance of the species random effect in a binomial
+linear mixed model for local survival was about 4.46 times larger than
+the variance of the sampling site random effect, in the model for
+colonization the variance among species was even 9.15 times bigger than
+the variance among sites.
+
+We inferred if we could explain differences in colonization and local
+survival with Ellenberg species values for temperature, humidity,
+nutrient or light (Table \ref{tab:resdifeffects}). We found that both
 colonization probability and local survival probability was best
-explained by temperature (i.e.~highest colonization or local survival
-probability of species with an Ellenberg value for Temperature that
-corresponds to the annual average temperature of the site) followed by
-nutrient availability (i.e.~highest colonization or local survival
-probability of species with an Ellenberg value for nutrient availability
-that corresponds to the annual N deposition rate estimated for the site;
-Table \ref{tab:resdifeffects}).
+explained by a model containing an interaction between species
+temperature and the average annual temperature of a site: highest
+colonization probability and local survival were found for the species
+with an Ellenberg value for Temperature that corresponded to the annual
+average temperature of the site (Fig. \ref{fig:Teff}).. While the
+colonization and local survival probability of cold living species
+(Ellenberg T = 2) declindes along the temperature gradient, the
+colonization and local survival probability of warm living species
+(Ellenberg T = 4) increases along the temperature gradient (Fig.
+\ref{fig:Teff}).
+
+The second best model contained an interaction between species Ellenberg
+value for nutrient availability and the average annual N deposition rate
+of a site (Table \ref{tab:resdifeffects}): the highest colonization
+probability and local survival were performed by species with an
+Ellenberg value for nutrient availability that corresponded to the
+annual N deposition rate of the site. While the colonization and local
+survival probability of oligotrophic species (Ellenberg N = 2) declindes
+along the N deposition gradient, the colonization and local survival
+probability of eutrophic species (N = 4) increases along the N
+deposition gradient (Fig. \ref{fig:Neff}).
 
@@
 
-While the colonization and local survival probability of cold living
-species (Ellenberg T = 2) declindes along the temperature gradient, the
-colonization and local survival probability of warm living species
-(Ellenberg T = 4) increases along the temperature gradient (Fig.
-\ref{fig:Teff}). Similarly for nutrient values: while the colonization
-and local survival probability of oligotrophic species (Ellenberg N = 2)
-declindes along the N deposition gradient, the colonization and local
-survival probability of eutrophic species (N = 4) increases along the N
-deposition gradient (Fig. \ref{fig:Neff}).
+For the temperature driven and nitrogen driven models in Table
+\ref{tab:resdifeffects} we additionally tested whether the strength of
+the effects on colonization probability and local survival differed from
+the first/second to the second/third survey period. For the temperatur
+driven models we found that the differences how species with different
+Ellenberg T values responded to the average annual temperature of a site
+tended to decrease from the first/second study period to the
+second/third study period (colonization probability: period x Ellenberg
+T x annual mean temperature of site = -0.0098, p = 0.70; local survival:
+period x Ellenberg T x annual mean temperature of site = -0.072, p =
+0.051). In contrast, however, the differences how species with different
+Ellenberg N values responded to the average N deposition of a site
+tended to increase from the first/second study period to the
+second/third study period (colonization probability: period x Ellenberg
+N x N deposition at site = 0.025, p = 0.74; local survival: period x
+Ellenberg N x N deposition at site = 0.075, p = 0.50).
 
-For the temperature driven and nitrogen driven models in table
-\ref{tab:resdifeffects} we also tested whether the strength of the
-effects differed from the first/second to the second/third survey
-period. For the temperatur driven models we found that the differences
-how species with different Ellenberg T values responded to the average
-annual temperature of a site tended to decrease from the first/second
-study period to the second/third study period (colonization probability:
-period x Ellenberg T x annual mean temperature of site = -0.0098, p =
-0.70; local survival: period x Ellenberg T x annual mean temperature of
-site = -0.072, p = 0.051). In contrast, however, the differences how
-species with different Ellenberg N values responded to the average N
-deposition of a site tended to increase from the first/second study
-period to the second/third study period (colonization probability:
-period x Ellenberg N x N deposition at site = 0.025, p = 0.74; local
-survival: period x Ellenberg N x N deposition at site = 0.075, p =
-0.50).
+\begin{verbatim}
+## Data: tt
+## Models:
+## m2: CI ~ Type + yr + (yr | aID_STAO)
+## m1: CI ~ Type * yr + (yr | aID_STAO)
+##    Df     AIC     BIC logLik deviance  Chisq Chi Df Pr(>Chisq)  
+## m2  7 -246.90 -214.28 130.45  -260.90                           
+## m1  8 -248.44 -211.16 132.22  -264.44 3.5419      1    0.05984 .
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+\end{verbatim}
+
+\begin{verbatim}
+## Generalized linear mixed model fit by maximum likelihood (Laplace
+##   Approximation) [glmerMod]
+##  Family: poisson  ( log )
+## Formula: SR ~ yr * NTOT * Temperature + (1 | aID_STAO)
+##    Data: surv
+## 
+##      AIC      BIC   logLik deviance df.resid 
+##   2718.9   2754.6  -1350.5   2700.9      381 
+## 
+## Scaled residuals: 
+##      Min       1Q   Median       3Q      Max 
+## -2.51161 -0.38950 -0.00348  0.37725  2.19699 
+## 
+## Random effects:
+##  Groups   Name        Variance Std.Dev.
+##  aID_STAO (Intercept) 0.03294  0.1815  
+## Number of obs: 390, groups:  aID_STAO, 130
+## 
+## Fixed effects:
+##                       Estimate Std. Error z value Pr(>|z|)    
+## (Intercept)          3.8808142  0.0268047 144.781   <2e-16 ***
+## yr                   0.0029242  0.0027364   1.069   0.2852    
+## NTOT                -0.0221202  0.0580005  -0.381   0.7029    
+## Temperature         -0.0259552  0.0112875  -2.299   0.0215 *  
+## yr:NTOT             -0.0079678  0.0059621  -1.336   0.1814    
+## yr:Temperature       0.0002304  0.0011735   0.196   0.8444    
+## NTOT:Temperature    -0.0148305  0.0154851  -0.958   0.3382    
+## yr:NTOT:Temperature  0.0016841  0.0015971   1.054   0.2917    
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##             (Intr) yr     NTOT   Tmprtr yr:NTOT yr:Tmp NTOT:T
+## yr           0.019                                           
+## NTOT        -0.612 -0.023                                    
+## Temperature  0.249  0.014 -0.366                             
+## yr:NTOT     -0.024 -0.626  0.044  0.005                      
+## yr:Tempertr  0.014  0.310  0.005  0.030 -0.393               
+## NTOT:Tmprtr  0.274 -0.002 -0.798 -0.063 -0.039  -0.045       
+## yr:NTOT:Tmp  0.000  0.255 -0.040 -0.046 -0.776  -0.064  0.089
+## convergence code: 0
+## Model failed to converge with max|grad| = 0.00224997 (tol = 0.001, component 1)
+## Model is nearly unidentifiable: very large eigenvalue
+##  - Rescale variables?
+\end{verbatim}
 
@@
 temperature correlation x temporal trend of nitrogen correlation =
--0.0015, p = 0.067), however the change in the correlation over the
-study period remained week and the temperature correlation remained
-stronger than the nitrogen correlation over the entire study period
-(Left panel of Fig. \ref{fig:figconsequences}). More important, however,
-was the effect on our model that we used to infer whether and how strong
-the spatial variation in N deposition is correlated with Nitrogen
+-0.0018, p = 0.16), however the change in the correlation over the study
+period remained week and the temperature correlation remained stronger
+than the nitrogen correlation over the entire study period (Left panel
+of Fig. \ref{fig:figconsequences}). More important, however, was the
+effect on our model that we used to infer whether and how strong the
+spatial variation in N deposition is correlated with Nitrogen
 deposition. If we apply that model the estimated effect of N depositon
@@
 \begin{itemize}
+\tightlist
 \item
@@
   depends on species characteristics.
+\end{itemize}
+
+-The rather large spatial turnover might be partly explained by species
+that remained undetected in one of the surveys, but it might also be the
+result of species that newly colonized sites (species gains) and species
+that truely disappeared from sites (species losses).
+
+\begin{itemize}
 \item
@@
 
+\subsubsection{Space for time
+substitution}\label{space-for-time-substitution}
+
+Often observational studies infer the change of plant diversity along a
+gradient of N deposi- tion. Thus, they infer how the spatial variation
+in species richness is related to N deposition and assume that this
+spatial variation in species richness arose because over time some areas
+lost more species than others because they chronically experienced
+higher N deposition. Alt- hough there is evidence supporting the use of
+such a `space for time substitution' for detecting the effects of N
+deposition on plant diversity (Stevens et al. 2010), they can not
+replace stud- ies that relate temporal patterns in species with N
+deposition (De Schrijver et al. 2011). While recovery of acidified
+surface waters has been well investigated (De Vries et al. 2015), there
+are only a limited number of studies inferring temporal trends of plant
+species diversity relat- ed to varying amounts of N-deposition. Storkey
+et al. (2015) demonstrated a positive response of biodiversity to
+reducing N addition from either atmospheric pollution or fertilizers in
+the Park Grass Experiment: «The proportion of legumes, species richness
+and diversity increased across the experiment between 1991 and 2012 as
+N-deposition declined». For forest floor vegetation in permanent plots
+across Europe the exceedance of critical loads of N over a peri- od from
+9 to 42 years had negative effects on the cover of oligotrophic plant
+species, i.e spe- cies that prefer nutrient-poor soils, although species
+richness remained constant (Dirnböck et al. 2014). Another example of
+recovery in eutrophicated habitats gives the recovery of species
+richness in previously fertilized plots (Clark and Tilman 2008). In this
+study, the recorded recovery in species richness within one or two
+decade was likely due to the species rich vege- tation surrounding the
+experimental plots, from where immigration was easily feasible.
+
+\begin{quote}
+Hier sagen, dass der CRL identisch wie der CRL von räumlichen
+Zusammenhängen ist.
+\end{quote}
+
 \section*{Conclusions}\label{conclusions}
@@
 
-xxx
+We thank the dedicated and qualified botanists who conducted fieldwork.
+The Swiss Federal Office for the Environment (FOEN) kindly provided
+biodiversity monitoring data and topographic data. This work was
+supported by the FOEN, the Swiss National Science Foundation (grant no.
+31003A\_156294), the Swiss Association Pro Petite Camargue Alsacienne,
+the Fondation de bienfaisance Jeanne Lovioz, and the MAVA Foundation.
 

### 2018-05-21 R/Settings.R

@@ -15,6 +15,7 @@ library(knitr)
 library(kableExtra)
 library(xtable)
 library(tidyverse)
+library(broom)
 library(arm)
 library(nlme)
 library(simba)
@@ -96,5 +97,6 @@ multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
     }
   }
 }
+
 
 

### 2018-05-21 Manuscript.Rmd
Are new or extinct species a random selection from the species that were recorded at the site oder are them specific in terms of Ellenberg value?

@@
 ```{r}
-x <- 28
-
-getsamples <- function(s, nspec, d) {
-  sample(d$T, nspec, replace = FALSE) %>% mean(na.rm = TRUE)
+# Functions 
+getsamples <- function(s, nspec, d, EL) {
+  sample(d %>% pull(paste(EL)), nspec, replace = FALSE) %>% mean(na.rm = TRUE)
 }
 
-getsim <- function(x) {
+getsim <- function(x, EL, gain) {
   tt <- pl %>% filter(aID_STAO == x) %>% 
@@
               Occ3 = sum(Visit == 3))
-  gain <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
-  lost <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
-  tmpgain <- map_dbl(1:nsim, getsamples, nspec = sum(gain), d = tt)
-  tmplost <- map_dbl(1:nsim, getsamples, nspec = sum(lost), d = tt)
+  if(gain)  ausw <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
+  if(!gain) ausw <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
+  tmp <- map_dbl(1:nsim, getsamples, nspec = sum(ausw), d = tt, EL = EL)
   list(
-    gain = mean(tt$T[gain], na.rm = TRUE),
-    gainran = mean(tmpgain, na.rm = TRUE),
-    gainransd = sd(tmpgain, na.rm = TRUE),
-    lost = mean(tt$T[lost], na.rm = TRUE),
-    lostran = mean(tmplost, na.rm = TRUE),
-    lostransd = sd(tmplost, na.rm = TRUE))
+    nchange = sum(ausw),
+    meanreal = mean(tt[ausw, paste(EL)] %>% unlist, na.rm = TRUE),
+    mean = mean(tmp, na.rm = TRUE),
+    lower = quantile(tmp, probs = 0.025, na.rm = TRUE),
+    upper = quantile(tmp, probs = 0.975, na.rm = TRUE))
 }
 
-tt <- map(sites$aID_STAO, getsim)
+res <- sites %>% 
+  filter(NTOT < 4) %>% 
+  transmute(aID_STAO = aID_STAO, Sitemeasure = NTOT)
+tmp <- map(res$aID_STAO, getsim, EL = "N", gain = TRUE)
+res$meanreal <- tmp %>% map_dbl("meanreal")
+res$mean <- tmp %>% map_dbl("mean")
+res$lower <- tmp %>% map_dbl("lower")
+res$upper <- tmp %>% map_dbl("upper")
+ggplot(res, aes(x = Sitemeasure, y = meanreal)) +
+  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey") +
+  geom_point(shape = 4, colour = "black", cex = 2) +
+  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black", alpha = 0.4) +
+  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "grey", alpha = 0.4)
 
-sites$ngain <-  tt %>% map_dbl(function(x) (x$gain - x$gainran))
-sites$nlost <-  tt %>% map_dbl(function(x) (x$lost - x$lostran))
+res1 <- rbind(
+  res %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = mean, type = "Sample"),
+  res %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = meanreal, type = "Real"))
+mod1 <- lmer(mean ~ Sitemeasure * type + (1|aID_STAO), data = res1)
+mod2 <- lmer(mean ~ Sitemeasure + type + (1|aID_STAO), data = res1)
+anova(mod1, mod2)
+
+summary(glmer(SR ~ yr * NTOT + (yr|aID_STAO), data = surv, family = "poisson"))
+mod1 <- lmer(N ~ yr * NTOT + (yr|aID_STAO), data = surv)
+mod2 <- lmer(N ~ yr + NTOT + (yr|aID_STAO), data = surv)
+anova(mod1, mod2)
 

### 2018-05-21 R/Data_processing.R

@@ -47,7 +47,7 @@ surv <- tbl(db, "KD_Z9") %>%
                        Temperature = bio1, Precipitation = bio12), copy = TRUE) %>% 
   left_join(tbl(db, "RAUMDATEN_Z9_NDEP") %>% 
               transmute(aID_STAO = aID_STAO, NTOT = NTOT2010b), copy = TRUE)
-  
+
 # Add community measures to 'surv'
 surv <- surv %>% left_join(
   tbl(db, "PL") %>% 

### 2018-05-22 Manuscript.Rmd
Grafik mit Abweichung von Zufalls-Aussterben und Zufalls-Einwandern.

@@
 
-# Average of functional traits
-res[ 8, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$CH, surv$Visit, mean), 2)
-tmp <- summary(lmer(CH ~ yr + (yr|aID_STAO), data = surv))$coefficients
-res[ 8, "Temporal-Trend"] <- formatC(tmp["yr", "Estimate"], format = "f", digits = 3)
-tmp <- anova(lmer(CH ~ yr + (yr|aID_STAO), data = surv), lmer(CH ~ 1 + (yr|aID_STAO), data = surv))[2, "Pr(>Chisq)"]
-res[ 8, "P-value"] <- formatC(tmp, format = "f", digits = 3)
-
-res[9, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$SLA, surv$Visit, mean), 2)
-tmp <- summary(lme(SLA ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
-res[9, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
-res[9, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)
-
-res[10, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$SM, surv$Visit, mean), 2)
-tmp <- summary(lme(SM ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
-res[10, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
-res[10, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)
-
 print(xtable(res, align = c("l", "l", rep("r", 5)), label = "temporaltrends", caption = "Average measures of community structure for the three survey periods (in each period all sites are surveyed once).
@@
 
-The different measures of total community structure suggested that plant communities in mountain hay meadows were stable between 2003 and 2017 (Table \@ref(temporaltrends)): for each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity, average Ellenberg values for temperature, nutrients, light and huminity, and average of species' canopy height, specific leaf area and seed mass did not vary much among the three sampling periods.
+The different measures of total community structure suggested that plant communities in mountain hay meadows were stable between 2003 and 2017 (Table \@ref(temporaltrends)): for each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity and the average Ellenberg values for temperature, nutrients, light and huminity did not vary much among the three sampling periods.
 For all measures, all p-values of the average temperal trends per site were $>0.05$ and only in the case of the average temperature value of the species the p-value was $<0.1$.
@@
 There is some evidence that this slight decrease in the temporal turnover from the first and second survey to the temporal turnover of the second and third survey is systematic and was not entirly caused by change (Binomial generalized linear model; effect size = `r coef(mod)[2]`; p = `r summary(mod)$coefficients["period", "Pr(>|z|)"]`).
+
+```{r diffromrandom}
+# Function to sample community and callculate CM
+getsamples <- function(s, nspec, d, EL) {
+  sample(d %>% pull(paste(EL)), nspec, replace = FALSE) %>% mean(na.rm = TRUE)
+}
+
+# Prepare data from one site and calculate mean and CrI for sample and mean for real community
+getsim <- function(x, EL, gain) {
+  tt <- pl %>% filter(aID_STAO == x) %>% 
+    group_by(aID_SP, T, F, N, L) %>% 
+    summarise(Occ1 = sum(Visit == 1),
+              Occ2 = sum(Visit == 2),
+              Occ3 = sum(Visit == 3))
+  if(gain)  ausw <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
+  if(!gain) ausw <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
+  tmp <- map_dbl(1:nsim, getsamples, nspec = sum(ausw), d = tt, EL = EL)
+  list(
+    nchange = sum(ausw),
+    meanreal = mean(tt[ausw, paste(EL)] %>% unlist, na.rm = TRUE),
+    mean = mean(tmp, na.rm = TRUE),
+    lower = quantile(tmp, probs = 0.025, na.rm = TRUE),
+    upper = quantile(tmp, probs = 0.975, na.rm = TRUE))
+}
+
+# Function to prepare the res file for plotting
+makesim <- function(EL, gain) {
+  res <- sites %>% 
+    filter(NTOT < 4) %>% 
+    transmute(aID_STAO = aID_STAO, Sitemeasure = NTOT)
+  tmp <- map(res$aID_STAO, getsim, EL = EL, gain = gain)
+  res$meanreal <- tmp %>% map_dbl("meanreal")
+  res$mean <- tmp %>% map_dbl("mean")
+  res$lower <- tmp %>% map_dbl("lower")
+  res$upper <- tmp %>% map_dbl("upper")
+  res
+}
+
+# Make calculations and plot
+tres <- makesim("N", gain = FALSE)
+txlab <- seq(0, 40, 5)
+tylab <- seq(2, 4, 0.5)
+pNloss <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
+  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
+  geom_point(shape = 4, colour = "black", cex = 2) +
+  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
+  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "grey60", lty = 2) +
+  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
+  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1.5, 1.5)) +
+  labs(x = "Nitrogen deposition", y = "Ellenberg nutrient value\n(community mean)",
+       title = "(A) Species losses")
+
+tres <- makesim("N", gain = TRUE)
+txlab <- seq(0, 40, 5)
+tylab <- seq(2, 4, 0.5)
+pNgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
+  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
+  geom_point(shape = 4, colour = "black", cex = 2) +
+  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
+  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "grey60", lty = 2) +
+  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
+  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1.5, 1.5)) +
+  labs(x = "Nitrogen deposition", y = "Ellenberg nutrient value\n(community mean)",
+       title = "(B) Species gains")
+
+multiplot(pNloss, pNgain, cols = 2)
+
+res1 <- rbind(
+  res %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = mean, type = "Sample"),
+  res %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = meanreal, type = "Real"))
+mod1 <- lmer(mean ~ Sitemeasure * type + (1|aID_STAO), data = res1)
+mod2 <- lmer(mean ~ Sitemeasure + type + (1|aID_STAO), data = res1)
+anova(mod1, mod2)
+
+summary(glmer(SR ~ yr * NTOT + (yr|aID_STAO), data = surv, family = "poisson"))
+mod1 <- lmer(N ~ yr * NTOT + (yr|aID_STAO), data = surv)
+mod2 <- lmer(N ~ yr + NTOT + (yr|aID_STAO), data = surv)
+anova(mod1, mod2)
+
+t.test(sites$ngain)
+t.test(sites$nlost)
+
+lm(ngain ~ Inclination + Temperature + Precipitation  + NTOT, data = sites) %>% 
+  tidy
+
+plot(sites$ngain, sites$nlost)
+
+tapply()
+
+```
 
@@
 
-```{r}
-# Calculate trend for all species
-gettrend <- function(s) {
-  tt <- surv %>% left_join(pl %>% filter(aID_SP == s) %>% dplyr::select(aID_KD, Occ)) %>% 
-    replace_na(list(Occ = 0))
-  trend <- NA
-  try(trend <- glmer(Occ ~ yr + (1|aID_STAO), data = tt, family = binomial) %>% 
-    tidy() %>% 
-    filter(term == "yr") %>% 
-    pull(estimate))
-trend
-}
-res <- pl %>% group_by(aID_SP, T, F, N, L) %>% 
-  summarise(Nsites = n_distinct(aID_STAO),
-            Nrecords = n())
-res$Trend <- map_dbl(res$aID_SP, gettrend)
-
-# Is trend of species related to Ellenberg values
-gls(Trend ~ N, data = res) %>% 
-  tidy
-
-library("nlme")
-vf <- varExp(form=~Nrecords)
-gls(Trend ~ T, data = res %>% filter(!is.na(T)), weights=vf) %>% summary
-gls(Trend ~ F, data = res %>% filter(!is.na(F)), weights=vf) %>% summary
-gls(Trend ~ N, data = res %>% filter(!is.na(N)), weights=vf) %>% summary
-gls(Trend ~ L, data = res %>% filter(!is.na(L)), weights=vf) %>% summary
-
-cor.test(res$T, res$N)
-```
-
-```{r}
-# Borrowoing ideas from community assembly
-library(geometry)
-
-getFric <- function(x) {
-  tt <- pl %>% 
-  filter(!is.na(T) & !is.na(F) & !is.na(N) & !is.na(L)) %>% 
-  filter(aID_KD == x) %>% dplyr::select(T, F, N, L) 
-  geometry::convhulln(tt, "FA")$vol
-}
-
-surv$FRic <- map_dbl(surv$aID_KD, getFric)
-
-summary(lmer(FRic ~ yr + (yr|aID_STAO), data = surv))
-tapply(surv$FRic, surv$Visit, mean)
-
-ex1 <- dbFD(dummy$trait, dummy$abun)
-ex1
-```
-
---> Es gibt keine Einheitlichen Entwicklungen der Communities. Das heisst aber nicht, dass es keine systematischen Veränderungen gibt!
-
-```{r}
-# Functions 
-getsamples <- function(s, nspec, d, EL) {
-  sample(d %>% pull(paste(EL)), nspec, replace = FALSE) %>% mean(na.rm = TRUE)
-}
-
-getsim <- function(x, EL, gain) {
-  tt <- pl %>% filter(aID_STAO == x) %>% 
-    group_by(aID_SP, T, F, N, L) %>% 
-    summarise(Occ1 = sum(Visit == 1),
-              Occ2 = sum(Visit == 2),
-              Occ3 = sum(Visit == 3))
-  if(gain)  ausw <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
-  if(!gain) ausw <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
-  tmp <- map_dbl(1:nsim, getsamples, nspec = sum(ausw), d = tt, EL = EL)
-  list(
-    nchange = sum(ausw),
-    meanreal = mean(tt[ausw, paste(EL)] %>% unlist, na.rm = TRUE),
-    mean = mean(tmp, na.rm = TRUE),
-    lower = quantile(tmp, probs = 0.025, na.rm = TRUE),
-    upper = quantile(tmp, probs = 0.975, na.rm = TRUE))
-}
-
-res <- sites %>% 
-  filter(NTOT < 4) %>% 
-  transmute(aID_STAO = aID_STAO, Sitemeasure = NTOT)
-tmp <- map(res$aID_STAO, getsim, EL = "N", gain = TRUE)
-res$meanreal <- tmp %>% map_dbl("meanreal")
-res$mean <- tmp %>% map_dbl("mean")
-res$lower <- tmp %>% map_dbl("lower")
-res$upper <- tmp %>% map_dbl("upper")
-ggplot(res, aes(x = Sitemeasure, y = meanreal)) +
-  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey") +
-  geom_point(shape = 4, colour = "black", cex = 2) +
-  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black", alpha = 0.4) +
-  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "grey", alpha = 0.4)
-
-res1 <- rbind(
-  res %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = mean, type = "Sample"),
-  res %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = meanreal, type = "Real"))
-mod1 <- lmer(mean ~ Sitemeasure * type + (1|aID_STAO), data = res1)
-mod2 <- lmer(mean ~ Sitemeasure + type + (1|aID_STAO), data = res1)
-anova(mod1, mod2)
-
-summary(glmer(SR ~ yr * NTOT + (yr|aID_STAO), data = surv, family = "poisson"))
-mod1 <- lmer(N ~ yr * NTOT + (yr|aID_STAO), data = surv)
-mod2 <- lmer(N ~ yr + NTOT + (yr|aID_STAO), data = surv)
-anova(mod1, mod2)
-
-t.test(sites$ngain)
-t.test(sites$nlost)
-
-lm(ngain ~ Inclination + Temperature + Precipitation  + NTOT, data = sites) %>% 
-  tidy
-
-plot(sites$ngain, sites$nlost)
-
-tapply()
-
-```
-
 # Discussion {-}

### 2018-05-22 Manuscript.Rmd
Start changing to rstanarm

@@
 # Results {-}
-```{r temporaltrends, results='asis', }
-res <- tibble(Measures = c("Alpha-diversity", "Beta-diversity", "Gamma-Diversity", 
-                           "Temperature value", "Huminity value", "Nutrients value", "Light value", 
-                           "Canopy height", "Specific leaf area", "Seed mass"))
-res$'Period 1' <- ""; res$'Period 2' <- ""; res$'Period 3' <- ""
-res$'Temporal-Trend' <- ""; res$'P-value' <- ""
+
+```{r communitytrends, results='asis', }
+res <- tibble(Measures = c("Alpha-diversity", "Beta-diversity", "Gamma-Diversity", "Temperature value", "Huminity value", "Nutrients value", "Light value"),
+              'Period 1' = NA,
+              'Period 2' = NA,
+              'Period 3' = NA,
+              'Temporal-Trend' = NA,
+              '2.5%' = NA,
+              '97.5%' = NA)
 
 # Alpha diversity
-res[1, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$SR, surv$Visit, mean), 2)
-tmp <- summary(glmer(SR ~ yr + (yr|aID_STAO), data = surv, family = poisson))$coefficients
-res[1, "Temporal-Trend"] <- formatC(tmp["yr", "Estimate"], format = "f", digits = 3)
-res[1, "P-value"] <- formatC(tmp["yr", "Pr(>|z|)"], format = "f", digits = 3)
+res[1, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$SR, surv$Visit, mean)
+tmp <- stan_glmer(SR ~ yr + (yr|aID_STAO), data = surv, family = poisson,
+                  cores = ncores, iter = niter)
+res[1, "Temporal-Trend"] <- fixef(tmp)["yr"]
+res[1, c('2.5%', '97.5%')] <- posterior_interval(tmp, pars = "yr")
 
@@
              listin = TRUE, listout = TRUE)
-  formatC(tmp$simpson, format = "f", digits = 2)
+  mean(tmp$simpson)
 }
@@
 
-print(xtable(res, align = c("l", "l", rep("r", 5)), label = "temporaltrends", caption = "Average measures of community structure for the three survey periods (in each period all sites are surveyed once).
+print(xtable(res, align = c("l", "l", rep("r", 6)), digits = ndigits, label = "temporaltrends", caption = "Average measures of community structure for the three survey periods (in each period all sites are surveyed once).
 The temporal trends and p-values are based on linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution) with site-ID as random effect.
@@
 
-The different measures of total community structure suggested that plant communities in mountain hay meadows were stable between 2003 and 2017 (Table \@ref(temporaltrends)): for each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity and the average Ellenberg values for temperature, nutrients, light and huminity did not vary much among the three sampling periods.
+The different measures of total community structure suggested that plant communities in mountain hay meadows were stable between 2003 and 2017 (Table \@ref(communitytrends)): for each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity and the average Ellenberg values for temperature, nutrients, light and huminity did not vary much among the three sampling periods.
 For all measures, all p-values of the average temperal trends per site were $>0.05$ and only in the case of the average temperature value of the species the p-value was $<0.1$.
@@
 
-```{r modturnover}
+```{r turnover}
 tt <- rbind(
@@
 The average percentage $\pm$ SD of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
-There is some evidence that this slight decrease in the temporal turnover from the first and second survey to the temporal turnover of the second and third survey is systematic and was not entirly caused by change (Binomial generalized linear model; effect size = `r coef(mod)[2]`; p = `r summary(mod)$coefficients["period", "Pr(>|z|)"]`).
+Thus, it seems that the turnover from the first/second survey to the temporal turnover of the second/third survey slightly decreased (95% Credible interval of the change in turnover estimated from a Binomial generalized linear model: `r coef(mod)[2]`; p = `r summary(mod)$coefficients["period", "Pr(>|z|)"]`).
 

### 2018-05-22 R/Data_processing.R

@@ -81,7 +81,7 @@ pl <- tbl(db, "PL") %>%
             Visit = as.integer(floor((yearP - 1998) / 5)),
             Occ = as.integer(1)) %>% 
   left_join(tbl(db, "Traits_PL") %>% dplyr::select(aID_SP, T, F, N, L), copy = TRUE) 
-  
+
 #------------------------------------------------------------------------------------------------------
 # Sites data 
 #------------------------------------------------------------------------------------------------------

### 2018-05-22 R/Settings.R

@@ -16,8 +16,7 @@ library(kableExtra)
 library(xtable)
 library(tidyverse)
 library(broom)
-library(arm)
-library(nlme)
+library(rstanarm)
 library(simba)
 
 #------------------------------------------------------------------------------------------------------
@@ -36,7 +35,7 @@ load("RData/survdat.RData")
 opts_chunk$set(echo = FALSE, hide = TRUE, cache = TRUE, warning = FALSE, message = FALSE,
                fig.asp = .4, fig.width = 8, out.width = "100%")
 
-# Darstellung von R-Berechnungen innerhalb von Text
+# Stting for withing text juncs
 inline_hook <- function(x) {
   if (is.numeric(x)) {
     x <- format(x, nsmall = 2, digits = 2)
@@ -45,7 +44,7 @@ inline_hook <- function(x) {
 }
 knit_hooks$set(inline = inline_hook)
 
-# Plot settings
+# GGplot settings
 theme_set(theme_classic() + 
             theme(legend.position="bottom") + 
             theme(legend.title = element_blank()))
@@ -53,50 +52,14 @@ pd <- position_dodge(0.2)
 tcol <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", 
           "#D55E00", "#CC79A7")
 
-# Remove comment for xtable
+# Settings for xtable()
 options(xtable.comment = FALSE)
+ndigits = 2
 
-# Number of posterior samples
-nsim <- 100
+# Settings for rstanarm
+niter = 500
+ncores = 4
 
-#------------------------------------------------------------------------------------------------------
-# Hilfsfunktionen
-#------------------------------------------------------------------------------------------------------
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-  
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-  
-  numPlots = length(plots)
-  
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                     ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-  
-  if (numPlots == 1) {
-    print(plots[[1]])
-    
-  } else {
-    # Set up the page
[…]

### 2018-05-22 Appendix_A.Rmd
Compare species loss and gain with randomly selected species from same community --> Added as Appendix A

---
title: "Appendix A"
output:
  bookdown::pdf_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r settings, echo = FALSE, message=FALSE, cache=FALSE}
source("R/Settings.R")
```

# Random communities {-}
```{r ransimfunctions}
# Number of simulations
nsim <- 100

# Function to sample community and callculate CM
getsamples <- function(s, nspec, d, EL) {
  sample(d %>% pull(paste(EL)), nspec, replace = FALSE) %>% mean(na.rm = TRUE)
}

# Prepare data from one site and calculate mean and CrI for sample and mean for real community
getsim <- function(x, EL, gain) {
  tt <- pl %>% filter(aID_STAO == x) %>% 
    group_by(aID_SP, T, F, N, L) %>% 
    dplyr::summarise(Occ1 = sum(Visit == 1),
                     Occ2 = sum(Visit == 2),
                     Occ3 = sum(Visit == 3))
  if(gain)  ausw <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
  if(!gain) ausw <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
  tmp <- map_dbl(1:nsim, getsamples, nspec = sum(ausw), d = tt, EL = EL)
  list(
    nchange = sum(ausw),
    meanreal = mean(tt[ausw, paste(EL)] %>% unlist, na.rm = TRUE),
    mean = mean(tmp, na.rm = TRUE),
    lower = quantile(tmp, probs = 0.05, na.rm = TRUE),
    upper = quantile(tmp, probs = 0.95, na.rm = TRUE))
}

# Function to prepare the res file for plotting
makesim <- function(EL, gain, sitem) {
  res <- tibble(aID_STAO = sites$aID_STAO, Sitemeasure = pull(sites[, paste(sitem)]))
  tmp <- map(res$aID_STAO, getsim, EL = EL, gain = gain)
  res$meanreal <- tmp %>% map_dbl("meanreal")
  res$mean <- tmp %>% map_dbl("mean")
  res$lower <- tmp %>% map_dbl("lower")
  res$upper <- tmp %>% map_dbl("upper")
  res
}
```

```{r Tplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{temperature} of lost (A) and newly appearing (B) species compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e. random communities). Crosses indicate the community mean of lost and newly appearing species; the grey vertical lines indicate the range of the 5\\% and 95\\% quantiles of community means calculated from the 1000 random communities for each site. The curves show the regression lines with third order polynomials for the community means of lost and newly appearing species (black solid lines) and for the average community means of random communities (orange dotted lines).}"}
# Make calculations and plot for Ellenberg T value
tres <- makesim(EL = "T", gain = FALSE, sitem = "Temperature")
txlab <- seq(0, 100, 10)
tylab <- seq(1, 5, 0.5)
ploss <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 50) / 10, labels = txlab / 10, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1, 1)) +
  labs(x = "Annual mean temperature", y = "Ellenberg temperature value\n(community mean)",
       title = "(A) Species losses") 
tres <- makesim(EL = "T", gain = TRUE, sitem = "Temperature")
pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 50) / 10, labels = txlab/10, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1, 1)) +
  labs(x = "Annual mean temperature", y = "Ellenberg temperature value\n(community mean)",
       title = "(B) Species gains") 
multiplot(ploss, pgain, cols = 2)
```

```{r Fplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{humidity} of lost (A) and newly appearing (B) species compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e. random communities). Crosses indicate the community mean of lost and newly appearing species; the grey vertical lines indicate the range of the 5\\% and 95\\% quantiles of community means calculated from the 1000 random communities for each site. The curves show the regression lines with third order polynomials for the community means of lost and newly appearing species (black solid lines) and for the average community means of random communities (orange dotted lines).}"}
# Make calculations and plot for Ellenberg F value
tres <- makesim(EL = "F", gain = FALSE, sitem = "Precipitation")
txlab <- seq(800, 2000, 200)
tylab <- seq(1, 4, 0.5)
ploss <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 1000) / 200, labels = txlab, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1, 1)) +
  labs(x = "Annual mean precipitation", y = "Ellenberg humidity value\n(community mean)",
       title = "(A) Species losses") 
tres <- makesim(EL = "F", gain = TRUE, sitem = "Precipitation")
pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 1000) / 200, labels = txlab, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1, 1)) +
  labs(x = "Annual mean precipitation", y = "Ellenberg humidity value\n(community mean)",
       title = "(B) Species gains") 
multiplot(ploss, pgain, cols = 2)
```

```{r Nplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{nutrients} of lost (A) and newly appearing (B) species compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e. random communities). Crosses indicate the community mean of lost and newly appearing species; the grey vertical lines indicate the range of the 5\\% and 95\\% quantiles of community means calculated from the 1000 random communities for each site. The curves show the regression lines with third order polynomials for the community means of lost and newly appearing species (black solid lines) and for the average community means of random communities (orange dotted lines).}"}
# Make calculations and plot for Ellenberg N value
tres <- makesim(EL = "N", gain = FALSE, sitem = "NTOT")
txlab <- seq(0, 40, 5)
tylab <- seq(1, 4, 0.5)
ploss <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1.2, 1)) +
  labs(x = "Nitrogen deposition", y = "Ellenberg nutrient value\n(community mean)",
       title = "(A) Species losses") 
tres <- makesim("N", gain = TRUE, sitem = "NTOT")
pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1.2, 1)) +
  labs(x = "Nitrogen deposition", y = "Ellenberg nutrient value\n(community mean)",
       title = "(B) Species gains") 
multiplot(ploss, pgain, cols = 2)
```

```{r Lplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{light} of lost (A) and newly appearing (B) species compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e. random communities). Crosses indicate the community mean of lost and newly appearing species; the grey vertical lines indicate the range of the 5\\% and 95\\% quantiles of community means calculated from the 1000 random communities for each site. The curves show the regression lines with third order polynomials for the community means of lost and newly appearing species (black solid lines) and for the average community means of random communities (orange dotted lines).}"}
# Make calculations and plot for Ellenberg N value
tres <- makesim(EL = "L", gain = FALSE, sitem = "Inclination")
txlab <- seq(0, 60, 10)
tylab <- seq(1, 5, 0.5)
ploss <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-0.5, 1.5)) +
  labs(x = "Inclination", y = "Ellenberg light value\n(community mean)",
       title = "(A) Species losses") 
tres <- makesim("L", gain = TRUE, sitem = "Inclination")
pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-0.5, 1.5)) +
  labs(x = "Inclination", y = "Ellenberg light value\n(community mean)",
       title = "(B) Species gains") 
multiplot(ploss, pgain, cols = 2)
```

### 2018-05-22 R/Run_models.R
do rstanarm models in a separate file and store results in Modelfit

@@ -1,0 +1,61 @@
+#======================================================================================================
+#
+# Plant community change in mountain hay meadows
+#
+# Run the models using rstanarm
+#
+#======================================================================================================
+
+rm(list=ls(all=TRUE))
+
+#------------------------------------------------------------------------------------------------------
+# Settings
+#------------------------------------------------------------------------------------------------------
+# Libraries
+library(tidyverse)
+library(rstanarm)
+
+# Load data
+load("RData/surv.RData")
+load("RData/sites.RData")
+
+# Settings for rstanarm
+ncores = 4
+
+#------------------------------------------------------------------------------------------------------
+# Temporal trend of community measures
+#------------------------------------------------------------------------------------------------------
+# Temporal trend of species richness
+mod <- stan_glmer(SR ~ yr + (yr|aID_STAO), data = surv, family = poisson, cores = ncores)
+save(mod, file = "Modelfit/SR-trend.RData")
+
+# Temporal trend of Ellenberg temperature value
+mod <- stan_lmer(T ~ yr + (yr|aID_STAO), data = surv, cores = ncores)
+save(mod, file = "Modelfit/T-trend.RData")
+
+# Temporal trend of Ellenberg humidity value
+mod <- stan_lmer(F ~ yr + (yr|aID_STAO), data = surv, cores = ncores)
+save(mod, file = "Modelfit/F-trend.RData")
+
+# Temporal trend of Ellenberg nutrient value
+mod <- stan_lmer(N ~ yr + (yr|aID_STAO), data = surv, cores = ncores)
+save(mod, file = "Modelfit/N-trend.RData")
+
+# Temporal trend of Ellenberg light value
+mod <- stan_lmer(L ~ yr + (yr|aID_STAO), data = surv, cores = ncores)
+save(mod, file = "Modelfit/L-trend.RData")
+
+#------------------------------------------------------------------------------------------------------
+# Difference in tunover between first/second and second/third survey period
+#------------------------------------------------------------------------------------------------------
+tt <- rbind(
+  sites %>% transmute(
+    change = change1, 
+    nochange = nochange1,
+    period = 1),
+  sites %>% transmute(
+    change = change2, 
+    nochange = nochange2,
+    period = 2))
+mod <- stan_glm(cbind(change, nochange) ~ period, data = tt, family = binomial, cores = ncores)
+save(mod, file = "Modelfit/turnover-trend.RData")

### 2018-05-22 Manuscript.Rmd
Edits in Manuscript. Remove site with very high deposition rate.

@@
 ```{r communitytrends, results='asis', }
-res <- tibble(Measures = c("Alpha-diversity", "Beta-diversity", "Gamma-Diversity", "Temperature value", "Huminity value", "Nutrients value", "Light value"),
+res <- tibble(Measures = c("Alpha-diversity", "Beta-diversity", "Temperature value", "Huminity value", "Nutrients value", "Light value"),
               'Period 1' = NA,
@@
               'Period 3' = NA,
-              'Temporal-Trend' = NA,
-              '2.5%' = NA,
-              '97.5%' = NA)
+              'Trend' = NA,
+              '5%' = NA,
+              '95%' = NA,
+              'Trend > 0' = NA)
 
@@
 res[1, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$SR, surv$Visit, mean)
-tmp <- stan_glmer(SR ~ yr + (yr|aID_STAO), data = surv, family = poisson,
-                  cores = ncores, iter = niter)
-res[1, "Temporal-Trend"] <- fixef(tmp)["yr"]
-res[1, c('2.5%', '97.5%')] <- posterior_interval(tmp, pars = "yr")
+load("Modelfit/SR-trend.RData")
+res[1, "Trend"] <- fixef(mod)["yr"]
+res[1, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
+res[1, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
@@
 getbeta <- function(x) {
-  tmp <- simba::sim(pl %>% filter(Visit == x) %>% dplyr::select(aID_STAO, aID_SP, Occ), method = "simpson", 
-             listin = TRUE, listout = TRUE)
+  tmp <- simba::sim(pl %>% filter(Visit == x) %>% dplyr::select(aID_STAO, aID_SP, Occ), 
+                    method = "simpson", listin = TRUE, listout = TRUE)
   mean(tmp$simpson)
@@
 
-# Gamma diversity
-res[3, c("Period 1", "Period 2", "Period 3")] <- tapply(pl$aID_SP, pl$Visit, function(x) length(unique(x)))
+# Mean community index
+res[3, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$T, surv$Visit, mean)
+load("Modelfit/T-trend.RData")
+res[3, "Trend"] <- fixef(mod)["yr"]
+res[3, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
+res[3, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
-# Mean community index
-res[4, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$T, surv$Visit, mean), 2)
-tmp <- summary(lme(T ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
-res[4, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
-res[4, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)
+res[4, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$F, surv$Visit, mean), 2)
+load("Modelfit/F-trend.RData")
+res[4, "Trend"] <- fixef(mod)["yr"]
+res[4, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
+res[4, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
-res[5, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$F, surv$Visit, mean), 2)
-tmp <- summary(lme(F ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
-res[5, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
-res[5, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)
+res[5, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$N, surv$Visit, mean), 2)
+load("Modelfit/N-trend.RData")
+res[5, "Trend"] <- fixef(mod)["yr"]
+res[5, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
+res[5, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
-res[6, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$N, surv$Visit, mean), 2)
-tmp <- summary(lme(N ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
-res[6, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
-res[6, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)
+res[6, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$L, surv$Visit, mean), 2)
+load("Modelfit/L-trend.RData")
+res[6, "Trend"] <- fixef(mod)["yr"]
+res[6, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
+res[6, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
-res[7, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$L, surv$Visit, mean), 2)
-tmp <- summary(lme(L ~ yr, random = ~ yr|aID_STAO, data = surv))$tTable
-res[7, "Temporal-Trend"] <- formatC(tmp["yr", "Value"], format = "f", digits = 3)
-res[7, "P-value"] <- formatC(tmp["yr", "p-value"], format = "f", digits = 3)
-
-print(xtable(res, align = c("l", "l", rep("r", 6)), digits = ndigits, label = "temporaltrends", caption = "Average measures of community structure for the three survey periods (in each period all sites are surveyed once).
-The temporal trends and p-values are based on linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution) with site-ID as random effect.
-Temporal-trends are given per 10 years.
-Linear mixed models could not be applied for beta- and gamma-diversity because measures are not available for the single sites."), include.rownames=FALSE)
+print(xtable(res, align = c("l", "l", rep("r", 7)), digits = ndigits, caption = "Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
+The temporal trends are given as  change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function) with site-ID as random intercept and slope effect.
+The measure of precision for the temporal trend is given as the 5% and 95% quantiles of the marginal posterior distribution.
+Finally, the probability that the linear trend is <0 is given.
+Linear mixed models were not applicable for beta-diversity because measures for beta-diversity were not available for the single sites."), include.rownames=FALSE)
 ```
 
-The different measures of total community structure suggested that plant communities in mountain hay meadows were stable between 2003 and 2017 (Table \@ref(communitytrends)): for each of the three 5-year survey periods the averages of alpha-, beta- and gamma-diversity and the average Ellenberg values for temperature, nutrients, light and huminity did not vary much among the three sampling periods.
-For all measures, all p-values of the average temperal trends per site were $>0.05$ and only in the case of the average temperature value of the species the p-value was $<0.1$.
-Note that beta- and gamma-diversity are note available for single sites and thus mixed models to calculate the p-value could not be applied.
+The different measures of total community structure suggested that plant communities in mountain hay meadows were stable between 2003 and 2017 (Table \@ref(communitytrends)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, nutrients, light and huminity did not vary much among the three sampling periods.
+Except for average Ellenberg value for temperature, the 90% credible-interval of the temporal trend contained zero and the probability of a trend $>0$ was rather close to 0.5.
 
 ```{r turnover}
-tt <- rbind(
-  sites %>% transmute(
-    change = change1, 
-    nochange = nochange1,
-    period = 1),
-  sites %>% transmute(
-    change = change2, 
-    nochange = nochange2,
-    period = 2))
-mod <- glm(cbind(change, nochange) ~ period, data = tt, family = binomial)
+load("Modelfit/turnover-trend.RData")
+tmp <-  posterior_interval(mod, pars = "period")
+CrI <- paste0(round(tmp[1,1], ndigits), " - ", round(tmp[1,2], ndigits))
 ```
@@
 The average percentage $\pm$ SD of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
-Thus, it seems that the turnover from the first/second survey to the temporal turnover of the second/third survey slightly decreased (95% Credible interval of the change in turnover estimated from a Binomial generalized linear model: `r coef(mod)[2]`; p = `r summary(mod)$coefficients["period", "Pr(>|z|)"]`).
+Thus, it seems that the turnover from the first/second survey to the temporal turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from a Binomial generalized linear model: `r CrI`).
 
 ```{r diffromrandom}
+# Number of simulations
+nsim <- 100
+
 # Function to sample community and callculate CM
@@
     group_by(aID_SP, T, F, N, L) %>% 
-    summarise(Occ1 = sum(Visit == 1),
+    dplyr::summarise(Occ1 = sum(Visit == 1),
                      Occ2 = sum(Visit == 2),
@@
 
-# Make calculations and plot
-tres <- makesim("N", gain = FALSE)
+# Make calculations and plot for T value
+tres <- makesim(EL = "T", gain = FALSE)
+
+tres <- makesim("N", gain = TRUE)
+txlab <- seq(0, 40, 5)
+tylab <- seq(2, 4, 0.5)
+pNgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
+  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
+  geom_point(shape = 4, colour = "black", cex = 2) +
+  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
+  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "grey60", lty = 2) +
+  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
+  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1.5, 1.5)) +
+  labs(x = "Nitrogen deposition", y = "Ellenberg nutrient value\n(community mean)",
+       title = "(B) Species gains") 
+# Make calculations and plot for N value
+tres <- makesim(EL = "N", gain = FALSE)
 txlab <- seq(0, 40, 5)
@@
 
-multiplot(pNloss, pNgain, cols = 2)
-
 res1 <- rbind(
@@
 
-t.test(sites$ngain)
-t.test(sites$nlost)
+```
 
-lm(ngain ~ Inclination + Temperature + Precipitation  + NTOT, data = sites) %>% 
-  tidy
-
-plot(sites$ngain, sites$nlost)
-
-tapply()
-
+```{r, fig.cap="Noch einfügen."}
+multiplot(pNloss, pNgain, cols = 2)
 ```

### 2018-05-22 R/Data_processing.R

@@ -31,7 +31,7 @@ tmp <- tbl(db, "KD_Z9") %>%
   as.tibble() %>% 
   filter(!is.na(match(aID_STAO, sel$aID_STAO)) & !is.na(yearPl) & yearP >= 2003 & yearP <= 2017) %>% 
   dplyr::select(aID_KD, aID_STAO, yearPl) %>% 
-  left_join(tbl(db, "PL") %>% group_by(aID_KD) %>% summarise(SR = n()), copy = TRUE) %>% 
+  left_join(tbl(db, "PL") %>% group_by(aID_KD) %>% dplyr::summarise(SR = n()), copy = TRUE) %>% 
   filter(SR >= 0)
 tmp <- table(tmp$aID_STAO)
 sum(tmp != 3)
@@ -54,20 +54,21 @@ surv <- surv %>% left_join(
     filter(!Z7) %>% 
     left_join(tbl(db, "Traits_PL")) %>%
     group_by(aID_KD) %>% 
-    summarise(SR = n(),
-              SR_oligo = sum(N <= 2, na.rm = TRUE),
-              SR_eutro = sum(N >= 4, na.rm = TRUE),
-              SLA = mean(log(SLA), na.rm = TRUE),
-              CH = mean(log(CH), na.rm = TRUE),
-              SM = mean(log(SM), na.rm = TRUE),
-              SLA_sd = sd(log(SLA), na.rm = TRUE),
-              CH_sd = sd(log(CH), na.rm = TRUE),
-              SM_sd = sd(log(SM), na.rm = TRUE),
-              T = mean(T, na.rm = TRUE), 
-              F = mean(F, na.rm = TRUE),
-              N = mean(N, na.rm = TRUE),
-              L = mean(L, na.rm = TRUE),
-              MV = mean(MV, na.rm = TRUE)), copy = TRUE)
+    dplyr::summarise(
+      SR = n(),
+      SR_oligo = sum(N <= 2, na.rm = TRUE),
+      SR_eutro = sum(N >= 4, na.rm = TRUE),
+      SLA = mean(log(SLA), na.rm = TRUE),
+      CH = mean(log(CH), na.rm = TRUE),
+      SM = mean(log(SM), na.rm = TRUE),
+      SLA_sd = sd(log(SLA), na.rm = TRUE),
+      CH_sd = sd(log(CH), na.rm = TRUE),
+      SM_sd = sd(log(SM), na.rm = TRUE),
[…]

### 2018-05-22 R/Settings.R

@@ -18,6 +18,7 @@ library(tidyverse)
 library(broom)
 library(rstanarm)
 library(simba)
+library(Rmisc)
 
 #------------------------------------------------------------------------------------------------------
 # Load data
@@ -44,7 +45,7 @@ inline_hook <- function(x) {
 }
 knit_hooks$set(inline = inline_hook)
 
-# GGplot settings
+# ggplot settings
 theme_set(theme_classic() + 
             theme(legend.position="bottom") + 
             theme(legend.title = element_blank()))
@@ -55,10 +56,6 @@ tcol <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
 # Settings for xtable()
 options(xtable.comment = FALSE)
 ndigits = 2
-
-# Settings for rstanarm
-niter = 500
-ncores = 4
 
 
 

### 2018-05-23 R/Data_processing.R
Add model on colonization and local survival (for Appendix B)

@@ -120,11 +120,6 @@ sites$Precipitation <- (sites$Precipitation - 1000) / 200
 sites$NTOT <- (sites$NTOT - 10) / 10
 sites$Inclination <- (sites$Inclination - 10) / 10
 
-pl$T <- (pl$T - 3)
-pl$F <- (pl$F - 3)
-pl$N <- (pl$N - 3)
-pl$L <- (pl$L - 3)
-
 #------------------------------------------------------------------------------------------------------
 # Compile data for colonization and local survival calculations
 #------------------------------------------------------------------------------------------------------

### 2018-05-23 R/Run_models.R

@@ -11,13 +11,8 @@ rm(list=ls(all=TRUE))
 #------------------------------------------------------------------------------------------------------
 # Settings
 #------------------------------------------------------------------------------------------------------
-# Libraries
-library(tidyverse)
-library(rstanarm)
-
-# Load data
-load("RData/surv.RData")
-load("RData/sites.RData")
+# Source overall settings
+source("R/Settings.R")
 
 # Settings for rstanarm
 ncores = 4
@@ -50,12 +45,29 @@ save(mod, file = "Modelfit/L-trend.RData")
 #------------------------------------------------------------------------------------------------------
 tt <- rbind(
   sites %>% transmute(
+    aID_STAO = aID_STAO,
     change = change1, 
     nochange = nochange1,
+    Temperature = Temperature,
+    Precipitation = Precipitation,
+    NTOT = NTOT,
+    Inclination = Inclination, 
     period = 1),
   sites %>% transmute(
+    aID_STAO = aID_STAO,
     change = change2, 
     nochange = nochange2,
+    Temperature = Temperature,
+    Precipitation = Precipitation,
+    NTOT = NTOT,
+    Inclination = Inclination, 
     period = 2))
-mod <- stan_glm(cbind(change, nochange) ~ period, data = tt, family = binomial, cores = ncores)
+mod <- stan_glmer(cbind(change, nochange) ~ period + Temperature + Precipitation + NTOT + Inclination + (1|aID_STAO), 
+                data = tt, family = binomial, cores = ncores)
 save(mod, file = "Modelfit/turnover-trend.RData")
+
+#------------------------------------------------------------------------------------------------------
+# 
+#------------------------------------------------------------------------------------------------------
+mod <- stan_glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial, cores = ncores)
+

### 2018-05-23 R/Settings.R

@@ -57,6 +57,44 @@ tcol <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
 options(xtable.comment = FALSE)
 ndigits = 2
 
+#------------------------------------------------------------------------------------------------------
+# Functions to sample random communities and compare with species gained or lost 
+#------------------------------------------------------------------------------------------------------
+# Number of simulations
+nsim <- 100
 
+# Function to sample community and callculate CM
+getsamples <- function(s, nspec, d, EL) {
+  sample(d %>% pull(paste(EL)), nspec, replace = FALSE) %>% mean(na.rm = TRUE)
+}
+
+# Prepare data from one site and calculate mean and CrI for sample and mean for real community
+getsim <- function(x, EL, gain) {
+  tt <- pl %>% filter(aID_STAO == x) %>% 
+    group_by(aID_SP, T, F, N, L) %>% 
+    dplyr::summarise(Occ1 = sum(Visit == 1),
+                     Occ2 = sum(Visit == 2),
+                     Occ3 = sum(Visit == 3))
+  if(gain)  ausw <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
+  if(!gain) ausw <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
+  tmp <- map_dbl(1:nsim, getsamples, nspec = sum(ausw), d = tt, EL = EL)
+  list(
+    nchange = sum(ausw),
+    meanreal = mean(tt[ausw, paste(EL)] %>% unlist, na.rm = TRUE),
+    mean = mean(tmp, na.rm = TRUE),
+    lower = quantile(tmp, probs = 0.05, na.rm = TRUE),
+    upper = quantile(tmp, probs = 0.95, na.rm = TRUE))
+}
+
+# Function to prepare the res file for plotting
+makesim <- function(EL, gain, sitem) {
+  res <- tibble(aID_STAO = sites$aID_STAO, Sitemeasure = pull(sites[, paste(sitem)]))
+  tmp <- map(res$aID_STAO, getsim, EL = EL, gain = gain)
+  res$meanreal <- tmp %>% map_dbl("meanreal")
+  res$mean <- tmp %>% map_dbl("mean")
+  res$lower <- tmp %>% map_dbl("lower")
+  res$upper <- tmp %>% map_dbl("upper")
+  res
+}
 
 

### 2018-05-23 R/Run_models.R
Models on colonization and local survival added.

@@ -70,4 +70,9 @@ save(mod, file = "Modelfit/turnover-trend.RData")
 # 
 #------------------------------------------------------------------------------------------------------
 mod <- stan_glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial, cores = ncores)
+save(mod, file = "Modelfit/colonization-T.RData")
+
+mod <- stan_glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial, cores = ncores)
+save(mod, file = "Modelfit/localsurvival-T.RData")
+
 

### 2018-05-23 R/Calc_species_trends.R
Verschiedene Tests: noch immer offen bleibt warum mittlere Temperaturzahlen eher zunehmen. Trends von Einzelarten berechnet.

@@ -1,0 +1,42 @@
+#======================================================================================================
+#
+# Plant community change in mountain hay meadows
+#
+# Calc temporal trend for all species
+#
+#======================================================================================================
+
+rm(list=ls(all=TRUE))
+
+#------------------------------------------------------------------------------------------------------
+# Settings
+#------------------------------------------------------------------------------------------------------
+# Source overall settings
+source("R/Settings.R")
+
+# Settings for rstanarm
+ncores = 4
+
+#------------------------------------------------------------------------------------------------------
+# Calculate trends for all species
+#------------------------------------------------------------------------------------------------------
+# spec <- pl %>% group_by(aID_SP, T, F, N, L) %>%
+#   dplyr::summarise(Nsites = n_distinct(aID_STAO),
+#                    Nrecords = n())
+# save(spec, file = "RData/spec.RData")
+load("RData/spec.RData")
+
+for(s in 1:nrow(spec)) {
+  if(spec$Nsites[s] >= 5) {try({
+    tt <- surv %>% left_join(pl %>% filter(aID_SP == spec$aID_SP[s]) %>% dplyr::select(aID_KD, Occ)) %>% 		
+      replace_na(list(Occ = 0))
+    mod <- stan_glmer(Occ ~ yr + (1|aID_STAO), data = tt, family = binomial, 
+                      prior_intercept = normal(location = 0, scale = 10), prior = normal(location = 0, scale = 2.5), 
+                      cores = ncores)
+    trend <- as.data.frame(mod, pars = "yr")$yr
+    save(trend, file = paste0("SpeciesTrends/", spec$aID_SP[s], ".RData"))
+    spec[s, "Filename"] <- paste0("SpeciesTrends/", spec$aID_SP[s], ".RData")
+  })}
+}
+
+save(spec, file = "RData/spec.RData")

### 2018-05-23 Appendix_A.Rmd

@@
 ---
-title: "Appendix A"
+title: "Appendix A: Random community change"
+toc: no
 output:
@@
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
 
-# Random communities {-}
-```{r ransimfunctions}
-# Number of simulations
-nsim <- 100
+# Introduction and mehtods {-}
+Add text to explain the ideas and methods.
 
-# Function to sample community and callculate CM
-getsamples <- function(s, nspec, d, EL) {
-  sample(d %>% pull(paste(EL)), nspec, replace = FALSE) %>% mean(na.rm = TRUE)
-}
+Auch noch auf Interaktion texten:
+```{r}
+library(arm)
 
-# Prepare data from one site and calculate mean and CrI for sample and mean for real community
-getsim <- function(x, EL, gain) {
-  tt <- pl %>% filter(aID_STAO == x) %>% 
-    group_by(aID_SP, T, F, N, L) %>% 
-    dplyr::summarise(Occ1 = sum(Visit == 1),
-                     Occ2 = sum(Visit == 2),
-                     Occ3 = sum(Visit == 3))
-  if(gain)  ausw <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
-  if(!gain) ausw <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
-  tmp <- map_dbl(1:nsim, getsamples, nspec = sum(ausw), d = tt, EL = EL)
-  list(
-    nchange = sum(ausw),
-    meanreal = mean(tt[ausw, paste(EL)] %>% unlist, na.rm = TRUE),
-    mean = mean(tmp, na.rm = TRUE),
-    lower = quantile(tmp, probs = 0.05, na.rm = TRUE),
-    upper = quantile(tmp, probs = 0.95, na.rm = TRUE))
-}
+res1 <- rbind(
+  tres %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = mean, type = "Sample"),
+  tres %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = meanreal, type = "Real"))
+mod1 <- lmer(mean ~ Sitemeasure * type + (1|aID_STAO), data = res1)
+mod2 <- lmer(mean ~ Sitemeasure + type + (1|aID_STAO), data = res1)
+anova(mod1, mod2)
 
-# Function to prepare the res file for plotting
-makesim <- function(EL, gain, sitem) {
-  res <- tibble(aID_STAO = sites$aID_STAO, Sitemeasure = pull(sites[, paste(sitem)]))
-  tmp <- map(res$aID_STAO, getsim, EL = EL, gain = gain)
-  res$meanreal <- tmp %>% map_dbl("meanreal")
-  res$mean <- tmp %>% map_dbl("mean")
-  res$lower <- tmp %>% map_dbl("lower")
-  res$upper <- tmp %>% map_dbl("upper")
-  res
-}
+summary(glmer(SR ~ yr * NTOT + (yr|aID_STAO), data = surv, family = "poisson"))
+mod1 <- lmer(N ~ yr * NTOT + (yr|aID_STAO), data = surv)
+mod2 <- lmer(N ~ yr + NTOT + (yr|aID_STAO), data = surv)
+anova(mod1, mod2)
 ```
+
+# Results {-}
 
@@
   scale_x_continuous(breaks = (txlab - 50) / 10, labels = txlab / 10, limits = c(-1, NA)) +
-  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1, 1)) +
+  scale_y_continuous(breaks = tylab, labels = tylab, limits = c(2, 4)) +
   labs(x = "Annual mean temperature", y = "Ellenberg temperature value\n(community mean)",
@@
   scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
-  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1.2, 1)) +
+  scale_y_continuous(breaks = tylab, labels = tylab, limits = c(1.8, 4)) +
   labs(x = "Nitrogen deposition", y = "Ellenberg nutrient value\n(community mean)",

### 2018-05-23 Appendix_B.Rmd

---
title: "Appendix A: Colonization probability and local survival along the gradient"
toc: no
output:
  bookdown::pdf_document2
---

```{r settings, echo = FALSE, message=FALSE, cache=FALSE}
source("R/Settings.R")
```

# Introduction and mehtods {-}
Add text to explain the ideas and methods.

# Results {-}

```{r Tplot, fig.cap="Colonization (a) and local survival (b) of low temperature species (Ellenberg T = 2; red line) and high temperature species (Ellengerg T = 4) species along the temperature gradient. Given are means and 95\\%-Credible Intervals from logistic linear mixed models."}
X <- seq(0, 10, 0.2)
Xsd <- (X - 5)

# Colonization
mod <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial)
simdat <- arm::sim(mod, nsim)@fixef
post <- map(1:nsim, getposterior)
res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
p1 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
  geom_line() +
    labs(title = "(A) Colonization", 
         y = "Colonization probability", 
         x = "Temperature gradient")

# Local survival
mod <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial)
simdat <- arm::sim(mod, nsim)@fixef
post <- map(1:nsim, getposterior)
res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
p2 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
  geom_line() +
    labs(title = "(B) Local survival", 
         y = "Local survival probability", 
         x = "Temperature gradient")
multiplot(p1, p2, cols = 2)
```

```{r Tplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{temperature} of lost (A) and newly appearing (B) species compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e. random communities). Crosses indicate the community mean of lost and newly appearing species; the grey vertical lines indicate the range of the 5\\% and 95\\% quantiles of community means calculated from the 1000 random communities for each site. The curves show the regression lines with third order polynomials for the community means of lost and newly appearing species (black solid lines) and for the average community means of random communities (orange dotted lines).}"}
# Make calculations and plot for Ellenberg T value
tres <- makesim(EL = "T", gain = FALSE, sitem = "Temperature")
txlab <- seq(0, 100, 10)
tylab <- seq(1, 5, 0.5)
ploss <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 50) / 10, labels = txlab / 10, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1, 1)) +
  labs(x = "Annual mean temperature", y = "Ellenberg temperature value\n(community mean)",
       title = "(A) Species losses") 
tres <- makesim(EL = "T", gain = TRUE, sitem = "Temperature")
pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
  geom_point(shape = 4, colour = "black", cex = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "orange", lty = 2) +
  scale_x_continuous(breaks = (txlab - 50) / 10, labels = txlab/10, limits = c(-1, NA)) +
  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1, 1)) +
  labs(x = "Annual mean temperature", y = "Ellenberg temperature value\n(community mean)",
       title = "(B) Species gains") 
multiplot(ploss, pgain, cols = 2)
```

`

### 2018-05-23 Manuscript.Rmd

@@
 
-res[4, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$F, surv$Visit, mean), 2)
+res[4, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$F, surv$Visit, mean)
 load("Modelfit/F-trend.RData")
@@
 
-res[5, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$N, surv$Visit, mean), 2)
+res[5, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$N, surv$Visit, mean)
 load("Modelfit/N-trend.RData")
@@
 
-res[6, c("Period 1", "Period 2", "Period 3")] <- round(tapply(surv$L, surv$Visit, mean), 2)
+res[6, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$L, surv$Visit, mean)
 load("Modelfit/L-trend.RData")
@@
 
-The different measures of total community structure suggested that plant communities in mountain hay meadows were stable between 2003 and 2017 (Table \@ref(communitytrends)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, nutrients, light and huminity did not vary much among the three sampling periods.
-Except for average Ellenberg value for temperature, the 90% credible-interval of the temporal trend contained zero and the probability of a trend $>0$ was rather close to 0.5.
+The different measures of total community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a consistent increase or decrease over time (Table \@ref(communitytrends)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, nutrients, light and huminity did not vary much among the three sampling periods and the estimated trends were rather small.
+Except for average Ellenberg value for temperature, the 90% credible-interval of the temporal trend contained zero.
+The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the Ellenberg value for temperature (probability for an increase: `r res[res$Measures == "Temperature value", "Trend > 0"]`), followed by the community mean of the Ellenberg light value (probaility for a decrease: `r 1 - res[res$Measures == "Light value", "Trend > 0"]`) and it was least likely for the alpha-diversity (probaility for an increase: `r res[res$Measures == "Alpha-diversity", "Trend > 0"]`).
+The chance that the community mean of the nutrient value decreased between 2003 and 2017 was `r 1- res[res$Measures == "Nutrients value", "Trend > 0"]`.
 
+### Site perspecitve
 ```{r turnover}
+# Load Binomial-GLMM for turnover
 load("Modelfit/turnover-trend.RData")
+
+# Calculate the probability of change
+probchange <- as.data.frame(mod, pars = c("Temperature", "Precipitation", "NTOT", "Inclination")) %>% 
+  apply(2, function(x) mean(x > 0))
+decreasing <- fixef(mod)[c("Temperature", "Precipitation", "NTOT", "Inclination")] < 0
+probchange[decreasing] <-  1 - probchange[decreasing]
+
+# 90% CrI for period effect
 tmp <-  posterior_interval(mod, pars = "period")
 CrI <- paste0(round(tmp[1,1], ndigits), " - ", round(tmp[1,2], ndigits))
+
 ```
+The temporal stability as inferred from the community measures were, however, in contrast to a rather large observed temporal turnover of recorded species.
+The average percentage $\pm$ SD of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
+Thus, it seems that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model that also contained the site variables: `r CrI`).
 
-The temporal stability as inferred from the community measures were, however, in contrast to the large  observed temporal turnover of recorded species.
-The average percentage $\pm$ SD of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
-Thus, it seems that the turnover from the first/second survey to the temporal turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from a Binomial generalized linear model: `r CrI`).
-
-```{r diffromrandom}
-# Number of simulations
-nsim <- 100
-
-# Function to sample community and callculate CM
-getsamples <- function(s, nspec, d, EL) {
-  sample(d %>% pull(paste(EL)), nspec, replace = FALSE) %>% mean(na.rm = TRUE)
+### Species perspective
+```{r}
+# Get posteriors from all single species trend models
+filenames <- list.files("SpeciesTrends", full.names = TRUE)
+aID_SP <- str_split(list.files("SpeciesTrends"), "\\.") %>% map(1) %>% unlist
+load(filenames[1])
+posttrend <- array(NA, dim = c(length(trend), length(aID_SP)))
+for(i in 1:length(filenames)){
+  load(filenames[i])
+  posttrend[, i] <- trend
 }
 
-# Prepare data from one site and calculate mean and CrI for sample and mean for real community
-getsim <- function(x, EL, gain) {
-  tt <- pl %>% filter(aID_STAO == x) %>% 
-    group_by(aID_SP, T, F, N, L) %>% 
-    dplyr::summarise(Occ1 = sum(Visit == 1),
-                     Occ2 = sum(Visit == 2),
-                     Occ3 = sum(Visit == 3))
-  if(gain)  ausw <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
-  if(!gain) ausw <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
-  tmp <- map_dbl(1:nsim, getsamples, nspec = sum(ausw), d = tt, EL = EL)
-  list(
-    nchange = sum(ausw),
-    meanreal = mean(tt[ausw, paste(EL)] %>% unlist, na.rm = TRUE),
-    mean = mean(tmp, na.rm = TRUE),
-    lower = quantile(tmp, probs = 0.025, na.rm = TRUE),
-    upper = quantile(tmp, probs = 0.975, na.rm = TRUE))
+# Get list of species for which trend estimates are available
+spec <- pl %>% group_by(aID_SP, T, F, N, L) %>%
+  dplyr::summarise(Nsites = n_distinct(aID_STAO),
+                   Nrecords = n())
+spec <- spec[!is.na(match(spec$aID_SP, aID_SP)), ]
+
+# Function to calculate effect of Ellenberg values on trend
+trendef <- function(x) {
+  spec$Trend <- posttrend[x,]
+  coef(lm(Trend ~ T + F + N + L, data = spec))
 }
 
-# Function to prepare the res file for plotting
-makesim <- function(EL, gain) {
-  res <- sites %>% 
-    filter(NTOT < 4) %>% 
-    transmute(aID_STAO = aID_STAO, Sitemeasure = NTOT)
-  tmp <- map(res$aID_STAO, getsim, EL = EL, gain = gain)
-  res$meanreal <- tmp %>% map_dbl("meanreal")
-  res$mean <- tmp %>% map_dbl("mean")
-  res$lower <- tmp %>% map_dbl("lower")
-  res$upper <- tmp %>% map_dbl("upper")
-  res
-}
+spec$Trend <- apply(posttrend, 2, median)
+spec$TrendSD <- apply(posttrend, 2, sd)
 
-# Make calculations and plot for T value
-tres <- makesim(EL = "T", gain = FALSE)
+summary(lm(Trend ~ T + F + N + L, data = spec, weights = 1/TrendSD))
 
-tres <- makesim("N", gain = TRUE)
-txlab <- seq(0, 40, 5)
-tylab <- seq(2, 4, 0.5)
-pNgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
-  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
-  geom_point(shape = 4, colour = "black", cex = 2) +
-  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
-  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "grey60", lty = 2) +
-  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
-  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1.5, 1.5)) +
-  labs(x = "Nitrogen deposition", y = "Ellenberg nutrient value\n(community mean)",
-       title = "(B) Species gains") 
-# Make calculations and plot for N value
-tres <- makesim(EL = "N", gain = FALSE)
-txlab <- seq(0, 40, 5)
-tylab <- seq(2, 4, 0.5)
-pNloss <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
-  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
-  geom_point(shape = 4, colour = "black", cex = 2) +
-  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
-  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "grey60", lty = 2) +
-  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
-  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1.5, 1.5)) +
-  labs(x = "Nitrogen deposition", y = "Ellenberg nutrient value\n(community mean)",
-       title = "(A) Species losses")
+library("nlme")		
+vf <- varExp(form= ~ 1/TrendSD)		
+gls(Trend ~ T + F + N + L, data = spec %>% filter(!is.na(T) & !is.na(F) & !is.na(N) & !is.na(L)), weights=vf) %>% summary
 
-tres <- makesim("N", gain = TRUE)
-txlab <- seq(0, 40, 5)
-tylab <- seq(2, 4, 0.5)
-pNgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
-  geom_errorbar(aes(ymin = lower, ymax = upper), lwd = 0.5, colour = "grey80") +
-  geom_point(shape = 4, colour = "black", cex = 2) +
-  stat_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, col = "black") +
-  stat_smooth(aes(y = mean), method = "lm", formula = y ~ poly(x,3), se = FALSE, col = "grey60", lty = 2) +
-  scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
-  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1.5, 1.5)) +
-  labs(x = "Nitrogen deposition", y = "Ellenberg nutrient value\n(community mean)",
-       title = "(B) Species gains")
+# Calculate for all samples of the posterior distribution
+tt <- map(1:1000, trendef)
+map_dbl(tt, "T") %>% quantile(probs = c(0.5, 0.05, 0.95))
+map_dbl(tt, "F") %>% quantile(probs = c(0.5, 0.05, 0.95))
+map_dbl(tt, "N") %>% quantile(probs = c(0.5, 0.05, 0.95))
+map_dbl(tt, "L") %>% quantile(probs = c(0.5, 0.05, 0.95))
 
-res1 <- rbind(
-  res %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = mean, type = "Sample"),
-  res %>% transmute(aID_STAO = aID_STAO, Sitemeasure = Sitemeasure, mean = meanreal, type = "Real"))
-mod1 <- lmer(mean ~ Sitemeasure * type + (1|aID_STAO), data = res1)
-mod2 <- lmer(mean ~ Sitemeasure + type + (1|aID_STAO), data = res1)
-anova(mod1, mod2)
+# Überlebenswahrscheinlichkeit ist grösser
+glm(Occ ~ Temperature + Precipitation + NTOT + Inclination, data = survdat, family = binomial) %>% summary
 
-summary(glmer(SR ~ yr * NTOT + (yr|aID_STAO), data = surv, family = "poisson"))
-mod1 <- lmer(N ~ yr * NTOT + (yr|aID_STAO), data = surv)
-mod2 <- lmer(N ~ yr + NTOT + (yr|aID_STAO), data = surv)
-anova(mod1, mod2)
+# Einwanderwahrscheinlichkeit ist grösser
+glm(Occ ~ Temperature + Precipitation + NTOT + Inclination, data = coldat, family = binomial) %>% summary
+
+cbind(
+survdat %>% group_by(T) %>% 
+  dplyr::summarise(
+    Anz_Ausgestroben = sum(Occ == 0)),
+coldat %>% group_by(T) %>% 
+  dplyr::summarise(
+    N = n(),
+    surv = mean(Occ)))
+
+plot(tt$surv)
+
+table(survdat$Period)
 
@@
 
-```{r, fig.cap="Noch einfügen."}
-multiplot(pNloss, pNgain, cols = 2)
-```
+--> Temperature and N have an effect on how species generally react.
 
-```{r compofvarianceran}
-modcol <- summary(glmer(Occ ~ 1 + (1|aID_STAO) + (1|aID_SP), data = coldat, family = binomial))
-modls <- summary(glmer(Occ ~ 1 + (1|aID_STAO) + (1|aID_SP), data = survdat, family = binomial))
-getratio <- function(x) x[1] / x[2]
-```
+Temporal turnover could be high due to random processes including
 
-```{r compofvariance}
-# Local survival
-ref <- glmer(Occ ~ 1 + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial)
-modT <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), 
-              data = survdat, family = binomial)
-modF <- glmer(Occ ~ Precipitation * F + (1|aID_SP) + (1|aID_STAO), 
-              data = survdat, family = binomial)
-modN <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), 
-              data = survdat, family = binomial)
-modL <- glmer(Occ ~ Inclination * L + (1|aID_SP) + (1|aID_STAO), 
-              data = survdat, family = binomial)
-tt <- data.frame(anova(ref, modT, modF, modN, modL))
-tt$Model <- c("Random change", "Temperature driven", "Precipitation driven", "Nitrogen driven", "Land-use driven")
-tt <- tt[order(tt$AIC),]
-tt$'Delta-AIC' <- tt$AIC - tt$AIC[1]
-resls <- tt[, c("Model", "Df", "AIC", "Delta-AIC")]
-
-# Colonization probability
-ref <- glmer(Occ ~ 1 + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial)
-modT <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), 
-              data = coldat, family = binomial)
-modF <- glmer(Occ ~ Precipitation * F + (1|aID_SP) + (1|aID_STAO), 
-              data = coldat, family = binomial)
-modN <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), 
-              data = coldat, family = binomial)
-modL <- glmer(Occ ~ Inclination * L + (1|aID_SP) + (1|aID_STAO), 
-              data = coldat, family = binomial)
-tt <- data.frame(anova(ref, modT, modF, modN, modL))
-tt$Model <- c("Random change", "Temperature driven", "Precipitation driven", "Nitrogen driven", "Land-use driven")
-tt <- tt[order(tt$AIC),]
-tt$'Delta-AIC' <- tt$AIC - tt$AIC[1]
-rescol <- tt[, c("Model", "Df", "AIC", "Delta-AIC")]
-```
-
-```{r resdifeffects}
-tt <- rbind(rescol, resls)
-kable(tt, digits = 2, row.names = FALSE, format = "latex", caption = "Comparision of results from different drivers to explain colonization probability (a) and local survvival (b).
-See methods for the tested models.", booktabs = T) %>%
-    kable_styling() %>%
-    group_rows("(a) Colonization probability", 1, 5, bold = F, italic = T) %>%
-    group_rows("(b) Local survival probability", 6, 10, bold = F, italic = T)
-```
+Spatial turnover will be high if the probability a site is colonized by a species is high (i.e.
+high colonization probability) and/or if the probability that species survives at a site between two time points is low (i.e.
+low local survival).
+We found that the variation in local survival and colonization probability was largely due to differences between species and only to a lesser extend to the differences between locations: while the variance of the species random effect in a binomial linear mixed model for local survival was about `r getratio(unlist(modls$varcor))` times larger than the variance of the sampling site random effect, in the model for colonization the variance among species was even `r getratio(unlist(modcol$varcor))` times bigger than the variance among sites.
 
@@
 \@ref(fig:Neff)).
-
-```{r Teff, fig.cap="Colonization (a) and local survival (b) of low temperature species (Ellenberg T = 2; red line) and high temperature species (Ellengerg T = 4) species along the temperature gradient.
-Given are means and 95\\%-Credible Intervals from logistic linear mixed models."}
-X <- seq(0, 10, 0.2)
-Xsd <- (X - 5)
-
-# Colonization
-mod <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial)
-simdat <- arm::sim(mod, nsim)@fixef
-post <- map(1:nsim, getposterior)
-res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
-p1 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
-  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
-  geom_line() +
-    labs(title = "(a) Colonization", 
-         y = "Colonization probability", 
-         x = "Temperature gradient")
-
-# Local survival
-mod <- glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial)
-simdat <- arm::sim(mod, nsim)@fixef
-post <- map(1:nsim, getposterior)
-res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
-p2 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
-  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
-  geom_line() +
-    labs(title = "(b) Local survival", 
-         y = "Local survival probability", 
-         x = "Temperature gradient")
-multiplot(p1, p2, cols = 2)
-```
 

### 2018-05-23 Appendix_A.Rmd
Betrachtung der standartisierten Effekt Sizes (Mittelwert der loss o. gain Arten minus Mittelwerte der gleichen Anzahl simulierten Arten) kann viel erklären. --> Bei Temperatur gibt es einen Unterschied zwischen Loss und Gain Arten: Gain Arten haben einen leicht überdurchschnittlichen Wert und loss Arten einen leicht unterdurchschnittlichen Wert.

@@
 anova(mod1, mod2)
+
+los <- tibble()
+gain <- tibble()
+T <- tibble()
+
+library(arm)
+los$ef <- los$meanreal - los$mean
+summary(lmer(ef ~ EL * Sitemeasure + (1|aID_STAO), data = los))
+
+gain$ef <- gain$meanreal - gain$mean
+summary(lmer(ef ~ EL * Sitemeasure + (1|aID_STAO), data = gain))
+
+T$ef <- T$meanreal - T$mean
+mod1 <- lmer(ef ~ type + (1|aID_STAO), data = T)
+mod2 <- lmer(ef ~ 1 + (1|aID_STAO), data = T)
+anova(mod1, mod2)
 ```
@@
 tres <- makesim(EL = "T", gain = FALSE, sitem = "Temperature")
+
+tres$EL <- "T"; los <- rbind(los, tres)
+tres$type <- "loss"; T <- rbind(T, tres)
+
 txlab <- seq(0, 100, 10)
@@
 tres <- makesim(EL = "T", gain = TRUE, sitem = "Temperature")
+tres$EL <- "T"; gain <- rbind(gain, tres)
+tres$type <- "gain"; T <- rbind(T, tres)
+
 pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
@@
 tres <- makesim(EL = "N", gain = FALSE, sitem = "NTOT")
+tres$EL <- "N"; los <- rbind(los, tres)
 txlab <- seq(0, 40, 5)
@@
 tres <- makesim("N", gain = TRUE, sitem = "NTOT")
+tres$EL <- "N"; gain <- rbind(gain, tres)
 pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +

### 2018-05-23 R/Calc_species_trends.R
Aufräumaktion. Vor allem Berechnung des Bestandestrends wieder entfernt.

@@ -1,42 +1,0 @@
-#======================================================================================================
-#
-# Plant community change in mountain hay meadows
-#
-# Calc temporal trend for all species
-#
-#======================================================================================================
-
-rm(list=ls(all=TRUE))
-
-#------------------------------------------------------------------------------------------------------
-# Settings
-#------------------------------------------------------------------------------------------------------
-# Source overall settings
-source("R/Settings.R")
-
-# Settings for rstanarm
-ncores = 4
-
-#------------------------------------------------------------------------------------------------------
-# Calculate trends for all species
-#------------------------------------------------------------------------------------------------------
-# spec <- pl %>% group_by(aID_SP, T, F, N, L) %>%
-#   dplyr::summarise(Nsites = n_distinct(aID_STAO),
-#                    Nrecords = n())
-# save(spec, file = "RData/spec.RData")
-load("RData/spec.RData")
-
-for(s in 1:nrow(spec)) {
-  if(spec$Nsites[s] >= 5) {try({
-    tt <- surv %>% left_join(pl %>% filter(aID_SP == spec$aID_SP[s]) %>% dplyr::select(aID_KD, Occ)) %>% 		
-      replace_na(list(Occ = 0))
-    mod <- stan_glmer(Occ ~ yr + (1|aID_STAO), data = tt, family = binomial, 
-                      prior_intercept = normal(location = 0, scale = 10), prior = normal(location = 0, scale = 2.5), 
-                      cores = ncores)
-    trend <- as.data.frame(mod, pars = "yr")$yr
-    save(trend, file = paste0("SpeciesTrends/", spec$aID_SP[s], ".RData"))
-    spec[s, "Filename"] <- paste0("SpeciesTrends/", spec$aID_SP[s], ".RData")
-  })}
-}
-
-save(spec, file = "RData/spec.RData")

### 2018-05-23 Appendix_A.Rmd
Appendix A: Methodenerklärung eingefügt. Simulierte Ergebnisse werden Communitysim gespeichert.

--- 2018-05-22 Appendix_A.Rmd
@@
 ---
-title: "Appendix A"
+title: "Appendix A: Comparision of lost/gained species to randomly selected species"
+toc: no
 output:
@@
 
-# Random communities {-}
-```{r ransimfunctions}
+# Methods {-}
+```{r functions}
 # Number of simulations
@@
   tmp <- map(res$aID_STAO, getsim, EL = EL, gain = gain)
+  res$gain = gain
+  res$EL = paste(EL)
   res$meanreal <- tmp %>% map_dbl("meanreal")
@@
 ```
+Temporal turnover is caused by species that are lost between surveys (or overseen in the second survey) and gained between surveys (or overseen during the first survey).
+It is likely that different drivers such as Nitrogen deposition or climate warming would cause differnt species to colonize or disappear from sites.
+Also methodological issus such as imperfect detection is likely to result in species with particular characteristics that are overseen in one survey and thus result apparend species gains or losses.
+We here borrowed ideas from functional ecology in order to try to disentangle the potential drivers of species gains and losses: We compared the characteristics of the species that newly colonized or the species that disappeared from a site with the same amount of species that were randomly selected from all the species that were observed at the site (i.e.
+the species assemblage).
+Instead of the functional traits that are traditinally used in functional ecology such as species height, we used the Elleberg values that we though could best reflect the main drivers that are likely to cause shifts in mountain hay plant communities.
+This were the Ellenberg values for temperature (is likely to respond to climate warming), humidity (is likely to respond to changes in precipitation), nutrients (is likely to change in response to Nitrogen deposition and land-use intensity) and light (is likely to change in response to land-use intensity).
+
+# Results {-}
 
@@
 tres <- makesim(EL = "T", gain = FALSE, sitem = "Temperature")
+save(tres, file = "Communitysim/lossT.RData")
 txlab <- seq(0, 100, 10)
@@
 tres <- makesim(EL = "T", gain = TRUE, sitem = "Temperature")
+save(tres, file = "Communitysim/gainT.RData")
 pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
@@
   scale_x_continuous(breaks = (txlab - 50) / 10, labels = txlab/10, limits = c(-1, NA)) +
-  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1, 1)) +
+  scale_y_continuous(breaks = tylab, labels = tylab, limits = c(2, 4)) +
   labs(x = "Annual mean temperature", y = "Ellenberg temperature value\n(community mean)",
@@
 tres <- makesim(EL = "F", gain = FALSE, sitem = "Precipitation")
+save(tres, file = "Communitysim/lossF.RData")
 txlab <- seq(800, 2000, 200)
@@
   scale_x_continuous(breaks = (txlab - 1000) / 200, labels = txlab, limits = c(-1, NA)) +
-  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-1, 1)) +
+  scale_y_continuous(breaks = tylab, labels = tylab, limits = c(2, 4)) +
   labs(x = "Annual mean precipitation", y = "Ellenberg humidity value\n(community mean)",
@@
 tres <- makesim(EL = "F", gain = TRUE, sitem = "Precipitation")
+save(tres, file = "Communitysim/gainF.RData")
 pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
@@
 tres <- makesim(EL = "N", gain = FALSE, sitem = "NTOT")
+save(tres, file = "Communitysim/lossN.RData")
 txlab <- seq(0, 40, 5)
@@
 tres <- makesim("N", gain = TRUE, sitem = "NTOT")
+save(tres, file = "Communitysim/gainN.RData")
 pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +
@@
 tres <- makesim(EL = "L", gain = FALSE, sitem = "Inclination")
+save(tres, file = "Communitysim/lossL.RData")
 txlab <- seq(0, 60, 10)
@@
   scale_x_continuous(breaks = (txlab - 10) / 10, labels = txlab, limits = c(-1, NA)) +
-  scale_y_continuous(breaks = (tylab - 3), labels = tylab, limits = c(-0.5, 1.5)) +
+  scale_y_continuous(breaks = tylab, labels = tylab, limits = c(2.5, 4.5)) +
   labs(x = "Inclination", y = "Ellenberg light value\n(community mean)",
@@
 tres <- makesim("L", gain = TRUE, sitem = "Inclination")
+save(tres, file = "Communitysim/gainL.RData")
 pgain <- ggplot(tres, aes(x = Sitemeasure, y = meanreal)) +

### 2018-05-23 R/Settings.R

@@ -57,44 +57,5 @@ tcol <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
 options(xtable.comment = FALSE)
 ndigits = 2
 
-#------------------------------------------------------------------------------------------------------
-# Functions to sample random communities and compare with species gained or lost 
-#------------------------------------------------------------------------------------------------------
-# Number of simulations
-nsim <- 100
-
-# Function to sample community and callculate CM
-getsamples <- function(s, nspec, d, EL) {
-  sample(d %>% pull(paste(EL)), nspec, replace = FALSE) %>% mean(na.rm = TRUE)
-}
-
-# Prepare data from one site and calculate mean and CrI for sample and mean for real community
-getsim <- function(x, EL, gain) {
-  tt <- pl %>% filter(aID_STAO == x) %>% 
-    group_by(aID_SP, T, F, N, L) %>% 
-    dplyr::summarise(Occ1 = sum(Visit == 1),
-                     Occ2 = sum(Visit == 2),
-                     Occ3 = sum(Visit == 3))
-  if(gain)  ausw <- (tt$Occ1 == 0 & tt$Occ2 == 1) | (tt$Occ2 == 0 & tt$Occ3 == 1)
-  if(!gain) ausw <- (tt$Occ1 == 1 & tt$Occ2 == 0) | (tt$Occ2 == 1 & tt$Occ3 == 0)
-  tmp <- map_dbl(1:nsim, getsamples, nspec = sum(ausw), d = tt, EL = EL)
-  list(
-    nchange = sum(ausw),
-    meanreal = mean(tt[ausw, paste(EL)] %>% unlist, na.rm = TRUE),
-    mean = mean(tmp, na.rm = TRUE),
-    lower = quantile(tmp, probs = 0.05, na.rm = TRUE),
-    upper = quantile(tmp, probs = 0.95, na.rm = TRUE))
-}
-
-# Function to prepare the res file for plotting
-makesim <- function(EL, gain, sitem) {
-  res <- tibble(aID_STAO = sites$aID_STAO, Sitemeasure = pull(sites[, paste(sitem)]))
-  tmp <- map(res$aID_STAO, getsim, EL = EL, gain = gain)
-  res$meanreal <- tmp %>% map_dbl("meanreal")
-  res$mean <- tmp %>% map_dbl("mean")
-  res$lower <- tmp %>% map_dbl("lower")
-  res$upper <- tmp %>% map_dbl("upper")
-  res
-}
 
 

### 2018-05-28 R/Data_processing.R
Add new calculations of NDEP from Beat Rihm.

@@ -46,7 +46,7 @@ surv <- tbl(db, "KD_Z9") %>%
               transmute(aID_STAO = aID_STAO, Elevation = Hoehe, Inclination = Neig, 
                        Temperature = bio1, Precipitation = bio12), copy = TRUE) %>% 
   left_join(tbl(db, "RAUMDATEN_Z9_NDEP") %>% 
-              transmute(aID_STAO = aID_STAO, NTOT = NTOT2010b), copy = TRUE)
+              transmute(aID_STAO = aID_STAO, NTOT = NTOT2010_Wiesen), copy = TRUE)
 
 # Add community measures to 'surv'
 surv <- surv %>% left_join(
@@ -93,7 +93,7 @@ sites <- data.frame(aID_STAO = as.integer(sel)) %>%
               transmute(aID_STAO = aID_STAO, Elevation = Hoehe, Inclination = Neig, 
                             Temperature = bio1, Precipitation = bio12), copy = TRUE) %>% 
   left_join(tbl(db, "RAUMDATEN_Z9_NDEP") %>% 
-              transmute(aID_STAO = aID_STAO, NTOT = NTOT2010b), copy = TRUE)
+              transmute(aID_STAO = aID_STAO, NTOT = NTOT2010_Wiesen), copy = TRUE)
 
 for(i in 1:nrow(sites)) {
   tt <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit <= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", listin = TRUE, listout = TRUE)

### 2018-05-28 Manuscript.Rmd
Results neu strukturiert und Tabellen mit Resultaten eingefügt: vor allem Vergleich mit random species colonization und random species disappearance. Tabelle mit gradient Effekt auf zeitlichen Turnover.

@@
 
-```{r communitytrends, results='asis', }
+```{r communitytrendscalc}
 res <- tibble(Measures = c("Alpha-diversity", "Beta-diversity", "Temperature value", "Huminity value", "Nutrients value", "Light value"),
@@
               '95%' = NA,
-              'Trend > 0' = NA)
+              'Prob. for Trend' = NA)
 
@@
 res[1, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
-res[1, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
+res[1, 'Prob. for Trend'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
@@
 res[3, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
-res[3, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
+res[3, 'Prob. for Trend'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
@@
 res[4, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
-res[4, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
+res[4, 'Prob. for Trend'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
@@
 res[5, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
-res[5, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
+res[5, 'Prob. for Trend'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
@@
 res[6, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
-res[6, 'Trend > 0'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
-
-print(xtable(res, align = c("l", "l", rep("r", 7)), digits = ndigits, caption = "Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
-The temporal trends are given as  change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function) with site-ID as random intercept and slope effect.
-The measure of precision for the temporal trend is given as the 5% and 95% quantiles of the marginal posterior distribution.
-Finally, the probability that the linear trend is <0 is given.
-Linear mixed models were not applicable for beta-diversity because measures for beta-diversity were not available for the single sites."), include.rownames=FALSE)
+res[6, 'Prob. for Trend'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 ```
 
-The different measures of total community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a consistent increase or decrease over time (Table \@ref(communitytrends)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, nutrients, light and huminity did not vary much among the three sampling periods and the estimated trends were rather small.
+```{r communitytrends, echo=FALSE}
+options(knitr.kable.NA = '')
+kable(res, align = c("l", rep("r", 7)), digits = 2,  booktabs = T, col.names = c("Measures", "Period 1", "Period 2", "Period 3", "Trend", "5%", "95%", "Prob.
+for trend"), caption = "Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
+The temporal trends are given as  change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function) with site-ID as random intercept and slope effect.
+The measure of precision for the temporal trend is given as the 5\\% and 95\\% quantiles of the marginal posterior distribution.
+Finally, the probability that the linear trend is $> 0$ is given.
+Linear mixed models were not applicable for beta-diversity because measures for beta-diversity were not available for the single sites.")
+```
+
+The different measures of total community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a consistent increase or decrease over time (Table \@ref(tab:communitytrends)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, huminity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
 Except for average Ellenberg value for temperature, the 90% credible-interval of the temporal trend contained zero.
-The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the Ellenberg value for temperature (probability for an increase: `r res[res$Measures == "Temperature value", "Trend > 0"]`), followed by the community mean of the Ellenberg light value (probaility for a decrease: `r 1 - res[res$Measures == "Light value", "Trend > 0"]`) and it was least likely for the alpha-diversity (probaility for an increase: `r res[res$Measures == "Alpha-diversity", "Trend > 0"]`).
-The chance that the community mean of the nutrient value decreased between 2003 and 2017 was `r 1- res[res$Measures == "Nutrients value", "Trend > 0"]`.
+The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the Ellenberg value for temperature (probability for an increase: `r
+res %>% filter(Measures == "Temperature value") %>% pull("Prob.
+for Trend")`), followed by the community mean of the Ellenberg light value (probaility for a decrease: `r 1 - res %>% filter(Measures == "Light value") %>% pull("Prob.
+for Trend")`) and it was least likely for the alpha-diversity (probability for an increase: `r res %>% filter(Measures == "Alpha-diversity") %>% pull("Prob.
+for Trend")`).
+The chance that the community mean of the nutrient value decreased between 2003 and 2017 was `r 1- res %>% filter(Measures == "Nutrients value") %>% pull("Prob.
+for Trend")`.
 
-### Site perspecitve
 ```{r turnover}
@@
 
-# Calculate the probability of change
-probchange <- as.data.frame(mod, pars = c("Temperature", "Precipitation", "NTOT", "Inclination")) %>% 
-  apply(2, function(x) mean(x > 0))
-decreasing <- fixef(mod)[c("Temperature", "Precipitation", "NTOT", "Inclination")] < 0
-probchange[decreasing] <-  1 - probchange[decreasing]
+# Compile res for table
+res <- tibble()
+tparams <- c("Temperature", "Precipitation", "NTOT", "Inclination")
+res <- tibble(
+  Gradient = c("Annual mean temperature", "Annual mean precipitation", "Nitrogen deposition", "Inclination"),
+  Estimate = fixef(mod)[tparams],
+  `5%` = posterior_interval(mod, pars = tparams)[,1],
+  `95%` = posterior_interval(mod, pars = tparams)[,2])
 
@@
 tmp <-  posterior_interval(mod, pars = "period")
-CrI <- paste0(round(tmp[1,1], ndigits), " - ", round(tmp[1,2], ndigits))
+CrI <- paste0(round(tmp[1,1], 2), " - ", round(tmp[1,2], 2))
+```
 
+```{r turnovertab, echo=FALSE}
+options(knitr.kable.NA = '')
+kable(res, align = c("l", rep("r", 3)), digits = 2,  booktabs = T, col.names = c("Gradient", "Estimate", "5%", "95%"), caption = "Change of species turnover along the four gradients.
+The slopes along the gradients (estimate) are given as the change per 10 years of the logit-probability of species that differed between two surveys.
+Estimates and the 5\\% and 95\\% quantiles of the marginal posterior distribution obtained from a Binomial-GLMM with the proportion of species that differed between two surveys as dependent variable and the site gradients and period as predictors and site-ID as random effect.")
 ```
-The temporal stability as inferred from the community measures were, however, in contrast to a rather large observed temporal turnover of recorded species.
+
+This temporal stability as inferred from the community measures was, however, in contrast to a rather large observed temporal turnover of species.
 The average percentage $\pm$ SD of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
-Thus, it seems that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model that also contained the site variables: `r CrI`).
+Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model that also contained site variables: `r CrI`).
+Variation in species turnover was largest along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen depostion \@ref(tab:turnovertab).
+The other three gradients were less inportant to explain the variation in species turnover among sites.
 
-### Species perspective
-```{r}
-# Get posteriors from all single species trend models
-filenames <- list.files("SpeciesTrends", full.names = TRUE)
-aID_SP <- str_split(list.files("SpeciesTrends"), "\\.") %>% map(1) %>% unlist
-load(filenames[1])
-posttrend <- array(NA, dim = c(length(trend), length(aID_SP)))
-for(i in 1:length(filenames)){
-  load(filenames[i])
-  posttrend[, i] <- trend
+```{r difffromrandom}
+getres <- function(param, gradient, file) {
+  tt <- tibble()
+  load(paste0("Communitysim/", file))
+  tres$StandEf <- tres$meanreal - tres$mean
+  mod <- stan_glm(StandEf ~ Sitemeasure, data = tres , family = gaussian(), cores = 4)
+  tt[1, "Ellenberg value"] <- param
+  tt[1, "Gradient"] <- gradient
+  tt[1, "Estimate"] <- coef(mod)[1]
+  tmp <- posterior_interval(mod, pars = "(Intercept)")[1,] %>% as.vector()
+  tt[1, "5%"] <- tmp[1] 
+  tt[1, "90%"] <- tmp[2] 
+  tt[1, "Estimate_Slope"] <- coef(mod)[2]
+  tmp <- posterior_interval(mod, pars = "Sitemeasure")[1,] %>% as.vector()
+  tt[1, "5%_Slope"] <- tmp[1] 
+  tt[1, "90%_Slope"] <- tmp[2] 
+  tt
 }
 
-# Get list of species for which trend estimates are available
-spec <- pl %>% group_by(aID_SP, T, F, N, L) %>%
-  dplyr::summarise(Nsites = n_distinct(aID_STAO),
-                   Nrecords = n())
-spec <- spec[!is.na(match(spec$aID_SP, aID_SP)), ]
-
-# Function to calculate effect of Ellenberg values on trend
-trendef <- function(x) {
-  spec$Trend <- posttrend[x,]
-  coef(lm(Trend ~ T + F + N + L, data = spec))
-}
-
-spec$Trend <- apply(posttrend, 2, median)
-spec$TrendSD <- apply(posttrend, 2, sd)
-
-summary(lm(Trend ~ T + F + N + L, data = spec, weights = 1/TrendSD))
-
-library("nlme")		
-vf <- varExp(form= ~ 1/TrendSD)		
-gls(Trend ~ T + F + N + L, data = spec %>% filter(!is.na(T) & !is.na(F) & !is.na(N) & !is.na(L)), weights=vf) %>% summary
-
-# Calculate for all samples of the posterior distribution
-tt <- map(1:1000, trendef)
-map_dbl(tt, "T") %>% quantile(probs = c(0.5, 0.05, 0.95))
-map_dbl(tt, "F") %>% quantile(probs = c(0.5, 0.05, 0.95))
-map_dbl(tt, "N") %>% quantile(probs = c(0.5, 0.05, 0.95))
-map_dbl(tt, "L") %>% quantile(probs = c(0.5, 0.05, 0.95))
-
-# Überlebenswahrscheinlichkeit ist grösser
-glm(Occ ~ Temperature + Precipitation + NTOT + Inclination, data = survdat, family = binomial) %>% summary
-
-# Einwanderwahrscheinlichkeit ist grösser
-glm(Occ ~ Temperature + Precipitation + NTOT + Inclination, data = coldat, family = binomial) %>% summary
-
-cbind(
-survdat %>% group_by(T) %>% 
-  dplyr::summarise(
-    Anz_Ausgestroben = sum(Occ == 0)),
-coldat %>% group_by(T) %>% 
-  dplyr::summarise(
-    N = n(),
-    surv = mean(Occ)))
-
-plot(tt$surv)
-
-table(survdat$Period)
-
+res <- getres("Temperature", "Annual mean temperature", "lossT.RData")
+res <- rbind(res,  z <- getres("Humidity", "Annual mean precipitation", "lossF.RData"))
+res <- rbind(res,  getres("Nutrients", "Nitrogen deposition", "lossN.RData"))
+res <- rbind(res,  getres("Light", "Inclination", "lossL.RData"))
+res <- rbind(res, getres("Temperature", "Annual mean temperature", "gainT.RData"))
+res <- rbind(res,  z <- getres("Humidity", "Annual mean precipitation", "gainF.RData"))
+res <- rbind(res,  getres("Nutrients", "Nitrogen deposition", "gainN.RData"))
+res <- rbind(res,  getres("Light", "Inclination", "gainL.RData"))
 ```
 
---> Temperature and N have an effect on how species generally react.
-
-Temporal turnover could be high due to random processes including
-
-Spatial turnover will be high if the probability a site is colonized by a species is high (i.e.
-high colonization probability) and/or if the probability that species survives at a site between two time points is low (i.e.
-low local survival).
-We found that the variation in local survival and colonization probability was largely due to differences between species and only to a lesser extend to the differences between locations: while the variance of the species random effect in a binomial linear mixed model for local survival was about `r getratio(unlist(modls$varcor))` times larger than the variance of the sampling site random effect, in the model for colonization the variance among species was even `r getratio(unlist(modcol$varcor))` times bigger than the variance among sites.
-
- Spatial turnover will be high if the probability a site is colonized by a species is high (i.e.
-high colonization probability) and/or if the probability that species survives at a site between two time points is low (i.e.
-low local survival).
-We found that the variation in local survival and colonization probability was largely due to differences between species and only to a lesser extend to the differences between locations: while the variance of the species random effect in a binomial linear mixed model for local survival was about `r getratio(unlist(modls$varcor))` times larger than the variance of the sampling site random effect, in the model for colonization the variance among species was even `r getratio(unlist(modcol$varcor))` times bigger than the variance among sites.
-
- We inferred if we could explain differences in colonization and local survival with Ellenberg species values for temperature, humidity, nutrient or light (Table \@ref(tab:resdifeffects)).
-We found that both colonization probability and local survival probability was best explained by a model containing an interaction between species temperature and the average annual temperature of a site:  highest colonization probability and local survival were found  for the species with an Ellenberg value for Temperature that corresponded to the annual average temperature of the site (Fig.
-\@ref(fig:Teff))..
-While the colonization and local survival probability of cold living species (Ellenberg T = 2) declindes along the temperature gradient, the colonization and local survival probability of warm living species (Ellenberg T = 4) increases along the temperature gradient (Fig.
-\@ref(fig:Teff)).
-
-The second best model contained an interaction between species Ellenberg value for nutrient availability and the average annual N deposition rate of a site (Table \@ref(tab:resdifeffects)): the highest colonization probability and local survival were performed by species with an Ellenberg value for nutrient availability that corresponded to the annual N deposition rate of the site.
-While the colonization and local survival probability of oligotrophic species (Ellenberg N = 2) declindes along the N deposition gradient, the colonization and local survival probability of eutrophic species (N = 4) increases along the N deposition gradient (Fig.
-\@ref(fig:Neff)).
-
-```{r Neff, fig.cap="Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
-Given are means and 95\\%-Credible Intervals from logistic linear mixed models."}
-X <- 0:50
-Xsd <- (X - 10) / 10
-
-# Colonization
-mod <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial)
-simdat <- arm::sim(mod, nsim)@fixef
-post <- map(1:nsim, getposterior)
-res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
-p1 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
-  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
-  geom_line() +
-    labs(title = "(a) Colonization", 
-         y = "Colonization probability", 
-         x = "Nitrogen deposition")
-
-# Local survival
-mod <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial)
-simdat <- arm::sim(mod, nsim)@fixef
-post <- map(1:nsim, getposterior)
-res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
-p2 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
-  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
-  geom_line() +
-    labs(title = "(b) Local survival", 
-         y = "Local survival probability", 
-         x = "Nitrogen deposition")
-multiplot(p1, p2, cols = 2)
+```{r difffromrandomtab}
+kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Ellenberg value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average Ellenberg of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
+Shown are the results from linear model with the difference between disappeard/colonized species and random species as dependend variable and the sitemeasure (gradient) as predictor variable.") %>%
+  kable_styling() %>% 
+  group_rows("(a) Plants that disappeard from a site", 1, 4, bold = F, italic = T) %>%
+  group_rows("(b) Plants that newly colonized a site", 5, 8, bold = F, italic = T) %>%
+  add_header_above(c(" " = 2, "Difference from random" = 3, "Change along gradient" = 3))
 ```
 
-```{r periodeffect}
-# Temperature driven model
-modTcol <- glmer(Occ ~ factor(Period) * Temperature * T + (1|aID_SP) + (1|aID_STAO),
-               data = coldat, family = binomial)
-modTls <- glmer(Occ ~ factor(Period) * Temperature * T + (1|aID_SP) + (1|aID_STAO),
-               data = survdat, family = binomial)
-
-# Nitrogen driven model
-modNcol <- glmer(Occ ~ factor(Period) * NTOT * N + (1|aID_SP) + (1|aID_STAO),
-               data = coldat, family = binomial)
-modNls <- glmer(Occ ~ factor(Period) * NTOT * N + (1|aID_SP) + (1|aID_STAO),
-               data = survdat, family = binomial)
-```
-
-For the temperature driven and nitrogen driven models in Table \@ref(tab:resdifeffects) we additionally tested whether the strength of the effects on colonization probability and local survival differed from the first/second to the second/third survey period.
-For the temperatur driven models we found that the differences how species with different Ellenberg T values responded to the average annual temperature of a site tended to decrease from the first/second study period to the second/third study period (colonization probability: period x Ellenberg T x annual mean temperature of site = `r fixef(modTcol)["factor(Period)2:Temperature:T"]`, p = `r summary(modTcol)$coefficients["factor(Period)2:Temperature:T", "Pr(>|z|)"]`; local survival: period x Ellenberg T x annual mean temperature of site = `r fixef(modTls)["factor(Period)2:Temperature:T"]`, p = `r summary(modTls)$coefficients["factor(Period)2:Temperature:T", "Pr(>|z|)"]`).
-In contrast, however, the differences how species with different Ellenberg N values responded to the average N deposition of a site tended to increase from the first/second study period to the second/third study period (colonization probability: period x Ellenberg N x N deposition at site = `r fixef(modNcol)["factor(Period)2:NTOT:N"]`, p = `r summary(modNcol)$coefficients["factor(Period)2:NTOT:N", "Pr(>|z|)"]`; local survival: period x Ellenberg N x N deposition at site = `r fixef(modNls)["factor(Period)2:NTOT:N"]`, p = `r summary(modNls)$coefficients["factor(Period)2:NTOT:N", "Pr(>|z|)"]`).
-
-```{r figconsequences, fig.cap="Beschreibung einfügen."}
-# Change of correlation across study period
-getcor <- function(year) {
-  list(
-    corT = cor(surv %>% filter(yearP >= year & yearP <= (year + 4)) %>% dplyr::select(Temperature, T))[1,2],
-    corN = cor(surv %>% filter(yearP >= year & yearP <= (year + 4)) %>% dplyr::select(NTOT, N))[1,2])
-}
-tt <- map(2003:2013, getcor)
-res <- rbind(
-  tibble(cor = map_dbl(tt, "corN") / tt[[1]]$corN,
-              driver = "NTOT",
-              year = 2003:2013),
-  tibble(cor = map_dbl(tt, "corT") / tt[[1]]$corT,
-              driver = "Temperature",
-              year = 2003:2013))
-
-tt <- rbind(surv %>% transmute(CI = T, yr = yearPl - 2010, Type = "T", aID_STAO = aID_STAO),
-            surv %>% transmute(CI = N, yr = yearPl - 2010, Type = "N", aID_STAO = aID_STAO))
-m1 <- lmer(CI ~ Type * yr + (yr|aID_STAO), data = tt)
-m2 <- lmer(CI ~ Type + yr + (yr|aID_STAO), data = tt)
-anova(m1, m2)
-
-m3 <- lmer(CI ~ Type + yr + (yr|aID_STAO), data = tt)
-
-p1 <- ggplot(res, aes(x = year, y = cor, col = driver)) +
-  geom_point() +
-  geom_line() +
-  ylim(0.95, 1.05)
-res$yr <- res$year - 2010 
-cormod <- lm(cor ~ yr * driver, data = res)
-
-# Effect on N deposition model
-surv$yr <- surv$yearPl - 2010
-summary(glmer(SR ~ yr * NTOT * Temperature + (1|aID_STAO), data = surv, family = poisson))
-
-makemod <- function(year) {
-  mod <- glm(SR ~ Temperature + Precipitation + NTOT + Inclination, family = poisson,
-      data = surv %>% filter(yearP >= year & yearP <= (year + 4)))
-  arm::sim(mod, nsim)@coef
-}
-getestim <- function(x) {
-  list(meanN = mean(x[, "NTOT"]),
-       lowN = quantile(x[, "NTOT"], probs = 0.025),
-       upN = quantile(x[, "NTOT"], probs = 0.975),
-       meanT = mean(x[, "Temperature"]),
-       lowT = quantile(x[, "Temperature"], probs = 0.25),
-       upT = quantile(x[, "Temperature"], probs = 0.75))
-}
-tt <- map(2003:2013, makemod) %>% map(getestim)
-res <- rbind(
-  tibble(mean = map_dbl(tt, "meanN"),
-              low = map_dbl(tt, "lowN"),
-              up = map_dbl(tt, "upN"),
-              factor = "NTOT",
-              period = 2003:2013),
-    tibble(mean = map_dbl(tt, "meanT"),
-              low = map_dbl(tt, "lowT"),
-              up = map_dbl(tt, "upT"),
-              factor = "Temperature",
-              period = 2003:2013))
-pd <- position_dodge(0.5)
-p2 <- ggplot(res, aes(x = period, y = mean, col = factor)) +
-  geom_errorbar(aes(ymin = low, ymax = up), width = 0.5, position=pd) +
-  geom_point(position=pd) +
-  geom_line(position=pd) +
-  geom_abline(intercept = 0, slope = 0) +
-  ylim(-0.2, 0.05)
-multiplot(p1, p2, cols = 2)
- # summary(lm(mean ~ factor * period, data = res))
-```
-
-It thus seems that temperature is the main driver of species turnover in the first part of the study period and its importance is slightly decreasing over time and nitrogen deposition becomes relatively more important as a driver that shapes the species communities.
-We thus expect to see that the correlation of the mean temperature value of a community with the yearly annual temperature of the site (temperature correlation) is decreasing over the years while the correlation of the mean nutrient value of a community with the N deposition of the site (nitrogen correlation) is increasing over the years.
-Our results tent to be in line with this expectation (linear model: temporal trend of temperature correlation x temporal trend of nitrogen correlation = `r coef(cormod)["yr:driverTemperature"]`, p = `r  summary(cormod)$coefficients["yr:driverTemperature", "Pr(>|t|)"]`), however the change in the correlation over the study period remained week and the temperature correlation remained stronger than the nitrogen correlation over the entire study period (Left panel of Fig.
-\@ref(fig:figconsequences)).
-More important, however, was the effect on our model that we used to infer whether and how strong the spatial variation in N deposition is correlated with Nitrogen deposition.
-If we apply that model the estimated effect of N depositon on total species richness is increasing over the study period (right panel of \@ref(fig:figconsequences)).
+High species turnover at a site is the result of species that disappeared from the site and species that newly colonized the site.
+To better understand the factors that drive these changes we are particularly interested whether the species that disappeared or colonized the sites differed in Ellenberg values compared to what would be expected if the same number of species randomly disappeared or colonized the sites and whether there is a change along the site gradients.
 

### 2018-05-28 Manuscript.tex
Results neu strukturiert und Tabellen mit Resultaten eingefügt: vor allem Vergleich mit random species colonization und random species disappearance. Tabelle mit gradient Effekt auf zeitlichen Turnover.

--- 2018-05-07 Manuscript.tex
@@
 
-\title{xxx}
+\title{No sign of recovery}
 
-\author[1]{Tobias Roth}
+\author[1, 2]{Tobias Roth}
 
-\corrauthor[1]{Tobias Roth}{\href{mailto:t.roth@unibas.ch}{\nolinkurl{t.roth@unibas.ch}}}
+\corrauthor[1, 2]{Tobias Roth}{\href{mailto:t.roth@unibas.ch}{\nolinkurl{t.roth@unibas.ch}}}
 \author[2]{Lukas Kohli}
@@
 \begin{abstract}
-xxx
+Nitrogen (N) deposition is a major threat to biodiversity of many
+habitats. The recent intro- duction of cleaner technologies in
+Switzerland has let to reductions in the emissions of nitro- gen oxides,
+with affiliated decrease in N-deposition. We infered different drivers
+of community change (i.e.~Nitrogen deposition, climate warming, land-use
+change) in Swiss mountain hay meadows. The data were obtained from the
+Swiss biodiversity monitoring. Species gain and losses were best
+explainde by differences in Ellenberg value for temperature suggesting
+that climate warming was the most important driver of community change
+in Mountain hay meadows.
 % Dummy abstract text.
@@
 
-xxx
+Nitrogen (N) deposition is a major threat to biodiversity. The recent
+introduction of cleaner technologies has let to reductions in the
+emissions of nitrogen oxides, with affiliated decrease in N-deposition
+in many parts of Europe. However, it is an open question whether and how
+fast the reduction in N deposition rates will lead to the recovey of
+extant plant communities.
+
+One useful approach to understanding biodiversity change is through
+estimates of biodiversity turnover reflecting both immigration and
+extinction, often in a closed range of values
+(Hillebrand\_et\_al-2018.pdf).
+
+Here we infered mountain hay meadows in Switzerland. Explain why
+mountain hay meadows are important. Also explain other threats to
+mountain hay meadows (climate change, land-use change).
 
@@
 
+\subsubsection*{Community measures}\label{community-measures}
+\addcontentsline{toc}{subsubsection}{Community measures}
+
+The temporal turnover (i.e.~species exchange ratio sensu Hillebrand et
+al. (2018)) is the proportion of species that differ between two time
+points calculated as
+
+\(\text{Spatial turnover} = \frac{\text{Species gained} + \text{Species lost}}{\text{Total species observed in both timepoints}}\).
+
 \subsection*{Statistical analyses}\label{statistical-analyses}
@@
 
-Different measures of total community structure suggested that plant
-communities of mountain hay meadow were rather stable between 2003 and
-2017: For each of the three 5-year survey periods the averages of
-alpha-, beta- and gamma-diversity, average ellenberg values for
-temperature, nutrients, light, huminity and reaction, and average of
-species' canopy height, specific leaf area and seed mass did not vary
-much. For all measures, the average temperal trend per site did not
-differ significantly from zero (Table \ref{temporaltrends}). Note that
-beta- and gamma-diversity are note available for single sites and thus
-mixed models could not be applied.
+\begin{table}
 
-\begin{table}[ht]
+\caption{\label{tab:communitytrends}Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
+The temporal trends are given as  change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function) with site-ID as random intercept and slope effect.
+The measure of precision for the temporal trend is given as the 5\% and 95\% quantiles of the marginal posterior distribution.
+Finally, the probability that the linear trend is $> 0$ is given.
+Linear mixed models were not applicable for beta-diversity because measures for beta-diversity were not available for the single sites.}
 \centering
-\begin{tabular}{lrrrrr}
-  \hline
-Measures & Period 1 & Period 2 & Period 3 & Temporal-Trend & P-value \\ 
-  \hline
-Alpha-diversity & 46.36 & 46.72 & 46.45 & 0.002 & 0.896 \\ 
-  Beta-diversity & 0.68 & 0.65 & 0.65 &  &  \\ 
-  Gamma-Diversity & 517 & 529 & 517 &  &  \\ 
-  Temperature value & 3.12 & 3.14 & 3.14 & 0.013 & 0.060 \\ 
-  Huminity value & 2.99 & 2.98 & 2.99 & 0.006 & 0.405 \\ 
-  Nutrients value & 3.22 & 3.22 & 3.22 & -0.004 & 0.698 \\ 
-  Light value & 3.57 & 3.56 & 3.56 & -0.010 & 0.196 \\ 
-  Canopy height & -1.24 & -1.22 & -1.23 & 0.013 & 0.307 \\ 
-  Specific leaf area & 8.21 & 8.27 & 8.24 & 0.030 & 0.621 \\ 
-  Seed mass & -0.34 & -0.32 & -0.33 & 0.010 & 0.596 \\ 
-   \hline
+\begin{tabular}[t]{lrrrrrrr}
+\toprule
+Measures & Period 1 & Period 2 & Period 3 & Trend & 5\% & 95\% & Prob. for trend\\
+\midrule
+Alpha-diversity & 45.72 & 46.02 & 45.74 & 0.00 & -0.03 & 0.03 & 0.53\\
+Beta-diversity & 0.60 & 0.60 & 0.60 &  &  &  & \\
+Temperature value & 3.11 & 3.13 & 3.13 & 0.01 & 0.00 & 0.02 & 0.97\\
+Huminity value & 2.99 & 2.98 & 2.99 & 0.01 & -0.01 & 0.02 & 0.78\\
+Nutrients value & 3.20 & 3.20 & 3.20 & -0.01 & -0.02 & 0.01 & 0.32\\
+Light value & 3.56 & 3.55 & 3.55 & -0.01 & -0.02 & 0.00 & 0.09\\
+\bottomrule
 \end{tabular}
-\caption{Average measures of community structure for the three survey periods (in each period all sites are surveyed once).
-The temporal trends and p-values are based on linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution) with site-ID as random effect.
-Temporal-trends are given per 10 years.
-Linear mixed models could not be applied for beta- and gamma-diversity because measures are not available for the single sites.}
-\label{temporaltrends}
 \end{table}
 
-The temporal stability suggested by the community measures were,
-however, in contrast to the temporal turn-over of recorded species
-(i.e.~species exchange ratio sensu Hillebrand et al. (2018)). In average
-\(\pm\) SD the proportion of species that differed between two surveys
-was 18.7 \(\pm\) 6.8.
+The different measures of total community structure suggested that plant
+communities in mountain hay meadows were rather stable between 2003 and
+2017 and did not show a consistent increase or decrease over time (Table
+\ref{tab:communitytrends}): for each of the three 5-year survey periods
+the averages of alpha- and beta-diversity and the average Ellenberg
+values for temperature, huminity, nutrients and light did not vary much
+among the three sampling periods and the estimated trends were rather
+small. Except for average Ellenberg value for temperature, the 90\%
+credible-interval of the temporal trend contained zero. The results from
+the linear mixed models suggest that a linear temporal change was most
+likely for the community mean of the Ellenberg value for temperature
+(probability for an increase: 0.97), followed by the community mean of
+the Ellenberg light value (probaility for a decrease: 0.91) and it was
+least likely for the alpha-diversity (probability for an increase:
+0.53). The chance that the community mean of the nutrient value
+decreased between 2003 and 2017 was 0.68.
 
-\begin{verbatim}
-## 
-## Call:
-## lm(formula = Turnover ~ NTOT2007 + Hoehe + Neig + Expos, data = sites)
-## 
-## Residuals:
-##       Min        1Q    Median        3Q       Max 
-## -0.148019 -0.040275 -0.007297  0.042223  0.205389 
-## 
-## Coefficients:
-##               Estimate Std. Error t value Pr(>|t|)    
-## (Intercept)  3.468e-01  5.191e-02   6.682 6.98e-10 ***
-## NTOT2007    -4.629e-03  1.498e-03  -3.091  0.00246 ** 
-## Hoehe       -7.279e-05  2.263e-05  -3.216  0.00165 ** 
-## Neig        -1.816e-04  7.590e-04  -0.239  0.81125    
-## Expos        2.433e-05  2.401e-05   1.014  0.31278    
-## ---
-## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
-## 
-## Residual standard error: 0.06522 on 125 degrees of freedom
-## Multiple R-squared:  0.1085, Adjusted R-squared:  0.07997 
-## F-statistic: 3.803 on 4 and 125 DF,  p-value: 0.005939
-\end{verbatim}
+\begin{table}
+
+\caption{\label{tab:turnovertab}Change of species turnover along the four gradients.
+The slopes along the gradients (estimate) are given as the change per 10 years of the logit-probability of species that differed between two surveys.
+Estimates and the 5\% and 95\% quantiles of the marginal posterior distribution obtained from a Binomial-GLMM with the proportion of species that differed between two surveys as dependent variable and the site gradients and period as predictors and site-ID as random effect.}
+\centering
+\begin{tabular}[t]{lrrr}
+\toprule
+Gradient & Estimate & 5\% & 95\%\\
+\midrule
+Annual mean temperature & 0.06 & 0.00 & 0.11\\
+Annual mean precipitation & -0.02 & -0.11 & 0.08\\
+Nitrogen deposition & -0.20 & -0.35 & -0.05\\
+Inclination & -0.04 & -0.10 & 0.03\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+This temporal stability as inferred from the community measures was,
+however, in contrast to a rather large observed temporal turnover of
+species. The average percentage \(\pm\) SD of species that differed
+between the first and second survey at a site was 37.65 \(\pm\) 10.43\%
+and the percentage of species that differ between the second and third
+survey was 35.66 \(\pm\) 10.36\%. Thus, it seemed that the turnover from
+the first/second survey to the turnover of the second/third survey
+moderately decreased (90\% Credible interval of the change in turnover
+estimated from the Binomial generalized linear mixed model that also
+contained site variables: -0.15 - -0.02). The ``
+
+\begin{table}
+
+\caption{\label{tab:difffromrandomtab}Difference in the average Ellenberg of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
+Shown are the results from linear model with the difference between disappeard/colonized species and random species as dependend variable and the sitemeasure (gradient) as predictor variable.}
+\centering
+\begin{tabular}[t]{llcccccc}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Difference from random} & \multicolumn{3}{c}{Change along gradient} \\
+\cmidrule(l{2pt}r{2pt}){3-5} \cmidrule(l{2pt}r{2pt}){6-8}
+Ellenberg value & Gradient & Estimate & 5\% & 90\% & Estimate & 5\% & 90\%\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textit{(a) Plants that disappeard from a site}}\\
+\hspace{1em}Temperature & Annual mean temperature & -0.012 & -0.034 & 0.009 & 0.007 & -0.002 & 0.016\\
+\hspace{1em}Humidity & Annual mean precipitation & -0.003 & -0.038 & 0.033 & 0.008 & -0.012 & 0.029\\
+\hspace{1em}Nutrients & Nitrogen deposition & -0.012 & -0.052 & 0.028 & 0.000 & -0.046 & 0.046\\
+\hspace{1em}Light & Inclination & -0.022 & -0.047 & 0.004 & -0.002 & -0.024 & 0.022\\
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textit{(b) Plants that newly colonized a site}}\\
+\hspace{1em}Temperature & Annual mean temperature & 0.018 & 0.001 & 0.034 & -0.001 & -0.008 & 0.006\\
+\hspace{1em}Humidity & Annual mean precipitation & 0.023 & -0.010 & 0.054 & -0.004 & -0.023 & 0.014\\
+\hspace{1em}Nutrients & Nitrogen deposition & -0.082 & -0.117 & -0.048 & 0.062 & 0.024 & 0.102\\
+\hspace{1em}Light & Inclination & -0.039 & -0.062 & -0.016 & 0.010 & -0.009 & 0.031\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-xxx
+\begin{itemize}
+\item
+  Any empty space that could be caused by any disturbance that let to
+  the local disappearance of species is likely to be fiellied by
+  eutrophic species. Disturbance depends on the site, while colonization
+  depends on species characteristics.
+\item
+  The rather large spatial turnover might be partly explained by species
+  that remained undetected in one of the surveys, but it might also be
+  the result of species that newly colonized sites (species gains) and
+  species that truely disappeared from sites (species losses).
+\item
+  Altough N deposition considerabely declinded between 2005 and 2015, we
+  could not detect major shifts in plant community structure during the
+  same time period.
+\item
+  Eutrophic species have rather high local survival across the entire
+  deposition gradient, while oligotrophic species have much reduced
+  local survival at high N deposition. This suggests that it takes much
+  more time to replace eutrophic by oligotrophic species than replacing
+  oligotrophic by eutrophic species.
+\item
+  Our data on colonization and local survival (i.e.~temporal variation)
+  confirm the empirical critical loads that we infered from anlysing
+  spatial co-variation of N deposition and species richness.
+\item
+  Local survival is higher for low temperature plants --\textgreater{}
+  This could explain the decrease in cumminity change along elevation.
+  This could also explain the differences at mount summets were space
+  was empty in the beginning.
+\item
+  Climatic effects are more likely to be reversed thant effects due to
+  fertilization.
+\end{itemize}
+
+\subsubsection{Space for time
+substitution}\label{space-for-time-substitution}
+
+Often observational studies infer the change of plant diversity along a
+gradient of N deposi- tion. Thus, they infer how the spatial variation
+in species richness is related to N deposition and assume that this
+spatial variation in species richness arose because over time some areas
+lost more species than others because they chronically experienced
+higher N deposition. Alt- hough there is evidence supporting the use of
+such a `space for time substitution' for detecting the effects of N
+deposition on plant diversity (Stevens et al. 2010), they can not
+replace stud- ies that relate temporal patterns in species with N
+deposition (De Schrijver et al. 2011). While recovery of acidified
+surface waters has been well investigated (De Vries et al. 2015), there
+are only a limited number of studies inferring temporal trends of plant
+species diversity relat- ed to varying amounts of N-deposition. Storkey
+et al. (2015) demonstrated a positive response of biodiversity to
+reducing N addition from either atmospheric pollution or fertilizers in
+the Park Grass Experiment: «The proportion of legumes, species richness
+and diversity increased across the experiment between 1991 and 2012 as
+N-deposition declined». For forest floor vegetation in permanent plots
+across Europe the exceedance of critical loads of N over a peri- od from
+9 to 42 years had negative effects on the cover of oligotrophic plant
+species, i.e spe- cies that prefer nutrient-poor soils, although species
+richness remained constant (Dirnböck et al. 2014). Another example of
+recovery in eutrophicated habitats gives the recovery of species
+richness in previously fertilized plots (Clark and Tilman 2008). In this
+study, the recorded recovery in species richness within one or two
+decade was likely due to the species rich vege- tation surrounding the
+experimental plots, from where immigration was easily feasible.
+
+\begin{quote}
+Hier sagen, dass der CRL identisch wie der CRL von räumlichen
+Zusammenhängen ist.
+\end{quote}
 

### 2018-05-28 R/Settings.R

@@ -13,7 +13,7 @@ rm(list=ls(all=TRUE))
 #------------------------------------------------------------------------------------------------------
 library(knitr)
 library(kableExtra)
-library(xtable)
+# library(xtable)
 library(tidyverse)
 library(broom)
 library(rstanarm)
@@ -53,9 +53,6 @@ pd <- position_dodge(0.2)
 tcol <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", 
           "#D55E00", "#CC79A7")
 
-# Settings for xtable()
-options(xtable.comment = FALSE)
-ndigits = 2
 
 
 

### 2018-05-28 Manuscript.tex
Manuskript neu berechnet.

@@
 estimated from the Binomial generalized linear mixed model that also
-contained site variables: -0.15 - -0.02). The ``
+contained site variables: -0.15 - -0.02). Variation in species turnover
+was largest along the Nitrogen deposition gradient with highest species
+turnover at sites with low Nitrogen depostion \ref{tab:turnovertab}. The
+other three gradients were less inportant to explain the variation in
+species turnover among sites.
 
@@
 \end{table}
+
+High species turnover at a site is the result of species that
+disappeared from the site and species that newly colonized the site. To
+better understand the factors that drive these changes we are
+particularly interested whether the species that disappeared or
+colonized the sites differed in Ellenberg values compared to what would
+be expected if the same number of species randomly disappeared or
+colonized the sites and whether there is a change along the site
+gradients.
 

### 2018-05-29 Manuscript.Rmd
Figure with colonization probability and loacal survival of oligotrophic species vs eutrophic species added. Figure include an estimate of the Critical load.

@@
 Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model that also contained site variables: `r CrI`).
-Variation in species turnover was largest along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen depostion \@ref(tab:turnovertab).
+Variation in species turnover was largest along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition (Table \@ref(tab:turnovertab)).
 The other three gradients were less inportant to explain the variation in species turnover among sites.
@@
 ```{r difffromrandomtab}
-kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Ellenberg value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average Ellenberg of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
+kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Ellenberg value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average Ellenberg value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
 Shown are the results from linear model with the difference between disappeard/colonized species and random species as dependend variable and the sitemeasure (gradient) as predictor variable.") %>%
@@
 High species turnover at a site is the result of species that disappeared from the site and species that newly colonized the site.
-To better understand the factors that drive these changes we are particularly interested whether the species that disappeared or colonized the sites differed in Ellenberg values compared to what would be expected if the same number of species randomly disappeared or colonized the sites and whether there is a change along the site gradients.
+To better understand the factors that drive these changes we are particularly interested whether the species that disappeared or colonized the sites differed in Ellenberg values compared to what would be expected if the same number of species randomly disappeared or colonized the sites (i.e.
+random disappearance and random colonization) and whether there is a change along the gradients.
+It seems that the Ellenberg values of newly colonizing species differened more from random colonization than the Ellenberg values of disappearing species from random disappearance (Table \@ref(tab:difffromrandomtab)).
+For colonizing species, we found the largest differences from random colonization in the Ellenberg value for nutrients: at sites with nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower Ellenberg value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreases with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
+
+While colonizing species had higher temperature values compared to what we would be expected under random colonization, the differences between colonizing species and random colonization was about four times smaller compared to the Ellenberg value for nutrients.
+Nevertheless, the variation in Ellenberg value for temperature was impportant to explain the total species turnover.
+This is because, disappearing species tend to have lower temperature value than randomly dieappearing species leading to an overall replacement of species with lower temperature value with species with higher temperature values.
+This was not the case for the Ellenberg value for nutrients: it was also species with lower nutrients values that tended to be more likely to disappear than randomly disappearing species (Table \@ref(tab:difffromrandomtab)).
+See also Appendix A were we present detailed results for the comparision between colonizating and disappearing species with randomly selected species.
+
+```{r neff, fig.cap="Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
+Given are means and 95\\%-Credible Intervals from logistic linear mixed models."}
+nsim <- 100
+X <- 5:30
+Xsd <- (X - 10) / 10
+
+library(sp)
+library(rgeos)
+
+# Help functions to get posterior samples
+getposterior <- function(x) {
+  EL1 <- plogis(simdat[x,1] + Xsd * simdat[x,2] + 2 * simdat[x,3] +  Xsd * 2 * simdat[x,4])
+  EL2 <- plogis(simdat[x,1] + Xsd * simdat[x,2] + 4 * simdat[x,3] +  Xsd * 4 * simdat[x,4])
+  ll1 <- Lines(list(Line(matrix(c(X, EL1), nc = 2, byrow = F))), ID = "1")
+  ll2 <- Lines(list(Line(matrix(c(X, EL2), nc = 2, byrow = F))), ID = "1")
+  sl1 <- SpatialLines(list(ll1), proj4string = CRS("+init=epsg:4269"))
+  sl2 <- SpatialLines(list(ll2), proj4string = CRS("+init=epsg:4269"))
+  tCL <- NA
+  try(tCL <- gIntersection(sl1, sl2, byid = TRUE)@coords[1, "x"])
+  list( 
+    `EL = 2` = EL1,
+    `EL = 4` = EL2,
+    CL = tCL)
+}
+
+compileres <- function(EL) {
+  tt <- map(1:length(X), function(s) map(post, paste(EL)) %>% map_dbl(s))
+  tibble(mean = map_dbl(tt, mean),
+         low = map_dbl(tt, quantile, probs = 0.025),
+         high = map_dbl(tt, quantile, probs = 0.975),
+         X = X,
+         value = paste(EL)
+  )
+}
+
+# Colonization
+load("Modelfit/colonization-N.RData")
+simdat <- arm::sim(mod, nsim)@fixef
+post <- map(1:nsim, getposterior)
+res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
+CLs <- map_dbl(post, "CL")
+p1 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
+  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
+  geom_line() +
+    labs(title = "(a) Colonization", 
+         y = "Colonization probability", 
+         x = "Nitrogen deposition") +
+  geom_vline(xintercept = mean(CLs, na.rm = TRUE)) +
+  geom_vline(xintercept = quantile(CLs, probs = 0.05, na.rm = TRUE), lty = 2) +
+  geom_vline(xintercept = quantile(CLs, probs = 0.95, na.rm = TRUE), lty = 2)
+
+# Local survival
+load("Modelfit/localsurvival-N.RData")
+simdat <- arm::sim(mod, nsim)@fixef
+post <- map(1:nsim, getposterior)
+res <- rbind(compileres("EL = 2"), compileres("EL = 4"))
+CLs <- map_dbl(post, "CL") 
+p2 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
+  geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
+  geom_line() +
+  labs(title = "(b) Local survival", 
+       y = "Local survival probability", 
+       x = "Nitrogen deposition")+
+  geom_vline(xintercept = mean(CLs, na.rm = TRUE)) +
+  geom_vline(xintercept = quantile(CLs, probs = 0.05, na.rm = TRUE), lty = 2) +
+  geom_vline(xintercept = quantile(CLs, probs = 0.95, na.rm = TRUE), lty = 2)
+multiplot(p1, p2, cols = 2)
+```
 

### 2018-05-29 Manuscript.tex
Figure with colonization probability and loacal survival of oligotrophic species vs eutrophic species added. Figure include an estimate of the Critical load.

--- 2018-05-28 Manuscript.tex
@@
 estimated from the Binomial generalized linear mixed model that also
-contained site variables: -0.15 - -0.02). The ``
+contained site variables: -0.15 - -0.02). Variation in species turnover
+was largest along the Nitrogen deposition gradient with highest species
+turnover at sites with low Nitrogen deposition (Table
+\ref{tab:turnovertab}). The other three gradients were less inportant to
+explain the variation in species turnover among sites.
 
@@
 
-\caption{\label{tab:difffromrandomtab}Difference in the average Ellenberg of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
+\caption{\label{tab:difffromrandomtab}Difference in the average Ellenberg value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
 Shown are the results from linear model with the difference between disappeard/colonized species and random species as dependend variable and the sitemeasure (gradient) as predictor variable.}
@@
 \end{table}
+
+High species turnover at a site is the result of species that
+disappeared from the site and species that newly colonized the site. To
+better understand the factors that drive these changes we are
+particularly interested whether the species that disappeared or
+colonized the sites differed in Ellenberg values compared to what would
+be expected if the same number of species randomly disappeared or
+colonized the sites (i.e.~random disappearance and random colonization)
+and whether there is a change along the gradients. It seems that the
+Ellenberg values of newly colonizing species differened more from random
+colonization than the Ellenberg values of disappearing species from
+random disappearance (Table \ref{tab:difffromrandomtab}). For colonizing
+species, we found the largest differences from random colonization in
+the Ellenberg value for nutrients: at sites with nitrogen deposition of
+10 kg ha\(^{-1}\) yr\(^{-1}\) the newly colonizing species had in
+average a lower Ellenberg value for nutrients than species under random
+disappearance (column ``Difference from random'' in Table
+\ref{tab:difffromrandomtab}), but this differences between colonizing
+species and random colonization decreases with increasing N deposition
+(column ``Change along gradient'' in Table \ref{tab:difffromrandomtab}).
+
+While colonizing species had higher temperature values compared to what
+we would be expected under random colonization, the differences between
+colonizing species and random colonization was about four times smaller
+compared to the Ellenberg value for nutrients. Nevertheless, the
+variation in Ellenberg value for temperature was impportant to explain
+the total species turnover. This is because, disappearing species tend
+to have lower temperature value than randomly dieappearing species
+leading to an overall replacement of species with lower temperature
+value with species with higher temperature values. This was not the case
+for the Ellenberg value for nutrients: it was also species with lower
+nutrients values that tended to be more likely to disappear than
+randomly disappearing species (Table \ref{tab:difffromrandomtab}). See
+also Appendix A were we present detailed results for the comparision
+between colonizating and disappearing species with randomly selected
+species.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Manuscript_files/figure-latex/neff-1} \caption{Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
+Given are means and 95\%-Credible Intervals from logistic linear mixed models.}\label{fig:neff}
+\end{figure}
 

### 2018-05-29 R/Run_models.R

@@ -67,12 +67,12 @@ mod <- stan_glmer(cbind(change, nochange) ~ period + Temperature + Precipitation
 save(mod, file = "Modelfit/turnover-trend.RData")
 
 #------------------------------------------------------------------------------------------------------
-# 
+# Modell for the colonization and local survival probabilities
 #------------------------------------------------------------------------------------------------------
-mod <- stan_glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial, cores = ncores)
-save(mod, file = "Modelfit/colonization-T.RData")
+mod <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial)
+save(mod, file = "Modelfit/colonization-N.RData")
 
-mod <- stan_glmer(Occ ~ Temperature * T + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial, cores = ncores)
-save(mod, file = "Modelfit/localsurvival-T.RData")
+mod <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial)
+save(mod, file = "Modelfit/localsurvival-N.RData")
 
 

### 2018-05-29 R/Settings.R

@@ -13,10 +13,10 @@ rm(list=ls(all=TRUE))
 #------------------------------------------------------------------------------------------------------
 library(knitr)
 library(kableExtra)
-# library(xtable)
 library(tidyverse)
 library(broom)
 library(rstanarm)
+library(arm)
 library(simba)
 library(Rmisc)
 

### 2018-05-29 Manuscript.Rmd
Figure with consequences for the analyses of the nitrogen deposition effect using current state data.

@@
 
-library(sp)
-library(rgeos)
-
 # Help functions to get posterior samples
@@
 
+```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of Nitrogen deposition on total species richness estimated from applying the Poisson-GLM with species richness as dependend variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from a five-year interval.
+Within every five-year interval all plots were sampled once."}
+# Effect on N deposition model
+makemod <- function(year) {
+  mod <- glm(SR ~ Elevation + I(Elevation^2) + Precipitation + NTOT + Inclination, family = poisson,
+      data = surv %>% filter(yearP >= year & yearP <= (year + 4)))
+  arm::sim(mod, nsim)@coef
+}
+
+getestim <- function(x) {
+  list(meanN = mean(x[, "NTOT"]),
+       lowN = quantile(x[, "NTOT"], probs = 0.05),
+       upN = quantile(x[, "NTOT"], probs = 0.95))
+}
+
+tt <- map(2003:2013, makemod) %>% map(getestim)
+res <- tibble(mean = map_dbl(tt, "meanN"),
+              low = map_dbl(tt, "lowN"),
+              up = map_dbl(tt, "upN"),
+              factor = "NTOT",
+              period = 2003:2013)
+
+summary(lm(mean ~ period, data = res))
+
+ggplot(res, aes(x = period, y = mean)) +
+  geom_errorbar(aes(ymin = low, ymax = up), width = 0.2, lty = 2, col = "grey65") +
+  geom_point() +
+  geom_line() +
+  geom_smooth(method = lm) +
+  geom_abline(intercept = 0, slope = 0, lty = 3) +
+  labs(x = "Study period", y = "Effect of Nitrogen deposition")
+```
+
 # Discussion {-}

### 2018-05-29 Manuscript.tex
Figure with consequences for the analyses of the nitrogen deposition effect using current state data.

@@
 
+\begin{verbatim}
+## 
+## Call:
+## lm(formula = mean ~ period, data = res)
+## 
+## Residuals:
+##        Min         1Q     Median         3Q        Max 
+## -0.0233852 -0.0119052  0.0008674  0.0105638  0.0238631 
+## 
+## Coefficients:
+##              Estimate Std. Error t value Pr(>|t|)  
+## (Intercept)  6.804642   3.297836   2.063   0.0691 .
+## period      -0.003415   0.001642  -2.079   0.0674 .
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Residual standard error: 0.01723 on 9 degrees of freedom
+## Multiple R-squared:  0.3245, Adjusted R-squared:  0.2494 
+## F-statistic: 4.323 on 1 and 9 DF,  p-value: 0.06736
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=0.5\linewidth]{Manuscript_files/figure-latex/figconsequences-1} \caption{Effect size of Nitrogen deposition on total species richness estimated from applying the Poisson-GLM with species richness as dependend variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from a five-year interval.
+Within every five-year interval all plots were sampled once.}\label{fig:figconsequences}
+\end{figure}
+
 \section*{Discussion}\label{discussion}

### 2018-05-29 R/Settings.R

@@ -19,6 +19,8 @@ library(rstanarm)
 library(arm)
 library(simba)
 library(Rmisc)
+library(sp)
+library(rgeos)
 
 #------------------------------------------------------------------------------------------------------
 # Load data

### 2018-05-29 Manuscript.Rmd
Edits in results section and small improvements of figures.

@@
 # Results {-}
-
 ```{r communitytrendscalc}
@@
 
-```{r communitytrends, echo=FALSE}
+```{r communitytrendstab, echo=FALSE}
 options(knitr.kable.NA = '')
@@
 
-The different measures of total community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a consistent increase or decrease over time (Table \@ref(tab:communitytrends)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, huminity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
+The different measures of total community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a consistent increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, huminity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
 Except for average Ellenberg value for temperature, the 90% credible-interval of the temporal trend contained zero.
@@
 For colonizing species, we found the largest differences from random colonization in the Ellenberg value for nutrients: at sites with nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower Ellenberg value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreases with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
+Thus at high Nitrogen deposition colonizing species did not differ from random species (see Figure 3b in Appendix A).
 
-While colonizing species had higher temperature values compared to what we would be expected under random colonization, the differences between colonizing species and random colonization was about four times smaller compared to the Ellenberg value for nutrients.
-Nevertheless, the variation in Ellenberg value for temperature was impportant to explain the total species turnover.
-This is because, disappearing species tend to have lower temperature value than randomly dieappearing species leading to an overall replacement of species with lower temperature value with species with higher temperature values.
+While colonizing species had higher temperature values compared to what we would be expected under random colonization, the differences between colonizing species and random colonization was about four times smaller compared to the difference in Ellenberg value for nutrients between colonizing species and random species.
+Nevertheless, the variation in Ellenberg value for temperature seemed impportant to explain the total species turnover.
+This is because, disappearing species tend to have lower temperature value than randomly disappearing species leading to an overall replacement of species with lower temperature value with species with higher temperature values.
 This was not the case for the Ellenberg value for nutrients: it was also species with lower nutrients values that tended to be more likely to disappear than randomly disappearing species (Table \@ref(tab:difffromrandomtab)).
-See also Appendix A were we present detailed results for the comparision between colonizating and disappearing species with randomly selected species.
+See also Appendix A were we present detailed results for the comparision between colonizing or disappearing species with randomly selected species.
 
-```{r neff, fig.cap="Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
+```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
 Given are means and 95\\%-Credible Intervals from logistic linear mixed models."}
-nsim <- 100
-X <- 5:30
+X <- 1:30
 Xsd <- (X - 10) / 10
@@
        y = "Colonization probability", 
-       x = "Nitrogen deposition") +
+       x = expression(paste("Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")"))) +
   geom_vline(xintercept = mean(CLs, na.rm = TRUE)) +
   geom_vline(xintercept = quantile(CLs, probs = 0.05, na.rm = TRUE), lty = 2) +
-  geom_vline(xintercept = quantile(CLs, probs = 0.95, na.rm = TRUE), lty = 2)
+  geom_vline(xintercept = quantile(CLs, probs = 0.95, na.rm = TRUE), lty = 2) +
+  xlim(0,30)
 
@@
 CLs <- map_dbl(post, "CL")
+
 p2 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
@@
        y = "Local survival probability", 
-       x = "Nitrogen deposition")+
+       x = expression(paste("Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")"))) +
   geom_vline(xintercept = mean(CLs, na.rm = TRUE)) +
   geom_vline(xintercept = quantile(CLs, probs = 0.05, na.rm = TRUE), lty = 2) +
-  geom_vline(xintercept = quantile(CLs, probs = 0.95, na.rm = TRUE), lty = 2)
+  geom_vline(xintercept = quantile(CLs, probs = 0.95, na.rm = TRUE), lty = 2) +
+  xlim(0,30)
+
 multiplot(p1, p2, cols = 2)
 ```
+
+In Fig.
+\@ref(fig:cl) we compare the colonization and local survival probability of oligotrophic (Ellenberg value of nutrients = 2) and eutrophic (Ellenberg value of nutrients = 4) species along the Nitrogen deposition gradient.
+Colonization probability is the sa
+
+While we could not detect a consistent decrease in average Ellenberg value for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate affects the spatial variation of spcies richness: sites with low Nitrogen deposition become more species rich over time resulting in steeper slope of the negative relationship between N deposition and species richness.
+Consequently, if we apply a model as in @Roth2013 to infer the effects of N deposition on the spatial variation of species richness at different time points, the resulting effect size (i.e.
+the slope) becomes more negative over time (Fig.
+\@ref(fig:figconsequences)).
 
@@
 
-summary(lm(mean ~ period, data = res))
-
 ggplot(res, aes(x = period, y = mean)) +
@@
   geom_abline(intercept = 0, slope = 0, lty = 3) +
-  labs(x = "Study period", y = "Effect of Nitrogen deposition")
+  labs(x = "Study period", y = "Effect of Nitrogen deposition") + 
+  scale_x_continuous(breaks = seq(2004, 2012, 2), labels = paste0(seq(2004, 2012, 2), "-\n", seq(2008, 2016, 2)))
 ```

### 2018-05-29 Manuscript.tex
Edits in results section and small improvements of figures.

--- 2018-05-29 Manuscript.tex
@@
 
-\caption{\label{tab:communitytrends}Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
+\caption{\label{tab:communitytrendstab}Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
 The temporal trends are given as  change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function) with site-ID as random intercept and slope effect.
@@
 2017 and did not show a consistent increase or decrease over time (Table
-\ref{tab:communitytrends}): for each of the three 5-year survey periods
-the averages of alpha- and beta-diversity and the average Ellenberg
-values for temperature, huminity, nutrients and light did not vary much
-among the three sampling periods and the estimated trends were rather
-small. Except for average Ellenberg value for temperature, the 90\%
-credible-interval of the temporal trend contained zero. The results from
-the linear mixed models suggest that a linear temporal change was most
-likely for the community mean of the Ellenberg value for temperature
-(probability for an increase: 0.97), followed by the community mean of
-the Ellenberg light value (probaility for a decrease: 0.91) and it was
-least likely for the alpha-diversity (probability for an increase:
-0.53). The chance that the community mean of the nutrient value
-decreased between 2003 and 2017 was 0.68.
+\ref{tab:communitytrendstab}): for each of the three 5-year survey
+periods the averages of alpha- and beta-diversity and the average
+Ellenberg values for temperature, huminity, nutrients and light did not
+vary much among the three sampling periods and the estimated trends were
+rather small. Except for average Ellenberg value for temperature, the
+90\% credible-interval of the temporal trend contained zero. The results
+from the linear mixed models suggest that a linear temporal change was
+most likely for the community mean of the Ellenberg value for
+temperature (probability for an increase: 0.97), followed by the
+community mean of the Ellenberg light value (probaility for a decrease:
+0.91) and it was least likely for the alpha-diversity (probability for
+an increase: 0.53). The chance that the community mean of the nutrient
+value decreased between 2003 and 2017 was 0.68.
 
@@
 (column ``Change along gradient'' in Table \ref{tab:difffromrandomtab}).
+Thus at high Nitrogen deposition colonizing species did not differ from
+random species (see Figure 3b in Appendix A).
 
@@
 colonizing species and random colonization was about four times smaller
-compared to the Ellenberg value for nutrients. Nevertheless, the
-variation in Ellenberg value for temperature was impportant to explain
-the total species turnover. This is because, disappearing species tend
-to have lower temperature value than randomly dieappearing species
-leading to an overall replacement of species with lower temperature
-value with species with higher temperature values. This was not the case
-for the Ellenberg value for nutrients: it was also species with lower
-nutrients values that tended to be more likely to disappear than
-randomly disappearing species (Table \ref{tab:difffromrandomtab}). See
-also Appendix A were we present detailed results for the comparision
-between colonizating and disappearing species with randomly selected
-species.
+compared to the difference in Ellenberg value for nutrients between
+colonizing species and random species. Nevertheless, the variation in
+Ellenberg value for temperature seemed impportant to explain the total
+species turnover. This is because, disappearing species tend to have
+lower temperature value than randomly disappearing species leading to an
+overall replacement of species with lower temperature value with species
+with higher temperature values. This was not the case for the Ellenberg
+value for nutrients: it was also species with lower nutrients values
+that tended to be more likely to disappear than randomly disappearing
+species (Table \ref{tab:difffromrandomtab}). See also Appendix A were we
+present detailed results for the comparision between colonizing or
+disappearing species with randomly selected species.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{Manuscript_files/figure-latex/neff-1} \caption{Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
-Given are means and 95\%-Credible Intervals from logistic linear mixed models.}\label{fig:neff}
+\includegraphics[width=1\linewidth]{Manuscript_files/figure-latex/cl-1} \caption{Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
+Given are means and 95\%-Credible Intervals from logistic linear mixed models.}\label{fig:cl}
+\end{figure}
+
+In Fig. \ref{fig:cl} we compare the colonization and local survival
+probability of oligotrophic (Ellenberg value of nutrients = 2) and
+eutrophic (Ellenberg value of nutrients = 4) species along the Nitrogen
+deposition gradient. Colonization probability is the sa
+
+While we could not detect a consistent decrease in average Ellenberg
+value for nutrients (Table \ref{tab:communitytrendstab}), the higher
+colonization rate of species with low nutrient value at sites with low
+deposition rate affects the spatial variation of spcies richness: sites
+with low Nitrogen deposition become more species rich over time
+resulting in steeper slope of the negative relationship between N
+deposition and species richness. Consequently, if we apply a model as in
+Roth et al. (2013) to infer the effects of N deposition on the spatial
+variation of species richness at different time points, the resulting
+effect size (i.e.~the slope) becomes more negative over time (Fig.
+\ref{fig:figconsequences}).
+
+\begin{figure}
+\includegraphics[width=0.5\linewidth]{Manuscript_files/figure-latex/figconsequences-1} \caption{Effect size of Nitrogen deposition on total species richness estimated from applying the Poisson-GLM with species richness as dependend variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from a five-year interval.
+Within every five-year interval all plots were sampled once.}\label{fig:figconsequences}
 \end{figure}
@@
 
+\hypertarget{ref-Roth2013}{}
+Roth, Tobias, Lukas Kohli, Beat Rihm, and Beat Achermann. 2013.
+``Nitrogen Deposition Is Negatively Related to Species Richness and
+Species Composition of Vascular Plants and Bryophytes in Swiss Mountain
+Grassland.'' \emph{Agriculture, Ecosystems \& Environment} 178: 121--26.
+doi:\href{https://doi.org/10.1016/j.agee.2013.07.002}{10.1016/j.agee.2013.07.002}.
+
 \end{document}

### 2018-05-29 R/Settings.R

@@ -55,6 +55,7 @@ pd <- position_dodge(0.2)
 tcol <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", 
           "#D55E00", "#CC79A7")
 
-
+# Numer of simulations from posterior distirbution
+nsim <- 100
 
 

### 2018-05-30 Manuscript.Rmd
Mit Methodenkapitel begonnen.

@@
 ---
-title: No sign of recovery 
+title: Recovery of mountain plant communities in response to reductions in Nitrogen emissions is hidden by other drivers of global change
 preprint: false
@@
   Nitrogen (N) deposition is a major threat to biodiversity of many habitats.
-The recent intro- duction of cleaner technologies in Switzerland has let to reductions in the emissions of nitro- gen oxides, with affiliated decrease in N-deposition.
+The recent introduction of cleaner technologies in Switzerland has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition.
 We infered different drivers of community change (i.e.
@@
 The data were obtained from the Swiss biodiversity monitoring.
-  Species gain and losses were best explainde by differences in Ellenberg value for temperature suggesting that climate warming was the most important driver of community change in Mountain hay meadows.
 
@@
 ## Monitoring data {-}
-- Selection of sample sites based on 1366 K_Standort.csv column "E23_1366".
-- Three surveys 2003-2007, 2008-2012 and 2013 - 2017.
+We analysed the presence/absence of vascular plants sampled within the scope of Switzerland's Biodiversity Monitoring (BDM) programme that was launched in 2001 to monitor Switzerland's biodiversity and to comply with the Convention on Biological Diversity of Rio de Janeiro [@Weber2004].
+The sampling sites were circles with a size of 10 m$^2$ and data collection was carried out by qualified botanists who visited each sampling site twice within the same season.
+During each visit all the vascular plant species detected on the plot were recorded.
+After the sampling of the plant data the botanists also assigned a habitat type to each sampling site according to the classification system developed for Switzerland [@Delarze2008].For more details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
 
-## Plant traits {-}
-Functional traits:
+We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all `r nrow(sites)` sampling sites in mountain hay meadows (EUNIS E2.3).
+We analyced the data from 2003 to 2017.
+During that study period each sampling site was surveyed once per five-year period: the first period lasted from 2003 to 2007, the second from 2008 to 2012 and the third from 2013 to 2017.
 
-- SLA: specific leaf area
-- CH: canopy height
-- SM: Seed mass
+## Community measures {-}
+For each survey (a survey conists of the two visits within the same season) we calculated the following community measures: (1) The number of recorded species (species richness), (2) the mean pair-wise disimilarity index (beta-diversity), (3) the community mean of the Ellenberg temperature values of recorded species (community temperature index), (4) the community mean of the Ellenberg humidity values of recorded species (community humidity index), (5) the community mean of the Ellenberg nutrients values of recorded species (community nutrients index) and (6) the community mean of the Ellenberg light values of recorded species (community light index).
 
+Additionally we measured the temporal turnover (i.e. species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points calculated as
+
+$\text{Temporal turnover} = \frac{\text{Species gained} + \text{Species lost}}{\text{Total species observed in both timepoints}}$.
+
+## Ellenberg values and gradients {-}
 Ellenberg indicator values:
 
-- L: light
-- N: Nutrient contentent
 - T: Temperature
 - F: Huminity
+- N: Nutrient contentent
+- L: light
 
-Community measures:
+Gradients among studied sites:
 
-- Species richness: number of recorded species per $10m^2$.
-- Spatial turnover (beta-diversity): Average turnover between all pair-wise combinations of study plots.
-- gamma diversity: Total number of species recorded in all study plots.
-
-### Community measures {-}
-
-The temporal turnover (i.e. species exchange ratio sensu @Hillebrand2018) is the proportion of species that differ between two time points calculated as
-
-$\text{Spatial turnover} = \frac{\text{Species gained} + \text{Species lost}}{\text{Total species observed in both timepoints}}$.
+- Mean annual temperature
+- Mean annual precipitation
+- Nitrogen deposition
+- Inclination: The steaper the site the less intensive a site. 
+- Environmental variables were standardized.
 
 ## Statistical analyses {-}
-Environmental variables were standardized.
-
-```{r helpfunctions}
-# Help functions to get posterior samples
-getposterior <- function(x) {
-  list( 
-    `EL = 2` = plogis(simdat[x,1] + Xsd * simdat[x,2] - simdat[x,3] -  Xsd * simdat[x,4]),
-    `EL = 4` = plogis(simdat[x,1] + Xsd * simdat[x,2] + simdat[x,3] +  Xsd * simdat[x,4]))
-}
-
-compileres <- function(EL) {
-  tt <- map(1:length(X), function(s) map(post, paste(EL)) %>% map_dbl(s))
-  tibble(mean = map_dbl(tt, mean),
-         low = map_dbl(tt, quantile, probs = 0.025),
-         high = map_dbl(tt, quantile, probs = 0.975),
-         X = X,
-         value = paste(EL)
-  )
-}
-```
+- Comparision of colonizating and disappearing species with randomly selected species (refer to Appendix A).
+- Gradient study applied at diffferent time steps (describe covariates used in @Roth2013).
 
 # Results {-}
+## Temporal change in community structures {-}
 ```{r communitytrendscalc}
@@
 
-The different measures of total community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a consistent increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, huminity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
+The different measures of total community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, huminity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
 Except for average Ellenberg value for temperature, the 90% credible-interval of the temporal trend contained zero.
-The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the Ellenberg value for temperature (probability for an increase: `r
+The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the Ellenberg value for temperature (probability of increase: `r
 res %>% filter(Measures == "Temperature value") %>% pull("Prob.
-for Trend")`), followed by the community mean of the Ellenberg light value (probaility for a decrease: `r 1 - res %>% filter(Measures == "Light value") %>% pull("Prob.
-for Trend")`) and it was least likely for the alpha-diversity (probability for an increase: `r res %>% filter(Measures == "Alpha-diversity") %>% pull("Prob.
+for Trend")`), followed by the community mean of the Ellenberg light value (probability of decrease: `r 1 - res %>% filter(Measures == "Light value") %>% pull("Prob.
+for Trend")`) and it was least likely for the alpha-diversity (probability of increase: `r res %>% filter(Measures == "Alpha-diversity") %>% pull("Prob.
 for Trend")`).
@@
 
+## Different drivers of species turnover {-}
 ```{r turnover}
@@
 The average percentage $\pm$ SD of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
-Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model that also contained site variables: `r CrI`).
+Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model: `r CrI`).
 Variation in species turnover was largest along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition (Table \@ref(tab:turnovertab)).
@@
 random disappearance and random colonization) and whether there is a change along the gradients.
-It seems that the Ellenberg values of newly colonizing species differened more from random colonization than the Ellenberg values of disappearing species from random disappearance (Table \@ref(tab:difffromrandomtab)).
-For colonizing species, we found the largest differences from random colonization in the Ellenberg value for nutrients: at sites with nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower Ellenberg value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreases with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
+It seems that the Ellenberg values of newly colonizing species differened more from random colonization than the Ellenberg values of disappearing species (Table \@ref(tab:difffromrandomtab)).
+For colonizing species, we found the largest differences from random colonization in the Ellenberg value for nutrients: at sites with nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower Ellenberg value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
 Thus at high Nitrogen deposition colonizing species did not differ from random species (see Figure 3b in Appendix A).
 
-While colonizing species had higher temperature values compared to what we would be expected under random colonization, the differences between colonizing species and random colonization was about four times smaller compared to the difference in Ellenberg value for nutrients between colonizing species and random species.
+While colonizing species had higher temperature values compared to what we would expect under random colonization, the differences between colonizing species and random colonization was about four times smaller compared to the difference in Ellenberg value for nutrients between colonizing species and random species.
 Nevertheless, the variation in Ellenberg value for temperature seemed impportant to explain the total species turnover.
-This is because, disappearing species tend to have lower temperature value than randomly disappearing species leading to an overall replacement of species with lower temperature value with species with higher temperature values.
-This was not the case for the Ellenberg value for nutrients: it was also species with lower nutrients values that tended to be more likely to disappear than randomly disappearing species (Table \@ref(tab:difffromrandomtab)).
+This is because, disappearing species tend to have lower temperature value than random species as well as colonizing species tend to have hihger temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
+This was not the case for the Ellenberg value for nutrients: species with lower nutrients values tended to be more likely to disappear from as well as to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
 See also Appendix A were we present detailed results for the comparision between colonizing or disappearing species with randomly selected species.
 
+## Potential effects of reduction in Nitrogen emmissions {-}
 ```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
-Given are means and 95\\%-Credible Intervals from logistic linear mixed models."}
+Given are means and 95\\%-Credible Intervals from logistic linear mixed models.
+The vertical lines indicat the deposition rate with equal colonization or survival probabilities for oligotrophic and eutrophic species with the solid line indicating the median and the dashed lines the 5\\% and 95\\% quantiels of the margional posterior distribution."}
 X <- 1:30
@@
 CLs <- map_dbl(post, "CL")
+CL_col <- median(CLs, na.rm = TRUE)
 p1 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
@@
        x = expression(paste("Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")"))) +
-  geom_vline(xintercept = mean(CLs, na.rm = TRUE)) +
+  geom_vline(xintercept = median(CLs, na.rm = TRUE)) +
   geom_vline(xintercept = quantile(CLs, probs = 0.05, na.rm = TRUE), lty = 2) +
@@
 CLs <- map_dbl(post, "CL") 
-
+CL_surv <- median(CLs, na.rm = TRUE)
 p2 <- ggplot(res, aes(x = X, y = mean, colour = value, shape = value)) +
@@
 \@ref(fig:cl) we compare the colonization and local survival probability of oligotrophic (Ellenberg value of nutrients = 2) and eutrophic (Ellenberg value of nutrients = 4) species along the Nitrogen deposition gradient.
-Colonization probability is the sa
+Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for for oligotrophic and euthrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
+In only `r mean((sites$NTOT * 10 + 10) < 11.5)`% of the sites the deposition rate was below 11.5 kg N ha$^{-1}$ yr$^{-1}$ where the replacement of eutrophic with oligotrophic species is likely.
 
-While we could not detect a consistent decrease in average Ellenberg value for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate affects the spatial variation of spcies richness: sites with low Nitrogen deposition become more species rich over time resulting in steeper slope of the negative relationship between N deposition and species richness.
-Consequently, if we apply a model as in @Roth2013 to infer the effects of N deposition on the spatial variation of species richness at different time points, the resulting effect size (i.e.
+While we could not detect a consistent decrease in average Ellenberg value for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial variation of species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between N deposition and species richness.
+Indeed, if we apply at different time points a similar model as in @Roth2013 to infer the effects of N deposition on the spatial variation of species richness, the resulting effect size (i.e.
 the slope) becomes more negative over time (Fig.
@@
 
-```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of Nitrogen deposition on total species richness estimated from applying the Poisson-GLM with species richness as dependend variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from a five-year interval.
-Within every five-year interval all plots were sampled once."}
+```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of Nitrogen deposition on total species richness estimated from applying the Poisson-GLM with species richness as dependend variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from one five-year interval.
+Note that within every five-year interval all plots were sampled once."}
 # Effect on N deposition model
@@
 # Discussion {-}
-- Any empty space that could be caused by any disturbance that let to the local disappearance of species is likely to be fiellied by eutrophic species.
-Disturbance depends on the site, while colonization depends on species characteristics.
+- *General points*: Altough N deposition considerabely declinded between 2005 and 2015, we could not detect major shifts in plant community structure during the same time period.
 
-- The rather large spatial turnover might be partly explained by species that remained undetected in one of the surveys, but it might also be the result of species that newly colonized sites (species gains) and species that truely disappeared from sites (species losses).
+- *Replacement of oligotrophic with eutrophic species is faster than the oposite direction*: Eutrophic species have rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at high N deposition.
+This suggests that it takes
+more time to replace eutrophic by oligotrophic species than replacing oligotrophic by eutrophic species. Climatic effects may be more likely to be reversed thant effects due to fertilization.
 
-- Altough N deposition considerabely declinded between 2005 and 2015, we could not detect major shifts in plant community structure during the same time period.
+- *Methodological point*: The rather large spatial turnover might be partly explained by species that remained undetected in one of the surveys.
+However, our results suggest that turnover is caused at least partly by species with specific Ellenberg values.
+These deviation from what we would expect under random species turnover is unlikely to be explained by species that remained undetected.
 
-- Eutrophic species have rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at high N deposition.
-This suggests that it takes much more time to replace eutrophic by oligotrophic species than replacing oligotrophic by eutrophic species.
+- *Empirical critical loads*: Our data on colonization and local survival (i.e.
+temporal variation) confirm the empirical critical loads that we infered from anlysing spatial co-variation of N deposition and species richness.
 
-- Our data on colonization and local survival (i.e. temporal variation) confirm the empirical critical loads that we infered from anlysing spatial co-variation of N deposition and species richness.
-
-- Local survival is higher for low temperature plants --> This could explain the decrease in cumminity change along elevation.
-This could also explain the differences at mount summets were space was empty in the beginning.
-
-- Climatic effects are more likely to be reversed thant effects due to fertilization.
-
-### Space for time substitution
-Often observational studies infer the change of plant diversity along a gradient of N deposi- tion.
+- *Space for time substitution*: Often observational studies infer the change of plant diversity along a gradient of N deposi- tion.
 Thus, they infer how the spatial variation in species richness is related to N deposition and assume that this spatial variation in species richness arose because over time some areas lost more species than others because they chronically experienced higher N deposition.
@@
 In this study, the recorded recovery in species richness within one or two decade was likely due to the species rich vege- tation surrounding the experimental plots, from where immigration was easily feasible.
-
-> Hier sagen, dass der CRL identisch wie der CRL von räumlichen Zusammenhängen ist.
 

### 2018-05-30 Manuscript.tex
Mit Methodenkapitel begonnen.

@@
 
-\title{No sign of recovery}
+\title{Recovery of mountain plant communities in response to reductions in
+Nitrogen emissions is hidden by other drivers of global change}
 
@@
 Nitrogen (N) deposition is a major threat to biodiversity of many
-habitats. The recent intro- duction of cleaner technologies in
-Switzerland has let to reductions in the emissions of nitro- gen oxides,
-with affiliated decrease in N-deposition. We infered different drivers
+habitats. The recent introduction of cleaner technologies in Switzerland
+has let to reductions in the emissions of nitrogen oxides, with
+affiliated decrease in Nitrogen deposition. We infered different drivers
 of community change (i.e.~Nitrogen deposition, climate warming, land-use
 change) in Swiss mountain hay meadows. The data were obtained from the
-Swiss biodiversity monitoring. Species gain and losses were best
-explainde by differences in Ellenberg value for temperature suggesting
-that climate warming was the most important driver of community change
-in Mountain hay meadows.
+Swiss biodiversity monitoring.
 % Dummy abstract text.
@@
 
+We analysed the presence/absence of vascular plants sampled within the
+scope of Switzerland's Biodiversity Monitoring (BDM) programme that was
+launched in 2001 to monitor Switzerland's biodiversity and to comply
+with the Convention on Biological Diversity of Rio de Janeiro (Weber,
+Hintermann, and Zangger 2004). Data collection was carried out by
+qualified botanists who visited each sampling site twice per season.
+During each visit all the vascular plant species detected on the plot
+were recorded. After the sampling of the plant data the botanists also
+assigned a habitat type to each sampling site according to the
+classification system developed for Switzerland (Delarze and Gonseth
+2008).For more details on the field methods see Plattner, Birrer, and
+Weber (2004), Roth et al. (2013) and Roth et al. (2017).
+
+We matched the habitat types of the Swiss classification system with the
+categories from the EUNIS system (level-3 classification; Davies, Moss,
+and Hill 2004).
+
+In the present study, we analysed the 129 sampling sites in mountain hay
+meadows (EUNIS E2.3).
+
+The sampling sites were circles with a size of 10 m\(^2\) that were
+arranged on a regular sampling grid covering the whole of Switzerland.
+Each sampling site was surveyed once in each survey period: first survey
+period lasted from 2003 to 2007, the second from 2008 to 2012 and the
+third from 2013 to 2017.
+
 \begin{itemize}
@@
   Selection of sample sites based on 1366 K\_Standort.csv column
-  ``E23\_1366''.
+  ``E23\_1366''. The site with exceptionally high nitrogen deposition
+  was excluded from that list.
 \item
-  Three surveys 2003-2007, 2008-2012 and 2013 - 2017.
+  Three surveys 2003 - 2007, 2008 - 2012 and 2013 - 2017.
 \end{itemize}
 
-\subsection*{Plant traits}\label{plant-traits}
-\addcontentsline{toc}{subsection}{Plant traits}
+\subsection*{Community measures}\label{community-measures}
+\addcontentsline{toc}{subsection}{Community measures}
 
-Functional traits:
-
-\begin{itemize}
-\tightlist
-\item
-  SLA: specific leaf area
-\item
-  CH: canopy height
-\item
-  SM: Seed mass
-\end{itemize}
-
-Ellenberg indicator values:
-
-\begin{itemize}
-\tightlist
-\item
-  L: light
-\item
-  N: Nutrient contentent
-\item
-  T: Temperature
-\item
-  F: Huminity
-\end{itemize}
-
-Community measures:
+To describe the plant communities we used the following measures:
 
@@
   gamma diversity: Total number of species recorded in all study plots.
-\end{itemize}
-
-\subsubsection*{Community measures}\label{community-measures}
-\addcontentsline{toc}{subsubsection}{Community measures}
-
+\item
   The temporal turnover (i.e.~species exchange ratio sensu Hillebrand et
@@
   points calculated as
+\end{itemize}
 
-\(\text{Spatial turnover} = \frac{\text{Species gained} + \text{Species lost}}{\text{Total species observed in both timepoints}}\).
+\(\text{Temporal turnover} = \frac{\text{Species gained} + \text{Species lost}}{\text{Total species observed in both timepoints}}\).
+
+\subsection*{Ellenberg values and
+gradients}\label{ellenberg-values-and-gradients}
+\addcontentsline{toc}{subsection}{Ellenberg values and gradients}
+
+Ellenberg indicator values:
+
+\begin{itemize}
+\tightlist
+\item
+  T: Temperature
+\item
+  F: Huminity
+\item
+  N: Nutrient contentent
+\item
+  L: light
+\end{itemize}
+
+Gradients among studied sites:
+
+\begin{itemize}
+\tightlist
+\item
+  Mean annual temperature
+\item
+  Mean annual precipitation
+\item
+  Nitrogen deposition
+\item
+  Inclination: The steaper the site the less intensive a site.
+\item
+  Environmental variables were standardized.
+\end{itemize}
 
@@
 
-Environmental variables were standardized.
+\begin{itemize}
+\tightlist
+\item
+  Comparision of colonizating and disappearing species with randomly
+  selected species (refer to Appendix A).
+\item
+  Gradient study applied at diffferent time steps (describe covariates
+  used in Roth et al. (2013)).
+\end{itemize}
 
@@
 \addcontentsline{toc}{section}{Results}
+
+\subsection*{Temporal change in community
+structures}\label{temporal-change-in-community-structures}
+\addcontentsline{toc}{subsection}{Temporal change in community
+structures}
 
@@
 communities in mountain hay meadows were rather stable between 2003 and
-2017 and did not show a consistent increase or decrease over time (Table
+2017 and did not show a clear increase or decrease over time (Table
 \ref{tab:communitytrendstab}): for each of the three 5-year survey
@@
 most likely for the community mean of the Ellenberg value for
-temperature (probability for an increase: 0.97), followed by the
-community mean of the Ellenberg light value (probaility for a decrease:
-0.91) and it was least likely for the alpha-diversity (probability for
-an increase: 0.53). The chance that the community mean of the nutrient
-value decreased between 2003 and 2017 was 0.68.
+temperature (probability of increase: 0.97), followed by the community
+mean of the Ellenberg light value (probability of decrease: 0.91) and it
+was least likely for the alpha-diversity (probability of increase:
+0.53). The chance that the community mean of the nutrient value
+decreased between 2003 and 2017 was 0.68.
+
+\subsection*{Different drivers of species
+turnover}\label{different-drivers-of-species-turnover}
+\addcontentsline{toc}{subsection}{Different drivers of species turnover}
 
@@
 moderately decreased (90\% Credible interval of the change in turnover
-estimated from the Binomial generalized linear mixed model that also
-contained site variables: -0.15 - -0.02). Variation in species turnover
-was largest along the Nitrogen deposition gradient with highest species
-turnover at sites with low Nitrogen deposition (Table
-\ref{tab:turnovertab}). The other three gradients were less inportant to
-explain the variation in species turnover among sites.
+estimated from the Binomial generalized linear mixed model: -0.15 -
+-0.02). Variation in species turnover was largest along the Nitrogen
+deposition gradient with highest species turnover at sites with low
+Nitrogen deposition (Table \ref{tab:turnovertab}). The other three
+gradients were less inportant to explain the variation in species
+turnover among sites.
 
@@
 Ellenberg values of newly colonizing species differened more from random
-colonization than the Ellenberg values of disappearing species from
-random disappearance (Table \ref{tab:difffromrandomtab}). For colonizing
-species, we found the largest differences from random colonization in
-the Ellenberg value for nutrients: at sites with nitrogen deposition of
-10 kg ha\(^{-1}\) yr\(^{-1}\) the newly colonizing species had in
-average a lower Ellenberg value for nutrients than species under random
-disappearance (column ``Difference from random'' in Table
+colonization than the Ellenberg values of disappearing species (Table
+\ref{tab:difffromrandomtab}). For colonizing species, we found the
+largest differences from random colonization in the Ellenberg value for
+nutrients: at sites with nitrogen deposition of 10 kg ha\(^{-1}\)
+yr\(^{-1}\) the newly colonizing species had in average a lower
+Ellenberg value for nutrients than species under random disappearance
+(column ``Difference from random'' in Table
 \ref{tab:difffromrandomtab}), but this differences between colonizing
-species and random colonization decreases with increasing N deposition
+species and random colonization decreased with increasing N deposition
 (column ``Change along gradient'' in Table \ref{tab:difffromrandomtab}).
@@
 While colonizing species had higher temperature values compared to what
-we would be expected under random colonization, the differences between
+we would expect under random colonization, the differences between
 colonizing species and random colonization was about four times smaller
@@
 species turnover. This is because, disappearing species tend to have
-lower temperature value than randomly disappearing species leading to an
-overall replacement of species with lower temperature value with species
-with higher temperature values. This was not the case for the Ellenberg
-value for nutrients: it was also species with lower nutrients values
-that tended to be more likely to disappear than randomly disappearing
-species (Table \ref{tab:difffromrandomtab}). See also Appendix A were we
-present detailed results for the comparision between colonizing or
-disappearing species with randomly selected species.
+lower temperature value than random species as well as colonizing
+species tend to have hihger temperature values than random species; both
+processes lead to an overall replacement of species with lower
+temperature value with species with higher temperature values. This was
+not the case for the Ellenberg value for nutrients: species with lower
+nutrients values tended to be more likely to disappear from as well as
+to colonize sites compared to random species (Table
+\ref{tab:difffromrandomtab}). See also Appendix A were we present
+detailed results for the comparision between colonizing or disappearing
+species with randomly selected species.
+
+\subsection*{Potential effects of reduction in Nitrogen
+emmissions}\label{potential-effects-of-reduction-in-nitrogen-emmissions}
+\addcontentsline{toc}{subsection}{Potential effects of reduction in
+Nitrogen emmissions}
 
@@
 \includegraphics[width=1\linewidth]{Manuscript_files/figure-latex/cl-1} \caption{Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
-Given are means and 95\%-Credible Intervals from logistic linear mixed models.}\label{fig:cl}
+Given are means and 95\%-Credible Intervals from logistic linear mixed models.
+The vertical lines indicat the deposition rate with equal colonization or survival probabilities for oligotrophic and eutrophic species with the solid line indicating the median and the dashed lines the 5\% and 95\% quantiels of the margional posterior distribution.}\label{fig:cl}
 \end{figure}
@@
 eutrophic (Ellenberg value of nutrients = 4) species along the Nitrogen
-deposition gradient. Colonization probability is the sa
+deposition gradient. Local survival probability was the same for
+oligotrophic and euthrophic species at a deposition rate of 11.89 kg N
+ha\(^{-1}\) yr\(^{-1}\); colonization probability was the same for for
+oligotrophic and euthrophic species at a deposition rate of 12.15 kg N
+ha\(^{-1}\) yr\(^{-1}\). In only 0.30\% of the sites the deposition rate
+was below 11.5 kg N ha\(^{-1}\) yr\(^{-1}\) where the replacement of
+eutrophic with oligotrophic species is likely.
 
@@
 colonization rate of species with low nutrient value at sites with low
-deposition rate affects the spatial variation of spcies richness: sites
-with low Nitrogen deposition become more species rich over time
-resulting in steeper slope of the negative relationship between N
-deposition and species richness. Consequently, if we apply a model as in
-Roth et al. (2013) to infer the effects of N deposition on the spatial
-variation of species richness at different time points, the resulting
-effect size (i.e.~the slope) becomes more negative over time (Fig.
-\ref{fig:figconsequences}).
+deposition rate seems to affect the spatial variation of species
+richness: sites with low Nitrogen deposition are likely to become more
+species rich over time likely resulting in steeper slope of the negative
+relationship between N deposition and species richness. Indeed, if we
+apply at different time points a similar model as in Roth et al. (2013)
+to infer the effects of N deposition on the spatial variation of species
+richness, the resulting effect size (i.e.~the slope) becomes more
+negative over time (Fig. \ref{fig:figconsequences}).
 
 \begin{figure}
-\includegraphics[width=0.5\linewidth]{Manuscript_files/figure-latex/figconsequences-1} \caption{Effect size of Nitrogen deposition on total species richness estimated from applying the Poisson-GLM with species richness as dependend variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from a five-year interval.
-Within every five-year interval all plots were sampled once.}\label{fig:figconsequences}
+\includegraphics[width=0.5\linewidth]{Manuscript_files/figure-latex/figconsequences-1} \caption{Effect size of Nitrogen deposition on total species richness estimated from applying the Poisson-GLM with species richness as dependend variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from one five-year interval.
+Note that within every five-year interval all plots were sampled once.}\label{fig:figconsequences}
 \end{figure}
@@
 \item
-  Any empty space that could be caused by any disturbance that let to
-  the local disappearance of species is likely to be fiellied by
-  eutrophic species. Disturbance depends on the site, while colonization
-  depends on species characteristics.
+  \emph{General points}: Altough N deposition considerabely declinded
+  between 2005 and 2015, we could not detect major shifts in plant
+  community structure during the same time period.
 \item
-  The rather large spatial turnover might be partly explained by species
-  that remained undetected in one of the surveys, but it might also be
-  the result of species that newly colonized sites (species gains) and
-  species that truely disappeared from sites (species losses).
+  \emph{Replacement of oligotrophic with eutrophic species is faster
+  than the oposite direction}: Eutrophic species have rather high local
+  survival across the entire deposition gradient, while oligotrophic
+  species have much reduced local survival at high N deposition. This
+  suggests that it takes more time to replace eutrophic by oligotrophic
+  species than replacing oligotrophic by eutrophic species. Climatic
+  effects may be more likely to be reversed thant effects due to
+  fertilization.
 \item
-  Altough N deposition considerabely declinded between 2005 and 2015, we
-  could not detect major shifts in plant community structure during the
-  same time period.
+  \emph{Methodological point}: The rather large spatial turnover might
+  be partly explained by species that remained undetected in one of the
+  surveys. However, our results suggest that turnover is caused at least
+  partly by species with specific Ellenberg values. These deviation from
+  what we would expect under random species turnover is unlikely to be
+  explained by species that remained undetected.
 \item
-  Eutrophic species have rather high local survival across the entire
-  deposition gradient, while oligotrophic species have much reduced
-  local survival at high N deposition. This suggests that it takes much
-  more time to replace eutrophic by oligotrophic species than replacing
-  oligotrophic by eutrophic species.
+  \emph{Empirical critical loads}: Our data on colonization and local
+  survival (i.e.~temporal variation) confirm the empirical critical
+  loads that we infered from anlysing spatial co-variation of N
+  deposition and species richness.
 \item
-  Our data on colonization and local survival (i.e.~temporal variation)
-  confirm the empirical critical loads that we infered from anlysing
-  spatial co-variation of N deposition and species richness.
-\item
-  Local survival is higher for low temperature plants --\textgreater{}
-  This could explain the decrease in cumminity change along elevation.
-  This could also explain the differences at mount summets were space
-  was empty in the beginning.
-\item
-  Climatic effects are more likely to be reversed thant effects due to
-  fertilization.
+  \emph{Space for time substitution}: Often observational studies infer
+  the change of plant diversity along a gradient of N deposi- tion.
+  Thus, they infer how the spatial variation in species richness is
+  related to N deposition and assume that this spatial variation in
+  species richness arose because over time some areas lost more species
+  than others because they chronically experienced higher N deposition.
+  Alt- hough there is evidence supporting the use of such a `space for
+  time substitution' for detecting the effects of N deposition on plant
+  diversity (Stevens et al. 2010), they can not replace stud- ies that
+  relate temporal patterns in species with N deposition (De Schrijver et
+  al. 2011). While recovery of acidified surface waters has been well
+  investigated (De Vries et al. 2015), there are only a limited number
+  of studies inferring temporal trends of plant species diversity relat-
+  ed to varying amounts of N-deposition. Storkey et al. (2015)
+  demonstrated a positive response of biodiversity to reducing N
+  addition from either atmospheric pollution or fertilizers in the Park
+  Grass Experiment: «The proportion of legumes, species richness and
+  diversity increased across the experiment between 1991 and 2012 as
+  N-deposition declined». For forest floor vegetation in permanent plots
+  across Europe the exceedance of critical loads of N over a peri- od
+  from 9 to 42 years had negative effects on the cover of oligotrophic
+  plant species, i.e spe- cies that prefer nutrient-poor soils, although
+  species richness remained constant (Dirnböck et al. 2014). Another
+  example of recovery in eutrophicated habitats gives the recovery of
+  species richness in previously fertilized plots (Clark and Tilman
+  2008). In this study, the recorded recovery in species richness within
+  one or two decade was likely due to the species rich vege- tation
+  surrounding the experimental plots, from where immigration was easily
+  feasible.
 \end{itemize}
-
-\subsubsection{Space for time
-substitution}\label{space-for-time-substitution}
-
-Often observational studies infer the change of plant diversity along a
-gradient of N deposi- tion. Thus, they infer how the spatial variation
-in species richness is related to N deposition and assume that this
-spatial variation in species richness arose because over time some areas
-lost more species than others because they chronically experienced
-higher N deposition. Alt- hough there is evidence supporting the use of
-such a `space for time substitution' for detecting the effects of N
-deposition on plant diversity (Stevens et al. 2010), they can not
-replace stud- ies that relate temporal patterns in species with N
-deposition (De Schrijver et al. 2011). While recovery of acidified
-surface waters has been well investigated (De Vries et al. 2015), there
-are only a limited number of studies inferring temporal trends of plant
-species diversity relat- ed to varying amounts of N-deposition. Storkey
-et al. (2015) demonstrated a positive response of biodiversity to
-reducing N addition from either atmospheric pollution or fertilizers in
-the Park Grass Experiment: «The proportion of legumes, species richness
-and diversity increased across the experiment between 1991 and 2012 as
-N-deposition declined». For forest floor vegetation in permanent plots
-across Europe the exceedance of critical loads of N over a peri- od from
-9 to 42 years had negative effects on the cover of oligotrophic plant
-species, i.e spe- cies that prefer nutrient-poor soils, although species
-richness remained constant (Dirnböck et al. 2014). Another example of
-recovery in eutrophicated habitats gives the recovery of species
-richness in previously fertilized plots (Clark and Tilman 2008). In this
-study, the recorded recovery in species richness within one or two
-decade was likely due to the species rich vege- tation surrounding the
-experimental plots, from where immigration was easily feasible.
-
-\begin{quote}
-Hier sagen, dass der CRL identisch wie der CRL von räumlichen
-Zusammenhängen ist.
-\end{quote}
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Davies2004}{}
+Davies, CE, D Moss, and MO Hill. 2004. ``EUNIS Habitat Classification,
+Revised 2004. Report to European Environment Agency, European Topic
+Centre on Nature Protection and Biodiversity.'' \emph{Google Scholar}.
+
+\hypertarget{ref-Delarze2008}{}
+Delarze, Raymond, and Yves Gonseth. 2008. \emph{Lebensräume Der Schweiz:
+Ökologie - Gefährdung - Kennarten}. Ott.
+\url{https://books.google.ch/books?id=8kFrAAAACAAJ}.
+
 \hypertarget{ref-Hillebrand2018}{}
@@
 
+\hypertarget{ref-Plattner2004}{}
+Plattner, Matthias, Stefan Birrer, and Darius Weber. 2004. ``Data
+Quality in Monitoring Plant Species Richness in Switzerland.''
+\emph{Community Ecology} 5 (1): 135--43.
+
 \hypertarget{ref-Roth2013}{}
@@
 
+\hypertarget{ref-Roth2017}{}
+Roth, Tobias, Lukas Kohli, Beat Rihm, Reto Meier, and Beat Achermann.
+2017. ``Using Change-Point Models to Estimate Empirical Critical Loads
+for Nitrogen in Mountain Ecosystems.'' \emph{Environmental Pollution}
+220: 1480--7.
+
+\hypertarget{ref-Weber2004}{}
+Weber, Darius, Urs Hintermann, and Adrian Zangger. 2004. ``Scale and
+Trends in Species Richness: Considerations for Monitoring Biological
+Diversity for Political Purposes.'' \emph{Global Ecology and
+Biogeography} 13 (2): 97--104.
+
 \end{document}

### 2018-05-30 R/Data_processing.R
Tidy up data processing.

@@ -37,16 +37,36 @@ tmp <- table(tmp$aID_STAO)
 sum(tmp != 3)
 sel <- names(tmp[tmp == 3])
 
-# Select surveys and add site data
+#------------------------------------------------------------------------------------------------------
+# Prepare site data 
+#------------------------------------------------------------------------------------------------------
+# Add Turnover to 'surv'
+sites <- tibble(aID_STAO = as.integer(sel)) %>% 
+  left_join(tbl(db, "RAUMDATEN_Z9") %>% 
+              transmute(aID_STAO = aID_STAO, Elevation = Hoehe, Inclination = Neig, Aspect = Expos, CACO3 = CACO3, 
+                        Temperature = bio1, Precipitation = bio12), copy = TRUE) %>% 
+  left_join(tbl(db, "RAUMDATEN_Z9_NABO") %>% 
+              transmute(aID_STAO = aID_STAO, pH = pH), by = c("aID_STAO" = "aID_STAO"), copy = TRUE) %>% 
+  left_join(tbl(db, "RAUMDATEN_Z9_NDEP") %>% 
+              transmute(aID_STAO = aID_STAO, NTOT_1980 = NTOT1980_Wiesen, NTOT_1990 = NTOT1990_Wiesen, 
+                        NTOT_2000 = NTOT2000_Wiesen, NTOT_2010 = NTOT2010_Wiesen, NTOT_2015 = NTOT2015_Wiesen), copy = TRUE)
+
+# Standardize covariates
+sites$Elevation <- (sites$Elevation - 1000) / 500
+sites$Temperature <- (sites$Temperature - 50) / 10
+sites$Precipitation <- (sites$Precipitation - 1000) / 200
+sites$NTOT <- (sites$NTOT_2010 - 10) / 10
+sites$Inclination <- (sites$Inclination - 10) / 10
+sites$pH <- sites$pH - 6
+
+#------------------------------------------------------------------------------------------------------
+# Prepare survey data 
+#------------------------------------------------------------------------------------------------------
+# Select surveys
 surv <- tbl(db, "KD_Z9") %>% 
   as.tibble() %>% 
   filter(!is.na(match(aID_STAO, sel)) & !is.na(yearPl) & yearP >= 2003 & yearP <= 2017) %>% 
[…]

### 2018-05-30 Manuscript.Rmd
Rerun models for community measures. Add some description to the methods. Small edits in data processing.

@@
 
-## Monitoring data {-}
+## Monitoring data and community measures{-}
 We analysed the presence/absence of vascular plants sampled within the scope of Switzerland's Biodiversity Monitoring (BDM) programme that was launched in 2001 to monitor Switzerland's biodiversity and to comply with the Convention on Biological Diversity of Rio de Janeiro [@Weber2004].
@@
 During each visit all the vascular plant species detected on the plot were recorded.
-After the sampling of the plant data the botanists also assigned a habitat type to each sampling site according to the classification system developed for Switzerland [@Delarze2008].For more details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
+After the sampling of the plant data the botanists also assigned a habitat type to each sampling site according to the classification system developed for Switzerland [@Delarze2008].
+For more details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
 
-We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all `r nrow(sites)` sampling sites in mountain hay meadows (EUNIS E2.3).
-We analyced the data from 2003 to 2017.
+We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all sampling sites in mountain hay meadows (EUNIS E2.3).
+We analysed the data from 2003 to 2017.
 During that study period each sampling site was surveyed once per five-year period: the first period lasted from 2003 to 2007, the second from 2008 to 2012 and the third from 2013 to 2017.
+These selection criterias resulted in sample of `r nrow(sites)` sites.
 
-## Community measures {-}
-For each survey (a survey conists of the two visits within the same season) we calculated the following community measures: (1) The number of recorded species (species richness), (2) the mean pair-wise disimilarity index (beta-diversity), (3) the community mean of the Ellenberg temperature values of recorded species (community temperature index), (4) the community mean of the Ellenberg humidity values of recorded species (community humidity index), (5) the community mean of the Ellenberg nutrients values of recorded species (community nutrients index) and (6) the community mean of the Ellenberg light values of recorded species (community light index).
+For each survey (a survey conists of the two visits within the same season) we calculated the following community measures: (1) The number of recorded species (species richness), (2) the community mean of the Ellenberg temperature values of recorded species (community temperature index), (3) the community mean of the Ellenberg humidity values of recorded species (community humidity index), (4) the community mean of the Ellenberg nutrients values of recorded species (community nutrients index) and (5) the community mean of the Ellenberg light values of recorded species (community light index).
+The Ellenberg values were obtained from the recalibrated indicator values for the Swiss Flora [@Landolt2010].
+Additionally to these five community measures that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
+species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points to describe the community change between two time points.
 
-Additionally we measured the temporal turnover (i.e. species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points calculated as
-
-$\text{Temporal turnover} = \frac{\text{Species gained} + \text{Species lost}}{\text{Total species observed in both timepoints}}$.
-
-## Ellenberg values and gradients {-}
-Ellenberg indicator values:
-
-- T: Temperature
-- F: Huminity
-- N: Nutrient contentent
-- L: light
-
-Gradients among studied sites:
-
+## Environmental gradients {-}
 - Mean annual temperature
@@
 ## Statistical analyses {-}
+To estimate the linear trend over time for each of the five community measures we applied linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and the logoraithm as link function).
+We used the year as predictor variable and site-ID as random intercept and slope variable to account for the fact that every site was sampled three times.
+Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
+
 - Comparision of colonizating and disappearing species with randomly selected species (refer to Appendix A).
-- Gradient study applied at diffferent time steps (describe covariates used in @Roth2013).
+
+To estimate the effect of N deposition at a given time point .
+We described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarihm as link function.
+The selection of predictor variables was based on an earlier study in Swiss mountain grasslands [@Roth2013].
+Except for N deposition we thus used elevation (linear and squared term), inclination (linear and squared term), mean annual precipitation (linear and squared term), pH, average moisture value, average light value and the aspect.
+For the analyses the variables were standardized.
 
@@
 ```{r communitytrendscalc}
-res <- tibble(Measures = c("Alpha-diversity", "Beta-diversity", "Temperature value", "Huminity value", "Nutrients value", "Light value"),
+res <- tibble(Measures = c("Alpha-diversity", "Temperature value", "Huminity value", "Nutrients value", "Light value"),
               'Period 1' = NA,
@@
 
-# Beta diversity
-getbeta <- function(x) {
-  tmp <- simba::sim(pl %>% filter(Visit == x) %>% dplyr::select(aID_STAO, aID_SP, Occ), 
-                    method = "simpson", listin = TRUE, listout = TRUE)
-  mean(tmp$simpson)
-}
-res[2, c("Period 1", "Period 2", "Period 3")] <- lapply(1:3, getbeta)
+# Mean community index
+res[2, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$T, surv$Visit, mean)
+load("Modelfit/T-trend.RData")
+res[2, "Trend"] <- fixef(mod)["yr"]
+res[2, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
+res[2, 'Prob. for Trend'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 
-# Mean community index
-res[3, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$T, surv$Visit, mean)
-load("Modelfit/T-trend.RData")
+res[3, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$F, surv$Visit, mean)
+load("Modelfit/F-trend.RData")
 res[3, "Trend"] <- fixef(mod)["yr"]
@@
 
-res[4, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$F, surv$Visit, mean)
-load("Modelfit/F-trend.RData")
+res[4, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$N, surv$Visit, mean)
+load("Modelfit/N-trend.RData")
 res[4, "Trend"] <- fixef(mod)["yr"]
@@
 
-res[5, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$N, surv$Visit, mean)
-load("Modelfit/N-trend.RData")
+res[5, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$L, surv$Visit, mean)
+load("Modelfit/L-trend.RData")
 res[5, "Trend"] <- fixef(mod)["yr"]
@@
 res[5, 'Prob. for Trend'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
-
-res[6, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$L, surv$Visit, mean)
-load("Modelfit/L-trend.RData")
-res[6, "Trend"] <- fixef(mod)["yr"]
-res[6, c('5%', '95%')] <- posterior_interval(mod, pars = "yr")
-res[6, 'Prob. for Trend'] <- mean(as.data.frame(mod, pars = "yr")$yr > 0)
 ```
@@
 for trend"), caption = "Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
-The temporal trends are given as  change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function) with site-ID as random intercept and slope effect.
-The measure of precision for the temporal trend is given as the 5\\% and 95\\% quantiles of the marginal posterior distribution.
-Finally, the probability that the linear trend is $> 0$ is given.
-Linear mixed models were not applicable for beta-diversity because measures for beta-diversity were not available for the single sites.")
+The temporal trends are given as change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function).
+The measure of precision for the temporal trend is given as the 5\\% and 95\\% quantiles of the marginal posterior distribution of the linear trend.
+The column 'Prob.
+for trend' gives the probability that the linear trend is $> 0$.")
 ```
 
-The different measures of total community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha- and beta-diversity and the average Ellenberg values for temperature, huminity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
+The five measures of plant community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha-diversity and the average Ellenberg values for temperature, huminity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
 Except for average Ellenberg value for temperature, the 90% credible-interval of the temporal trend contained zero.
@@
 Note that within every five-year interval all plots were sampled once."}
-# Effect on N deposition model
+# Add site covariates to 'surv'
+dat <- surv %>% left_join(sites)
+
+# Estimate effect of N deposition at a given time point
 makemod <- function(year) {
-  mod <- glm(SR ~ Elevation + I(Elevation^2) + Precipitation + NTOT + Inclination, family = poisson,
-      data = surv %>% filter(yearP >= year & yearP <= (year + 4)))
+  mod <- glm(SR ~ Temperature + Precipitation + NTOT + Inclination,
+             data = dat %>% filter(yearP >= year & yearP <= (year + 4)), family = poisson)
   arm::sim(mod, nsim)@coef
@@
 
+# Get estimates from the marginal the posterior distribution of the effect of N deposition
 getestim <- function(x) {

### 2018-05-30 Manuscript.tex
Rerun models for community measures. Add some description to the methods. Small edits in data processing.

@@
 
-\subsection*{Monitoring data}\label{monitoring-data}
-\addcontentsline{toc}{subsection}{Monitoring data}
+\subsection*{Monitoring data and community
+measures}\label{monitoring-data-and-community-measures}
+\addcontentsline{toc}{subsection}{Monitoring data and community
+measures}
 
@@
 with the Convention on Biological Diversity of Rio de Janeiro (Weber,
-Hintermann, and Zangger 2004). Data collection was carried out by
-qualified botanists who visited each sampling site twice per season.
+Hintermann, and Zangger 2004). The sampling sites were circles with a
+size of 10 m\(^2\) and data collection was carried out by qualified
+botanists who visited each sampling site twice within the same season.
 During each visit all the vascular plant species detected on the plot
@@
 classification system developed for Switzerland (Delarze and Gonseth
-2008).For more details on the field methods see Plattner, Birrer, and
+2008). For more details on the field methods see Plattner, Birrer, and
 Weber (2004), Roth et al. (2013) and Roth et al. (2017).
@@
 categories from the EUNIS system (level-3 classification; Davies, Moss,
-and Hill 2004).
+and Hill 2004) and selected all sampling sites in mountain hay meadows
+(EUNIS E2.3). We analysed the data from 2003 to 2017. During that study
+period each sampling site was surveyed once per five-year period: the
+first period lasted from 2003 to 2007, the second from 2008 to 2012 and
+the third from 2013 to 2017. These selection criterias resulted in
+sample of 129 sites.
 
-In the present study, we analysed the 129 sampling sites in mountain hay
-meadows (EUNIS E2.3).
+For each survey (a survey conists of the two visits within the same
+season) we calculated the following community measures: (1) The number
+of recorded species (species richness), (2) the community mean of the
+Ellenberg temperature values of recorded species (community temperature
+index), (3) the community mean of the Ellenberg humidity values of
+recorded species (community humidity index), (4) the community mean of
+the Ellenberg nutrients values of recorded species (community nutrients
+index) and (5) the community mean of the Ellenberg light values of
+recorded species (community light index). The Ellenberg values were
+obtained from the recalibrated indicator values for the Swiss Flora
+(Landolt et al. 2010). Additionally to these five community measures
+that describe the state of plant communities for each site at a given
+time point, we also estimated the temporal turnover (i.e.~species
+exchange ratio sensu Hillebrand et al. (2018)) as the proportion of
+species that differ between two time points to describe the community
+change between two time points.
 
-The sampling sites were circles with a size of 10 m\(^2\) that were
-arranged on a regular sampling grid covering the whole of Switzerland.
-Each sampling site was surveyed once in each survey period: first survey
-period lasted from 2003 to 2007, the second from 2008 to 2012 and the
-third from 2013 to 2017.
-
-\begin{itemize}
-\tightlist
-\item
-  Selection of sample sites based on 1366 K\_Standort.csv column
-  ``E23\_1366''. The site with exceptionally high nitrogen deposition
-  was excluded from that list.
-\item
-  Three surveys 2003 - 2007, 2008 - 2012 and 2013 - 2017.
-\end{itemize}
-
-\subsection*{Community measures}\label{community-measures}
-\addcontentsline{toc}{subsection}{Community measures}
-
-To describe the plant communities we used the following measures:
-
-\begin{itemize}
-\tightlist
-\item
-  Species richness: number of recorded species per \(10m^2\).
-\item
-  Spatial turnover (beta-diversity): Average turnover between all
-  pair-wise combinations of study plots.
-\item
-  gamma diversity: Total number of species recorded in all study plots.
-\item
-  The temporal turnover (i.e.~species exchange ratio sensu Hillebrand et
-  al. (2018)) is the proportion of species that differ between two time
-  points calculated as
-\end{itemize}
-
-\(\text{Temporal turnover} = \frac{\text{Species gained} + \text{Species lost}}{\text{Total species observed in both timepoints}}\).
-
-\subsection*{Ellenberg values and
-gradients}\label{ellenberg-values-and-gradients}
-\addcontentsline{toc}{subsection}{Ellenberg values and gradients}
-
-Ellenberg indicator values:
-
-\begin{itemize}
-\tightlist
-\item
-  T: Temperature
-\item
-  F: Huminity
-\item
-  N: Nutrient contentent
-\item
-  L: light
-\end{itemize}
-
-Gradients among studied sites:
+\subsection*{Environmental gradients}\label{environmental-gradients}
+\addcontentsline{toc}{subsection}{Environmental gradients}
 
@@
 
+To estimate the linear trend over time for each of the five community
+measures we applied linear mixed models with normal distribution (except
+for alpha-diversity with Poisson distribution and the logoraithm as link
+function). We used the year as predictor variable and site-ID as random
+intercept and slope variable to account for the fact that every site was
+sampled three times. Model parameters were estimated in a Bayesian
+framework using the R-Package \emph{rstanarm} (Stan Development Team
+2016, Muth, Oravecz, and Gabry (2018)).
+
 \begin{itemize}
@@
   selected species (refer to Appendix A).
-\item
-  Gradient study applied at diffferent time steps (describe covariates
-  used in Roth et al. (2013)).
 \end{itemize}
+
+To estimate the effect of N deposition at a given time point . We
+described the plant species richness at the sites using a generalized
+linear model with Poisson distribution and the logarihm as link
+function. The selection of predictor variables was based on an earlier
+study in Swiss mountain grasslands (Roth et al. 2013). Except for N
+deposition we thus used elevation (linear and squared term), inclination
+(linear and squared term), mean annual precipitation (linear and squared
+term), pH, average moisture value, average light value and the aspect.
+For the analyses the variables were standardized.
 
@@
 \caption{\label{tab:communitytrendstab}Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
-The temporal trends are given as  change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function) with site-ID as random intercept and slope effect.
-The measure of precision for the temporal trend is given as the 5\% and 95\% quantiles of the marginal posterior distribution.
-Finally, the probability that the linear trend is $> 0$ is given.
-Linear mixed models were not applicable for beta-diversity because measures for beta-diversity were not available for the single sites.}
+The temporal trends are given as change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function).
+The measure of precision for the temporal trend is given as the 5\% and 95\% quantiles of the marginal posterior distribution of the linear trend.
+The column 'Prob.
+for trend' gives the probability that the linear trend is $> 0$.}
 \centering
@@
 Alpha-diversity & 45.72 & 46.02 & 45.74 & 0.00 & -0.03 & 0.03 & 0.53\\
-Beta-diversity & 0.60 & 0.60 & 0.60 &  &  &  & \\
-Temperature value & 3.11 & 3.13 & 3.13 & 0.01 & 0.00 & 0.02 & 0.97\\
-Huminity value & 2.99 & 2.98 & 2.99 & 0.01 & -0.01 & 0.02 & 0.78\\
-Nutrients value & 3.20 & 3.20 & 3.20 & -0.01 & -0.02 & 0.01 & 0.32\\
-Light value & 3.56 & 3.55 & 3.55 & -0.01 & -0.02 & 0.00 & 0.09\\
+Temperature value & 3.11 & 3.13 & 3.13 & 0.01 & 0.00 & 0.03 & 0.97\\
+Huminity value & 2.99 & 2.98 & 2.99 & 0.01 & -0.01 & 0.02 & 0.80\\
+Nutrients value & 3.20 & 3.20 & 3.20 & 0.00 & -0.02 & 0.01 & 0.33\\
+Light value & 3.56 & 3.55 & 3.55 & -0.01 & -0.02 & 0.00 & 0.07\\
 \bottomrule
@@
 
-The different measures of total community structure suggested that plant
+The five measures of plant community structure suggested that plant
 communities in mountain hay meadows were rather stable between 2003 and
@@
 \ref{tab:communitytrendstab}): for each of the three 5-year survey
-periods the averages of alpha- and beta-diversity and the average
-Ellenberg values for temperature, huminity, nutrients and light did not
-vary much among the three sampling periods and the estimated trends were
-rather small. Except for average Ellenberg value for temperature, the
-90\% credible-interval of the temporal trend contained zero. The results
-from the linear mixed models suggest that a linear temporal change was
-most likely for the community mean of the Ellenberg value for
-temperature (probability of increase: 0.97), followed by the community
-mean of the Ellenberg light value (probability of decrease: 0.91) and it
-was least likely for the alpha-diversity (probability of increase:
-0.53). The chance that the community mean of the nutrient value
-decreased between 2003 and 2017 was 0.68.
+periods the averages of alpha-diversity and the average Ellenberg values
+for temperature, huminity, nutrients and light did not vary much among
+the three sampling periods and the estimated trends were rather small.
+Except for average Ellenberg value for temperature, the 90\%
+credible-interval of the temporal trend contained zero. The results from
+the linear mixed models suggest that a linear temporal change was most
+likely for the community mean of the Ellenberg value for temperature
+(probability of increase: 0.97), followed by the community mean of the
+Ellenberg light value (probability of decrease: 0.93) and it was least
+likely for the alpha-diversity (probability of increase: 0.53). The
+chance that the community mean of the nutrient value decreased between
+2003 and 2017 was 0.67.
 
@@
 species. The average percentage \(\pm\) SD of species that differed
-between the first and second survey at a site was 37.65 \(\pm\) 10.43\%
-and the percentage of species that differ between the second and third
-survey was 35.66 \(\pm\) 10.36\%. Thus, it seemed that the turnover from
-the first/second survey to the turnover of the second/third survey
+between the first and second survey at a site was NaN \(\pm\) NA\% and
+the percentage of species that differ between the second and third
+survey was NaN \(\pm\) NA\%. Thus, it seemed that the turnover from the
+first/second survey to the turnover of the second/third survey
 moderately decreased (90\% Credible interval of the change in turnover
@@
 Revised 2004. Report to European Environment Agency, European Topic
-Centre on Nature Protection and Biodiversity.'' \emph{Google Scholar}.
+Centre on Nature Protection and Biodiversity.''
 
@@
 doi:\href{https://doi.org/10.1111/1365-2664.12959}{10.1111/1365-2664.12959}.
+
+\hypertarget{ref-Landolt2010}{}
+Landolt, Elias, Beat Bäumler, Andreas Erhardt, Otto Hegg, Frank Klötzli,
+Walter Lämmler, Michael Nobis, et al. 2010. \emph{Flora Indicativa.
+Ecological Inicator Values and Biological Attributes of the Flora of
+Switzerland and the Alps.} Haupt Verlag.
+
+\hypertarget{ref-Muth2018}{}
+Muth, Chelsea, Zita Oravecz, and Jonah Gabry. 2018. ``User-Friendly
+Bayesian Regression Modeling: A Tutorial with Rstanarm and Shinystan.''
+\emph{The Quantitative Methods for Psychology} 14 (2): 99--119.
 
@@
 
+\hypertarget{ref-Stan2016}{}
+Stan Development Team. 2016. ``Rstanarm: Bayesian Applied Regression
+Modeling via Stan.'' \url{http://mc-stan.org/}.
+
 \hypertarget{ref-Weber2004}{}

### 2018-05-30 R/Data_processing.R

@@ -81,6 +81,9 @@ surv <- surv %>% left_join(
       N = mean(N, na.rm = TRUE),
       L = mean(L, na.rm = TRUE)), copy = TRUE)
 
+# Standardize years
+surv$yr <- (surv$yearPl - 2010) / 10
+
 #------------------------------------------------------------------------------------------------------
 # Prepare plant data
 #------------------------------------------------------------------------------------------------------

### 2018-05-30 Manuscript.Rmd
Tidy up everything. First version of Material and methods. Spell check. Description of data in Readme.

@@
 The recent introduction of cleaner technologies has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in N-deposition in many parts of Europe.
-However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovey of extant plant communities.
+However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities.
 
@@
 
-Here we infered mountain hay meadows in Switzerland. Explain why mountain hay meadows are important. Also explain other threats to mountain hay meadows (climate change, land-use change).
+Here we inferred mountain hay meadows in Switzerland. Explain why mountain hay meadows are important. Also explain other threats to mountain hay meadows (climate change, land-use change).
 
@@
 During that study period each sampling site was surveyed once per five-year period: the first period lasted from 2003 to 2007, the second from 2008 to 2012 and the third from 2013 to 2017.
-These selection criterias resulted in sample of `r nrow(sites)` sites.
+These selection criteria resulted in sample of `r nrow(sites)` sites.
 
-For each survey (a survey conists of the two visits within the same season) we calculated the following community measures: (1) The number of recorded species (species richness), (2) the community mean of the Ellenberg temperature values of recorded species (community temperature index), (3) the community mean of the Ellenberg humidity values of recorded species (community humidity index), (4) the community mean of the Ellenberg nutrients values of recorded species (community nutrients index) and (5) the community mean of the Ellenberg light values of recorded species (community light index).
+For each survey (a survey consists of the two visits within the same season) we calculated the following community measures: (1) The number of recorded species (species richness), (2) the community mean of the Ellenberg temperature values of recorded species (community temperature index), (3) the community mean of the Ellenberg humidity values of recorded species (community humidity index), (4) the community mean of the Ellenberg nutrients values of recorded species (community nutrients index) and (5) the community mean of the Ellenberg light values of recorded species (community light index).
 The Ellenberg values were obtained from the recalibrated indicator values for the Swiss Flora [@Landolt2010].
@@
 ## Environmental gradients {-}
-- Mean annual temperature
-- Mean annual precipitation
-- Nitrogen deposition
-- Inclination: The steaper the site the less intensive a site. 
-- Environmental variables were standardized.
+We expected different global change patterns to simultaneously cause temporal change in mountain hey communities.
+To better disentangle the importance of these mechanisms we ordered the sites along four main environmental gradients.
+First we expected communities to response to climate warming and the response might be different depending along the temperature gradient [@Roth2014].
+To describe the temperature gradient we used the annual mean temperature per site from the WordClim database [@Fick2017].
+Another key drivers for plants that is likely to be affected by climate change is precipitation [@Beier2012].
+We used the annual mean precipitation per site from the WordClim database [@Fick2017].Further, we estimated atmospheric N deposition for each site using a pragmatic approach described in @Rihm2001 that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models.
+For details on the estimation of N deposition see @Rihm2001 and @Roth2017.
+Finally we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively used.
+We expect that well-accessible sites become more intensively managed, whereas management intensity will rather decrease at remote sites [@Strebel2015].
 
 ## Statistical analyses {-}
-To estimate the linear trend over time for each of the five community measures we applied linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and the logoraithm as link function).
-We used the year as predictor variable and site-ID as random intercept and slope variable to account for the fact that every site was sampled three times.
+To estimate the linear trend over time for each of the five community measures we applied linear mixed models (LMM) with normal distribution except for alpha-diversity with Poisson distribution and the logarithm as link function.
+We specified site specific trends with the assumption that the between-site differences in intercepts and slopes can be described with normal distributions [i.e.
+a random intercept random slope model, @German2006].
 Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
 
-- Comparision of colonizating and disappearing species with randomly selected species (refer to Appendix A).
+To infer whether species turnover was changing along the gradient, we used a Binomial-LMM with the proportion of species that differed between two surveys as dependent variable and the site gradients and period (first/second vs.
+second/third surveys) as predictors and site-ID as random effect.
+Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
 
-To estimate the effect of N deposition at a given time point .
-We described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarihm as link function.
-The selection of predictor variables was based on an earlier study in Swiss mountain grasslands [@Roth2013].
-Except for N deposition we thus used elevation (linear and squared term), inclination (linear and squared term), mean annual precipitation (linear and squared term), pH, average moisture value, average light value and the aspect.
-For the analyses the variables were standardized.
+To infer whether species that colonized or disappeared from a site had particular Ellenberg values that differed from the other species at that site, we produced for each site a list with all species that were recorded during the three surveys (total community).
+We then calculated the community mean (CM) of the Ellenberg value for all species that colonized the site during the three surveys (i.e.
+not recorded during first survey and recorded during second or not recorded during second and recorded during third).
+We then randomly selected the same number of species from the total community and also calculated the community mean of the Ellenberg value for these species (random-CM).
+We repeated the random selection of species 1000 times.
+Then we calculated the differences of the CM minus the average of the random-CMs to obtain a standardized measure (standardized-CM) of how different the disappearing species were from random expectation.
+A standardized-CM < 0 would suggest that the Ellenberg value of colonizing species were lower than the average species that were occurring at this site.
+We applied this method for both colonizing and disappearing species and for the Ellenberg values for temperature, humidity, nutrients and light (see Appendix A).
+We then inferred whether the standardized-CM is changing along the corresponding gradient using linear models.
+Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
+
+We used logistic-LMM to infer whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this changing depended on the species Ellenberg value for nutrients.
+To analyse the colonization probability we selected all species that were not observed during the first survey, and asked if they were observed (then Occ = 1) or not observed (then Occ = 0) during the second survey.
+The same was also done for all species that were not observed during the second survey.
+The variable 'occ' was then used as dependent variable in the logistic-LMM.
+As predictor variables the model contained the Nitrogen deposition of the site, the Ellenberg value for Nutrients of the species and the interaction of these two variables.
+Additionally we species-ID and site-ID were included as random effects.
+The same logistic_LMM was also used to infer local survival probability.
+In that case, however, we selected all species that were recorded during the first or second survey and the variable 'occ' then indicated whether or not the species was also observed during the next survey.
+Model parameters were estimated using an approximate Bayesian approach using the R-Package *arm* [@arm2018].
+
+To estimate the effect of N deposition on total species richness at a given time point, we described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarithm as link function.
+As predictors we used the four environmental gradients that we described above.
+Model parameters were estimated using an approximate Bayesian approach using the R-Package *arm* [@arm2018].
 
@@
 
-The five measures of plant community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha-diversity and the average Ellenberg values for temperature, huminity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
+The five measures of plant community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha-diversity and the average Ellenberg values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
 Except for average Ellenberg value for temperature, the 90% credible-interval of the temporal trend contained zero.
@@
 
-## Different drivers of species turnover {-}
+## Species turnover {-}
 ```{r turnover}
@@
 The slopes along the gradients (estimate) are given as the change per 10 years of the logit-probability of species that differed between two surveys.
-Estimates and the 5\\% and 95\\% quantiles of the marginal posterior distribution obtained from a Binomial-GLMM with the proportion of species that differed between two surveys as dependent variable and the site gradients and period as predictors and site-ID as random effect.")
+Estimates and the 5\\% and 95\\% quantiles of the marginal posterior distribution obtained from a Binomial-GLMM.")
 ```
@@
 Variation in species turnover was largest along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition (Table \@ref(tab:turnovertab)).
-The other three gradients were less inportant to explain the variation in species turnover among sites.
+The other three gradients were less important to explain the variation in species turnover among sites.
 
@@
 random disappearance and random colonization) and whether there is a change along the gradients.
-It seems that the Ellenberg values of newly colonizing species differened more from random colonization than the Ellenberg values of disappearing species (Table \@ref(tab:difffromrandomtab)).
+It seems that the Ellenberg values of newly colonizing species differed more from random colonization than the Ellenberg values of disappearing species (Table \@ref(tab:difffromrandomtab)).
 For colonizing species, we found the largest differences from random colonization in the Ellenberg value for nutrients: at sites with nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower Ellenberg value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
@@
 While colonizing species had higher temperature values compared to what we would expect under random colonization, the differences between colonizing species and random colonization was about four times smaller compared to the difference in Ellenberg value for nutrients between colonizing species and random species.
-Nevertheless, the variation in Ellenberg value for temperature seemed impportant to explain the total species turnover.
-This is because, disappearing species tend to have lower temperature value than random species as well as colonizing species tend to have hihger temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
+Nevertheless, the variation in Ellenberg value for temperature seemed important to explain the total species turnover.
+This is because, disappearing species tend to have lower temperature value than random species as well as colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
 This was not the case for the Ellenberg value for nutrients: species with lower nutrients values tended to be more likely to disappear from as well as to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
-See also Appendix A were we present detailed results for the comparision between colonizing or disappearing species with randomly selected species.
+See also Appendix A were we present detailed results for the comparison between colonizing or disappearing species with randomly selected species.
 
@@
 Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for for oligotrophic and euthrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
-In only `r mean((sites$NTOT * 10 + 10) < 11.5)`% of the sites the deposition rate was below 11.5 kg N ha$^{-1}$ yr$^{-1}$ where the replacement of eutrophic with oligotrophic species is likely.
+In only `r mean((sites$NTOT * 10 + 10) < 12.5)`% of the sites the deposition rate was below 12.5 kg N ha$^{-1}$ yr$^{-1}$ where the replacement of eutrophic with oligotrophic species is likely.
 
@@
 # Discussion {-}
-- *General points*: Altough N deposition considerabely declinded between 2005 and 2015, we could not detect major shifts in plant community structure during the same time period.
+- *General points*: Although N deposition considerably declined between 2005 and 2015, we could not detect major shifts in plant community structure during the same time period.
 
@@
 This suggests that it takes
-more time to replace eutrophic by oligotrophic species than replacing oligotrophic by eutrophic species. Climatic effects may be more likely to be reversed thant effects due to fertilization.
+more time to replace eutrophic by oligotrophic species than replacing oligotrophic by eutrophic species. Climatic effects may be more likely to be reversed than effects due to fertilization.
 
@@
 - *Empirical critical loads*: Our data on colonization and local survival (i.e.
-temporal variation) confirm the empirical critical loads that we infered from anlysing spatial co-variation of N deposition and species richness.
+temporal variation) confirm the empirical critical loads that we inferred from analysing spatial co-variation of N deposition and species richness.
 

### 2018-05-30 Manuscript.tex
Tidy up everything. First version of Material and methods. Spell check. Description of data in Readme.

@@
 in many parts of Europe. However, it is an open question whether and how
-fast the reduction in N deposition rates will lead to the recovey of
+fast the reduction in N deposition rates will lead to the recovery of
 extant plant communities.
@@
 
-Here we infered mountain hay meadows in Switzerland. Explain why
+Here we inferred mountain hay meadows in Switzerland. Explain why
 mountain hay meadows are important. Also explain other threats to
@@
 first period lasted from 2003 to 2007, the second from 2008 to 2012 and
-the third from 2013 to 2017. These selection criterias resulted in
-sample of 129 sites.
+the third from 2013 to 2017. These selection criteria resulted in sample
+of 129 sites.
 
-For each survey (a survey conists of the two visits within the same
+For each survey (a survey consists of the two visits within the same
 season) we calculated the following community measures: (1) The number
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Mean annual temperature
-\item
-  Mean annual precipitation
-\item
-  Nitrogen deposition
-\item
-  Inclination: The steaper the site the less intensive a site.
-\item
-  Environmental variables were standardized.
-\end{itemize}
+We expected different global change patterns to simultaneously cause
+temporal change in mountain hey communities. To better disentangle the
+importance of these mechanisms we ordered the sites along four main
+environmental gradients. First we expected communities to response to
+climate warming and the response might be different depending along the
+temperature gradient (Roth, Plattner, and Amrhein 2014). To describe the
+temperature gradient we used the annual mean temperature per site from
+the WordClim database (Fick and Hijmans 2017). Another key drivers for
+plants that is likely to be affected by climate change is precipitation
+(Beier et al. 2012). We used the annual mean precipitation per site from
+the WordClim database (Fick and Hijmans 2017).Further, we estimated
+atmospheric N deposition for each site using a pragmatic approach
+described in Rihm and Kurz (2001) that combines monitoring data, spatial
+interpolation methods, emission inventories, statistical dispersion
+models and inferential deposition models. For details on the estimation
+of N deposition see Rihm and Kurz (2001) and Roth et al. (2017). Finally
+we used inclination as proxy for land-use intensity because we expected
+that steeper sites are likely to be less intensively used. We expect
+that well-accessible sites become more intensively managed, whereas
+management intensity will rather decrease at remote sites (Strebel and
+Bühler 2015).
 
@@
 To estimate the linear trend over time for each of the five community
-measures we applied linear mixed models with normal distribution (except
-for alpha-diversity with Poisson distribution and the logoraithm as link
-function). We used the year as predictor variable and site-ID as random
-intercept and slope variable to account for the fact that every site was
-sampled three times. Model parameters were estimated in a Bayesian
-framework using the R-Package \emph{rstanarm} (Stan Development Team
-2016, Muth, Oravecz, and Gabry (2018)).
+measures we applied linear mixed models (LMM) with normal distribution
+except for alpha-diversity with Poisson distribution and the logarithm
+as link function. We specified site specific trends with the assumption
+that the between-site differences in intercepts and slopes can be
+described with normal distributions (i.e.~a random intercept random
+slope model, {\textbf{???}}). Model parameters were estimated in a
+Bayesian framework using the R-Package \emph{rstanarm} (Stan Development
+Team 2016, Muth, Oravecz, and Gabry (2018)).
 
-\begin{itemize}
-\tightlist
-\item
-  Comparision of colonizating and disappearing species with randomly
-  selected species (refer to Appendix A).
-\end{itemize}
+To infer whether species turnover was changing along the gradient, we
+used a Binomial-LMM with the proportion of species that differed between
+two surveys as dependent variable and the site gradients and period
+(first/second vs.~second/third surveys) as predictors and site-ID as
+random effect. Model parameters were estimated in a Bayesian framework
+using the R-Package \emph{rstanarm} (Stan Development Team 2016, Muth,
+Oravecz, and Gabry (2018)).
 
-To estimate the effect of N deposition at a given time point . We
-described the plant species richness at the sites using a generalized
-linear model with Poisson distribution and the logarihm as link
-function. The selection of predictor variables was based on an earlier
-study in Swiss mountain grasslands (Roth et al. 2013). Except for N
-deposition we thus used elevation (linear and squared term), inclination
-(linear and squared term), mean annual precipitation (linear and squared
-term), pH, average moisture value, average light value and the aspect.
-For the analyses the variables were standardized.
+To infer whether species that colonized or disappeared from a site had
+particular Ellenberg values that differed from the other species at that
+site, we produced for each site a list with all species that were
+recorded during the three surveys (total community). We then calculated
+the community mean (CM) of the Ellenberg value for all species that
+colonized the site during the three surveys (i.e.~not recorded during
+first survey and recorded during second or not recorded during second
+and recorded during third). We then randomly selected the same number of
+species from the total community and also calculated the community mean
+of the Ellenberg value for these species (random-CM). We repeated the
+random selection of species 1000 times. Then we calculated the
+differences of the CM minus the average of the random-CMs to obtain a
+standardized measure (standardized-CM) of how different the disappearing
+species were from random expectation. A standardized-CM \textless{} 0
+would suggest that the Ellenberg value of colonizing species were lower
+than the average species that were occurring at this site. We applied
+this method for both colonizing and disappearing species and for the
+Ellenberg values for temperature, humidity, nutrients and light (see
+Appendix A). We then inferred whether the standardized-CM is changing
+along the corresponding gradient using linear models. Model parameters
+were estimated in a Bayesian framework using the R-Package
+\emph{rstanarm} (Stan Development Team 2016, Muth, Oravecz, and Gabry
+(2018)).
+
+We used logistic-LMM to infer whether the colonization probability or
+local survival probability was changing along the Nitrogen deposition
+gradient and whether this changing depended on the species Ellenberg
+value for nutrients. To analyse the colonization probability we selected
+all species that were not observed during the first survey, and asked if
+they were observed (then Occ = 1) or not observed (then Occ = 0) during
+the second survey. The same was also done for all species that were not
+observed during the second survey. The variable `occ' was then used as
+dependent variable in the logistic-LMM. As predictor variables the model
+contained the Nitrogen deposition of the site, the Ellenberg value for
+Nutrients of the species and the interaction of these two variables.
+Additionally we species-ID and site-ID were included as random effects.
+The same logistic\_LMM was also used to infer local survival
+probability. In that case, however, we selected all species that were
+recorded during the first or second survey and the variable `occ' then
+indicated whether or not the species was also observed during the next
+survey. Model parameters were estimated using an approximate Bayesian
+approach using the R-Package \emph{arm} (Gelman and Su 2018).
+
+To estimate the effect of N deposition on total species richness at a
+given time point, we described the plant species richness at the sites
+using a generalized linear model with Poisson distribution and the
+logarithm as link function. As predictors we used the four environmental
+gradients that we described above. Model parameters were estimated using
+an approximate Bayesian approach using the R-Package \emph{arm} (Gelman
+and Su 2018).
 
@@
 periods the averages of alpha-diversity and the average Ellenberg values
-for temperature, huminity, nutrients and light did not vary much among
+for temperature, humidity, nutrients and light did not vary much among
 the three sampling periods and the estimated trends were rather small.
@@
 
-\subsection*{Different drivers of species
-turnover}\label{different-drivers-of-species-turnover}
-\addcontentsline{toc}{subsection}{Different drivers of species turnover}
+\subsection*{Species turnover}\label{species-turnover}
+\addcontentsline{toc}{subsection}{Species turnover}
 
@@
 The slopes along the gradients (estimate) are given as the change per 10 years of the logit-probability of species that differed between two surveys.
-Estimates and the 5\% and 95\% quantiles of the marginal posterior distribution obtained from a Binomial-GLMM with the proportion of species that differed between two surveys as dependent variable and the site gradients and period as predictors and site-ID as random effect.}
+Estimates and the 5\% and 95\% quantiles of the marginal posterior distribution obtained from a Binomial-GLMM.}
 \centering
@@
 \midrule
-Annual mean temperature & 0.06 & 0.00 & 0.11\\
-Annual mean precipitation & -0.02 & -0.11 & 0.08\\
-Nitrogen deposition & -0.20 & -0.35 & -0.05\\
-Inclination & -0.04 & -0.10 & 0.03\\
+Annual mean temperature & 0.04 & -0.02 & 0.10\\
+Annual mean precipitation & -0.03 & -0.12 & 0.07\\
+Nitrogen deposition & -0.14 & -0.30 & 0.02\\
+Inclination & -0.03 & -0.10 & 0.03\\
 \bottomrule
@@
 species. The average percentage \(\pm\) SD of species that differed
-between the first and second survey at a site was NaN \(\pm\) NA\% and
-the percentage of species that differ between the second and third
-survey was NaN \(\pm\) NA\%. Thus, it seemed that the turnover from the
-first/second survey to the turnover of the second/third survey
+between the first and second survey at a site was 37.65 \(\pm\) 10.43\%
+and the percentage of species that differ between the second and third
+survey was 35.66 \(\pm\) 10.36\%. Thus, it seemed that the turnover from
+the first/second survey to the turnover of the second/third survey
 moderately decreased (90\% Credible interval of the change in turnover
-estimated from the Binomial generalized linear mixed model: -0.15 -
--0.02). Variation in species turnover was largest along the Nitrogen
+estimated from the Binomial generalized linear mixed model: -0.14 -
+-0.03). Variation in species turnover was largest along the Nitrogen
 deposition gradient with highest species turnover at sites with low
 Nitrogen deposition (Table \ref{tab:turnovertab}). The other three
-gradients were less inportant to explain the variation in species
+gradients were less important to explain the variation in species
 turnover among sites.
@@
 \hspace{1em}Temperature & Annual mean temperature & -0.012 & -0.034 & 0.009 & 0.007 & -0.002 & 0.016\\
-\hspace{1em}Humidity & Annual mean precipitation & -0.003 & -0.038 & 0.033 & 0.008 & -0.012 & 0.029\\
-\hspace{1em}Nutrients & Nitrogen deposition & -0.012 & -0.052 & 0.028 & 0.000 & -0.046 & 0.046\\
-\hspace{1em}Light & Inclination & -0.022 & -0.047 & 0.004 & -0.002 & -0.024 & 0.022\\
+\hspace{1em}Humidity & Annual mean precipitation & -0.006 & -0.042 & 0.029 & 0.011 & -0.010 & 0.032\\
+\hspace{1em}Nutrients & Nitrogen deposition & -0.025 & -0.059 & 0.010 & 0.013 & -0.031 & 0.058\\
+\hspace{1em}Light & Inclination & -0.023 & -0.050 & 0.002 & -0.003 & -0.026 & 0.021\\
 \addlinespace[0.3em]
 \multicolumn{8}{l}{\textit{(b) Plants that newly colonized a site}}\\
-\hspace{1em}Temperature & Annual mean temperature & 0.018 & 0.001 & 0.034 & -0.001 & -0.008 & 0.006\\
-\hspace{1em}Humidity & Annual mean precipitation & 0.023 & -0.010 & 0.054 & -0.004 & -0.023 & 0.014\\
-\hspace{1em}Nutrients & Nitrogen deposition & -0.082 & -0.117 & -0.048 & 0.062 & 0.024 & 0.102\\
-\hspace{1em}Light & Inclination & -0.039 & -0.062 & -0.016 & 0.010 & -0.009 & 0.031\\
+\hspace{1em}Temperature & Annual mean temperature & 0.017 & 0.000 & 0.034 & -0.001 & -0.008 & 0.006\\
+\hspace{1em}Humidity & Annual mean precipitation & 0.019 & -0.013 & 0.051 & -0.002 & -0.020 & 0.016\\
+\hspace{1em}Nutrients & Nitrogen deposition & -0.076 & -0.108 & -0.045 & 0.060 & 0.023 & 0.098\\
+\hspace{1em}Light & Inclination & -0.041 & -0.063 & -0.018 & 0.010 & -0.010 & 0.029\\
 \bottomrule
@@
 and whether there is a change along the gradients. It seems that the
-Ellenberg values of newly colonizing species differened more from random
+Ellenberg values of newly colonizing species differed more from random
 colonization than the Ellenberg values of disappearing species (Table
@@
 colonizing species and random species. Nevertheless, the variation in
-Ellenberg value for temperature seemed impportant to explain the total
+Ellenberg value for temperature seemed important to explain the total
 species turnover. This is because, disappearing species tend to have
 lower temperature value than random species as well as colonizing
-species tend to have hihger temperature values than random species; both
+species tend to have higher temperature values than random species; both
 processes lead to an overall replacement of species with lower
@@
 \ref{tab:difffromrandomtab}). See also Appendix A were we present
-detailed results for the comparision between colonizing or disappearing
+detailed results for the comparison between colonizing or disappearing
 species with randomly selected species.
@@
 deposition gradient. Local survival probability was the same for
-oligotrophic and euthrophic species at a deposition rate of 11.89 kg N
+oligotrophic and euthrophic species at a deposition rate of 11.66 kg N
 ha\(^{-1}\) yr\(^{-1}\); colonization probability was the same for for
-oligotrophic and euthrophic species at a deposition rate of 12.15 kg N
-ha\(^{-1}\) yr\(^{-1}\). In only 0.30\% of the sites the deposition rate
-was below 11.5 kg N ha\(^{-1}\) yr\(^{-1}\) where the replacement of
+oligotrophic and euthrophic species at a deposition rate of 12.09 kg N
+ha\(^{-1}\) yr\(^{-1}\). In only 0.36\% of the sites the deposition rate
+was below 12.5 kg N ha\(^{-1}\) yr\(^{-1}\) where the replacement of
 eutrophic with oligotrophic species is likely.
@@
 \item
-  \emph{General points}: Altough N deposition considerabely declinded
+  \emph{General points}: Although N deposition considerably declined
   between 2005 and 2015, we could not detect major shifts in plant
@@
   species than replacing oligotrophic by eutrophic species. Climatic
-  effects may be more likely to be reversed thant effects due to
+  effects may be more likely to be reversed than effects due to
   fertilization.
@@
   survival (i.e.~temporal variation) confirm the empirical critical
-  loads that we infered from anlysing spatial co-variation of N
+  loads that we inferred from analysing spatial co-variation of N
   deposition and species richness.
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Beier2012}{}
+Beier, Claus, Carl Beierkuhnlein, Thomas Wohlgemuth, Josep Penuelas,
+Bridget Emmett, Christian Körner, Hans Boeck, et al. 2012.
+``Precipitation Manipulation Experiments--challenges and Recommendations
+for the Future.'' \emph{Ecology Letters} 15 (8): 899--911.
+
 \hypertarget{ref-Davies2004}{}
@@
 \url{https://books.google.ch/books?id=8kFrAAAACAAJ}.
+
+\hypertarget{ref-Fick2017}{}
+Fick, Stephen E, and Robert J Hijmans. 2017. ``WorldClim 2: New 1-Km
+Spatial Resolution Climate Surfaces for Global Land Areas.''
+\emph{International Journal of Climatology} 37 (12): 4302--15.
+
+\hypertarget{ref-arm2018}{}
+Gelman, Andrew, and Yu-Sung Su. 2018. \emph{Arm: Data Analysis Using
+Regression and Multilevel/Hierarchical Models}.
+\url{https://CRAN.R-project.org/package=arm}.
 
@@
 
+\hypertarget{ref-Rihm2001}{}
+Rihm, Beat, and Daniel Kurz. 2001. ``Deposition and Critical Loads of
+Nitrogen in Switzerland.'' In \emph{Acid Rain 2000}, 1223--8.
+
 \hypertarget{ref-Roth2013}{}
@@
 
+\hypertarget{ref-Roth2014}{}
+Roth, Tobias, Matthias Plattner, and Valentin Amrhein. 2014. ``Plants,
+Birds and Butterflies: Short-Term Responses of Species Communities to
+Climate Warming Vary by Taxon and with Altitude.'' \emph{PloS One} 9
+(1): e82490.
+
 \hypertarget{ref-Stan2016}{}
@@
 Modeling via Stan.'' \url{http://mc-stan.org/}.
+
+\hypertarget{ref-Strebel2015}{}
+Strebel, Nicolas, and Christoph Bühler. 2015. ``Recent Shifts in Plant
+Species Suggest Opposing Land-Use Changes in Alpine Pastures.''
+\emph{Alpine Botany} 125 (1): 1--9.
 

### 2018-05-30 R/Data_processing.R

@@ -90,7 +90,7 @@ surv$yr <- (surv$yearPl - 2010) / 10
 pl <- tbl(db, "PL") %>% 
   filter(Z7 == 0 & !is.na(aID_SP)) %>% 
   as.tibble() %>% 
-  filter(!is.na(match(aID_KD, sites$aID_KD))) %>% 
+  filter(!is.na(match(aID_KD, surv$aID_KD))) %>% 
   left_join(tbl(db, "KD_Z9"), copy = TRUE) %>% 
   transmute(aID_KD = aID_KD, aID_STAO = aID_STAO, aID_SP = aID_SP, yearP = yearP, yearPl = yearPl, 
             Visit = as.integer(floor((yearP - 1998) / 5)),
@@ -120,38 +120,29 @@ d <- expand.grid(aID_STAO = sites$aID_STAO, aID_SP = unique(pl$aID_SP)) %>% as.t
               transmute(aID_STAO = aID_STAO, aID_SP = aID_SP, Occ2 = Occ)) %>% 
   left_join(pl %>% filter(yearP >= 2013 & yearP <= 2017) %>% 
               transmute(aID_STAO = aID_STAO, aID_SP = aID_SP, Occ3 = Occ)) %>% 
-  replace_na(replace = list(Occ1 = 0, Occ2 = 0, Occ3 = 0)) %>% 
-  left_join(sites) 
+  replace_na(replace = list(Occ1 = 0, Occ2 = 0, Occ3 = 0))
 
 # Prepare colonization data
 coldat <- rbind(
   d %>% filter(Occ1 == 0) %>% 
     transmute(aID_STAO = aID_STAO, aID_SP =aID_SP, 
-              Temperature = Temperature, Precipitation = Precipitation,
-              NTOT = NTOT, Inclination = Inclination,
               T = T, F = F, N = N, L = L, Occ = Occ2, Period = 1),
   d %>% filter(Occ2 == 0) %>% 
-    transmute(aID_STAO = aID_STAO, aID_SP =aID_SP, 
-              Temperature = Temperature, Precipitation = Precipitation,
-              NTOT = NTOT, Inclination = Inclination,
-              T = T, F = F, N = N, L = L, Occ = Occ3, Period = 2))%>% 
-  filter(!is.na(T) & !is.na(F) & !is.na(N) & !is.na(L)) 
[…]

### 2018-05-30 R/Run_models.R

@@ -67,11 +67,13 @@ mod <- stan_glmer(cbind(change, nochange) ~ period + Temperature + Precipitation
 save(mod, file = "Modelfit/turnover-trend.RData")
 
 #------------------------------------------------------------------------------------------------------
-# Modell for the colonization and local survival probabilities
+# Model for the colonization and local survival probabilities
 #------------------------------------------------------------------------------------------------------
+coldat <- coldat %>% left_join(sites) %>% dplyr::filter(!is.na(N))
 mod <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), data = coldat, family = binomial)
 save(mod, file = "Modelfit/colonization-N.RData")
 
+survdat <- survdat %>% left_join(sites) %>% dplyr::filter(!is.na(N))
 mod <- glmer(Occ ~ NTOT * N + (1|aID_SP) + (1|aID_STAO), data = survdat, family = binomial)
 save(mod, file = "Modelfit/localsurvival-N.RData")
 

### 2018-06-12 Manuscript.Rmd
Update Manuscript.Rmd

@@
     affiliation: 2
+ - name: Beat Rihm
+    affiliation: 3
+ - name: Reto Meier
+    affiliation: 4    
 affiliation:
@@
     address: Hintermann  Weber AG, Austrasse 2a, 4153 Reinach, Switzerland
+ - code: 3
+    address: Meteotest,Fabrikstrasse 14, 3012 Bern, Switzerland 
+ - code: 4
+    address: Federal Office for the Environment (FOEN), Air Pollution Control and Chemicals Division, 3003 Bern, Switzerland
+
 abstract: >
@@
 
-Here we inferred mountain hay meadows in Switzerland. Explain why mountain hay meadows are important. Also explain other threats to mountain hay meadows (climate change, land-use change).
+Here we inferred mountain hay meadows (EUNIS E2.3) in Switzerland. Explain why mountain hay meadows are important. Also explain other threats to mountain hay meadows (climate change, land-use change).
 
@@
 After the sampling of the plant data the botanists also assigned a habitat type to each sampling site according to the classification system developed for Switzerland [@Delarze2008].
-For more details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
 
-We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all sampling sites in mountain hay meadows (EUNIS E2.3).
+We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all sampling sites in mountain hay meadows (EUNIS E2.3) For more details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
 We analysed the data from 2003 to 2017.
@@
 
-For each survey (a survey consists of the two visits within the same season) we calculated the following community measures: (1) The number of recorded species (species richness), (2) the community mean of the Ellenberg temperature values of recorded species (community temperature index), (3) the community mean of the Ellenberg humidity values of recorded species (community humidity index), (4) the community mean of the Ellenberg nutrients values of recorded species (community nutrients index) and (5) the community mean of the Ellenberg light values of recorded species (community light index).
-The Ellenberg values were obtained from the recalibrated indicator values for the Swiss Flora [@Landolt2010].
+For each survey (a survey consists of the two visits per season) we calculated the following community measures: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for  temperature (2), humidity (3), nutrients (4) and light (5).
+The indicator values were obtained from the recalibrated indicator values for the Swiss Flora [@Landolt2010].
 Additionally to these five community measures that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
@@
 For details on the estimation of N deposition see @Rihm2001 and @Roth2017.
-Finally we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively used.
-We expect that well-accessible sites become more intensively managed, whereas management intensity will rather decrease at remote sites [@Strebel2015].
+Finally we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed.
+(folgenden Satz streichen?) We expect that well-accessible sites become more intensively managed, whereas management intensity will rather decrease at remote sites [@Strebel2015].
 
@@
 
-To infer whether species that colonized or disappeared from a site had particular Ellenberg values that differed from the other species at that site, we produced for each site a list with all species that were recorded during the three surveys (total community).
-We then calculated the community mean (CM) of the Ellenberg value for all species that colonized the site during the three surveys (i.e.
+To infer whether species that colonized or disappeared from a site had particular indicator values that differed from the other species at that site, we produced for each site a list with all species that were recorded during the three surveys (total community).
+We then calculated the community mean (CM) of the indicator value for all species that colonized the site during the three surveys (i.e.
 not recorded during first survey and recorded during second or not recorded during second and recorded during third).
-We then randomly selected the same number of species from the total community and also calculated the community mean of the Ellenberg value for these species (random-CM).
+We then randomly selected the same number of species from the total community and also calculated the community mean of the value for these species (random-CM).
 We repeated the random selection of species 1000 times.
 Then we calculated the differences of the CM minus the average of the random-CMs to obtain a standardized measure (standardized-CM) of how different the disappearing species were from random expectation.
-A standardized-CM < 0 would suggest that the Ellenberg value of colonizing species were lower than the average species that were occurring at this site.
-We applied this method for both colonizing and disappearing species and for the Ellenberg values for temperature, humidity, nutrients and light (see Appendix A).
+A standardized-CM < 0 would suggest that the indicator value of colonizing species were lower than the average species that were occurring at this site.
+We applied this method for both colonizing and disappearing species and for the indicator values for temperature, humidity, nutrients and light (see Appendix A).
 We then inferred whether the standardized-CM is changing along the corresponding gradient using linear models.
@@
 
-We used logistic-LMM to infer whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this changing depended on the species Ellenberg value for nutrients.
+We used logistic-LMM to infer whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this changing depended on the species indicator value for nutrients.
 To analyse the colonization probability we selected all species that were not observed during the first survey, and asked if they were observed (then Occ = 1) or not observed (then Occ = 0) during the second survey.
@@
 The variable 'occ' was then used as dependent variable in the logistic-LMM.
-As predictor variables the model contained the Nitrogen deposition of the site, the Ellenberg value for Nutrients of the species and the interaction of these two variables.
-Additionally we species-ID and site-ID were included as random effects.
+As predictor variables the model contained the Nitrogen deposition of the site, the indicator value for nutrients of the species and the interaction of these two variables.
+Additionally  species-ID and site-ID were included as random effects.
 The same logistic_LMM was also used to infer local survival probability.
@@
 
-To estimate the effect of N deposition on total species richness at a given time point, we described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarithm as link function.
-As predictors we used the four environmental gradients that we described above.
+To estimate the effect of Nitrogen deposition on total species richness at a given time point, we described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarithm as link function.
+As predictors we used the four environmental gradients as described above.
 Model parameters were estimated using an approximate Bayesian approach using the R-Package *arm* [@arm2018].
@@
 
-# Alpha diversity
+# Alpha diversity (species richness) 
 res[1, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$SR, surv$Visit, mean)
@@
 for trend"), caption = "Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
-The temporal trends are given as change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function).
+The temporal trends are given as change per 10 years and were estimated from linear mixed models with normal distribution (except for species richness with Poisson distribution and a log-link function).
 The measure of precision for the temporal trend is given as the 5\\% and 95\\% quantiles of the marginal posterior distribution of the linear trend.
@@
 
-The five measures of plant community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha-diversity and the average Ellenberg values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
-Except for average Ellenberg value for temperature, the 90% credible-interval of the temporal trend contained zero.
-The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the Ellenberg value for temperature (probability of increase: `r
+The five measures of plant community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha-diversity and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
+Except for average indicator value for temperature, the 90% credible-interval of the temporal trend contained zero.
+The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the indicator value for temperature (probability of increase: `r
 res %>% filter(Measures == "Temperature value") %>% pull("Prob.
-for Trend")`), followed by the community mean of the Ellenberg light value (probability of decrease: `r 1 - res %>% filter(Measures == "Light value") %>% pull("Prob.
-for Trend")`) and it was least likely for the alpha-diversity (probability of increase: `r res %>% filter(Measures == "Alpha-diversity") %>% pull("Prob.
+for Trend")`), followed by the community mean of the indicator value for light (probability of decrease: `r 1 - res %>% filter(Measures == "Light value") %>% pull("Prob.
+for Trend")`) and it was least likely for the species richness (probability of increase: `r res %>% filter(Measures == "Alpha-diversity") %>% pull("Prob.
 for Trend")`).
@@
 High species turnover at a site is the result of species that disappeared from the site and species that newly colonized the site.
-To better understand the factors that drive these changes we are particularly interested whether the species that disappeared or colonized the sites differed in Ellenberg values compared to what would be expected if the same number of species randomly disappeared or colonized the sites (i.e.
+To better understand the factors that drive these changes we are particularly interested whether the species that disappeared or colonized the sites differed in indicator values compared to what would be expected if the same number of species randomly disappeared or colonized the sites (i.e.
 random disappearance and random colonization) and whether there is a change along the gradients.
-It seems that the Ellenberg values of newly colonizing species differed more from random colonization than the Ellenberg values of disappearing species (Table \@ref(tab:difffromrandomtab)).
-For colonizing species, we found the largest differences from random colonization in the Ellenberg value for nutrients: at sites with nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower Ellenberg value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
+It seems that the indicator values of newly colonizing species differed more from random colonization than the indicator values of disappearing species (Table \@ref(tab:difffromrandomtab)).
+For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower indicator value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing Nitrogen deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
 Thus at high Nitrogen deposition colonizing species did not differ from random species (see Figure 3b in Appendix A).
 
-While colonizing species had higher temperature values compared to what we would expect under random colonization, the differences between colonizing species and random colonization was about four times smaller compared to the difference in Ellenberg value for nutrients between colonizing species and random species.
-Nevertheless, the variation in Ellenberg value for temperature seemed important to explain the total species turnover.
+While colonizing species had higher temperature values compared to what we would expect under random colonization, the differences between colonizing species and random colonization was about four times smaller compared to the difference in indicator value for nutrients between colonizing species and random species.
+Nevertheless, the variation in indicator value for temperature seemed important to explain the total species turnover.
 This is because, disappearing species tend to have lower temperature value than random species as well as colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
-This was not the case for the Ellenberg value for nutrients: species with lower nutrients values tended to be more likely to disappear from as well as to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
+This was not the case for the indicator value for nutrients: species with lower nutrients values tended to be more likely to disappear from as well as to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
 See also Appendix A were we present detailed results for the comparison between colonizing or disappearing species with randomly selected species.
@@
 In Fig.
-\@ref(fig:cl) we compare the colonization and local survival probability of oligotrophic (Ellenberg value of nutrients = 2) and eutrophic (Ellenberg value of nutrients = 4) species along the Nitrogen deposition gradient.
+\@ref(fig:cl) we compare the colonization and local survival probability of oligotrophic (indicator value of nutrients = 2) and eutrophic (indicator value of nutrients = 4) species along the Nitrogen deposition gradient.
 Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for for oligotrophic and euthrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
@@
 
-While we could not detect a consistent decrease in average Ellenberg value for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial variation of species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between N deposition and species richness.
-Indeed, if we apply at different time points a similar model as in @Roth2013 to infer the effects of N deposition on the spatial variation of species richness, the resulting effect size (i.e.
+While we could not detect a consistent decrease in average indicator value for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial variation of species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between Nitrogen deposition and species richness.
+Indeed, if we apply at different time points a similar model as in @Roth2013 to infer the effects of Nitrogen deposition on the spatial variation of species richness, the resulting effect size (i.e.
 the slope) becomes more negative over time (Fig.

### 2018-06-12 Manuscript.Rmd
Update Manuscript.Rmd
Ellenberg ersetzt
Introduction eingefügt
Diskussion

@@
 # Introduction {-}
-Nitrogen (N) deposition is a major threat to biodiversity.
-The recent introduction of cleaner technologies has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in N-deposition in many parts of Europe.
+Nitrogen deposition is supposed to be one of the major threats to biodiversity (Sala 2000, Bobbink 2010, Menier et al.
+2016).
+There is strong evidence for the impairment of biodiversity by nitrogen deposition for species-rich acid and calcareous grasslands in Europe (Stevens et al.
+2006, Maskell et al.
+2010, Duprè et al.
+2010).
+Fertilizing experiments show that plant biodiversity in mountain grasslands is negatively affected by nitrogen addition, too (Humbert et al.
+2016).
+Since the late 1980s measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe (Tørseth et al., 2012, Maas and Grennfelt, 2016) [Tørseth et al.,: Atmos.
+Chem.
+Phys., 12, 5447–5481, 2012].
+
 However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities.
+Quantifying the potential recovery of biodiversity in response to reducing air pollution requires an alternative to the space-for-time approach, ideally monitoring long-term community dynamics on permanent plots (Duprã, C.
+et al.
+Changes in species richness and composition in European acidic grasslands over the past 70 years: the contribution of cumulative atmospheric nitrogen deposition.
+Glob.
+Change Biol.
+16, 344–357, (2010).
 
-One useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both immigration and extinction, often in a closed range of values (Hillebrand_et_al-2018.pdf).
+One useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both immigration and extinction, often in a closed range of values (Hillebrand et al-2018.pdf).
 
-Here we inferred mountain hay meadows (EUNIS E2.3) in Switzerland. Explain why mountain hay meadows are important. Also explain other threats to mountain hay meadows (climate change, land-use change).
+Observation data of vascular plants from field recordings shows negative correlations of Nitrogen deposition with species richness and species composition in mountain hay meadows (Roth et al. 2013). 
+In Switzerland, grassland accounts for 70 per cent of the agricultural land.
+With extensive cultivation, permanent grassland has a very high biodiversity.
+This applies in particular to the meadows of the Swiss mountain region (Schlup et al.
+2013).
+In the alpine region, meadows with a high plant diversity are of great importance agronomically (Leiber et al.
+2005), for the image of the products and for the conservation of biodiversity (Hohl, 2006).
+The meta-analysis of Humbert et al.
+(2016) showed management and abiotic factors to interact with nitrogen addition to drive plant community changes: The overall effects of N addition on plant species richness and diversity (Shannon index) were more pronounced in managed experimental sites with warmer summers.
+Management regimes of accessible mountainous areas are intensified, while poorly accessible mountainous areas are being abandoned (Tasser and Tappeiner, 2002; Graf et al., 2014, Strebel and Bühler, 2015).
+This change in traditional management regimes is considered as a major driver that endangers grassland biodiversity in mountainous regions of Europe (Niedrist et al., 2009; Homburger and Hofer, 2012).
+
+On European mountain summits, plants shifted upwards in the last decades (Rumpf et al.
+2018).
+The authors conclude that «These results suggest that recent climate warming interacted with airborne nitrogen deposition in driving the observed dynamics.
+As changes increased more for species from lower elevations and eutrophic plant species showed the strongest increase in abundance, these findings might be relevant for mountain hay meadows, too.
 
@@
 ## Environmental gradients {-}
-We expected different global change patterns to simultaneously cause temporal change in mountain hey communities.
+We expected different global change patterns to simultaneously cause temporal change in mountain hay communities.
 To better disentangle the importance of these mechanisms we ordered the sites along four main environmental gradients.
@@
 ```{r difffromrandomtab}
-kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Ellenberg value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average Ellenberg value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
+kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Ellenberg value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average indicator value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
 Shown are the results from linear model with the difference between disappeard/colonized species and random species as dependend variable and the sitemeasure (gradient) as predictor variable.") %>%
@@
 - *General points*: Although N deposition considerably declined between 2005 and 2015, we could not detect major shifts in plant community structure during the same time period. 
+Although NOx emissions in Switzerland decreased by 46% between 1990 and 2010 and NH3 emissions by 14% (Maas and Grennfelt, 2016), nitrogen deposition on our plots decreased by only 2.7 kg ha$^{-1}$ yr$^{-1}$ on average between 2000 and 2015 (Meteotest, 2018).
+This is only about one tenth of the decrease in England in the same period, where Nitrogen deposition decreased by 24kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 (Storkey et al.
+2015).
+In 2015, average N deposition was 14.8 kg ha$^{-1}$ yr$^{-1}$, about 85% of 2000 deposition, and due to the rather short period of time and the small decrease in nitrogen deposition, it is not surprising that species richness has not reacted significantly.
 
@@
 - *Methodological point*: The rather large spatial turnover might be partly explained by species that remained undetected in one of the surveys.
-However, our results suggest that turnover is caused at least partly by species with specific Ellenberg values.
-These deviation from what we would expect under random species turnover is unlikely to be explained by species that remained undetected.
+From the randomly selected quality control of the BDM 17 mountain hay meadows were investigated from 2003 to 2017.
+Species richness was surveyed in the same year by two independent botanists.
+The species match is (mean ± SD) 85 ± 5.2 per cent, suggesting that at least a portion of the turn over is due to agent effects.
+However, our results suggest that turnover is caused at least partly by species with specific indicator values.
+This deviation from what we would expect under random species turnover is unlikely to be explained by species that remained undetected.
 

### 2018-06-14 Manuscript.Rmd
Kleiner Anpassungen. Verschiedene Zwischenprodukte löschen.

@@
     address: Federal Office for the Environment (FOEN), Air Pollution Control and Chemicals Division, 3003 Bern, Switzerland
-
 abstract: >
@@
 # Introduction {-}
-Nitrogen deposition is supposed to be one of the major threats to biodiversity (Sala 2000, Bobbink 2010, Menier et al.
+Nitrogen deposition is one of the major threats to biodiversity (Sala 2000, Bobbink 2010, Menier et al.
 2016).
-There is strong evidence for the impairment of biodiversity by nitrogen deposition for species-rich acid and calcareous grasslands in Europe (Stevens et al.
+Especially, there is strong evidence for the reduction in diversity of species-rich grasslands in Europe (Stevens et al.
 2006, Maskell et al.
@@
 2010).
-Fertilizing experiments show that plant biodiversity in mountain grasslands is negatively affected by nitrogen addition, too (Humbert et al.
-2016).
+
+Fertilizing experiments show that plant biodiversity in mountain grasslands is negatively affected by nitrogen addition, too (Humbert et al. 2016).
+
 Since the late 1980s measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe (Tørseth et al., 2012, Maas and Grennfelt, 2016) [Tørseth et al.,: Atmos.
@@
 Phys., 12, 5447–5481, 2012].
+However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities.
 
-However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities.
 Quantifying the potential recovery of biodiversity in response to reducing air pollution requires an alternative to the space-for-time approach, ideally monitoring long-term community dynamics on permanent plots (Duprã, C.
@@
 Observation data of vascular plants from field recordings shows negative correlations of Nitrogen deposition with species richness and species composition in mountain hay meadows (Roth et al. 2013).
+
 In Switzerland, grassland accounts for 70 per cent of the agricultural land.
@@
 2005), for the image of the products and for the conservation of biodiversity (Hohl, 2006).
+
 The meta-analysis of Humbert et al.
@@
 First we expected communities to response to climate warming and the response might be different depending along the temperature gradient [@Roth2014].
-To describe the temperature gradient we used the annual mean temperature per site from the WordClim database [@Fick2017].
+To describe the temperature gradient we used the mean annual temperature per site from the WordClim database [@Fick2017].
 Another key drivers for plants that is likely to be affected by climate change is precipitation [@Beier2012].
-We used the annual mean precipitation per site from the WordClim database [@Fick2017].Further, we estimated atmospheric N deposition for each site using a pragmatic approach described in @Rihm2001 that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models.
+We used the annual precipitation per site from the WordClim database [@Fick2017].Further, we estimated atmospheric N deposition for each site using a pragmatic approach described in @Rihm2001 that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models.
 For details on the estimation of N deposition see @Rihm2001 and @Roth2017.
-Finally we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed.
-(folgenden Satz streichen?) We expect that well-accessible sites become more intensively managed, whereas management intensity will rather decrease at remote sites [@Strebel2015].
+Finally we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed [@Strebel2015].
 
 ## Statistical analyses {-}
-To estimate the linear trend over time for each of the five community measures we applied linear mixed models (LMM) with normal distribution except for alpha-diversity with Poisson distribution and the logarithm as link function.
+To estimate the linear trend over time for each of the five community measures we applied linear mixed models (LMM) with normal distribution except for species-richness with Poisson distribution and the logarithm as link function.
 We specified site specific trends with the assumption that the between-site differences in intercepts and slopes can be described with normal distributions [i.e.
@@
 ```{r communitytrendscalc}
-res <- tibble(Measures = c("Alpha-diversity", "Temperature value", "Huminity value", "Nutrients value", "Light value"),
+res <- tibble(Measures = c("Species-richness", "Temperature value", "Huminity value", "Nutrients value", "Light value"),
               'Period 1' = NA,
@@
 
-# Alpha diversity (species richness) 
+# Species richness
 res[1, c("Period 1", "Period 2", "Period 3")] <- tapply(surv$SR, surv$Visit, mean)
@@
 
-The five measures of plant community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of alpha-diversity and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
+The five measures of plant community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of species-richness and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
 Except for average indicator value for temperature, the 90% credible-interval of the temporal trend contained zero.
@@
 for Trend")`), followed by the community mean of the indicator value for light (probability of decrease: `r 1 - res %>% filter(Measures == "Light value") %>% pull("Prob.
-for Trend")`) and it was least likely for the species richness (probability of increase: `r res %>% filter(Measures == "Alpha-diversity") %>% pull("Prob.
+for Trend")`) and it was least likely for the species richness (probability of increase: `r res %>% filter(Measures == "Species-richness") %>% pull("Prob.
 for Trend")`).
@@
 Species richness was surveyed in the same year by two independent botanists.
-The species match is (mean ± SD) 85 ± 5.2 per cent, suggesting that at least a portion of the turn over is due to agent effects.
+The species match is (mean $\pm$ SD) 85 $\pm$ 5.2 per cent, suggesting that at least a portion of the turn over is due to agent effects.
 However, our results suggest that turnover is caused at least partly by species with specific indicator values.
@@
 # The MS was build with the following R sesstion:
-writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
+writeLines(capture.output(sessionInfo()), "Settings/sessionInfo.txt")
 ```

### 2018-06-14 Manuscript.tex
Kleiner Anpassungen. Verschiedene Zwischenprodukte löschen.

@@
 
+\author[3]{Beat Rihm}
+
+\author[4]{Reto Meier}
+
 \affil[1]{Zoological Institute, University of Basel, Basel, Switzerland}
 \affil[2]{Hintermann Weber AG, Austrasse 2a, 4153 Reinach, Switzerland}
+\affil[3]{Meteotest,Fabrikstrasse 14, 3012 Bern, Switzerland}
+\affil[4]{Federal Office for the Environment (FOEN), Air Pollution Control and
+Chemicals Division, 3003 Bern, Switzerland}
 
@@
 
-Nitrogen (N) deposition is a major threat to biodiversity. The recent
-introduction of cleaner technologies has let to reductions in the
-emissions of nitrogen oxides, with affiliated decrease in N-deposition
-in many parts of Europe. However, it is an open question whether and how
+Nitrogen deposition is one of the major threats to biodiversity (Sala
+2000, Bobbink 2010, Menier et al. 2016). Especially, there is strong
+evidence for the reduction in diversity of species-rich grasslands in
+Europe (Stevens et al. 2006, Maskell et al. 2010, Duprè et al. 2010).
+
+Fertilizing experiments show that plant biodiversity in mountain
+grasslands is negatively affected by nitrogen addition, too (Humbert et
+al. 2016).
+
+Since the late 1980s measures to reduce atmospheric pollution have
+successfully reduced emissions of nitrogen oxides, with affiliated
+decrease in Nitrogen deposition in many parts of Europe (Tørseth et al.,
+2012, Maas and Grennfelt, 2016) {[}Tørseth et al.,: Atmos. Chem. Phys.,
+12, 5447--5481, 2012{]}. However, it is an open question whether and how
 fast the reduction in N deposition rates will lead to the recovery of
@@
 
+Quantifying the potential recovery of biodiversity in response to
+reducing air pollution requires an alternative to the space-for-time
+approach, ideally monitoring long-term community dynamics on permanent
+plots (Duprã, C. et al. Changes in species richness and composition in
+European acidic grasslands over the past 70 years: the contribution of
+cumulative atmospheric nitrogen deposition. Glob. Change Biol. 16,
+344--357, (2010).
+
 One useful approach to understanding biodiversity change is through
 estimates of biodiversity turnover reflecting both immigration and
-extinction, often in a closed range of values
-(Hillebrand\_et\_al-2018.pdf).
+extinction, often in a closed range of values (Hillebrand et
+al-2018.pdf).
 
-Here we inferred mountain hay meadows in Switzerland. Explain why
-mountain hay meadows are important. Also explain other threats to
-mountain hay meadows (climate change, land-use change).
+Observation data of vascular plants from field recordings shows negative
+correlations of Nitrogen deposition with species richness and species
+composition in mountain hay meadows (Roth et al. 2013).
+
+In Switzerland, grassland accounts for 70 per cent of the agricultural
+land. With extensive cultivation, permanent grassland has a very high
+biodiversity. This applies in particular to the meadows of the Swiss
+mountain region (Schlup et al. 2013). In the alpine region, meadows with
+a high plant diversity are of great importance agronomically (Leiber et
+al. 2005), for the image of the products and for the conservation of
+biodiversity (Hohl, 2006).
+
+The meta-analysis of Humbert et al. (2016) showed management and abiotic
+factors to interact with nitrogen addition to drive plant community
+changes: The overall effects of N addition on plant species richness and
+diversity (Shannon index) were more pronounced in managed experimental
+sites with warmer summers. Management regimes of accessible mountainous
+areas are intensified, while poorly accessible mountainous areas are
+being abandoned (Tasser and Tappeiner, 2002; Graf et al., 2014, Strebel
+and Bühler, 2015). This change in traditional management regimes is
+considered as a major driver that endangers grassland biodiversity in
+mountainous regions of Europe (Niedrist et al., 2009; Homburger and
+Hofer, 2012).
+
+On European mountain summits, plants shifted upwards in the last decades
+(Rumpf et al. 2018). The authors conclude that «These results suggest
+that recent climate warming interacted with airborne nitrogen deposition
+in driving the observed dynamics. As changes increased more for species
+from lower elevations and eutrophic plant species showed the strongest
+increase in abundance, these findings might be relevant for mountain hay
+meadows, too.
 
@@
 classification system developed for Switzerland (Delarze and Gonseth
-2008). For more details on the field methods see Plattner, Birrer, and
-Weber (2004), Roth et al. (2013) and Roth et al. (2017).
+2008).
 
@@
 and Hill 2004) and selected all sampling sites in mountain hay meadows
-(EUNIS E2.3). We analysed the data from 2003 to 2017. During that study
-period each sampling site was surveyed once per five-year period: the
-first period lasted from 2003 to 2007, the second from 2008 to 2012 and
-the third from 2013 to 2017. These selection criteria resulted in sample
-of 129 sites.
+(EUNIS E2.3) For more details on the field methods see Plattner, Birrer,
+and Weber (2004), Roth et al. (2013) and Roth et al. (2017). We analysed
+the data from 2003 to 2017. During that study period each sampling site
+was surveyed once per five-year period: the first period lasted from
+2003 to 2007, the second from 2008 to 2012 and the third from 2013 to
+2017. These selection criteria resulted in sample of 129 sites.
 
-For each survey (a survey consists of the two visits within the same
-season) we calculated the following community measures: (1) The number
-of recorded species (species richness), (2) the community mean of the
-Ellenberg temperature values of recorded species (community temperature
-index), (3) the community mean of the Ellenberg humidity values of
-recorded species (community humidity index), (4) the community mean of
-the Ellenberg nutrients values of recorded species (community nutrients
-index) and (5) the community mean of the Ellenberg light values of
-recorded species (community light index). The Ellenberg values were
-obtained from the recalibrated indicator values for the Swiss Flora
-(Landolt et al. 2010). Additionally to these five community measures
-that describe the state of plant communities for each site at a given
-time point, we also estimated the temporal turnover (i.e.~species
-exchange ratio sensu Hillebrand et al. (2018)) as the proportion of
-species that differ between two time points to describe the community
-change between two time points.
+For each survey (a survey consists of the two visits per season) we
+calculated the following community measures: (1) The number of recorded
+species (species richness), the community mean of indicator values of
+recorded species for temperature (2), humidity (3), nutrients (4) and
+light (5). The indicator values were obtained from the recalibrated
+indicator values for the Swiss Flora (Landolt et al. 2010). Additionally
+to these five community measures that describe the state of plant
+communities for each site at a given time point, we also estimated the
+temporal turnover (i.e.~species exchange ratio sensu Hillebrand et al.
+(2018)) as the proportion of species that differ between two time points
+to describe the community change between two time points.
 
@@
 We expected different global change patterns to simultaneously cause
-temporal change in mountain hey communities. To better disentangle the
+temporal change in mountain hay communities. To better disentangle the
 importance of these mechanisms we ordered the sites along four main
@@
 temperature gradient (Roth, Plattner, and Amrhein 2014). To describe the
-temperature gradient we used the annual mean temperature per site from
+temperature gradient we used the mean annual temperature per site from
 the WordClim database (Fick and Hijmans 2017). Another key drivers for
 plants that is likely to be affected by climate change is precipitation
-(Beier et al. 2012). We used the annual mean precipitation per site from
-the WordClim database (Fick and Hijmans 2017).Further, we estimated
+(Beier et al. 2012). We used the annual precipitation per site from the
+WordClim database (Fick and Hijmans 2017).Further, we estimated
 atmospheric N deposition for each site using a pragmatic approach
@@
 we used inclination as proxy for land-use intensity because we expected
-that steeper sites are likely to be less intensively used. We expect
-that well-accessible sites become more intensively managed, whereas
-management intensity will rather decrease at remote sites (Strebel and
-Bühler 2015).
+that steeper sites are likely to be less intensively managed (Strebel
+and Bühler 2015).
 
@@
 measures we applied linear mixed models (LMM) with normal distribution
-except for alpha-diversity with Poisson distribution and the logarithm
+except for species-richness with Poisson distribution and the logarithm
 as link function. We specified site specific trends with the assumption
@@
 To infer whether species that colonized or disappeared from a site had
-particular Ellenberg values that differed from the other species at that
+particular indicator values that differed from the other species at that
 site, we produced for each site a list with all species that were
 recorded during the three surveys (total community). We then calculated
-the community mean (CM) of the Ellenberg value for all species that
+the community mean (CM) of the indicator value for all species that
 colonized the site during the three surveys (i.e.~not recorded during
@@
 species from the total community and also calculated the community mean
-of the Ellenberg value for these species (random-CM). We repeated the
-random selection of species 1000 times. Then we calculated the
-differences of the CM minus the average of the random-CMs to obtain a
-standardized measure (standardized-CM) of how different the disappearing
-species were from random expectation. A standardized-CM \textless{} 0
-would suggest that the Ellenberg value of colonizing species were lower
-than the average species that were occurring at this site. We applied
-this method for both colonizing and disappearing species and for the
-Ellenberg values for temperature, humidity, nutrients and light (see
-Appendix A). We then inferred whether the standardized-CM is changing
-along the corresponding gradient using linear models. Model parameters
-were estimated in a Bayesian framework using the R-Package
-\emph{rstanarm} (Stan Development Team 2016, Muth, Oravecz, and Gabry
-(2018)).
+of the value for these species (random-CM). We repeated the random
+selection of species 1000 times. Then we calculated the differences of
+the CM minus the average of the random-CMs to obtain a standardized
+measure (standardized-CM) of how different the disappearing species were
+from random expectation. A standardized-CM \textless{} 0 would suggest
+that the indicator value of colonizing species were lower than the
+average species that were occurring at this site. We applied this method
+for both colonizing and disappearing species and for the indicator
+values for temperature, humidity, nutrients and light (see Appendix A).
+We then inferred whether the standardized-CM is changing along the
+corresponding gradient using linear models. Model parameters were
+estimated in a Bayesian framework using the R-Package \emph{rstanarm}
+(Stan Development Team 2016, Muth, Oravecz, and Gabry (2018)).
 
@@
 local survival probability was changing along the Nitrogen deposition
-gradient and whether this changing depended on the species Ellenberg
+gradient and whether this changing depended on the species indicator
 value for nutrients. To analyse the colonization probability we selected
@@
 dependent variable in the logistic-LMM. As predictor variables the model
-contained the Nitrogen deposition of the site, the Ellenberg value for
-Nutrients of the species and the interaction of these two variables.
-Additionally we species-ID and site-ID were included as random effects.
-The same logistic\_LMM was also used to infer local survival
-probability. In that case, however, we selected all species that were
-recorded during the first or second survey and the variable `occ' then
-indicated whether or not the species was also observed during the next
-survey. Model parameters were estimated using an approximate Bayesian
-approach using the R-Package \emph{arm} (Gelman and Su 2018).
+contained the Nitrogen deposition of the site, the indicator value for
+nutrients of the species and the interaction of these two variables.
+Additionally species-ID and site-ID were included as random effects. The
+same logistic\_LMM was also used to infer local survival probability. In
+that case, however, we selected all species that were recorded during
+the first or second survey and the variable `occ' then indicated whether
+or not the species was also observed during the next survey. Model
+parameters were estimated using an approximate Bayesian approach using
+the R-Package \emph{arm} (Gelman and Su 2018).
 
-To estimate the effect of N deposition on total species richness at a
-given time point, we described the plant species richness at the sites
-using a generalized linear model with Poisson distribution and the
+To estimate the effect of Nitrogen deposition on total species richness
+at a given time point, we described the plant species richness at the
+sites using a generalized linear model with Poisson distribution and the
 logarithm as link function. As predictors we used the four environmental
-gradients that we described above. Model parameters were estimated using
-an approximate Bayesian approach using the R-Package \emph{arm} (Gelman
-and Su 2018).
+gradients as described above. Model parameters were estimated using an
+approximate Bayesian approach using the R-Package \emph{arm} (Gelman and
+Su 2018).
 
@@
 \caption{\label{tab:communitytrendstab}Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
-The temporal trends are given as change per 10 years and were estimated from linear mixed models with normal distribution (except for alpha-diversity with Poisson distribution and a log-link function).
+The temporal trends are given as change per 10 years and were estimated from linear mixed models with normal distribution (except for species richness with Poisson distribution and a log-link function).
 The measure of precision for the temporal trend is given as the 5\% and 95\% quantiles of the marginal posterior distribution of the linear trend.
@@
 \midrule
-Alpha-diversity & 45.72 & 46.02 & 45.74 & 0.00 & -0.03 & 0.03 & 0.53\\
+Species-richness & 45.72 & 46.02 & 45.74 & 0.00 & -0.03 & 0.03 & 0.53\\
 Temperature value & 3.11 & 3.13 & 3.13 & 0.01 & 0.00 & 0.03 & 0.97\\
@@
 \ref{tab:communitytrendstab}): for each of the three 5-year survey
-periods the averages of alpha-diversity and the average Ellenberg values
-for temperature, humidity, nutrients and light did not vary much among
-the three sampling periods and the estimated trends were rather small.
-Except for average Ellenberg value for temperature, the 90\%
+periods the averages of species-richness and the average indicator
+values for temperature, humidity, nutrients and light did not vary much
+among the three sampling periods and the estimated trends were rather
+small. Except for average indicator value for temperature, the 90\%
 credible-interval of the temporal trend contained zero. The results from
 the linear mixed models suggest that a linear temporal change was most
-likely for the community mean of the Ellenberg value for temperature
+likely for the community mean of the indicator value for temperature
 (probability of increase: 0.97), followed by the community mean of the
-Ellenberg light value (probability of decrease: 0.93) and it was least
-likely for the alpha-diversity (probability of increase: 0.53). The
-chance that the community mean of the nutrient value decreased between
-2003 and 2017 was 0.67.
+indicator value for light (probability of decrease: 0.93) and it was
+least likely for the species richness (probability of increase: 0.53).
+The chance that the community mean of the nutrient value decreased
+between 2003 and 2017 was 0.67.
 
@@
 
-\caption{\label{tab:difffromrandomtab}Difference in the average Ellenberg value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
+\caption{\label{tab:difffromrandomtab}Difference in the average indicator value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
 Shown are the results from linear model with the difference between disappeard/colonized species and random species as dependend variable and the sitemeasure (gradient) as predictor variable.}
@@
 \multicolumn{8}{l}{\textit{(a) Plants that disappeard from a site}}\\
-\hspace{1em}Temperature & Annual mean temperature & -0.012 & -0.034 & 0.009 & 0.007 & -0.002 & 0.016\\
-\hspace{1em}Humidity & Annual mean precipitation & -0.006 & -0.042 & 0.029 & 0.011 & -0.010 & 0.032\\
-\hspace{1em}Nutrients & Nitrogen deposition & -0.025 & -0.059 & 0.010 & 0.013 & -0.031 & 0.058\\
-\hspace{1em}Light & Inclination & -0.023 & -0.050 & 0.002 & -0.003 & -0.026 & 0.021\\
+\hspace{1em}Temperature & Annual mean temperature & -0.012 & -0.034 & 0.011 & 0.007 & -0.002 & 0.017\\
+\hspace{1em}Humidity & Annual mean precipitation & -0.006 & -0.042 & 0.032 & 0.011 & -0.010 & 0.032\\
+\hspace{1em}Nutrients & Nitrogen deposition & -0.024 & -0.059 & 0.011 & 0.013 & -0.028 & 0.054\\
+\hspace{1em}Light & Inclination & -0.023 & -0.050 & 0.003 & -0.002 & -0.026 & 0.021\\
 \addlinespace[0.3em]
 \multicolumn{8}{l}{\textit{(b) Plants that newly colonized a site}}\\
-\hspace{1em}Temperature & Annual mean temperature & 0.017 & 0.000 & 0.034 & -0.001 & -0.008 & 0.006\\
-\hspace{1em}Humidity & Annual mean precipitation & 0.019 & -0.013 & 0.051 & -0.002 & -0.020 & 0.016\\
-\hspace{1em}Nutrients & Nitrogen deposition & -0.076 & -0.108 & -0.045 & 0.060 & 0.023 & 0.098\\
-\hspace{1em}Light & Inclination & -0.041 & -0.063 & -0.018 & 0.010 & -0.010 & 0.029\\
+\hspace{1em}Temperature & Annual mean temperature & 0.016 & 0.000 & 0.033 & -0.001 & -0.008 & 0.006\\
+\hspace{1em}Humidity & Annual mean precipitation & 0.020 & -0.013 & 0.051 & -0.003 & -0.021 & 0.016\\
+\hspace{1em}Nutrients & Nitrogen deposition & -0.076 & -0.108 & -0.042 & 0.060 & 0.022 & 0.098\\
+\hspace{1em}Light & Inclination & -0.041 & -0.063 & -0.018 & 0.011 & -0.009 & 0.029\\
 \bottomrule
@@
 particularly interested whether the species that disappeared or
-colonized the sites differed in Ellenberg values compared to what would
+colonized the sites differed in indicator values compared to what would
 be expected if the same number of species randomly disappeared or
@@
 and whether there is a change along the gradients. It seems that the
-Ellenberg values of newly colonizing species differed more from random
-colonization than the Ellenberg values of disappearing species (Table
+indicator values of newly colonizing species differed more from random
+colonization than the indicator values of disappearing species (Table
 \ref{tab:difffromrandomtab}). For colonizing species, we found the
-largest differences from random colonization in the Ellenberg value for
+largest differences from random colonization in the indicator value for
 nutrients: at sites with nitrogen deposition of 10 kg ha\(^{-1}\)
 yr\(^{-1}\) the newly colonizing species had in average a lower
-Ellenberg value for nutrients than species under random disappearance
+indicator value for nutrients than species under random disappearance
 (column ``Difference from random'' in Table
 \ref{tab:difffromrandomtab}), but this differences between colonizing
-species and random colonization decreased with increasing N deposition
-(column ``Change along gradient'' in Table \ref{tab:difffromrandomtab}).
-Thus at high Nitrogen deposition colonizing species did not differ from
-random species (see Figure 3b in Appendix A).
+species and random colonization decreased with increasing Nitrogen
+deposition (column ``Change along gradient'' in Table
+\ref{tab:difffromrandomtab}). Thus at high Nitrogen deposition
+colonizing species did not differ from random species (see Figure 3b in
+Appendix A).
 
@@
 colonizing species and random colonization was about four times smaller
-compared to the difference in Ellenberg value for nutrients between
+compared to the difference in indicator value for nutrients between
 colonizing species and random species. Nevertheless, the variation in
-Ellenberg value for temperature seemed important to explain the total
+indicator value for temperature seemed important to explain the total
 species turnover. This is because, disappearing species tend to have
@@
 temperature value with species with higher temperature values. This was
-not the case for the Ellenberg value for nutrients: species with lower
+not the case for the indicator value for nutrients: species with lower
 nutrients values tended to be more likely to disappear from as well as
@@
 In Fig. \ref{fig:cl} we compare the colonization and local survival
-probability of oligotrophic (Ellenberg value of nutrients = 2) and
-eutrophic (Ellenberg value of nutrients = 4) species along the Nitrogen
+probability of oligotrophic (indicator value of nutrients = 2) and
+eutrophic (indicator value of nutrients = 4) species along the Nitrogen
 deposition gradient. Local survival probability was the same for
-oligotrophic and euthrophic species at a deposition rate of 11.66 kg N
+oligotrophic and euthrophic species at a deposition rate of 12.66 kg N
 ha\(^{-1}\) yr\(^{-1}\); colonization probability was the same for for
-oligotrophic and euthrophic species at a deposition rate of 12.09 kg N
+oligotrophic and euthrophic species at a deposition rate of 12.22 kg N
 ha\(^{-1}\) yr\(^{-1}\). In only 0.36\% of the sites the deposition rate
@@
 
-While we could not detect a consistent decrease in average Ellenberg
+While we could not detect a consistent decrease in average indicator
 value for nutrients (Table \ref{tab:communitytrendstab}), the higher
@@
 species rich over time likely resulting in steeper slope of the negative
-relationship between N deposition and species richness. Indeed, if we
-apply at different time points a similar model as in Roth et al. (2013)
-to infer the effects of N deposition on the spatial variation of species
-richness, the resulting effect size (i.e.~the slope) becomes more
-negative over time (Fig. \ref{fig:figconsequences}).
+relationship between Nitrogen deposition and species richness. Indeed,
+if we apply at different time points a similar model as in Roth et al.
+(2013) to infer the effects of Nitrogen deposition on the spatial
+variation of species richness, the resulting effect size (i.e.~the
+slope) becomes more negative over time (Fig. \ref{fig:figconsequences}).
 
@@
   between 2005 and 2015, we could not detect major shifts in plant
-  community structure during the same time period.
+  community structure during the same time period. Although NOx
+  emissions in Switzerland decreased by 46\% between 1990 and 2010 and
+  NH3 emissions by 14\% (Maas and Grennfelt, 2016), nitrogen deposition
+  on our plots decreased by only 2.7 kg ha\(^{-1}\) yr\(^{-1}\) on
+  average between 2000 and 2015 (Meteotest, 2018). This is only about
+  one tenth of the decrease in England in the same period, where
+  Nitrogen deposition decreased by 24kg ha\(^{-1}\) yr\(^{-1}\) from
+  1996 to 2011 (Storkey et al. 2015). In 2015, average N deposition was
+  14.8 kg ha\(^{-1}\) yr\(^{-1}\), about 85\% of 2000 deposition, and
+  due to the rather short period of time and the small decrease in
+  nitrogen deposition, it is not surprising that species richness has
+  not reacted significantly.
 \item
@@
   be partly explained by species that remained undetected in one of the
-  surveys. However, our results suggest that turnover is caused at least
-  partly by species with specific Ellenberg values. These deviation from
+  surveys. From the randomly selected quality control of the BDM 17
+  mountain hay meadows were investigated from 2003 to 2017. Species
+  richness was surveyed in the same year by two independent botanists.
+  The species match is (mean \(\pm\) SD) 85 \(\pm\) 5.2 per cent,
+  suggesting that at least a portion of the turn over is due to agent
+  effects. However, our results suggest that turnover is caused at least
+  partly by species with specific indicator values. This deviation from
   what we would expect under random species turnover is unlikely to be

### 2018-06-14 Manuscript.Rmd
Grafik zur Verteilung der NDep über die Plots (Zustand und Veränderung).

@@
 ## Potential effects of reduction in Nitrogen emmissions {-}
+```{r Ndepchange, fig.cap="(a) The nitrogen deposition in 2010 and (b) the change in Nitrogen deposition between 2000 and 2010 along the elevational gradient of the study sites."}
+p1 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = NTOT_2000)) +
+  geom_point(cex = 0.8) +
+  geom_smooth(method = loess) +
+  geom_abline(intercept = 0, slope = 0, lty = 3) +
+  labs(title = "(a) Nitrogen deposition in 2010", 
+       y = expression(paste("Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")")), 
+       x = "Elevation (m)") 
+p2 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = (NTOT_2010 - NTOT_2000) )) +
+  geom_point(cex = 0.8) +
+  geom_smooth(method = loess)  +
+  labs(title = "(b) Temporal change (2000 to 2010)", 
+       y = expression(paste(Delta, "Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")")), 
+       x = "Elevation (m)") +
+  geom_hline(yintercept = 0, lty = 2) 
+multiplot(p1, p2, cols = 2)
+```
+
+Nitrogen deposition decreased with increasing elevation (\@ref(fig:Ndepchange)a).
+In 2010 only `r 100 * mean(sites$NTOT_2010 < 10)`% of sites had a Nitrogen deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, which lied all above 1000m.
+Between 2000 and 2010 the Nitrogen deposition decreased in average around 2 kg N ha$^{-1}$ yr$^{-1}$ with slightly larger net decreases at lower elevation (\@ref(fig:Ndepchange)a).
+
 ```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
@@
   geom_vline(xintercept = quantile(CLs, probs = 0.95, na.rm = TRUE), lty = 2) +
-  xlim(0,30)
+  xlim(0,30) + theme(legend.position = c(0.75, 0.9)) + 
+  scale_color_manual(labels = c("Nutrients value = 2", "Nutrients value = 4"), 
+                     values = tcol[c(2,4)])
 

### 2018-06-14 R/Settings.R

@@ -56,6 +56,6 @@ tcol <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
           "#D55E00", "#CC79A7")
 
 # Numer of simulations from posterior distirbution
-nsim <- 100
+nsim <- 200
 
 

### 2018-06-14 Manuscript.Rmd
Claculation of pseudo-turnover between two survey from the same year.

@@
 
-Nitrogen deposition decreased with increasing elevation (\@ref(fig:Ndepchange)a).
+Nitrogen deposition decreased with increasing elevation (Fig.
+\@ref(fig:Ndepchange)a).
 In 2010 only `r 100 * mean(sites$NTOT_2010 < 10)`% of sites had a Nitrogen deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, which lied all above 1000m.
-Between 2000 and 2010 the Nitrogen deposition decreased in average around 2 kg N ha$^{-1}$ yr$^{-1}$ with slightly larger net decreases at lower elevation (\@ref(fig:Ndepchange)a).
+Between 2000 and 2010 the Nitrogen deposition decreased in average around 2 kg N ha$^{-1}$ yr$^{-1}$ with slightly larger net decreases at lower elevation (Fig.
+\@ref(fig:Ndepchange)a).
 

### 2018-06-14 R/Data_processing.R

@@ -110,6 +110,31 @@ for(i in 1:nrow(sites)) {
 }
 
 #------------------------------------------------------------------------------------------------------
+# Turnover between two survey from the same year
+#------------------------------------------------------------------------------------------------------
+# Select sites with two surveys per year
+turnover_double <- tbl(db, "KD_Z9") %>% 
+  filter(Aufnahmetyp == "Doppelaufnahme_Z9_Pflanzen") %>% 
+  filter(yearPl >= 2003 & yearPl <= 2017) %>% 
+  as.tibble() %>% 
+  filter(!is.na(match(aID_STAO, sites$aID_STAO))) %>% 
+  transmute(aID_KD_K = aID_KD,  aID_STAO = aID_STAO, yearPl = yearPl) %>% 
+  left_join(surv %>% transmute(aID_KD_R = aID_KD,  aID_STAO = aID_STAO, yearPl = yearP)) %>% 
+  filter(!is.na(aID_KD_R))
+
+getturnover <- function(x) {
+  tbl(db, "PL") %>% 
+    filter(aID_KD == turnover_double$aID_KD_K[x] | 
+             aID_KD == turnover_double$aID_KD_R[x]) %>% 
+    filter(!is.na(aID_SP)) %>% 
+    transmute(aID_KD = aID_KD, aID_SP = aID_SP, Occ = 1) %>% 
+    sim(method = "cocogaston", listin = TRUE, listout = TRUE) %>% 
+    pull(cocogaston)
+}
+
+turnover_double$turnover <- map_dbl(1:nrow(turnover_double), getturnover)
+
+#------------------------------------------------------------------------------------------------------
 # Compile data for colonization and local survival calculations
 #------------------------------------------------------------------------------------------------------
 d <- expand.grid(aID_STAO = sites$aID_STAO, aID_SP = unique(pl$aID_SP)) %>% as.tibble() %>% 
@@ -167,3 +192,5 @@ save(coldat, file = "RData/coldat.RData")
 survdat$aID_STAO <- siteIDs$siteID[match(survdat$aID_STAO, siteIDs$aID_STAO)]
 save(survdat, file = "RData/survdat.RData")
 
+turnover_double$aID_STAO <- siteIDs$siteID[match(turnover_double$aID_STAO, siteIDs$aID_STAO)]
+save(turnover_double, file = "RData/turnover_double.RData")

### 2018-06-14 Manuscript.Rmd
Teil der Diskussion geschrieben und Referenzen ergänzt.

@@
 
+NOx emissions in Switzerland decreased by 46% between 1990 and 2010 and NH3 emissions by 14% (Maas and Grennfelt, 2016)
+
 The meta-analysis of Humbert et al.
@@
 species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points to describe the community change between two time points.
+
+To test data quality, independent replicate surveys are routinely performed within the BDM programm by botansists who were not involved in the regular BDM surveys.
+The regular surveyors did not know if and which sites were replicated [@Plattner2004].
+We used the data from `r nrow(turnover_double)` such replicated surveys that were available to calculate the pseudo-turnover, which is the proportion of species that differed between two surveys that were conducted by two different surveyors during the same year on the same site.
 
@@
 
-This temporal stability as inferred from the community measures was, however, in contrast to a rather large observed temporal turnover of species.
-The average percentage $\pm$ SD of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
+The average $\pm$ SD percentage of species that differed between replicated surveys was `r mean(100 * turnover_double$turnover)` $\pm$ `r sd(100 * turnover_double$turnover)`.
+This pseud-turnover was lower than the observed temporal turnover: the average percentage of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
 Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model: `r CrI`).
@@
 # Discussion {-}
-- *General points*: Although N deposition considerably declined between 2005 and 2015, we could not detect major shifts in plant community structure during the same time period. 
-Although NOx emissions in Switzerland decreased by 46% between 1990 and 2010 and NH3 emissions by 14% (Maas and Grennfelt, 2016), nitrogen deposition on our plots decreased by only 2.7 kg ha$^{-1}$ yr$^{-1}$ on average between 2000 and 2015 (Meteotest, 2018).
-This is only about one tenth of the decrease in England in the same period, where Nitrogen deposition decreased by 24kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 (Storkey et al.
-2015).
-In 2015, average N deposition was 14.8 kg ha$^{-1}$ yr$^{-1}$, about 85% of 2000 deposition, and due to the rather short period of time and the small decrease in nitrogen deposition, it is not surprising that species richness has not reacted significantly.
+Although Nitrogen deposition considerably declined between 2000 and 2015 (Fig.
+\@ref(fig:Ndepchange)), we could not detect major shifts in plant community structure during the same time period (Table \@ref(tab:communitytrendstab)).
+While the slight increase in average temperature value suggests that plant communities adopted to increasing temperature, the constant aveage nutrients value suggests that the decrease in Nitrogen deposition did not yet affected plant communities.
+However, the stability in community structure was accompanied by marked turnover in species identities.
+Species turnover especially varied along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition.
+These results are in line with recent studies suggesting that species loss (or gain) is not necessarily apparent in assessments of contemporary communities [@Vellend2013, @Dornelas2014, @Hillebrand2018].
 
-- *Replacement of oligotrophic with eutrophic species is faster than the oposite direction*: Eutrophic species have rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at high N deposition.
-This suggests that it takes
-more time to replace eutrophic by oligotrophic species than replacing oligotrophic by eutrophic species. Climatic effects may be more likely to be reversed than effects due to fertilization.
+Species communities are shaped by a range of factors, including deterministic processes such as environmental filtering or competitive interactions [@Goetzenberger2012, @Janevcek2013].
+Such factors select for species with specific characteristics.
+Community assembly theories thus suggest that the factors that drive change in species communitis could be inferred from comparing the charactereistics of disappearing or colonizign species with random expectation [@Chase2011].
+We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site.
+The nutrient values of colonization species showed larges deviations from random expectation suggesting that Nitrogen deposition or other factors that change the nutrient content of the soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
 
-- *Methodological point*: The rather large spatial turnover might be partly explained by species that remained undetected in one of the surveys.
-From the randomly selected quality control of the BDM 17 mountain hay meadows were investigated from 2003 to 2017.
-Species richness was surveyed in the same year by two independent botanists.
-The species match is (mean $\pm$ SD) 85 $\pm$ 5.2 per cent, suggesting that at least a portion of the turn over is due to agent effects.
+Nitrogen deposition at our plots decreased by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ on average between 2000 and 2015.
+This is only about one tenth of the decrease in England in the same period, where Nitrogen deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
+This might explain why we observed no change in average nutrient value of communities.
+An other explanation might be that communtiy change due to the change in the rate of Nitrogen deposition was overruled by other anthropogenic pressures including climate change.
+In this study eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at high N deposition.
+This suggests that it takes more time to replace eutrophic by oligotrophic species than replacing oligotrophic by eutrophic species.
+This is in contrast to the species' indicator values for temperature: species with low and high temperature values had both optimal temperatures were colonization and local survival probabilities were high (see Fig.
+.
+Thus comunitiy change in response to climate warming might be fast because low temperature species have lower local survival and high temperature species have higher colonization probability.
+
+Observational studies along a gradient of Nitrogen deposition often infer how the spatial variation in species richness is related to Nitrogen deposition.
+They assume that the spatial variation in species richness arose because over time some areas lost more species than others because they chronically experienced higher Nitrogen deposition.
+Although there is evidence supporting the use of such a 'space for time substitution' for detecting the effects of Nitrogen deposition on plant diversity [@Stevens2010], they can not replace studies that relate temporal patterns in species with Nitrogen deposition [@deSchrijver2011].
+There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of N-deposition [@Storkey2015].
+In @Clark2008, the recorded recovery in species richness within one or two decade was likely due to the species rich vegetation surrounding the experimental plots, from where immigration was easily feasible.
+This is in line with our results sugggesting that it is mainly colonizing and not disappearing species are important to shift back from euttrophic to oligotrophic communities.
+
+- *Methological issue*: The rather large spatial turnover could be partly explained by species that remained undetected in one of the surveys.
 However, our results suggest that turnover is caused at least partly by species with specific indicator values.
@@
 
-- *Space for time substitution*: Often observational studies infer the change of plant diversity along a gradient of N deposi- tion.
-Thus, they infer how the spatial variation in species richness is related to N deposition and assume that this spatial variation in species richness arose because over time some areas lost more species than others because they chronically experienced higher N deposition.
-Alt- hough there is evidence supporting the use of such a 'space for time substitution' for detecting the effects of N deposition on plant diversity (Stevens et al.
-2010), they can not replace stud- ies that relate temporal patterns in species with N deposition (De Schrijver et al.
-2011).
-While recovery of acidified surface waters has been well investigated (De Vries et al.
-2015), there are only a limited number of studies inferring temporal trends of plant species diversity relat- ed to varying amounts of N-deposition.
-Storkey et al.
-(2015) demonstrated a positive response of biodiversity to reducing N addition from either atmospheric pollution or fertilizers in the Park Grass Experiment: «The proportion of legumes, species richness and diversity increased across the experiment between 1991 and 2012 as N-deposition declined».
-For forest floor vegetation in permanent plots across Europe the exceedance of critical loads of N over a peri- od from 9 to 42 years had negative effects on the cover of oligotrophic plant species, i.e spe- cies that prefer nutrient-poor soils, although species richness remained constant (Dirnböck et al.
-2014).
-Another example of recovery in eutrophicated habitats gives the recovery of species richness in previously fertilized plots (Clark and Tilman 2008).
-In this study, the recorded recovery in species richness within one or two decade was likely due to the species rich vege- tation surrounding the experimental plots, from where immigration was easily feasible.
-
 # Conclusions {-}
-xxx
+Our study adds to the understanding of contemporary biodiversity change that is currently not not well understood (Magurran et al 2018.pdf).
+Our results support the notion of Hillebrand_et_al-2018.pdf that biodiversity monitoring programmes that quantify species turnover will generate a far more reliable understanding of the biotic response to changing environments than programmes solely tracking species richness.
 

### 2018-06-14 R/Settings.R

@@ -30,6 +30,7 @@ load("RData/pl.RData")
 load("RData/sites.RData")
 load("RData/coldat.RData")
 load("RData/survdat.RData")
+load("RData/turnover_double.RData")
 
 #------------------------------------------------------------------------------------------------------
 # Settings 

### 2018-06-15 Manuscript.Rmd
Ologotrophic species richness ergänzt. Diskussion überarbeitet.

@@
 makemod <- function(year) {
-  mod <- glm(SR ~ Temperature + Precipitation + NTOT + Inclination,
+  mod <- glm(SR_olig ~ Temperature + Precipitation + NTOT + Inclination,
              data = dat %>% filter(yearP >= year & yearP <= (year + 4)), family = poisson)
@@
 Although Nitrogen deposition considerably declined between 2000 and 2015 (Fig.
-\@ref(fig:Ndepchange)), we could not detect major shifts in plant community structure during the same time period (Table \@ref(tab:communitytrendstab)).
-While the slight increase in average temperature value suggests that plant communities adopted to increasing temperature, the constant aveage nutrients value suggests that the decrease in Nitrogen deposition did not yet affected plant communities.
-However, the stability in community structure was accompanied by marked turnover in species identities.
-Species turnover especially varied along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition.
-These results are in line with recent studies suggesting that species loss (or gain) is not necessarily apparent in assessments of contemporary communities [@Vellend2013, @Dornelas2014, @Hillebrand2018].
+\@ref(fig:Ndepchange)), we could detect only weak shifts in plant community structure during the same time period (Table \@ref(tab:communitytrendstab).
+While the slight increase in average temperature value suggests that plant communities adopted to increasing temperatures, the constant aveage nutrients value suggests that the decrease in Nitrogen deposition did not yet affected plant communities.
+However, this apparent stability in community structure was accompanied by marked temporal turnover in species identities.
+It seems unlikely that this temporal turnover can entirly be explained by methodological issues such as overlooked species.
+First, pseudo-turnover of species entities in two independent surveys of the same site during the same season was smaller than the observed temporal turnover between two surveys from different years.
+Second, spatial variation of turnover showed patterns that can hardly be explained by methodological issues.
+For instance species turnover varied along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition.
+Taken together, our results adds to the increasing evidence that contemporary communities are often relatively stability at community level, but this apparent stability is often accompanied by marked turnover of species [@Vellend2013, @Dornelas2014, @Hillebrand2018].
 
@@
 Such factors select for species with specific characteristics.
-Community assembly theories thus suggest that the factors that drive change in species communitis could be inferred from comparing the charactereistics of disappearing or colonizign species with random expectation [@Chase2011].
+Community assembly theories thus suggest that the factors that drive change in species communities could be inferred from comparing the charactereistics of disappearing or colonizing species with random expectation [@Chase2011].
 We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site.
-The nutrient values of colonization species showed larges deviations from random expectation suggesting that Nitrogen deposition or other factors that change the nutrient content of the soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
+We found that the nutrient values of colonization species showed largest deviations from random expectation suggesting that Nitrogen deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
 
 Nitrogen deposition at our plots decreased by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ on average between 2000 and 2015.
-This is only about one tenth of the decrease in England in the same period, where Nitrogen deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
-This might explain why we observed no change in average nutrient value of communities.
-An other explanation might be that communtiy change due to the change in the rate of Nitrogen deposition was overruled by other anthropogenic pressures including climate change.
-In this study eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at high N deposition.
+This is only about one tenth of the decrease in England, where Nitrogen deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
+This comparatively low change in Nitrogen deposition might explain why we observed no change in average nutrient value of communities.
+Additionally, other anthropogenic pressures such as climate change might overrule Nitrogen depositin effects on community structure.
+Particularly, we found that the species disappearing from the sites tend to have below average indicator values for temperature, while species that newly colonize sites have above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
+Thus, the effect of disappearing and the effect of colonizating species on the community mean for temperature is additive resulting in increasing average temperature values (Tab.
+\@ref(tab:communitytrendstab)).
+This is in contrast to how Nitrogen deposition is affecting disappearing and colonization of species: It seems that both species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
+Thus, the effect of disappearing and colonizating species on the average community value for nutrients partially cancel each other out.
+Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at high Nitrogen deposition.
 This suggests that it takes more time to replace eutrophic by oligotrophic species than replacing oligotrophic by eutrophic species.
-This is in contrast to the species' indicator values for temperature: species with low and high temperature values had both optimal temperatures were colonization and local survival probabilities were high (see Fig.
-.
-Thus comunitiy change in response to climate warming might be fast because low temperature species have lower local survival and high temperature species have higher colonization probability.
+Taken toghether, these reasons might explain why the response of mountain hay medow communities responded stronger to to climate warming affects than to Nitrogen reduction, altough the reduction in Nitrogen resulted in above average colonization of oligotrophic species.
 
 Observational studies along a gradient of Nitrogen deposition often infer how the spatial variation in species richness is related to Nitrogen deposition.
-They assume that the spatial variation in species richness arose because over time some areas lost more species than others because they chronically experienced higher Nitrogen deposition.
+They assume that the spatial variation in species richness (or other metrics of community structure) arose because over time some areas lost more species than others because they chronically experienced higher Nitrogen deposition.
 Although there is evidence supporting the use of such a 'space for time substitution' for detecting the effects of Nitrogen deposition on plant diversity [@Stevens2010], they can not replace studies that relate temporal patterns in species with Nitrogen deposition [@deSchrijver2011].
-There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of N-deposition [@Storkey2015].
-In @Clark2008, the recorded recovery in species richness within one or two decade was likely due to the species rich vegetation surrounding the experimental plots, from where immigration was easily feasible.
-This is in line with our results sugggesting that it is mainly colonizing and not disappearing species are important to shift back from euttrophic to oligotrophic communities.
+There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of N-deposition [@Clark2008, @Storkey2015].
+In an recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the Nitrogen deposition gradient at which species richnes of oligotrophic species richness start to decrease with increasing Nitrogen deposition [@Roth2017].
+Using this spatial variation in species richness and nitrogen deposition we estimated a critical load for mountain hay meadows of 13.1 kg ha$^{-1}$ yr$^{-1}$.
+In the current study we estimated the rate of Nitrogen deposition at which local survival probability or colonization probability was equal for oligotrophic and eutrophic species.
+This rate could also be interpreted as a critical load and revealed very similar results as the spatial approach.
+Our results are thus validating the space for time approach.
 
-- *Methological issue*: The rather large spatial turnover could be partly explained by species that remained undetected in one of the surveys.
-However, our results suggest that turnover is caused at least partly by species with specific indicator values.
-This deviation from what we would expect under random species turnover is unlikely to be explained by species that remained undetected.
-
-- *Empirical critical loads*: Our data on colonization and local survival (i.e.
-temporal variation) confirm the empirical critical loads that we inferred from analysing spatial co-variation of N deposition and species richness.
+Hower, the results of Fig.
+\@ref(fig:) also shows that the results of spatial comparisons must be interpreted carefully.
+When we used the same model to infer how species richness is decreasing along the Nitrogen deposition gradient at differnt time points, except for the first two study periods the decrease became steeper (more negative) over time.
+Our first interpretation was that the Nitrogen deposition effect became stronger over time.
+This contraticted our prediction that the nitrogen deposition should become weaker over time since Nitrogen deposition decreased.
+However, we then realized that species turnover was highest in low deposition sites (Tab.
+\@ref(tab:turnovertab)) where colonizationg species were species with below average indicator values for Nutrients.
+It seems that the decrease in Nitrogen deposition resulted in higher numbers of oligotrophic species at sites with low deposition, thus leading to a steeper decline in oligotrophic species richness from sites with low to sites with high Nitrogen deposition.
 

### 2018-06-15 R/Data_processing.R

@@ -76,6 +76,7 @@ surv <- surv %>% left_join(
     group_by(aID_KD) %>% 
     dplyr::summarise(
       SR = n(),
+      SR_olig = sum(N <= 2, na.rm = TRUE),
       T = mean(T, na.rm = TRUE), 
       F = mean(F, na.rm = TRUE),
       N = mean(N, na.rm = TRUE),

### 2018-06-17 Manuscript.Rmd
Diskussion überarbeitet.

@@
 ---
-title: Recovery of mountain plant communities in response to reductions in Nitrogen emissions is hidden by other drivers of global change
+title: Species turnover reveal hidden effects of decreasing Nitrogen deposition in mountain hay meadows
 preprint: false
@@
 ## Potential effects of reduction in Nitrogen emmissions {-}
-```{r Ndepchange, fig.cap="(a) The nitrogen deposition in 2010 and (b) the change in Nitrogen deposition between 2000 and 2010 along the elevational gradient of the study sites."}
+```{r Ndepchange, fig.cap="(a) The nitrogen deposition in 2000 and (b) the change in Nitrogen deposition between 2000 and 2010 along the elevational gradient of the study sites."}
 p1 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = NTOT_2000)) +
@@
        x = "Elevation (m)") 
-p2 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = (NTOT_2010 - NTOT_2000) )) +
+p2 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = (NTOT_2015 - NTOT_2000) )) +
   geom_point(cex = 0.8) +
   geom_smooth(method = loess)  +
-  labs(title = "(b) Temporal change (2000 to 2010)", 
+  labs(title = "(b) Temporal change (2000 to 2015)", 
        y = expression(paste(Delta, "Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")")), 
@@
 \@ref(fig:Ndepchange)a).
-In 2010 only `r 100 * mean(sites$NTOT_2010 < 10)`% of sites had a Nitrogen deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, which lied all above 1000m.
-Between 2000 and 2010 the Nitrogen deposition decreased in average around 2 kg N ha$^{-1}$ yr$^{-1}$ with slightly larger net decreases at lower elevation (Fig.
+In 2000 only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a Nitrogen deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, which lied all above 1000m.
+Between 2000 and 2015 the Nitrogen deposition decreased in average `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$ with slightly higher net decreases at lower elevation (Fig.
 \@ref(fig:Ndepchange)a).
@@
 Although Nitrogen deposition considerably declined between 2000 and 2015 (Fig.
-\@ref(fig:Ndepchange)), we could detect only weak shifts in plant community structure during the same time period (Table \@ref(tab:communitytrendstab).
+\@ref(fig:Ndepchange)), we could detect only weak shifts in plant community structure during the same time period (Table \@ref(tab:communitytrendstab)).
 While the slight increase in average temperature value suggests that plant communities adopted to increasing temperatures, the constant aveage nutrients value suggests that the decrease in Nitrogen deposition did not yet affected plant communities.
@@
 It seems unlikely that this temporal turnover can entirly be explained by methodological issues such as overlooked species.
-First, pseudo-turnover of species entities in two independent surveys of the same site during the same season was smaller than the observed temporal turnover between two surveys from different years.
+First, pseudo-turnover of species entities in independent surveys of the same site during the same season was smaller than the observed temporal turnover between two surveys from different years.
 Second, spatial variation of turnover showed patterns that can hardly be explained by methodological issues.
-For instance species turnover varied along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition.
-Taken together, our results adds to the increasing evidence that contemporary communities are often relatively stability at community level, but this apparent stability is often accompanied by marked turnover of species [@Vellend2013, @Dornelas2014, @Hillebrand2018].
+For instance species turnover varied along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition (Table \@ref(tab:turnovertab)).
+Taken together, our results adds to the increasing evidence that contemporary communities are often relatively stable at community level, but this apparent stability is often accompanied by marked turnover of species [@Vellend2013; @Dornelas2014; @Hillebrand2018].
 
-Species communities are shaped by a range of factors, including deterministic processes such as environmental filtering or competitive interactions [@Goetzenberger2012, @Janevcek2013].
+Species communities are shaped by a range of factors, including deterministic processes such as environmental filtering or competitive interactions [@Goetzenberger2012; @Janevcek2013].
 Such factors select for species with specific characteristics.
-Community assembly theories thus suggest that the factors that drive change in species communities could be inferred from comparing the charactereistics of disappearing or colonizing species with random expectation [@Chase2011].
+Community assembly theories thus suggest that the factors that drive the composition of species in a community can be inferred from comparing the charactereistics of the species in the community with random expectation [@Chase2011].
 We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site.
@@
 This is only about one tenth of the decrease in England, where Nitrogen deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
-This comparatively low change in Nitrogen deposition might explain why we observed no change in average nutrient value of communities.
-Additionally, other anthropogenic pressures such as climate change might overrule Nitrogen depositin effects on community structure.
+This comparatively low change in Nitrogen deposition in Swiss mountain hay meadows might explain why we observed no change in average nutrient value of communities.
+Additionally, other anthropogenic pressures such as climate change might have overruled  Nitrogen depositin effects on community structure.
 Particularly, we found that the species disappearing from the sites tend to have below average indicator values for temperature, while species that newly colonize sites have above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
@@
 
-Observational studies along a gradient of Nitrogen deposition often infer how the spatial variation in species richness is related to Nitrogen deposition.
+Observational studies along a gradient of Nitrogen deposition often infer how the spatial variation in species richness is related to Nitrogen deposition [@Stevens2010gradient; chapter 4 in @deVries2015].
 They assume that the spatial variation in species richness (or other metrics of community structure) arose because over time some areas lost more species than others because they chronically experienced higher Nitrogen deposition.
 Although there is evidence supporting the use of such a 'space for time substitution' for detecting the effects of Nitrogen deposition on plant diversity [@Stevens2010], they can not replace studies that relate temporal patterns in species with Nitrogen deposition [@deSchrijver2011].
-There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of N-deposition [@Clark2008, @Storkey2015].
+There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of N-deposition in extant communities [@Clark2008; @Storkey2015].
 In an recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the Nitrogen deposition gradient at which species richnes of oligotrophic species richness start to decrease with increasing Nitrogen deposition [@Roth2017].
@@
 In the current study we estimated the rate of Nitrogen deposition at which local survival probability or colonization probability was equal for oligotrophic and eutrophic species.
-This rate could also be interpreted as a critical load and revealed very similar results as the spatial approach.
+This rate could also be interpreted as a critical load.
+With this aproach we obtained very similar results as the spatial approach.
 Our results are thus validating the space for time approach.
@@
 Hower, the results of Fig.
-\@ref(fig:) also shows that the results of spatial comparisons must be interpreted carefully.
+\@ref(fig:figconsequences) also shows that the results of spatial comparisons must be interpreted carefully.
 When we used the same model to infer how species richness is decreasing along the Nitrogen deposition gradient at differnt time points, except for the first two study periods the decrease became steeper (more negative) over time.
@@
 However, we then realized that species turnover was highest in low deposition sites (Tab.
-\@ref(tab:turnovertab)) where colonizationg species were species with below average indicator values for Nutrients.
-It seems that the decrease in Nitrogen deposition resulted in higher numbers of oligotrophic species at sites with low deposition, thus leading to a steeper decline in oligotrophic species richness from sites with low to sites with high Nitrogen deposition.
+\@ref(tab:turnovertab)) where the colonizating species were species with below average indicator values for Nutrients.
+It seems that the decrease in Nitrogen deposition resulted in oligotrophic species replacing eutrophic species especially at sites with low deposition, thus leading to a steeper decline in oligotrophic species richness from sites with low to sites with high Nitrogen deposition.
+The increasingly steeper decline over time in the eutrophic species richness along the spatial Nitrogen deposition gradient is thus a sign of the recovery of plant communities in low deposition sites rather than a sign that the Nitrogen deposition effects became stronger over time.
 
 # Conclusions {-}
-Our study adds to the understanding of contemporary biodiversity change that is currently not not well understood (Magurran et al 2018.pdf).
-Our results support the notion of Hillebrand_et_al-2018.pdf that biodiversity monitoring programmes that quantify species turnover will generate a far more reliable understanding of the biotic response to changing environments than programmes solely tracking species richness.
+Comparing the indicator values of colonizating and desappearing species with random expecation we found that oligotrphic species are currently more likeley to colonize mountain hay meadows than oligotrophic species, which might be the results of the recently decreasing atmospeheric Nitrogen deposition.
+However, our results also caution that re-transfering mountain hay maeadows to oligotrophic species communities after high nitrogen depostion might take much longer than transfering a and oligotrophic into an eutrophic community.
+This is because eutrophic species have high local survival probabilities also when Nitrogen deposition becomes lower again.
+All these results were, however, hidden when only inferring the temporal trends of species communities.
+Our study thus adds to the understanding of contemporary biodiversity change that is currently not well understood [@Magurran2018], and it supports the notion of @Hillebrand2018 that inferring species turnover will generate a far more reliable understanding of the biotic response to changing environments than solely tracking community composition.
 

### 2018-06-18 Manuscript.Rmd
Referenzen in Einleitung eingefügt.

@@
 We infered different drivers of community change (i.e.
-Nitrogen deposition, climate warming, land-use change) in Swiss mountain hay meadows.
-The data were obtained from the Swiss biodiversity monitoring.
+Nitrogen deposition, climate warming, land-use change) in Swiss mountain hay meadows using data from the Swiss biodiversity monitoring.
 
@@
 # Introduction {-}
-Nitrogen deposition is one of the major threats to biodiversity (Sala 2000, Bobbink 2010, Menier et al.
-2016).
-Especially, there is strong evidence for the reduction in diversity of species-rich grasslands in Europe (Stevens et al.
-2006, Maskell et al.
-2010, Duprè et al.
-2010).
+Nitrogen deposition is the entry of reactive nitrogen compounds into water and soil via deposition through the air.
+Toghether with land-use and climate change, Nitrogen deposition is one of the major threats to biodiversity [@Sala2000; @Bobbink2010].
+Since Nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low N availability [@Vitousek1997], the addition of Nitrogen strongly alter these communities [@@Sala2000].
+While there is strong evidence for the reduction in diversity of species-rich grasslands in Europe due to increased Nitrogen deposition [@Stevens2006; @Maskell2010; @Dupre2010], mountain grasslands have received less attention [@Humbert2016].
 
-Fertilizing experiments show that plant biodiversity in mountain grasslands is negatively affected by nitrogen addition, too (Humbert et al. 2016).
-
-Since the late 1980s measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe (Tørseth et al., 2012, Maas and Grennfelt, 2016) [Tørseth et al.,: Atmos.
-Chem.
-Phys., 12, 5447–5481, 2012].
+Since the late 1980s measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe [@Torseth2012](Maas and Grennfelt, 2016).
 However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities.

### 2018-06-18 Manuscript.Rmd
Einleitung ergänzt, Referenzen eingefügt.

@@
 ---
-title: Species turnover reveal hidden effects of decreasing Nitrogen deposition in mountain hay meadows
+title: Species turnover reveals hidden effects of decreasing Nitrogen deposition in mountain hay meadows
 preprint: false
@@
     affiliation: 4    
+  - name: Valentin Amrhein
+    affiliation: 1    
 affiliation:
@@
 Toghether with land-use and climate change, Nitrogen deposition is one of the major threats to biodiversity [@Sala2000; @Bobbink2010].
-Since Nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low N availability [@Vitousek1997], the addition of Nitrogen strongly alter these communities [@@Sala2000].
+Since Nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low N availability [@Vitousek1997], the addition of Nitrogen strongly alter these communities [@Sala2000].
 While there is strong evidence for the reduction in diversity of species-rich grasslands in Europe due to increased Nitrogen deposition [@Stevens2006; @Maskell2010; @Dupre2010], mountain grasslands have received less attention [@Humbert2016].
 
-Since the late 1980s measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe [@Torseth2012](Maas and Grennfelt, 2016).
-However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities.
+Since the late 1980s measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe [@Torseth2012; @Fowler2007].
+However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities (>>REF EINFÜGEN<<).
 
-Quantifying the potential recovery of biodiversity in response to reducing air pollution requires an alternative to the space-for-time approach, ideally monitoring long-term community dynamics on permanent plots (Duprã, C.
-et al.
-Changes in species richness and composition in European acidic grasslands over the past 70 years: the contribution of cumulative atmospheric nitrogen deposition.
-Glob.
-Change Biol.
-16, 344–357, (2010).
+>> Kapitel zu konstante Communities und viel Turnover einfügen und dann erklären, dass wir eine Idee von Community assembly verwendet haben.
 
-One useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both immigration and extinction, often in a closed range of values (Hillebrand et al-2018.pdf).
-
-Observation data of vascular plants from field recordings shows negative correlations of Nitrogen deposition with species richness and species composition in mountain hay meadows (Roth et al. 2013).
+One useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both colonization and local extinction [@Hillebrand2018].
 
@@
 With extensive cultivation, permanent grassland has a very high biodiversity.
-This applies in particular to the meadows of the Swiss mountain region (Schlup et al.
-2013).
-In the alpine region, meadows with a high plant diversity are of great importance agronomically (Leiber et al.
+This applies in particular to the meadows in the alpine region, where meadows with a high plant diversity are of great importance agronomically (Leiber et al.
 2005), for the image of the products and for the conservation of biodiversity (Hohl, 2006).
+Observation data of vascular plants from field recordings shows negative correlations of Nitrogen deposition with species richness and species composition in mountain hay meadows [@Roth2013].
+However, between 1990 and 2010 NOx emissions in Switzerland decreased by 46% and NH3 emissions by 14% (>>REF EINFÜGEN: Maas and Grennfelt, 2016<<), suggesting that mountain hay meadows should have recoverd.
 
-NOx emissions in Switzerland decreased by 46% between 1990 and 2010 and NH3 emissions by 14% (Maas and Grennfelt, 2016)
+However, mountain regions are threatened also by other global change drivers.
+In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35&degC per decade, which is about 1.6 times the northern hemispheric warm- ing rate (Ceppi et al.
+2010).
+In autumn and early winter, the warming rate depended on altitude with strongest trends at low altitudes, whereas in spring slightly stronger trends are found at altitudes close to the snow line.
+This climate warming is likely to directly affect or interact with Nitrogen deposition to drive plant community changes [@Humbert2016].
+At the landscape scale, changes of plant communities were found that could be attributed to climate warming [@Roth2014].
+Particularly, the shift of plant communities at mountain summits seems to be the restult of recent climate warming interacting with airborne nitrogen deposition [@Rumpf2018].
+Furthermore, traditional management regimes are currently changing wich also has major impacts on grassland biodiversity in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
+In Switzerland  Management regimes of accessible mountainous areas are intensified, while poorly accessible mountainous areas are being abandoned [@Tasser2002; @Strebel2015].
 
-The meta-analysis of Humbert et al.
-(2016) showed management and abiotic factors to interact with nitrogen addition to drive plant community changes: The overall effects of N addition on plant species richness and diversity (Shannon index) were more pronounced in managed experimental sites with warmer summers.
-Management regimes of accessible mountainous areas are intensified, while poorly accessible mountainous areas are being abandoned (Tasser and Tappeiner, 2002; Graf et al., 2014, Strebel and Bühler, 2015).
-This change in traditional management regimes is considered as a major driver that endangers grassland biodiversity in mountainous regions of Europe (Niedrist et al., 2009; Homburger and Hofer, 2012).
+Here, we used the data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions:
 
-On European mountain summits, plants shifted upwards in the last decades (Rumpf et al.
-2018).
-The authors conclude that «These results suggest that recent climate warming interacted with airborne nitrogen deposition in driving the observed dynamics.
-As changes increased more for species from lower elevations and eutrophic plant species showed the strongest increase in abundance, these findings might be relevant for mountain hay meadows, too.
+>>1.  community measures , 2. Compare to turnover and 3. compare incicator values of species that newly colonized or disappeard (local extinction) from sites with random expectation.
 

### 2018-06-18 Manuscript.Rmd
Einleitung angepasst.

@@
 Nitrogen deposition is the entry of reactive nitrogen compounds into water and soil via deposition through the air.
-Toghether with land-use and climate change, Nitrogen deposition is one of the major threats to biodiversity [@Sala2000; @Bobbink2010].
-Since Nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low N availability [@Vitousek1997], the addition of Nitrogen strongly alter these communities [@Sala2000].
-While there is strong evidence for the reduction in diversity of species-rich grasslands in Europe due to increased Nitrogen deposition [@Stevens2006; @Maskell2010; @Dupre2010], mountain grasslands have received less attention [@Humbert2016].
+Since Nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low Nitrogen availability [@Vitousek1997], the addition of Nitrogen is likely to alter these communities.
+Indeed, Nitrogen deposition toghether with land-use and climate change is one of the major threats to biodiversity [@Sala2000; @Bobbink2010].
+While there is strong evidence for the reduction in diversity of species-rich grasslands due to increased Nitrogen deposition [@Stevens2006; @Dupre2010; @Maskell2010; @Wesche2012], mountain grasslands have received less attention [@Humbert2016].
 
 Since the late 1980s measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe [@Torseth2012; @Fowler2007].
-However, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities (>>REF EINFÜGEN<<).
+However, there are a number of factors that may prevent the recovery of plant communities that sufered from increased Nitrogen deposition.
+First, even if Nitrogen deposition is reduced large areas might still be above the critical load (>>REF<<).
+Second, there is a potential that communities reach an alternative stable state after decades of increased Nitrogen deposition and these species are unlikely to disappear again after the reduction of Nitrogen depositon [@Stevens2016].
+Therefore, it is an open question whether and how fast the reduction in Nitrogen deposition rates will lead to the recovery of extant plant communities.
 
->> Kapitel zu konstante Communities und viel Turnover einfügen und dann erklären, dass wir eine Idee von Community assembly verwendet haben.
-
-One useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both colonization and local extinction [@Hillebrand2018].
+Recovry of extant plant communities from high Nitrogen depositon would imply that the state of communities measured at different time points is improving over time (i.e.
+improving biodiversity endpoints sensu @Rowe2017).
+Species richness is an easy to measure and communicate biodiversity entpoint and it is often negatively correlated with Nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
+Other metrics that are potentially more usful to reflect favourable changes can be derived from the traits of the species in a community.
+In Europe, environmental preference has often been expressed using incicator values assigned to each plant species [@Ellenberg1992].
+Examples of such metrics would be the number of oligotrophic species or the average indicator value of the species in a community [@Roth2013; @Rowe2017].
+However, the lack of a temporal trend in such biodiversity endpoints -- particularly in species richness -- does not necessarily mean that species composition remains unchanged.
+This is because immigration and extinction might be equally frequent and may cancel each other out [@Hillebrand2018].
+Thus an useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both colonization and local extinction [@Hillebrand2018], especially if colonization and local extinction is compared to random expectation [@Chase2011].
 
@@
 Observation data of vascular plants from field recordings shows negative correlations of Nitrogen deposition with species richness and species composition in mountain hay meadows [@Roth2013].
-However, between 1990 and 2010 NOx emissions in Switzerland decreased by 46% and NH3 emissions by 14% (>>REF EINFÜGEN: Maas and Grennfelt, 2016<<), suggesting that mountain hay meadows should have recoverd.
+Between 1990 and 2010 NOx emissions in Switzerland decreased by 46% and NH3 emissions by 14% (>>REF EINFÜGEN: Maas and Grennfelt, 2016<<), suggesting that Nitrogen deposition has decreased also in mountain hay meadows should and we expected plant communities to respond accordingly.
 
-However, mountain regions are threatened also by other global change drivers.
-In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35&degC per decade, which is about 1.6 times the northern hemispheric warm- ing rate (Ceppi et al.
-2010).
-In autumn and early winter, the warming rate depended on altitude with strongest trends at low altitudes, whereas in spring slightly stronger trends are found at altitudes close to the snow line.
+However, Swiss mountain ecosystems are threatened also by other global change drivers.
+In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35$^{\circ}$C per decade, which is about 1.6 times the northern hemispheric warm- ing rate [@Ceppi2012].
 This climate warming is likely to directly affect or interact with Nitrogen deposition to drive plant community changes [@Humbert2016].
-At the landscape scale, changes of plant communities were found that could be attributed to climate warming [@Roth2014].
-Particularly, the shift of plant communities at mountain summits seems to be the restult of recent climate warming interacting with airborne nitrogen deposition [@Rumpf2018].
-Furthermore, traditional management regimes are currently changing wich also has major impacts on grassland biodiversity in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
+Indeed, in an earlier study we found that at a landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014] and particularly the shift of plant communities at mountain summits are the restult of recent climate warming interacting with airborne nitrogen deposition [@Rumpf2018].
+Furthermore, traditional management regimes are currently changing wich also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
 In Switzerland  Management regimes of accessible mountainous areas are intensified, while poorly accessible mountainous areas are being abandoned [@Tasser2002; @Strebel2015].
 
-Here, we used the data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions:
-
->>1.  community measures , 2. Compare to turnover and 3. compare incicator values of species that newly colonized or disappeard (local extinction) from sites with random expectation.
+Here, we used the data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, nitrogen deposition or land-use intensity change over the last 15 years, (2) Was species turnover correlated with the avearage temperature, precipitation, Nitrogen deposition or stepness (we expect step areas to be less intensively managed), and (3) did species that newly colonized or disappeard (local extinction) from sites different from random expectation according to their incicator values for tempearture, humidity, nutrients or light.
 
@@
 Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at high Nitrogen deposition.
-This suggests that it takes more time to replace eutrophic by oligotrophic species than replacing oligotrophic by eutrophic species.
+This suggests that mountain hay meadow communities can reach alternative stable states with eutrophic species that are unlikely to disappear again even if Nitrogen depositon is decreasing again [@Stevens2016].
 Taken toghether, these reasons might explain why the response of mountain hay medow communities responded stronger to to climate warming affects than to Nitrogen reduction, altough the reduction in Nitrogen resulted in above average colonization of oligotrophic species.
@@
 Although there is evidence supporting the use of such a 'space for time substitution' for detecting the effects of Nitrogen deposition on plant diversity [@Stevens2010], they can not replace studies that relate temporal patterns in species with Nitrogen deposition [@deSchrijver2011].
-There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of N-deposition in extant communities [@Clark2008; @Storkey2015].
+There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of N-deposition in extant communities [@Clark2008; @Storkey2015; @Stevens2016].
 In an recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the Nitrogen deposition gradient at which species richnes of oligotrophic species richness start to decrease with increasing Nitrogen deposition [@Roth2017].

### 2018-06-18 Manuscript.Rmd
Update Manuscript.Rmd

@@
     email: t.roth@unibas.ch
+  - name: Valentin Amrhein
+    affiliation: 1 
   - name: Lukas Kohli
@@
     affiliation: 3
+  - name: Reto Meuli
+    affiliation: 4  
   - name: Reto Meier
-    affiliation: 4    
-  - name: Valentin Amrhein
-    affiliation: 1    
+    affiliation: 5
+
 affiliation:
@@
   - code: 3
-    address: Meteotest,Fabrikstrasse 14, 3012 Bern, Switzerland 
+    address: Meteotest, Fabrikstrasse 14, 3012 Bern, Switzerland
  - code: 4
+    address: Agroscope, Swiss Soil Monitoring Network NABO, 8046 Zurich, Switzerland
+  - code: 5
     address: Federal Office for the Environment (FOEN), Air Pollution Control and Chemicals Division, 3003 Bern, Switzerland
@@
 However, there are a number of factors that may prevent the recovery of plant communities that sufered from increased Nitrogen deposition.
-First, even if Nitrogen deposition is reduced large areas might still be above the critical load (>>REF<<).
+First, even if Nitrogen deposition is reduced, large areas might still be above the critical load (>>@Slootweg2015; @RihmAchermann<<).
 Second, there is a potential that communities reach an alternative stable state after decades of increased Nitrogen deposition and these species are unlikely to disappear again after the reduction of Nitrogen depositon [@Stevens2016].
@@
 Species richness is an easy to measure and communicate biodiversity entpoint and it is often negatively correlated with Nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
-Other metrics that are potentially more usful to reflect favourable changes can be derived from the traits of the species in a community.
+Other metrics that are potentially more useful to reflect favourable changes can be derived from the traits of the species in a community.
 In Europe, environmental preference has often been expressed using incicator values assigned to each plant species [@Ellenberg1992].
@@
 Observation data of vascular plants from field recordings shows negative correlations of Nitrogen deposition with species richness and species composition in mountain hay meadows [@Roth2013].
-Between 1990 and 2010 NOx emissions in Switzerland decreased by 46% and NH3 emissions by 14% (>>REF EINFÜGEN: Maas and Grennfelt, 2016<<), suggesting that Nitrogen deposition has decreased also in mountain hay meadows should and we expected plant communities to respond accordingly.
+Between 1990 and 2010 NOx emissions in Switzerland decreased by 46% and NH3 emissions by 14% (>>REF EINFÜGEN: Maas and Grennfelt, 2016<<), suggesting that Nitrogen deposition has decreased also in mountain hay meadows and we expected plant communities to respond accordingly.
 
 However, Swiss mountain ecosystems are threatened also by other global change drivers.
-In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35$^{\circ}$C per decade, which is about 1.6 times the northern hemispheric warm- ing rate [@Ceppi2012].
+In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35$^{\circ}$C per decade, which is about 1.6 times the northern hemispheric warming rate [@Ceppi2012].
 This climate warming is likely to directly affect or interact with Nitrogen deposition to drive plant community changes [@Humbert2016].
@@
 Furthermore, traditional management regimes are currently changing wich also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
-In Switzerland  Management regimes of accessible mountainous areas are intensified, while poorly accessible mountainous areas are being abandoned [@Tasser2002; @Strebel2015].
+Management regimes of accessible mountainous areas are frequently intensified, while poorly accessible mountainous areas are being abandoned [@Tasser2002; @Strebel2015].
 
-Here, we used the data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, nitrogen deposition or land-use intensity change over the last 15 years, (2) Was species turnover correlated with the avearage temperature, precipitation, Nitrogen deposition or stepness (we expect step areas to be less intensively managed), and (3) did species that newly colonized or disappeard (local extinction) from sites different from random expectation according to their incicator values for tempearture, humidity, nutrients or light.
+Here, we used the data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, Nitrogen deposition or land-use intensity change over the last 15 years, (2) Was species turnover correlated with the avearage temperature, precipitation, Nitrogen deposition or stepness (we expect step areas to be less intensively managed), and (3) did species that newly colonized or disappeard (local extinction) from sites different from random expectation according to their incicator values for tempearture, humidity, nutrients or light.
 
@@
 Another key drivers for plants that is likely to be affected by climate change is precipitation [@Beier2012].
-We used the annual precipitation per site from the WordClim database [@Fick2017].Further, we estimated atmospheric N deposition for each site using a pragmatic approach described in @Rihm2001 that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models.
+We used the annual precipitation per site from the WordClim database [@Fick2017].
+Further, we estimated atmospheric Nitrogen deposition for each site using a pragmatic approach described in @Rihm2001 that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models.
 For details on the estimation of N deposition see @Rihm2001 and @Roth2017.
@@
 ```{r communitytrendscalc}
-res <- tibble(Measures = c("Species-richness", "Temperature value", "Huminity value", "Nutrients value", "Light value"),
+res <- tibble(Measures = c("Species-richness", "Temperature value", "Humidity value", "Nutrients value", "Light value"),
               'Period 1' = NA,
@@
 ## Potential effects of reduction in Nitrogen emmissions {-}
-```{r Ndepchange, fig.cap="(a) The nitrogen deposition in 2000 and (b) the change in Nitrogen deposition between 2000 and 2010 along the elevational gradient of the study sites."}
+```{r Ndepchange, fig.cap="(a) The nitrogen deposition in 2000 and (b) the change in Nitrogen deposition between 2000 and 2015 along the elevational gradient of the study sites."}
 p1 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = NTOT_2000)) +
@@
 
-```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (Ellenberg N = 2; red line) and eutrophic (Ellengerg N = 4) species along the N deposition gradient.
+```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (indicator value for nutrients <3; red line) and eutrophic (indicator value for nutrients >3) species along the Nitrogen deposition gradient.
 Given are means and 95\\%-Credible Intervals from logistic linear mixed models.
@@
   xlim(0,30) + theme(legend.position = c(0.75, 0.9)) + 
-  scale_color_manual(labels = c("Nutrients value = 2", "Nutrients value = 4"), 
+  scale_color_manual(labels = c("oligotrophic species", "eutrophic species"), 
                      values = tcol[c(2,4)])

### 2018-06-18 Manuscript.Rmd
Update Manuscript.Rmd

@@
 This climate warming is likely to directly affect or interact with Nitrogen deposition to drive plant community changes [@Humbert2016].
-Indeed, in an earlier study we found that at a landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014] and particularly the shift of plant communities at mountain summits are the restult of recent climate warming interacting with airborne nitrogen deposition [@Rumpf2018].
+Indeed, in an earlier study we found that at a landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014] and particularly the shift of plant communities at mountain summits are the restult of recent climate warming interacting with airborne nitrogen deposition [@Rumpf2018](>> Steinbauer<<).
 Furthermore, traditional management regimes are currently changing wich also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
@@
 
-We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all sampling sites in mountain hay meadows (EUNIS E2.3) For more details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
+We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all sampling sites in mountain hay meadows (EUNIS E2.3).
+For more details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
 We analysed the data from 2003 to 2017.
@@
 Further, we estimated atmospheric Nitrogen deposition for each site using a pragmatic approach described in @Rihm2001 that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models.
-For details on the estimation of N deposition see @Rihm2001 and @Roth2017.
+For details on the estimation of Nitrogen deposition see @Rihm2001 and @Roth2017.
 Finally we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed [@Strebel2015].
@@
 res <- tibble(
-  Gradient = c("Annual mean temperature", "Annual mean precipitation", "Nitrogen deposition", "Inclination"),
+  Gradient = c("Mean annual temperature", "Mean annual precipitation", "Nitrogen deposition", "Inclination"),
   Estimate = fixef(mod)[tparams],
@@
 In Fig.
-\@ref(fig:cl) we compare the colonization and local survival probability of oligotrophic (indicator value of nutrients = 2) and eutrophic (indicator value of nutrients = 4) species along the Nitrogen deposition gradient.
+\@ref(fig:cl) we compare the colonization and local survival probability of oligotrophic (indicator value of nutrients <3) and eutrophic (indicator value of nutrients > 3) species along the Nitrogen deposition gradient.
 Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for for oligotrophic and euthrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
@@
 Thus, the effect of disappearing and colonizating species on the average community value for nutrients partially cancel each other out.
-Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at high Nitrogen deposition.
-This suggests that mountain hay meadow communities can reach alternative stable states with eutrophic species that are unlikely to disappear again even if Nitrogen depositon is decreasing again [@Stevens2016].
+Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at higher Nitrogen deposition rates.
+This suggests that mountain hay meadow communities can reach alternative stable states with eutrophic species that are unlikely to disappear even if Nitrogen depositon is reduced [@Stevens2016].
 Taken toghether, these reasons might explain why the response of mountain hay medow communities responded stronger to to climate warming affects than to Nitrogen reduction, altough the reduction in Nitrogen resulted in above average colonization of oligotrophic species.
@@
 Our first interpretation was that the Nitrogen deposition effect became stronger over time.
-This contraticted our prediction that the nitrogen deposition should become weaker over time since Nitrogen deposition decreased.
+This contraticted our prediction that the Nitrogen deposition should become weaker over time since Nitrogen deposition decreased.
 However, we then realized that species turnover was highest in low deposition sites (Tab.

### 2018-06-18 Manuscript.Rmd
Update Manuscript.Rmd

@@
 However, there are a number of factors that may prevent the recovery of plant communities that sufered from increased Nitrogen deposition.
-First, even if Nitrogen deposition is reduced, large areas might still be above the critical load (>>@Slootweg2015; @RihmAchermann<<).
+First, even if Nitrogen deposition is reduced, large areas might still be above the critical load [@Slootweg2015; @RihmAchermann2016].
 Second, there is a potential that communities reach an alternative stable state after decades of increased Nitrogen deposition and these species are unlikely to disappear again after the reduction of Nitrogen depositon [@Stevens2016].
@@
 Observation data of vascular plants from field recordings shows negative correlations of Nitrogen deposition with species richness and species composition in mountain hay meadows [@Roth2013].
-Between 1990 and 2010 NOx emissions in Switzerland decreased by 46% and NH3 emissions by 14% (>>REF EINFÜGEN: Maas and Grennfelt, 2016<<), suggesting that Nitrogen deposition has decreased also in mountain hay meadows and we expected plant communities to respond accordingly.
+Between 1990 and 2010 NOx emissions in Switzerland decreased by 46% and NH3 emissions by 14% [@MaasGrennfelt2016], suggesting that Nitrogen deposition has decreased also in mountain hay meadows and we expected plant communities to respond accordingly.
 
@@
 Further, we estimated atmospheric Nitrogen deposition for each site using a pragmatic approach described in @Rihm2001 that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models.
-For details on the estimation of Nitrogen deposition see @Rihm2001 and @Roth2017.
+For details on the estimation of Nitrogen deposition see @RihmAchermann2016 and @Roth2017.
 Finally we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed [@Strebel2015].

### 2018-06-18 Manuscript.Rmd
Update Manuscript.Rmd
Literatur ergänzt 
Abb. 2 oligo- und eutrophic species (statt N-Wert)

@@
 
-The average $\pm$ SD percentage of species that differed between replicated surveys was `r mean(100 * turnover_double$turnover)` $\pm$ `r sd(100 * turnover_double$turnover)`.
+The average $\pm$ SD percentage of species that differed between replicated surveys was `r mean(100 * turnover_double$turnover)` $\pm$ `r sd(100 * turnover_double$turnover)`%.
 This pseud-turnover was lower than the observed temporal turnover: the average percentage of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.

### 2018-06-25 Manuscript.Rmd
Corrections of typos.

@@
 The recent introduction of cleaner technologies in Switzerland has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition.
-We infered different drivers of community change (i.e.
+We inferred different drivers of community change (i.e.
 Nitrogen deposition, climate warming, land-use change) in Swiss mountain hay meadows using data from the Swiss biodiversity monitoring.
+While oligotrophic species were more likely to colonize compared to random expectation, we found only weak shifts in plant community structure.
+Particularly, the average nutrient value of plant communities remained stable over time.
+Apparently, the community effects of the replacement of eutrophic species with oligotrophic species was outweighed by climate warming.
+Our results add to the increasing evidence that inferring species turnover will generate a far more reliable understanding of the biotic response to changing environments than solely tracking average community composition.
 
@@
 Since Nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low Nitrogen availability [@Vitousek1997], the addition of Nitrogen is likely to alter these communities.
-Indeed, Nitrogen deposition toghether with land-use and climate change is one of the major threats to biodiversity [@Sala2000; @Bobbink2010].
+Indeed, Nitrogen deposition together with land-use and climate change is one of the major threats to biodiversity [@Sala2000; @Bobbink2010].
 While there is strong evidence for the reduction in diversity of species-rich grasslands due to increased Nitrogen deposition [@Stevens2006; @Dupre2010; @Maskell2010; @Wesche2012], mountain grasslands have received less attention [@Humbert2016].
@@
 Since the late 1980s measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe [@Torseth2012; @Fowler2007].
-However, there are a number of factors that may prevent the recovery of plant communities that sufered from increased Nitrogen deposition.
-First, even if Nitrogen deposition is reduced, large areas might still be above the critical load [@Slootweg2015; @RihmAchermann2016].
-Second, there is a potential that communities reach an alternative stable state after decades of increased Nitrogen deposition and these species are unlikely to disappear again after the reduction of Nitrogen depositon [@Stevens2016].
+However, there are a number of factors that may prevent the recovery of plant communities that suffered from increased Nitrogen deposition.
+First, even if Nitrogen deposition is reduced, large areas might still be above the critical load [@Slootweeg2015; @RihmAchermann2016].
+Second, there is a potential that communities reach an alternative stable state after decades of increased Nitrogen deposition and these species are unlikely to disappear again after the reduction of Nitrogen deposition [@Stevens2016].
 Therefore, it is an open question whether and how fast the reduction in Nitrogen deposition rates will lead to the recovery of extant plant communities.
 
-Recovry of extant plant communities from high Nitrogen depositon would imply that the state of communities measured at different time points is improving over time (i.e.
+Recovery of extant plant communities from high Nitrogen deposition would imply that the state of communities measured at different time points is improving over time (i.e.
 improving biodiversity endpoints sensu @Rowe2017).
-Species richness is an easy to measure and communicate biodiversity entpoint and it is often negatively correlated with Nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
+Species richness is an easy to measure and communicate biodiversity endpoint and it is often negatively correlated with Nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
 Other metrics that are potentially more useful to reflect favourable changes can be derived from the traits of the species in a community.
-In Europe, environmental preference has often been expressed using incicator values assigned to each plant species [@Ellenberg1992].
+In Europe, environmental preference has often been expressed using indicator values assigned to each plant species [@Ellenberg1992].
 Examples of such metrics would be the number of oligotrophic species or the average indicator value of the species in a community [@Roth2013; @Rowe2017].
@@
 This is because immigration and extinction might be equally frequent and may cancel each other out [@Hillebrand2018].
-Thus an useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both colonization and local extinction [@Hillebrand2018], especially if colonization and local extinction is compared to random expectation [@Chase2011].
+Thus a useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both colonization and local extinction [@Hillebrand2018], especially if colonization and local extinction is compared to random expectation [@Chase2011].
 
-In Switzerland, grassland accounts for 70 per cent of the agricultural land.
+In Switzerland, grassland accounts for 70% of the agricultural land.
 With extensive cultivation, permanent grassland has a very high biodiversity.
-This applies in particular to the meadows in the alpine region, where meadows with a high plant diversity are of great importance agronomically (Leiber et al.
-2005), for the image of the products and for the conservation of biodiversity (Hohl, 2006).
-Observation data of vascular plants from field recordings shows negative correlations of Nitrogen deposition with species richness and species composition in mountain hay meadows [@Roth2013].
-Between 1990 and 2010 NOx emissions in Switzerland decreased by 46% and NH3 emissions by 14% [@MaasGrennfelt2016], suggesting that Nitrogen deposition has decreased also in mountain hay meadows and we expected plant communities to respond accordingly.
+This applies in particular to the meadows in the alpine region, where meadows with a high plant diversity are also of importance agronomically [@Leiber2006].
+In mountain hay meadows, the spatial variation of species richness in vascular plants is negative correlated with Nitrogen deposition [@Roth2013], suggesting that mountain grasslands suffered from increased Nitrogen deposition.
+However, between 1990 and 2010 NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% [@MaasGrennfelt2016] and we expect plant communities to recover.
 
@@
 This climate warming is likely to directly affect or interact with Nitrogen deposition to drive plant community changes [@Humbert2016].
-Indeed, in an earlier study we found that at a landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014] and particularly the shift of plant communities at mountain summits are the restult of recent climate warming interacting with airborne nitrogen deposition [@Rumpf2018](>> Steinbauer<<).
-Furthermore, traditional management regimes are currently changing wich also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
+Indeed, in an earlier study we found that at the landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014] and particularly the shift of plant communities at mountain summits are the result of recent climate warming interacting with airborne nitrogen deposition [@Steinbauer2018].
+Furthermore, traditional management regimes are currently changing, which also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
 Management regimes of accessible mountainous areas are frequently intensified, while poorly accessible mountainous areas are being abandoned [@Tasser2002; @Strebel2015].
 
-Here, we used the data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, Nitrogen deposition or land-use intensity change over the last 15 years, (2) Was species turnover correlated with the avearage temperature, precipitation, Nitrogen deposition or stepness (we expect step areas to be less intensively managed), and (3) did species that newly colonized or disappeard (local extinction) from sites different from random expectation according to their incicator values for tempearture, humidity, nutrients or light.
+Here, we used the data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, Nitrogen deposition or land-use intensity change over the last 15 years, (2) was species turnover correlated with the average temperature, precipitation, Nitrogen deposition or inclination (we expect steep areas to be less intensively managed), and (3) did species that newly colonized or disappeared (local extinction) from sites differ from random expectation according to their indicator values for temperature, humidity, nutrients or light.
 
@@
 
-## Monitoring data and community measures{-}
+## Monitoring data and community measures {-}
 We analysed the presence/absence of vascular plants sampled within the scope of Switzerland's Biodiversity Monitoring (BDM) programme that was launched in 2001 to monitor Switzerland's biodiversity and to comply with the Convention on Biological Diversity of Rio de Janeiro [@Weber2004].
@@
 During each visit all the vascular plant species detected on the plot were recorded.
+For details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
+
 After the sampling of the plant data the botanists also assigned a habitat type to each sampling site according to the classification system developed for Switzerland [@Delarze2008].
-
 We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all sampling sites in mountain hay meadows (EUNIS E2.3).
-For more details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
 We analysed the data from 2003 to 2017.
@@
 
-For each survey (a survey consists of the two visits per season) we calculated the following community measures: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for  temperature (2), humidity (3), nutrients (4) and light (5).
+For each survey (a survey consists of the two visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for (2) temperature, (3) humidity, (4) nutrients and (5) light.
 The indicator values were obtained from the recalibrated indicator values for the Swiss Flora [@Landolt2010].
-Additionally to these five community measures that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
+Additionally, to these five biodiversity endpoints that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
 species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points to describe the community change between two time points.
 
-To test data quality, independent replicate surveys are routinely performed within the BDM programm by botansists who were not involved in the regular BDM surveys.
+To test data quality, independent replicate surveys are routinely performed within the BDM program by botanists who were not involved in the regular BDM surveys.
 The regular surveyors did not know if and which sites were replicated [@Plattner2004].
@@
 To better disentangle the importance of these mechanisms we ordered the sites along four main environmental gradients.
-First we expected communities to response to climate warming and the response might be different depending along the temperature gradient [@Roth2014].
-To describe the temperature gradient we used the mean annual temperature per site from the WordClim database [@Fick2017].
+First, we expected communities to response to climate warming and the response might be different along the temperature gradient [@Roth2014].
+To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
 Another key drivers for plants that is likely to be affected by climate change is precipitation [@Beier2012].
@@
 Further, we estimated atmospheric Nitrogen deposition for each site using a pragmatic approach described in @Rihm2001 that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models.
-For details on the estimation of Nitrogen deposition see @RihmAchermann2016 and @Roth2017.
-Finally we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed [@Strebel2015].
+For details on the estimation of Nitrogen deposition see also @RihmAchermann2016 and @Roth2017.
+Finally, we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed [@Strebel2015].
 
 ## Statistical analyses {-}
-To estimate the linear trend over time for each of the five community measures we applied linear mixed models (LMM) with normal distribution except for species-richness with Poisson distribution and the logarithm as link function.
+To estimate the linear trend over time for each of the five biodiversity endpoints we applied linear mixed models (LMM) with normal distribution except for species-richness with Poisson distribution.
 We specified site specific trends with the assumption that the between-site differences in intercepts and slopes can be described with normal distributions [i.e.
-a random intercept random slope model, @German2006].
+a random intercept random slope model, @Gelman2006].
 Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
@@
 Then we calculated the differences of the CM minus the average of the random-CMs to obtain a standardized measure (standardized-CM) of how different the disappearing species were from random expectation.
-A standardized-CM < 0 would suggest that the indicator value of colonizing species were lower than the average species that were occurring at this site.
+A standardized-CM < 0 would suggest that the indicator values of colonizing species were lower than the indicator values of average species that were occurring at this site.
 We applied this method for both colonizing and disappearing species and for the indicator values for temperature, humidity, nutrients and light (see Appendix A).
@@
 We used logistic-LMM to infer whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this changing depended on the species indicator value for nutrients.
-To analyse the colonization probability we selected all species that were not observed during the first survey, and asked if they were observed (then Occ = 1) or not observed (then Occ = 0) during the second survey.
-The same was also done for all species that were not observed during the second survey.
+To analyse the colonization probability, we selected all species that were not observed during the first survey and asked if they were observed (then Occ = 1) or not observed (then Occ = 0) during the second survey.
+The same was also done for all species that were not observed during the second survey if they were observed or not observed (then Occ = 0) during the third survey.
 The variable 'occ' was then used as dependent variable in the logistic-LMM.
 As predictor variables the model contained the Nitrogen deposition of the site, the indicator value for nutrients of the species and the interaction of these two variables.
-Additionally  species-ID and site-ID were included as random effects.
-The same logistic_LMM was also used to infer local survival probability.
+Additionally, species-ID and site-ID were included as random effects.
+The same logistic LMM was also used to infer local survival probability.
 In that case, however, we selected all species that were recorded during the first or second survey and the variable 'occ' then indicated whether or not the species was also observed during the next survey.
@@
 kable(res, align = c("l", rep("r", 7)), digits = 2,  booktabs = T, col.names = c("Measures", "Period 1", "Period 2", "Period 3", "Trend", "5%", "95%", "Prob.
-for trend"), caption = "Average measures of community structure for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
+for trend"), caption = "Average measures of the biodiversity endpoints for the three sampling periods (period 1: 2003-2007; period 2: 2008-2012; period: 2013-2017).
 The temporal trends are given as change per 10 years and were estimated from linear mixed models with normal distribution (except for species richness with Poisson distribution and a log-link function).
@@
 
-The five measures of plant community structure suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of species-richness and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
+The five measures of plant community structure (i.e.
+biodiversity endpoints) suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of species-richness and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
 Except for average indicator value for temperature, the 90% credible-interval of the temporal trend contained zero.
@@
 kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Ellenberg value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average indicator value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
-Shown are the results from linear model with the difference between disappeard/colonized species and random species as dependend variable and the sitemeasure (gradient) as predictor variable.") %>%
+Shown are the results from linear model with the difference between disappeared/colonized species and random species as dependent variable and the site-measure (gradient) as predictor variable.") %>%
   kable_styling() %>% 
@@
 High species turnover at a site is the result of species that disappeared from the site and species that newly colonized the site.
-To better understand the factors that drive these changes we are particularly interested whether the species that disappeared or colonized the sites differed in indicator values compared to what would be expected if the same number of species randomly disappeared or colonized the sites (i.e.
-random disappearance and random colonization) and whether there is a change along the gradients.
+To better understand the factors that drive these changes we were particularly interested whether the species that disappeared or colonized the sites differed in indicator values compared to what would be expected if the same number of species randomly disappeared or colonized the sites (i.e.
+random disappearance and random colonization) and whether there is a spatial variation along the gradients.
 It seems that the indicator values of newly colonizing species differed more from random colonization than the indicator values of disappearing species (Table \@ref(tab:difffromrandomtab)).
 For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower indicator value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing Nitrogen deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
-Thus at high Nitrogen deposition colonizing species did not differ from random species (see Figure 3b in Appendix A).
+Thus, at high Nitrogen deposition colonizing species did not differ from random species (see Figure 3b in Appendix A).
 
@@
 This was not the case for the indicator value for nutrients: species with lower nutrients values tended to be more likely to disappear from as well as to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
-See also Appendix A were we present detailed results for the comparison between colonizing or disappearing species with randomly selected species.
+See also Appendix A where we present detailed results for the comparison between colonizing or disappearing species with randomly selected species.
 
@@
 Given are means and 95\\%-Credible Intervals from logistic linear mixed models.
-The vertical lines indicat the deposition rate with equal colonization or survival probabilities for oligotrophic and eutrophic species with the solid line indicating the median and the dashed lines the 5\\% and 95\\% quantiels of the margional posterior distribution."}
+The vertical lines indicate the deposition rate with equal colonization or survival probabilities for oligotrophic and eutrophic species with the solid line indicating the median and the dashed lines the 5\\% and 95\\% quantiles of the marginal posterior distribution."}
 X <- 1:30
@@
 \@ref(fig:cl) we compare the colonization and local survival probability of oligotrophic (indicator value of nutrients <3) and eutrophic (indicator value of nutrients > 3) species along the Nitrogen deposition gradient.
-Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for for oligotrophic and euthrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
+Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for oligotrophic and eutrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
 In only `r mean((sites$NTOT * 10 + 10) < 12.5)`% of the sites the deposition rate was below 12.5 kg N ha$^{-1}$ yr$^{-1}$ where the replacement of eutrophic with oligotrophic species is likely.
 
-While we could not detect a consistent decrease in average indicator value for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial variation of species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between Nitrogen deposition and species richness.
-Indeed, if we apply at different time points a similar model as in @Roth2013 to infer the effects of Nitrogen deposition on the spatial variation of species richness, the resulting effect size (i.e.
-the slope) becomes more negative over time (Fig.
+While we could not detect a consistent decrease in average indicator value for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial variation of oligotrophic species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between Nitrogen deposition and oligotrophic species richness.
+Indeed, if we apply at different time points a similar model as in @Roth2013 to infer the effects of Nitrogen deposition on the spatial variation of oligotrophic species richness, the resulting effect size (i.e.
+the slope) became more negative over time (Fig.
 \@ref(fig:figconsequences)).
 
-```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of Nitrogen deposition on total species richness estimated from applying the Poisson-GLM with species richness as dependend variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from one five-year interval.
+```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of Nitrogen deposition on oligotrophic species richness estimated from applying the Poisson-GLM with species richness as dependent variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from one five-year interval.
 Note that within every five-year interval all plots were sampled once."}
@@
 Although Nitrogen deposition considerably declined between 2000 and 2015 (Fig.
-\@ref(fig:Ndepchange)), we could detect only weak shifts in plant community structure during the same time period (Table \@ref(tab:communitytrendstab)).
-While the slight increase in average temperature value suggests that plant communities adopted to increasing temperatures, the constant aveage nutrients value suggests that the decrease in Nitrogen deposition did not yet affected plant communities.
+\@ref(fig:Ndepchange)), we could detect only weak shifts in plant community structure (i.e.
+biodiversity endpoints sensu @Rowe2017) during the same time period (Table \@ref(tab:communitytrendstab)).
+While the slight increase in average temperature value suggests that plant communities adopted to increasing temperatures, the constant average nutrients value suggests that the decrease in Nitrogen deposition did not yet affect plant communities.
 However, this apparent stability in community structure was accompanied by marked temporal turnover in species identities.
-It seems unlikely that this temporal turnover can entirly be explained by methodological issues such as overlooked species.
+It seems unlikely that this temporal turnover can entirely be explained by methodological issues such as overlooked species.
 First, pseudo-turnover of species entities in independent surveys of the same site during the same season was smaller than the observed temporal turnover between two surveys from different years.
@@
 For instance species turnover varied along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition (Table \@ref(tab:turnovertab)).
-Taken together, our results adds to the increasing evidence that contemporary communities are often relatively stable at community level, but this apparent stability is often accompanied by marked turnover of species [@Vellend2013; @Dornelas2014; @Hillebrand2018].
+Taken together, our results add to the increasing evidence that contemporary communities are often relatively stable at community level, but this apparent stability is often accompanied by marked turnover of species [@Vellend2013; @Dornelas2014; @Hillebrand2018].
 
@@
 Such factors select for species with specific characteristics.
-Community assembly theories thus suggest that the factors that drive the composition of species in a community can be inferred from comparing the charactereistics of the species in the community with random expectation [@Chase2011].
-We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site.
+Community assembly theories thus suggest that the factors that drive the composition of species in a community can be inferred from comparing the characteristics of the species in the community with random expectation [@Chase2011].
+We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site (Appendix A).
 We found that the nutrient values of colonization species showed largest deviations from random expectation suggesting that Nitrogen deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
@@
 This comparatively low change in Nitrogen deposition in Swiss mountain hay meadows might explain why we observed no change in average nutrient value of communities.
-Additionally, other anthropogenic pressures such as climate change might have overruled  Nitrogen depositin effects on community structure.
+Additionally, other anthropogenic pressures such as climate change might have overruled Nitrogen depositing effects on community structure.
 Particularly, we found that the species disappearing from the sites tend to have below average indicator values for temperature, while species that newly colonize sites have above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
-Thus, the effect of disappearing and the effect of colonizating species on the community mean for temperature is additive resulting in increasing average temperature values (Tab.
+Thus, the effect of disappearing and the effect of colonizing species on the community mean for temperature is additive resulting in increasing average temperature values (Tab.
 \@ref(tab:communitytrendstab)).
 This is in contrast to how Nitrogen deposition is affecting disappearing and colonization of species: It seems that both species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
-Thus, the effect of disappearing and colonizating species on the average community value for nutrients partially cancel each other out.
+Thus, the effect of disappearing and colonizing  species on the average community value for nutrients partially cancel each other out.
 Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at higher Nitrogen deposition rates.
-This suggests that mountain hay meadow communities can reach alternative stable states with eutrophic species that are unlikely to disappear even if Nitrogen depositon is reduced [@Stevens2016].
-Taken toghether, these reasons might explain why the response of mountain hay medow communities responded stronger to to climate warming affects than to Nitrogen reduction, altough the reduction in Nitrogen resulted in above average colonization of oligotrophic species.
+This suggests that mountain hay meadow communities can reach alternative stable states with eutrophic species that are unlikely to disappear even if Nitrogen deposition is reduced [@Stevens2016].
+Taken together, these reasons might explain why the response of mountain hay meadow communities responded stronger to climate warming than to Nitrogen reduction, although the reduction in Nitrogen resulted in above average colonization of oligotrophic species.
 
@@
 Although there is evidence supporting the use of such a 'space for time substitution' for detecting the effects of Nitrogen deposition on plant diversity [@Stevens2010], they can not replace studies that relate temporal patterns in species with Nitrogen deposition [@deSchrijver2011].
-There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of N-deposition in extant communities [@Clark2008; @Storkey2015; @Stevens2016].
-In an recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the Nitrogen deposition gradient at which species richnes of oligotrophic species richness start to decrease with increasing Nitrogen deposition [@Roth2017].
+There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of Nitrogen deposition in extant communities [@Clark2008; @Storkey2015; @Stevens2016].
+In a recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the Nitrogen deposition gradient at which species richness of oligotrophic species richness start to decrease with increasing Nitrogen deposition [@Roth2017].
 Using this spatial variation in species richness and nitrogen deposition we estimated a critical load for mountain hay meadows of 13.1 kg ha$^{-1}$ yr$^{-1}$.
@@
 This rate could also be interpreted as a critical load.
-With this aproach we obtained very similar results as the spatial approach.
-Our results are thus validating the space for time approach.
+With this temporal approach we obtained very similar results as the spatial approach.
+Our results are thus validating the space for time approach, at least for Swiss mountain hay meadows.
 
-Hower, the results of Fig.
+However, the results of Fig.
 \@ref(fig:figconsequences) also shows that the results of spatial comparisons must be interpreted carefully.
-When we used the same model to infer how species richness is decreasing along the Nitrogen deposition gradient at differnt time points, except for the first two study periods the decrease became steeper (more negative) over time.
+When we used the same model to infer how oligotrophic species richness is decreasing along the Nitrogen deposition gradient at different time points, except for the first two study periods the decrease became steeper (more negative) over time.
 Our first interpretation was that the Nitrogen deposition effect became stronger over time.
-This contraticted our prediction that the Nitrogen deposition should become weaker over time since Nitrogen deposition decreased.
-However, we then realized that species turnover was highest in low deposition sites (Tab.
-\@ref(tab:turnovertab)) where the colonizating species were species with below average indicator values for Nutrients.
-It seems that the decrease in Nitrogen deposition resulted in oligotrophic species replacing eutrophic species especially at sites with low deposition, thus leading to a steeper decline in oligotrophic species richness from sites with low to sites with high Nitrogen deposition.
-The increasingly steeper decline over time in the eutrophic species richness along the spatial Nitrogen deposition gradient is thus a sign of the recovery of plant communities in low deposition sites rather than a sign that the Nitrogen deposition effects became stronger over time.
+This contradicted our prediction that the Nitrogen deposition should become weaker over time since Nitrogen deposition decreased.
+Only then we realized that species turnover was highest in low deposition sites (Tab.
+\@ref(tab:turnovertab)) where the colonizing species were species with below average indicator values for Nutrients.
+It seems that the decrease in Nitrogen deposition resulted in oligotrophic species replacing eutrophic species especially at sites with low deposition, thus leading to a steeper spatial decline in oligotrophic species richness from sites with low to sites with high Nitrogen deposition.
+The increasingly steeper decline over time in the eutrophic species richness along the spatial Nitrogen deposition gradient is thus a sign of the recovery of plant communities in low deposition sites rather than a sign that the negative Nitrogen deposition effects became stronger over time.
 
 # Conclusions {-}
-Comparing the indicator values of colonizating and desappearing species with random expecation we found that oligotrphic species are currently more likeley to colonize mountain hay meadows than oligotrophic species, which might be the results of the recently decreasing atmospeheric Nitrogen deposition.
-However, our results also caution that re-transfering mountain hay maeadows to oligotrophic species communities after high nitrogen depostion might take much longer than transfering a and oligotrophic into an eutrophic community.
+  Comparing the indicator values of colonizing and disappearing species with random expectation we found that oligotrophic species are currently more likely to colonize mountain hay meadows than oligotrophic species, which might be the results of the recently decreasing atmospheric Nitrogen deposition.
+However, our results also caution that re-transferring mountain hay meadows to oligotrophic species communities after high nitrogen deposition might take much longer than transferring an oligotrophic into an eutrophic community.
 This is because eutrophic species have high local survival probabilities also when Nitrogen deposition becomes lower again.

### 2018-06-25 Manuscript.Rmd
Bü als Koauthor eingefügt.

@@
   - name: Lukas Kohli
+    affiliation: 2
+  - name: Christoph Bühler
     affiliation: 2

### 2018-06-25 Manuscript.Rmd
Abbildungslegene korrigiert.

@@
 
-```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (indicator value for nutrients <3; red line) and eutrophic (indicator value for nutrients >3) species along the Nitrogen deposition gradient.
+```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (indicator value for nutrients $<3$; red line) and eutrophic (indicator value for nutrients $>3$) species along the Nitrogen deposition gradient.
 Given are means and 95\\%-Credible Intervals from logistic linear mixed models.

### 2018-06-28 Manuscript.Rmd
Update Manuscript.Rmd

@@
   Nitrogen (N) deposition is a major threat to biodiversity of many habitats.
-The recent introduction of cleaner technologies in Switzerland has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition.
-We inferred different drivers of community change (i.e.
-Nitrogen deposition, climate warming, land-use change) in Swiss mountain hay meadows using data from the Swiss biodiversity monitoring.
-While oligotrophic species were more likely to colonize compared to random expectation, we found only weak shifts in plant community structure.
+The recent introduction of cleaner technologies in Switzerland has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in N-deposition.
+We considered different drivers of plant community change (i.e.
+N-deposition, climate warming, land-use change) in Swiss mountain hay meadows using data from the Swiss biodiversity monitoring.
+While oligotrophic plant species were more likely to colonize compared to random expectation, we found only weak shifts in plant community structure.
 Particularly, the average nutrient value of plant communities remained stable over time.
@@
 # Introduction {-}
-Nitrogen deposition is the entry of reactive nitrogen compounds into water and soil via deposition through the air.
+Nitrogen deposition is the entry of reactive nitrogen compounds into water and soil via input from the atmosphere to the biosphere.
 Since Nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low Nitrogen availability [@Vitousek1997], the addition of Nitrogen is likely to alter these communities.
@@
 
-Since the late 1980s measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe [@Torseth2012; @Fowler2007].
+Since the late 1980s measures to reduce atmospheric pollution have successfully diminished emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe [@Torseth2012; @Fowler2007].
 However, there are a number of factors that may prevent the recovery of plant communities that suffered from increased Nitrogen deposition.
 First, even if Nitrogen deposition is reduced, large areas might still be above the critical load [@Slootweeg2015; @RihmAchermann2016].
-Second, there is a potential that communities reach an alternative stable state after decades of increased Nitrogen deposition and these species are unlikely to disappear again after the reduction of Nitrogen deposition [@Stevens2016].
-Therefore, it is an open question whether and how fast the reduction in Nitrogen deposition rates will lead to the recovery of extant plant communities.
+Second, there is a potential that communities reach an alternative stable state after decades of increased Nitrogen deposition and these species are unlikely to disappear subsequent to reduced N-deposition [@Stevens2016].
+Therefore, it is an open question whether and how fast the reduction in N-deposition rates will lead to the recovery of extant plant communities.
 
-Recovery of extant plant communities from high Nitrogen deposition would imply that the state of communities measured at different time points is improving over time (i.e.
+Recovery of extant plant communities from high N- deposition would imply that the state of communities measured at different points in time is improving over time (i.e.
 improving biodiversity endpoints sensu @Rowe2017).
@@
 This is because immigration and extinction might be equally frequent and may cancel each other out [@Hillebrand2018].
-Thus a useful approach to understanding biodiversity change is through estimates of biodiversity turnover reflecting both colonization and local extinction [@Hillebrand2018], especially if colonization and local extinction is compared to random expectation [@Chase2011].
+Thus a useful approach to understanding biodiversity change is through estimates of species turnover reflecting both colonization and local extinction [@Hillebrand2018], especially if colonization and local extinction is compared to random expectation [@Chase2011].
 
@@
 With extensive cultivation, permanent grassland has a very high biodiversity.
-This applies in particular to the meadows in the alpine region, where meadows with a high plant diversity are also of importance agronomically [@Leiber2006].
+This applies in particular to the meadows in the alpine region, where meadows with a high plant diversity are also of agronomical importance [@Leiber2006].
 In mountain hay meadows, the spatial variation of species richness in vascular plants is negative correlated with Nitrogen deposition [@Roth2013], suggesting that mountain grasslands suffered from increased Nitrogen deposition.
-However, between 1990 and 2010 NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% [@MaasGrennfelt2016] and we expect plant communities to recover.
+However, between 1990 and 2010 NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% [@MaasGrennfelt2016] and we would expect plant communities to recover.
 
-However, Swiss mountain ecosystems are threatened also by other global change drivers.
+In addition to N-deposition, Swiss mountain ecosystems are threatened also by other global change drivers.
 In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35$^{\circ}$C per decade, which is about 1.6 times the northern hemispheric warming rate [@Ceppi2012].
@@
 
-Here, we used the data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, Nitrogen deposition or land-use intensity change over the last 15 years, (2) was species turnover correlated with the average temperature, precipitation, Nitrogen deposition or inclination (we expect steep areas to be less intensively managed), and (3) did species that newly colonized or disappeared (local extinction) from sites differ from random expectation according to their indicator values for temperature, humidity, nutrients or light.
+Here, we used data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, Nitrogen deposition or land-use intensity change over the last 15 years, (2) was species turnover correlated with the average temperature, precipitation, Nitrogen deposition or inclination (we expect steep areas to be less intensively managed), and (3) did species that newly colonized or disappeared (local extinction) from sites differ from random expectation according to their indicator values for temperature, humidity, nutrients or light.
 
@@
 During that study period each sampling site was surveyed once per five-year period: the first period lasted from 2003 to 2007, the second from 2008 to 2012 and the third from 2013 to 2017.
-These selection criteria resulted in sample of `r nrow(sites)` sites.
+These selection criteria yielded a sample of `r nrow(sites)` sites.
 
-For each survey (a survey consists of the two visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for (2) temperature, (3) humidity, (4) nutrients and (5) light.
+For each survey (a survey consists of two  visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for (2) temperature, (3) humidity, (4) nutrients and (5) light.
 The indicator values were obtained from the recalibrated indicator values for the Swiss Flora [@Landolt2010].
-Additionally, to these five biodiversity endpoints that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
-species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points to describe the community change between two time points.
+In addittion to these five biodiversity endpoints that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
+species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points to describe the community change between two points in time.
 
@@
 To better disentangle the importance of these mechanisms we ordered the sites along four main environmental gradients.
-First, we expected communities to response to climate warming and the response might be different along the temperature gradient [@Roth2014].
+First, we expected communities to response to climate warming with this response being dependent on the position of the site along the temperature gradient [@Roth2014].
 To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
-Another key drivers for plants that is likely to be affected by climate change is precipitation [@Beier2012].
+Another key driver for plants that is likely to be affected by climate change is precipitation [@Beier2012].
 We used the annual precipitation per site from the WordClim database [@Fick2017].
@@
 Then we calculated the differences of the CM minus the average of the random-CMs to obtain a standardized measure (standardized-CM) of how different the disappearing species were from random expectation.
-A standardized-CM < 0 would suggest that the indicator values of colonizing species were lower than the indicator values of average species that were occurring at this site.
+For example, a standardized-CM < 0 would suggest that the indicator values of colonizing species were lower than might be expected from random colonisation from the species-pool for this site.
 We applied this method for both colonizing and disappearing species and for the indicator values for temperature, humidity, nutrients and light (see Appendix A).
-We then inferred whether the standardized-CM is changing along the corresponding gradient using linear models.
+We then tested whether the standardized-CM is changing along the corresponding gradient (inferred from independet datasets, see Environmental gradients section) using linear models.
 Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
 
-We used logistic-LMM to infer whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this changing depended on the species indicator value for nutrients.
-To analyse the colonization probability, we selected all species that were not observed during the first survey and asked if they were observed (then Occ = 1) or not observed (then Occ = 0) during the second survey.
+We used logistic-LMM to analyse whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
+To analyse the colonization probability of species, we selected all species that were not observed during the first survey.
+For each of these species we asked if they were observed (then Occ = 1) or not observed (then Occ = 0) during the second survey.
 The same was also done for all species that were not observed during the second survey if they were observed or not observed (then Occ = 0) during the third survey.
@@
 Additionally, species-ID and site-ID were included as random effects.
-The same logistic LMM was also used to infer local survival probability.
+The same logistic LMM was also used to investigate local survival probabilities  .
 In that case, however, we selected all species that were recorded during the first or second survey and the variable 'occ' then indicated whether or not the species was also observed during the next survey.
@@
 
-To estimate the effect of Nitrogen deposition on total species richness at a given time point, we described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarithm as link function.
+To estimate the effect of Nitrogen deposition on total species richness at a given point in time, we described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarithm as link function.
 As predictors we used the four environmental gradients as described above.

### 2018-06-28 Manuscript.Rmd
Update Manuscript.Rmd

@@
 
-The five measures of plant community structure (i.e.
+The five measures of plant community composition (i.e.
 biodiversity endpoints) suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of species-richness and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
@@
 for Trend")`), followed by the community mean of the indicator value for light (probability of decrease: `r 1 - res %>% filter(Measures == "Light value") %>% pull("Prob.
-for Trend")`) and it was least likely for the species richness (probability of increase: `r res %>% filter(Measures == "Species-richness") %>% pull("Prob.
+for Trend")`).
+A linear temporal change was least likely for the species richness (probability of increase: `r res %>% filter(Measures == "Species-richness") %>% pull("Prob.
 for Trend")`).
@@
 To better understand the factors that drive these changes we were particularly interested whether the species that disappeared or colonized the sites differed in indicator values compared to what would be expected if the same number of species randomly disappeared or colonized the sites (i.e.
-random disappearance and random colonization) and whether there is a spatial variation along the gradients.
+random disappearance and random colonization) and whether there is a spatial variation along the environmental gradients.
 It seems that the indicator values of newly colonizing species differed more from random colonization than the indicator values of disappearing species (Table \@ref(tab:difffromrandomtab)).
-For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower indicator value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing Nitrogen deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
+For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with relatively low nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower indicator value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing Nitrogen deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
 Thus, at high Nitrogen deposition colonizing species did not differ from random species (see Figure 3b in Appendix A).
@@
 Nevertheless, the variation in indicator value for temperature seemed important to explain the total species turnover.
-This is because, disappearing species tend to have lower temperature value than random species as well as colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
+This is because disappearing species tend to have lower temperature values than random species as well as colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
 This was not the case for the indicator value for nutrients: species with lower nutrients values tended to be more likely to disappear from as well as to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
@@
 \@ref(fig:Ndepchange)a).
-In 2000 only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a Nitrogen deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, which lied all above 1000m.
-Between 2000 and 2015 the Nitrogen deposition decreased in average `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$ with slightly higher net decreases at lower elevation (Fig.
+In 2000 only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a Nitrogen deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$.
+All of these sites are situated above 1000m.
+Between 2000 and 2015 the Nitrogen deposition decreased in average by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$ with slightly higher net decreases at lower elevation (Fig.
 \@ref(fig:Ndepchange)a).
@@
 
-While we could not detect a consistent decrease in average indicator value for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial variation of oligotrophic species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between Nitrogen deposition and oligotrophic species richness.
-Indeed, if we apply at different time points a similar model as in @Roth2013 to infer the effects of Nitrogen deposition on the spatial variation of oligotrophic species richness, the resulting effect size (i.e.
+While we could not detect a consistent decrease in average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial variation of oligotrophic species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between Nitrogen deposition and oligotrophic species richness.
+Indeed, if we apply a similar model as in @Roth2013 at different points in time to infer the effects of Nitrogen deposition on the spatial variation of oligotrophic species richness, the resulting effect size (i.e.
 the slope) became more negative over time (Fig.
@@
 biodiversity endpoints sensu @Rowe2017) during the same time period (Table \@ref(tab:communitytrendstab)).
-While the slight increase in average temperature value suggests that plant communities adopted to increasing temperatures, the constant average nutrients value suggests that the decrease in Nitrogen deposition did not yet affect plant communities.
-However, this apparent stability in community structure was accompanied by marked temporal turnover in species identities.
+While the slight increase in average temperature values  suggests that plant communities adopted to increasing temperatures, the constant average nutrients value suggests that the decrease in Nitrogen deposition did not yet affect plant communities.
+However, this apparent stability in community composition  was accompanied by marked temporal turnover in species identities.
 It seems unlikely that this temporal turnover can entirely be explained by methodological issues such as overlooked species.
@@
 We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site (Appendix A).
-We found that the nutrient values of colonization species showed largest deviations from random expectation suggesting that Nitrogen deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
+We found that the nutrient values of colonizing species showed largest deviations from random expectation suggesting that Nitrogen deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
 
@@
 This comparatively low change in Nitrogen deposition in Swiss mountain hay meadows might explain why we observed no change in average nutrient value of communities.
-Additionally, other anthropogenic pressures such as climate change might have overruled Nitrogen depositing effects on community structure.
+Additionally, other anthropogenic pressures such as climate change might have overruled Nitrogen depositing effects on community composition.
 Particularly, we found that the species disappearing from the sites tend to have below average indicator values for temperature, while species that newly colonize sites have above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
@@
 This suggests that mountain hay meadow communities can reach alternative stable states with eutrophic species that are unlikely to disappear even if Nitrogen deposition is reduced [@Stevens2016].
-Taken together, these reasons might explain why the response of mountain hay meadow communities responded stronger to climate warming than to Nitrogen reduction, although the reduction in Nitrogen resulted in above average colonization of oligotrophic species.
+Taken together, these reasons might explain why the composition of mountain hay meadow communities responded stronger to climate warming than to Nitrogen reduction, although the reduction in Nitrogen resulted in above average colonization of oligotrophic species.
 
 Observational studies along a gradient of Nitrogen deposition often infer how the spatial variation in species richness is related to Nitrogen deposition [@Stevens2010gradient; chapter 4 in @deVries2015].
-They assume that the spatial variation in species richness (or other metrics of community structure) arose because over time some areas lost more species than others because they chronically experienced higher Nitrogen deposition.
+They assume that the spatial variation in species richness (or other metrics of community composition) arose because over time some areas lost more species than others because they chronically experienced higher Nitrogen deposition.
 Although there is evidence supporting the use of such a 'space for time substitution' for detecting the effects of Nitrogen deposition on plant diversity [@Stevens2010], they can not replace studies that relate temporal patterns in species with Nitrogen deposition [@deSchrijver2011].

### 2018-06-29 Manuscript.Rmd
Update Manuscript.Rmd

@@
 
-Observational studies along a gradient of Nitrogen deposition often infer how the spatial variation in species richness is related to Nitrogen deposition [@Stevens2010gradient; chapter 4 in @deVries2015].
-They assume that the spatial variation in species richness (or other metrics of community composition) arose because over time some areas lost more species than others because they chronically experienced higher Nitrogen deposition.
-Although there is evidence supporting the use of such a 'space for time substitution' for detecting the effects of Nitrogen deposition on plant diversity [@Stevens2010], they can not replace studies that relate temporal patterns in species with Nitrogen deposition [@deSchrijver2011].
-There are only a limited number of studies inferring temporal trends of plant species diversity related to varying amounts of Nitrogen deposition in extant communities [@Clark2008; @Storkey2015; @Stevens2016].
+Observational studies along a gradient of Nitrogen deposition often conclude how the spatial variation in species richness is related to Nitrogen deposition [@Stevens2010gradient; chapter 4 in @deVries2015].
+They assume that the spatial variation in species richness (or other metrics of community composition) arose because of unequal species loss of different areas over time resulting from elevated N-deposition chronically experienced by some areas.
+Although there is evidence supporting the pertinency of such a 'space for time substitution' for detecting the effects of Nitrogen deposition on plant diversity [@Stevens2010], this approach can not replace studies that relate temporal patterns in species composition with Nitrogen deposition [@deSchrijver2011].
+There are only a limited number of studies directly relating temporal trends of plant species diversity to varying amounts of Nitrogen deposition in extant communities [@Clark2008; @Storkey2015; @Stevens2016].
 In a recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the Nitrogen deposition gradient at which species richness of oligotrophic species richness start to decrease with increasing Nitrogen deposition [@Roth2017].
@@
 However, the results of Fig.
-\@ref(fig:figconsequences) also shows that the results of spatial comparisons must be interpreted carefully.
-When we used the same model to infer how oligotrophic species richness is decreasing along the Nitrogen deposition gradient at different time points, except for the first two study periods the decrease became steeper (more negative) over time.
+\@ref(fig:figconsequences) also show that the results of spatial comparisons must be interpreted carefully.
+When we applied the same model as in Roth et al.
+(2017) for different study periods to infer how oligotrophic species richness is decreasing along the Nitrogen deposition gradient, the relationship showed to vary between study periods.
+The decrease gradually became steeper (more negative) over time, except for the first two study periods.
 Our first interpretation was that the Nitrogen deposition effect became stronger over time.
-This contradicted our prediction that the Nitrogen deposition should become weaker over time since Nitrogen deposition decreased.
-Only then we realized that species turnover was highest in low deposition sites (Tab.
-\@ref(tab:turnovertab)) where the colonizing species were species with below average indicator values for Nutrients.
-It seems that the decrease in Nitrogen deposition resulted in oligotrophic species replacing eutrophic species especially at sites with low deposition, thus leading to a steeper spatial decline in oligotrophic species richness from sites with low to sites with high Nitrogen deposition.
-The increasingly steeper decline over time in the eutrophic species richness along the spatial Nitrogen deposition gradient is thus a sign of the recovery of plant communities in low deposition sites rather than a sign that the negative Nitrogen deposition effects became stronger over time.
+This was against our prediction that the effect of N-deposition should become weaker over time since Nitrogen deposition actually decreased in the course of the study period.
+Then we realized that species turnover was highest at low N-deposition sites (Tab.
+\@ref(tab:turnovertab)).
+At low N-deposition rates colonizing species have below average indicator values for nutrients.
+It seems that the decrease in Nitrogen deposition resulted in oligotrophic species replacing eutrophic species especially at sites with low N-deposition.
+This explains why the decline in oligotrophic species richness inferred from spatial patterns of species richness and N-deposition is becoming steeper over time.
+This may be interpreted as evidence that plant communities are recovering at least at low deposition sites rather than inferring that the negative N-deposition effects have become stronger over time.
 
 # Conclusions {-}
-  Comparing the indicator values of colonizing and disappearing species with random expectation we found that oligotrophic species are currently more likely to colonize mountain hay meadows than oligotrophic species, which might be the results of the recently decreasing atmospheric Nitrogen deposition.
-However, our results also caution that re-transferring mountain hay meadows to oligotrophic species communities after high nitrogen deposition might take much longer than transferring an oligotrophic into an eutrophic community.
+  Comparing the indicator values of colonizing and disappearing species with random expectation we found that oligotrophic species are currently more likely to colonize mountain hay meadows than eutrophic species, which might be the result of decreasing atmospheric Nitrogen deposition recently observed.
+However, our results also indicate that re-transferring mountain hay meadows to oligotrophic species communities after high nitrogen deposition might take much longer than transferring an oligotrophic into an eutrophic community.
 This is because eutrophic species have high local survival probabilities also when Nitrogen deposition becomes lower again.
-All these results were, however, hidden when only inferring the temporal trends of species communities.
-Our study thus adds to the understanding of contemporary biodiversity change that is currently not well understood [@Magurran2018], and it supports the notion of @Hillebrand2018 that inferring species turnover will generate a far more reliable understanding of the biotic response to changing environments than solely tracking community composition.
+However, these results are hidden if only the temporal trends of species composition are analysed instead of species turnover, which can better reveal ongoing processes in this case.
+Our study thus adds to the understanding of contemporary biodiversity change where some key findings are still missing [@Magurran2018], and it supports the notion of @Hillebrand2018 that consicering species turnover will generate far more reliable view of the biotic response to changing environments than solely tracking community composition.
 
 # Acknowledgements {-}
-We thank the dedicated and qualified botanists who conducted fieldwork.
+We thank the dedicated and qualified botanists who conducted fieldwork for the Swiss Biodiversity Monitoring BDM.
 The Swiss Federal Office for the Environment (FOEN) kindly provided biodiversity monitoring data and topographic data.

### 2018-06-29 Manuscript.Rmd
Update Manuscript.Rmd

@@
 
-While we could not detect a consistent decrease in average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial variation of oligotrophic species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between Nitrogen deposition and oligotrophic species richness.
+While we could not detect a consistent decrease in average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial pattern of oligotrophic species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between Nitrogen deposition and oligotrophic species richness.
 Indeed, if we apply a similar model as in @Roth2013 at different points in time to infer the effects of Nitrogen deposition on the spatial variation of oligotrophic species richness, the resulting effect size (i.e.

### 2018-07-04 Manuscript.Rmd
Update Manuscript.Rmd

@@
   Nitrogen (N) deposition is a major threat to biodiversity of many habitats.
-The recent introduction of cleaner technologies in Switzerland has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in N-deposition.
+The recent introduction of cleaner technologies in Switzerland has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in N deposition.
 We considered different drivers of plant community change (i.e.
 N-deposition, climate warming, land-use change) in Swiss mountain hay meadows using data from the Swiss biodiversity monitoring.
+For this purpose, we compared indicator values of species that disappeared or colonized a site with the indicator values of randomly chosen species from the same site.
 While oligotrophic plant species were more likely to colonize compared to random expectation, we found only weak shifts in plant community structure.
-Particularly, the average nutrient value of plant communities remained stable over time.
+Particularly, the average nutrient value of plant communities remained stable over time (2003–2017).
+Despite this, we found that the nutrient values of colonization species showed largest deviations from random expectation suggesting that N deposition or other factors that change the nutrient content of soils were important drivers of the species composition change over the last 15 years in Swiss mountain hay meadows.
+In addition, we observed an overall replacement of species with lower indicator values for temperature by species with higher values.
 Apparently, the community effects of the replacement of eutrophic species with oligotrophic species was outweighed by climate warming.
@@
 # Introduction {-}
-Nitrogen deposition is the entry of reactive nitrogen compounds into water and soil via input from the atmosphere to the biosphere.
-Since Nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low Nitrogen availability [@Vitousek1997], the addition of Nitrogen is likely to alter these communities.
-Indeed, Nitrogen deposition together with land-use and climate change is one of the major threats to biodiversity [@Sala2000; @Bobbink2010].
+Nitrogen (N) deposition is the entry of reactive nitrogen compounds into soil, water and vegetation input from the atmosphere to the biosphere.
+Since nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low nitrogen availability [@Vitousek1997], the addition of nitrogen is likely to alter these communities.
+Indeed, N deposition together with land-use and climate change is one of the major threats to biodiversity [@Sala2000; @Bobbink2010; @Murphy2013].
 While there is strong evidence for the reduction in diversity of species-rich grasslands due to increased Nitrogen deposition [@Stevens2006; @Dupre2010; @Maskell2010; @Wesche2012], mountain grasslands have received less attention [@Humbert2016].
 
-Since the late 1980s measures to reduce atmospheric pollution have successfully diminished emissions of nitrogen oxides, with affiliated decrease in Nitrogen deposition in many parts of Europe [@Torseth2012; @Fowler2007].
-However, there are a number of factors that may prevent the recovery of plant communities that suffered from increased Nitrogen deposition.
-First, even if Nitrogen deposition is reduced, large areas might still be above the critical load [@Slootweeg2015; @RihmAchermann2016].
-Second, there is a potential that communities reach an alternative stable state after decades of increased Nitrogen deposition and these species are unlikely to disappear subsequent to reduced N-deposition [@Stevens2016].
-Therefore, it is an open question whether and how fast the reduction in N-deposition rates will lead to the recovery of extant plant communities.
+Since the late 1980s measures to reduce atmospheric pollution have successfully diminished emissions of nitrogen oxides, with affiliated decrease in N deposition in many parts of Europe [@Torseth2012; @Fowler2007].
+However, there are a number of factors that may prevent the recovery of plant communities that suffered from increased N deposition.
+First, N deposition is still high at many sites, since, in contrast to nitrogen oxides, the recent decrease of ammonia emissions was only minor.
+Thus, even if N deposition is reduced, large areas might still be above the critical load [@Slootweeg2015; @RihmAchermann2016].
+Second, there is a potential that communities reach an alternative stable state after decades of increased N deposition and these species are unlikely to disappear subsequent to reduced N deposition [@Stevens2016].
+Therefore, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities.
 
-Recovery of extant plant communities from high N- deposition would imply that the state of communities measured at different points in time is improving over time (i.e.
+Recovery of existant plant communities from high N deposition would imply that the state of communities measured at different points in time is improving over time (i.e.
 improving biodiversity endpoints sensu @Rowe2017).
-Species richness is an easy to measure and communicate biodiversity endpoint and it is often negatively correlated with Nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
+Species richness, a biodiversity endpoint that can be assessed relatively simple and communicated easily, is often negatively related to nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
 Other metrics that are potentially more useful to reflect favourable changes can be derived from the traits of the species in a community.
-In Europe, environmental preference has often been expressed using indicator values assigned to each plant species [@Ellenberg1992].
+In Europe, environmental preference of plants has often been expressed using indicator values assigned to each plant species [@Ellenberg1992; @Landolt2010].
 Examples of such metrics would be the number of oligotrophic species or the average indicator value of the species in a community [@Roth2013; @Rowe2017].
@@
 This applies in particular to the meadows in the alpine region, where meadows with a high plant diversity are also of agronomical importance [@Leiber2006].
-In mountain hay meadows, the spatial variation of species richness in vascular plants is negative correlated with Nitrogen deposition [@Roth2013], suggesting that mountain grasslands suffered from increased Nitrogen deposition.
-However, between 1990 and 2010 NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% [@MaasGrennfelt2016] and we would expect plant communities to recover.
+In mountain hay meadows, the spatial variation of species richness in vascular plants is negative correlated with N deposition [@Roth2013], suggesting that mountain grasslands suffered from increased N deposition.
+However, between 1990 and 2010 NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% [@MaasGrennfelt2016].
+What may have led to a partial recovery of plant communities.
 
-In addition to N-deposition, Swiss mountain ecosystems are threatened also by other global change drivers.
+In addition to N deposition, Swiss mountain ecosystems are threatened also by other global change drivers.
 In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35$^{\circ}$C per decade, which is about 1.6 times the northern hemispheric warming rate [@Ceppi2012].
-This climate warming is likely to directly affect or interact with Nitrogen deposition to drive plant community changes [@Humbert2016].
-Indeed, in an earlier study we found that at the landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014] and particularly the shift of plant communities at mountain summits are the result of recent climate warming interacting with airborne nitrogen deposition [@Steinbauer2018].
+This climate warming is likely to directly affect or interact with N deposition to drive plant community changes [@Humbert2016].
+Indeed, in an earlier study we found that at the landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014] and particularly the shift of plant communities at mountain summits are the result of recent climate warming interacting with airborne N deposition [@Steinbauer2018].
 Furthermore, traditional management regimes are currently changing, which also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
@@
 
-Here, we used data from the Swiss biodiversity monitoring [@Weber2004] and addressed the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, Nitrogen deposition or land-use intensity change over the last 15 years, (2) was species turnover correlated with the average temperature, precipitation, Nitrogen deposition or inclination (we expect steep areas to be less intensively managed), and (3) did species that newly colonized or disappeared (local extinction) from sites differ from random expectation according to their indicator values for temperature, humidity, nutrients or light.
+Here, we used data from the Swiss biodiversity monitoring [@Weber2004] to address the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, N deposition or land-use intensity change over the last 15 years, (2) was species turnover correlated with the average temperature, precipitation, N deposition or inclination (we expect steep areas to be less intensively managed), and (3) did species that newly colonized or disappeared (local extinction) from sites differ from random expectation according to their indicator values for temperature, humidity, nutrients or light.
 
@@
 We used the annual precipitation per site from the WordClim database [@Fick2017].
-Further, we estimated atmospheric Nitrogen deposition for each site using a pragmatic approach described in @Rihm2001 that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models.
-For details on the estimation of Nitrogen deposition see also @RihmAchermann2016 and @Roth2017.
+Further, we estimated atmospheric N deposition for each site using a pragmatic approach  that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models [@RihmAchermann2016].
 Finally, we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed [@Strebel2015].
@@
 We repeated the random selection of species 1000 times.
-Then we calculated the differences of the CM minus the average of the random-CMs to obtain a standardized measure (standardized-CM) of how different the disappearing species were from random expectation.
-For example, a standardized-CM < 0 would suggest that the indicator values of colonizing species were lower than might be expected from random colonisation from the species-pool for this site.
+Then we calculated the differences of the CM minus the average of the random-CMs to obtain a standardized difference (standardized-CM) of how different the disappearing species were from random expectation.
+For example, a difference < 0 would suggest that the indicator values of colonizing species were lower than might be expected from random colonisation from the species-pool for this site.
 We applied this method for both colonizing and disappearing species and for the indicator values for temperature, humidity, nutrients and light (see Appendix A).
-We then tested whether the standardized-CM is changing along the corresponding gradient (inferred from independet datasets, see Environmental gradients section) using linear models.
+We then tested whether this standardized difference is changing along the corresponding gradient (inferred from independet datasets, see Environmental gradients section) using linear models.
 Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
@@
 Additionally, species-ID and site-ID were included as random effects.
-The same logistic LMM was also used to investigate local survival probabilities  .
+The same logistic LMM was also used to investigate local survival probabilities.
 In that case, however, we selected all species that were recorded during the first or second survey and the variable 'occ' then indicated whether or not the species was also observed during the next survey.
@@
 
-To estimate the effect of Nitrogen deposition on total species richness at a given point in time, we described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarithm as link function.
+To estimate the effect of N deposition on total species richness at a given point in time, we described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarithm as link function.
 As predictors we used the four environmental gradients as described above.
@@
 The column 'Prob.
-for trend' gives the probability that the linear trend is $> 0$.")
+for trend' gives the probability that the linear trend is $> 0$.
+Indicator values according to @Landolt2010")
 ```
 
-The five measures of plant community composition (i.e.
+Species richness and four measures of plant community structure according to Landolt indicator values (i.e.
 biodiversity endpoints) suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of species-richness and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
@@
 ```{r difffromrandomtab}
-kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Ellenberg value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average indicator value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
+kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Indicator value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average indicator value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
 Shown are the results from linear model with the difference between disappeared/colonized species and random species as dependent variable and the site-measure (gradient) as predictor variable.") %>%
@@
 It seems that the indicator values of newly colonizing species differed more from random colonization than the indicator values of disappearing species (Table \@ref(tab:difffromrandomtab)).
-For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with relatively low nitrogen deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower indicator value for nutrients than species under random disappearance (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing Nitrogen deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
-Thus, at high Nitrogen deposition colonizing species did not differ from random species (see Figure 3b in Appendix A).
+For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with relatively low N deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower indicator value for nutrients than species under random colonization (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
+Thus, at high N deposition colonizing species did not differ from random species (see Figure 3b in Appendix A).
 
-While colonizing species had higher temperature values compared to what we would expect under random colonization, the differences between colonizing species and random colonization was about four times smaller compared to the difference in indicator value for nutrients between colonizing species and random species.
-Nevertheless, the variation in indicator value for temperature seemed important to explain the total species turnover.
+While colonizing species had higher temperature values compared to what we would expect under random colonization, the differences between colonizing species and random colonization was about four times smaller than the difference in indicator value for nutrients between colonizing species and random species.
+Nevertheless, the variation in indicator value for temperature is revealed to be important to explain the total species turnover.
 This is because disappearing species tend to have lower temperature values than random species as well as colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
@@
 ## Potential effects of reduction in Nitrogen emmissions {-}
-```{r Ndepchange, fig.cap="(a) The nitrogen deposition in 2000 and (b) the change in Nitrogen deposition between 2000 and 2015 along the elevational gradient of the study sites."}
+```{r Ndepchange, fig.cap="(a) The nitrogen (N) deposition in 2000 and (b) the change in N deposition between 2000 and 2015 along the elevational gradient of the study sites."}
 p1 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = NTOT_2000)) +
@@
   geom_abline(intercept = 0, slope = 0, lty = 3) +
-  labs(title = "(a) Nitrogen deposition in 2010", 
-       y = expression(paste("Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")")), 
+  labs(title = "(a) Nitrogen deposition in 2000", 
+       y = expression(paste("N deposition (kg ", ha^{-1}," ", yr^{-1}, ")")), 
        x = "Elevation (m)") 
@@
   labs(title = "(b) Temporal change (2000 to 2015)", 
-       y = expression(paste(Delta, "Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")")), 
+       y = expression(paste(Delta, "N deposition (kg ", ha^{-1}," ", yr^{-1}, ")")), 
        x = "Elevation (m)") +
@@
 \@ref(fig:Ndepchange)a).
-In 2000 only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a Nitrogen deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$.
-All of these sites are situated above 1000m.
-Between 2000 and 2015 the Nitrogen deposition decreased in average by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$ with slightly higher net decreases at lower elevation (Fig.
+In 2000 only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a N deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, all situated above 1000m.
+Between 2000 and 2015 the N deposition decreased in average by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$ with slightly higher net decreases at lower elevation (Fig.
 \@ref(fig:Ndepchange)a).
 
-```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (indicator value for nutrients $<3$; red line) and eutrophic (indicator value for nutrients $>3$) species along the Nitrogen deposition gradient.
+```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (indicator value for nutrients $<3$; red line) and eutrophic (indicator value for nutrients $>3$) species along the N deposition gradient.
 Given are means and 95\\%-Credible Intervals from logistic linear mixed models.
@@
        y = "Colonization probability", 
-       x = expression(paste("Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")"))) +
+       x = expression(paste("N deposition (kg ", ha^{-1}," ", yr^{-1}, ")"))) +
   geom_vline(xintercept = median(CLs, na.rm = TRUE)) +
@@
        y = "Local survival probability", 
-       x = expression(paste("Nitrogen deposition (kg ", ha^{-1}," ", yr^{-1}, ")"))) +
+       x = expression(paste("N deposition (kg ", ha^{-1}," ", yr^{-1}, ")"))) +
   geom_vline(xintercept = mean(CLs, na.rm = TRUE)) +
@@
 In Fig.
-\@ref(fig:cl) we compare the colonization and local survival probability of oligotrophic (indicator value of nutrients <3) and eutrophic (indicator value of nutrients > 3) species along the Nitrogen deposition gradient.
+\@ref(fig:cl) we compare the colonization and local survival probability of oligotrophic species (indicator value of nutrients <3) and eutrophic species (indicator value of nutrients > 3) along the N deposition gradient.
 Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for oligotrophic and eutrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
@@
 
-While we could not detect a consistent decrease in average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low deposition rate seems to affect the spatial pattern of oligotrophic species richness: sites with low Nitrogen deposition are likely to become more species rich over time likely resulting in steeper slope of the negative relationship between Nitrogen deposition and oligotrophic species richness.
-Indeed, if we apply a similar model as in @Roth2013 at different points in time to infer the effects of Nitrogen deposition on the spatial variation of oligotrophic species richness, the resulting effect size (i.e.
+While we could not detect a consistent decrease in average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low N deposition rate seems to affect the spatial pattern of oligotrophic species richness: sites with low N deposition are likely to become more species rich over time.
+This likely results in steeper slope of the negative relationship between N deposition and oligotrophic species richness, when comparing this relationshep with a spatial approach as in @Roth2013.
+Indeed, if we apply such an approach at different points in time to infer the effects of N deposition on the spatial variation of oligotrophic species richness, the resulting effect size (i.e.
 the slope) became more negative over time (Fig.
@@
 
-```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of Nitrogen deposition on oligotrophic species richness estimated from applying the Poisson-GLM with species richness as dependent variable and Nitrogen deposition plus other site covariates as predictors using only the surveys from one five-year interval.
+```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of N deposition on oligotrophic species richness estimated from applying the Poisson-GLM with species richness as dependent variable and N deposition plus other site covariates as predictors using only the surveys from one five-year interval.
 Note that within every five-year interval all plots were sampled once."}
@@
   geom_abline(intercept = 0, slope = 0, lty = 3) +
-  labs(x = "Study period", y = "Effect of Nitrogen deposition") + 
+  labs(x = "Study period", y = "Effect of N deposition") + 
   scale_x_continuous(breaks = seq(2004, 2012, 2), labels = paste0(seq(2004, 2012, 2), "-\n", seq(2008, 2016, 2)))
@@
 # Discussion {-}
-Although Nitrogen deposition considerably declined between 2000 and 2015 (Fig.
+Although N deposition declined between 2000 and 2015 (Fig.
 \@ref(fig:Ndepchange)), we could detect only weak shifts in plant community structure (i.e.
 biodiversity endpoints sensu @Rowe2017) during the same time period (Table \@ref(tab:communitytrendstab)).
-While the slight increase in average temperature values  suggests that plant communities adopted to increasing temperatures, the constant average nutrients value suggests that the decrease in Nitrogen deposition did not yet affect plant communities.
+While the slight increase in average temperature indicator values suggests that plant communities adopted to increasing temperatures, the constant average nutrients value suggests that the decrease in N deposition did not yet affect plant communities.
 However, this apparent stability in community composition  was accompanied by marked temporal turnover in species identities.
@@
 Second, spatial variation of turnover showed patterns that can hardly be explained by methodological issues.
-For instance species turnover varied along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition (Table \@ref(tab:turnovertab)).
+For instance species turnover varied along the N deposition gradient with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
 Taken together, our results add to the increasing evidence that contemporary communities are often relatively stable at community level, but this apparent stability is often accompanied by marked turnover of species [@Vellend2013; @Dornelas2014; @Hillebrand2018].
@@
 We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site (Appendix A).
-We found that the nutrient values of colonizing species showed largest deviations from random expectation suggesting that Nitrogen deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
+We found that the nutrient values of colonizing species showed largest deviations from random expectation suggesting that N deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
 
 Nitrogen deposition at our plots decreased by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ on average between 2000 and 2015.
-This is only about one tenth of the decrease in England, where Nitrogen deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
-This comparatively low change in Nitrogen deposition in Swiss mountain hay meadows might explain why we observed no change in average nutrient value of communities.
-Additionally, other anthropogenic pressures such as climate change might have overruled Nitrogen depositing effects on community composition.
+This is only about one tenth of the decrease in England, where N deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
+This comparatively low change in N deposition in Swiss mountain hay meadows might explain why we observed no change in average nutrient value of communities.
+Additionally, other anthropogenic pressures such as climate change might have overruled effects of N depositon on community composition.
 Particularly, we found that the species disappearing from the sites tend to have below average indicator values for temperature, while species that newly colonize sites have above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
@@
 \@ref(tab:communitytrendstab)).
-This is in contrast to how Nitrogen deposition is affecting disappearing and colonization of species: It seems that both species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
+This is in contrast to how N deposition is affecting disappearing and colonization of species: It seems that both species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
 Thus, the effect of disappearing and colonizing  species on the average community value for nutrients partially cancel each other out.
-Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at higher Nitrogen deposition rates.
-This suggests that mountain hay meadow communities can reach alternative stable states with eutrophic species that are unlikely to disappear even if Nitrogen deposition is reduced [@Stevens2016].
-Taken together, these reasons might explain why the composition of mountain hay meadow communities responded stronger to climate warming than to Nitrogen reduction, although the reduction in Nitrogen resulted in above average colonization of oligotrophic species.
+Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at higher N deposition rates.
+This suggests that mountain hay meadow communities can reach alternative stable states with eutrophic species that are unlikely to disappear even if N deposition is reduced [@Stevens2016].
+Taken together, these reasons might explain why the composition of mountain hay meadow communities responded stronger to climate warming than to nitrogen reduction, although the reduction in nitrogen resulted in above average colonization of oligotrophic species.
 
-Observational studies along a gradient of Nitrogen deposition often conclude how the spatial variation in species richness is related to Nitrogen deposition [@Stevens2010gradient; chapter 4 in @deVries2015].
-They assume that the spatial variation in species richness (or other metrics of community composition) arose because of unequal species loss of different areas over time resulting from elevated N-deposition chronically experienced by some areas.
-Although there is evidence supporting the pertinency of such a 'space for time substitution' for detecting the effects of Nitrogen deposition on plant diversity [@Stevens2010], this approach can not replace studies that relate temporal patterns in species composition with Nitrogen deposition [@deSchrijver2011].
-There are only a limited number of studies directly relating temporal trends of plant species diversity to varying amounts of Nitrogen deposition in extant communities [@Clark2008; @Storkey2015; @Stevens2016].
-In a recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the Nitrogen deposition gradient at which species richness of oligotrophic species richness start to decrease with increasing Nitrogen deposition [@Roth2017].
-Using this spatial variation in species richness and nitrogen deposition we estimated a critical load for mountain hay meadows of 13.1 kg ha$^{-1}$ yr$^{-1}$.
-In the current study we estimated the rate of Nitrogen deposition at which local survival probability or colonization probability was equal for oligotrophic and eutrophic species.
+Observational studies along a gradient of N deposition often conclude how the spatial variation in species richness is related to N deposition [@Stevens2010gradient; chapter 4 in @deVries2015].
+They assume that the spatial variation in species richness (or other metrics of community composition) arose because of unequal species loss of different areas over time resulting from elevated N deposition chronically experienced by some areas.
+Although there is evidence supporting the pertinency of such a 'space for time substitution' for detecting the effects of N deposition on plant diversity [@Stevens2010], this approach can not replace studies that relate temporal patterns in species composition with N deposition [@deSchrijver2011].
+There are only a limited number of studies directly relating temporal trends of plant species diversity to varying amounts of N deposition in existant communities [@Clark2008; @Storkey2015; @Stevens2016].
+In a recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the N deposition gradient at which species richness of oligotrophic species richness start to decrease with increasing N deposition [@Roth2017].
+Using this spatial variation in species richness and N deposition we estimated a critical load for mountain hay meadows of 13.1 kg ha$^{-1}$ yr$^{-1}$.
+In the current study we estimated the rate of N deposition at which local survival probability or colonization probability was equal for oligotrophic and eutrophic species.
 This rate could also be interpreted as a critical load.
@@
 At low N-deposition rates colonizing species have below average indicator values for nutrients.
-It seems that the decrease in Nitrogen deposition resulted in oligotrophic species replacing eutrophic species especially at sites with low N-deposition.
-This explains why the decline in oligotrophic species richness inferred from spatial patterns of species richness and N-deposition is becoming steeper over time.
-This may be interpreted as evidence that plant communities are recovering at least at low deposition sites rather than inferring that the negative N-deposition effects have become stronger over time.
+It seems that the decrease in N deposition resulted in oligotrophic species replacing eutrophic species especially at sites with low N deposition.
+This explains why the decline in oligotrophic species richness inferred from spatial patterns of species richness and N deposition is becoming steeper over time.
+This may be interpreted as evidence that plant communities are recovering at least at low deposition sites rather than inferring that the negative N deposition effects have become stronger over time.
 
 # Conclusions {-}
-  Comparing the indicator values of colonizing and disappearing species with random expectation we found that oligotrophic species are currently more likely to colonize mountain hay meadows than eutrophic species, which might be the result of decreasing atmospheric Nitrogen deposition recently observed.
-However, our results also indicate that re-transferring mountain hay meadows to oligotrophic species communities after high nitrogen deposition might take much longer than transferring an oligotrophic into an eutrophic community.
+  Comparing the indicator values of colonizing and disappearing species with random expectation we found that oligotrophic species are currently more likely to colonize mountain hay meadows than eutrophic species, which might be the result of decreasing atmospheric N deposition recently observed.
+However, our results also indicate that re-transferring mountain hay meadows to oligotrophic species communities after high N deposition might take much longer than transferring an oligotrophic into an eutrophic community.
 This is because eutrophic species have high local survival probabilities also when Nitrogen deposition becomes lower again.

### 2018-07-05 Manuscript.Rmd
Korrekturen von Lü übernommen.
Update Manuscript.Rmd

@@
 We considered different drivers of plant community change (i.e.
-N-deposition, climate warming, land-use change) in Swiss mountain hay meadows using data from the Swiss biodiversity monitoring.
+N deposition, climate warming, land-use change) in Swiss mountain hay meadows using data from the Swiss biodiversity monitoring.
 For this purpose, we compared indicator values of species that disappeared or colonized a site with the indicator values of randomly chosen species from the same site.
@@
 Indeed, N deposition together with land-use and climate change is one of the major threats to biodiversity [@Sala2000; @Bobbink2010; @Murphy2013].
-While there is strong evidence for the reduction in diversity of species-rich grasslands due to increased Nitrogen deposition [@Stevens2006; @Dupre2010; @Maskell2010; @Wesche2012], mountain grasslands have received less attention [@Humbert2016].
+While there is strong evidence for the reduction in diversity of species-rich grasslands due to increased N deposition [@Stevens2006; @Dupre2010; @Maskell2010; @Wesche2012], mountain grasslands have received less attention [@Humbert2016].
 
@@
 The indicator values were obtained from the recalibrated indicator values for the Swiss Flora [@Landolt2010].
-In addittion to these five biodiversity endpoints that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
+In addition to these five biodiversity endpoints that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
 species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points to describe the community change between two points in time.
@@
 We repeated the random selection of species 1000 times.
-Then we calculated the differences of the CM minus the average of the random-CMs to obtain a standardized difference (standardized-CM) of how different the disappearing species were from random expectation.
+Then we calculated the differences of the CM minus the average of the random-CMs to obtain a standardized difference (standardized-CM) of how different the colonizing species were from random expectation.
 For example, a difference < 0 would suggest that the indicator values of colonizing species were lower than might be expected from random colonisation from the species-pool for this site.
@@
 The variable 'occ' was then used as dependent variable in the logistic-LMM.
-As predictor variables the model contained the Nitrogen deposition of the site, the indicator value for nutrients of the species and the interaction of these two variables.
+As predictor variables the model contained the N deposition of the site, the indicator value for nutrients of the species and the interaction of these two variables.
 Additionally, species-ID and site-ID were included as random effects.
@@
 for trend' gives the probability that the linear trend is $> 0$.
-Indicator values according to @Landolt2010")
+Indicator values according to Landolt et al.
+(2010).")
 ```
@@
 Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model: `r CrI`).
-Variation in species turnover was largest along the Nitrogen deposition gradient with highest species turnover at sites with low Nitrogen deposition (Table \@ref(tab:turnovertab)).
+Variation in species turnover was largest along the N deposition gradient with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
 The other three gradients were less important to explain the variation in species turnover among sites.
@@
 
-```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic (indicator value for nutrients $<3$; red line) and eutrophic (indicator value for nutrients $>3$) species along the N deposition gradient.
+```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic species (indicator value for nutrients $<3$; red line) and eutrophic species (indicator value for nutrients $>3$) species along the N deposition gradient.
 Given are means and 95\\%-Credible Intervals from logistic linear mixed models.
@@
 When we applied the same model as in Roth et al.
-(2017) for different study periods to infer how oligotrophic species richness is decreasing along the Nitrogen deposition gradient, the relationship showed to vary between study periods.
+(2017) for different study periods to infer how oligotrophic species richness is decreasing along the N deposition gradient, the relationship showed to vary between study periods.
 The decrease gradually became steeper (more negative) over time, except for the first two study periods.
-Our first interpretation was that the Nitrogen deposition effect became stronger over time.
-This was against our prediction that the effect of N-deposition should become weaker over time since Nitrogen deposition actually decreased in the course of the study period.
+Our first interpretation was that the N deposition effect became stronger over time.
+This was against our prediction that the effect of N deposition should become weaker over time since N deposition actually decreased in the course of the study period.
 Then we realized that species turnover was highest at low N-deposition sites (Tab.
 \@ref(tab:turnovertab)).
-At low N-deposition rates colonizing species have below average indicator values for nutrients.
+At low N deposition rates colonizing species have below average indicator values for nutrients.
 It seems that the decrease in N deposition resulted in oligotrophic species replacing eutrophic species especially at sites with low N deposition.

### 2018-07-13 Manuscript.Rmd
Update Manuscript.Rmd
Korrekturen Rihm, Meuli eingearbeitet

@@
     affiliation: 3
-  - name: Reto Meuli
+  - name: Reto Giulio Meuli
     affiliation: 4  
@@
 N deposition, climate warming, land-use change) in Swiss mountain hay meadows using data from the Swiss biodiversity monitoring.
-For this purpose, we compared indicator values of species that disappeared or colonized a site with the indicator values of randomly chosen species from the same site.
+For this purpose, we compared indicator values of species that disappeared or colonized a site (turnover) with the indicator values of randomly chosen species from the same site.
 While oligotrophic plant species were more likely to colonize compared to random expectation, we found only weak shifts in plant community structure.
@@
 Second, there is a potential that communities reach an alternative stable state after decades of increased N deposition and these species are unlikely to disappear subsequent to reduced N deposition [@Stevens2016].
-Therefore, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of extant plant communities.
+Therefore, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of existant plant communities.
 
@@
 This climate warming is likely to directly affect or interact with N deposition to drive plant community changes [@Humbert2016].
-Indeed, in an earlier study we found that at the landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014] and particularly the shift of plant communities at mountain summits are the result of recent climate warming interacting with airborne N deposition [@Steinbauer2018].
+Indeed, in an earlier study we found that at the landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014].
+Particularly the shift of plant communities at mountain summits are the result of recent climate warming and the aothors assume an interacting with airborne N deposition [@Steinbauer2018].
 Furthermore, traditional management regimes are currently changing, which also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
@@
 To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
-Another key driver for plants that is likely to be affected by climate change is precipitation [@Beier2012].
-We used the annual precipitation per site from the WordClim database [@Fick2017].
+From the same database we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change  [@Beier2012].
 Further, we estimated atmospheric N deposition for each site using a pragmatic approach  that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models [@RihmAchermann2016].
@@
 The chance that the community mean of the nutrient value decreased between 2003 and 2017 was `r 1- res %>% filter(Measures == "Nutrients value") %>% pull("Prob.
-for Trend")`.
+for Trend")` (`r res %>% filter(Measures == "Nutrients value") %>% pull("Prob.
+for Trend")`gives the probability for an increase).
 
@@
 Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model: `r CrI`).
-Variation in species turnover was largest along the N deposition gradient with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
-The other three gradients were less important to explain the variation in species turnover among sites.
+Along the N deposition gradient variation in species turnover was largest, with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
+The other three gradients were less conclusive to explain the variation in species turnover among sites.
 
@@
 Between 2000 and 2015 the N deposition decreased in average by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$ with slightly higher net decreases at lower elevation (Fig.
-\@ref(fig:Ndepchange)a).
+\@ref(fig:Ndepchange)b).
 
@@
 
-While we could not detect a consistent decrease in average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at sites with low N deposition rate seems to affect the spatial pattern of oligotrophic species richness: sites with low N deposition are likely to become more species rich over time.
+While we could not detect a consistent decrease in average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at the few sites with low N deposition rate seems to affect the spatial pattern of oligotrophic species richness: sites with low N deposition are likely to become more species rich over time.
 This likely results in steeper slope of the negative relationship between N deposition and oligotrophic species richness, when comparing this relationshep with a spatial approach as in @Roth2013.
@@
 
-```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of N deposition on oligotrophic species richness estimated from applying the Poisson-GLM with species richness as dependent variable and N deposition plus other site covariates as predictors using only the surveys from one five-year interval.
+```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of N deposition on oligotrophic species richness estimated from applying the Poisson-GLM with species richness as dependent variable and N deposition plus other site covariates (elevation, precipitation, inclination, mean indicator values for humidity and light) as predictors using only the surveys from one five-year interval.
 Note that within every five-year interval all plots were sampled once."}
@@
   geom_abline(intercept = 0, slope = 0, lty = 3) +
-  labs(x = "Study period", y = "Effect of N deposition") + 
+  labs(x = "five-year interval", y = "Effect of N deposition") + 
   scale_x_continuous(breaks = seq(2004, 2012, 2), labels = paste0(seq(2004, 2012, 2), "-\n", seq(2008, 2016, 2)))
@@
 We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site (Appendix A).
-We found that the nutrient values of colonizing species showed largest deviations from random expectation suggesting that N deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
+We found that the nutrient values of colonizing species showed largest deviations from random expectation (Table  \@ref(tab:difffromrandomtab)) suggesting that N deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
 
@@
 
-Observational studies along a gradient of N deposition often conclude how the spatial variation in species richness is related to N deposition [@Stevens2010gradient; chapter 4 in @deVries2015].
+Observational studies along a gradient of N deposition often conclude how the spatial variation in species richness is related to N deposition [@Stevens2010gradient; @Roth2013; chapter 4 in @deVries2015].
 They assume that the spatial variation in species richness (or other metrics of community composition) arose because of unequal species loss of different areas over time resulting from elevated N deposition chronically experienced by some areas.
@@
 Our first interpretation was that the N deposition effect became stronger over time.
-This was against our prediction that the effect of N deposition should become weaker over time since N deposition actually decreased in the course of the study period.
-Then we realized that species turnover was highest at low N-deposition sites (Tab.
+This was against our prediction that the effect of N deposition should become weaker over time since N deposition is decreasing during the study period.
+Then we realized that species turnover was highest at low N deposition sites (Tab.
 \@ref(tab:turnovertab)).
@@
 However, our results also indicate that re-transferring mountain hay meadows to oligotrophic species communities after high N deposition might take much longer than transferring an oligotrophic into an eutrophic community.
-This is because eutrophic species have high local survival probabilities also when Nitrogen deposition becomes lower again.
+This is because eutrophic species have high local survival probabilities even after N deposition decreases again.
 However, these results are hidden if only the temporal trends of species composition are analysed instead of species turnover, which can better reveal ongoing processes in this case.

### 2018-08-09 Manuscript.Rmd
Korrekturen Co-Autoren
Boden bei Methoden eingeführt
Korrekturen von Kl übernommen, kleine Anpassungen.

--- 2018-07-04 Manuscript.Rmd
@@
     email: t.roth@unibas.ch
-  - name: Valentin Amrhein
-    affiliation: 1 
   - name: Lukas Kohli
@@
     affiliation: 5    
+  - name: Valentin Amrhein
+    affiliation: 1
 
@@
 To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
-Another key driver for plants that is likely to be affected by climate change is precipitation [@Beier2012].
-We used the annual precipitation per site from the WordClim database [@Fick2017].
+The average $\pm$ SD mean annual temperature at our sites was `r mean(sites$TemperatureR / 10)` $\pm$ `r sd(sites$TemperatureR / 10)` $^\circ$C. 
+Another key driver for plants that is likely to be affected by climate change is precipitation [@Beier2012]. We used the annual precipitation per site from the WordClim database [@Fick2017]. 
+The average annual precipitation at our sites was `r mean(sites$PrecipitationR)` $\pm$ `r sd(sites$PrecipitationR)` mm.
 Further, we estimated atmospheric N deposition for each site using a pragmatic approach  that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models [@RihmAchermann2016].
+The average nitrogen deposition at our sites was `r mean(sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ in 2000 and `r mean(sites$NTOT_2015)` $\pm$ `r sd(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$ in 2015.
 Finally, we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed [@Strebel2015].
+The average inclination at our sites was `r mean(sites$InclinationR)` $\pm$ `r sd(sites$InclinationR)`$^\circ$.
 
@@
 
-To infer whether species turnover was changing along the gradient, we used a Binomial-LMM with the proportion of species that differed between two surveys as dependent variable and the site gradients and period (first/second vs.
+To infer whether species turnover was changing along the gradient, we used a Binomial-GLMM with the proportion of species that differed between two surveys as dependent variable and the site gradients and period (first/second vs.
 second/third surveys) as predictors and site-ID as random effect.
@@
 
-We used logistic-LMM to analyse whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
+We used logistic-GLMM to analyse whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
 To analyse the colonization probability of species, we selected all species that were not observed during the first survey.
@@
 The same was also done for all species that were not observed during the second survey if they were observed or not observed (then Occ = 0) during the third survey.
-The variable 'occ' was then used as dependent variable in the logistic-LMM.
+The variable 'occ' was then used as dependent variable in the logistic-GLMM.
 As predictor variables the model contained the Nitrogen deposition of the site, the indicator value for nutrients of the species and the interaction of these two variables.
 Additionally, species-ID and site-ID were included as random effects.
-The same logistic LMM was also used to investigate local survival probabilities.
+The same logistic-GLMM was also used to investigate local survival probabilities.
 In that case, however, we selected all species that were recorded during the first or second survey and the variable 'occ' then indicated whether or not the species was also observed during the next survey.
@@
 Model parameters were estimated using an approximate Bayesian approach using the R-Package *arm* [@arm2018].
+
+Fully executable R code and a description to produce this manuscript are provided at https://github.com/TobiasRoth/NDep-Trend.
 
@@
 ## Potential effects of reduction in Nitrogen emmissions {-}
-```{r Ndepchange, fig.cap="(a) The nitrogen (N) deposition in 2000 and (b) the change in N deposition between 2000 and 2015 along the elevational gradient of the study sites."}
+```{r Ndepchange, fig.cap="(a) The nitrogen (N) deposition in 2000 and (b) the change in N deposition between 2000 and 2015 along the N deposition gradient of the study sites in 2000."}
 p1 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = NTOT_2000)) +
@@
   labs(title = "(a) Nitrogen deposition in 2000", 
-       y = expression(paste("N deposition (kg ", ha^{-1}," ", yr^{-1}, ")")), 
+       y = expression(paste("N deposition in 2000 (kg ", ha^{-1}," ", yr^{-1}, ")")), 
        x = "Elevation (m)") 
-p2 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = (NTOT_2015 - NTOT_2000) )) +
+p2 <- ggplot(sites, aes(x = NTOT_2000, y = (NTOT_2015 - NTOT_2000) )) +
   geom_point(cex = 0.8) +
@@
        y = expression(paste(Delta, "N deposition (kg ", ha^{-1}," ", yr^{-1}, ")")), 
-       x = "Elevation (m)") +
+       x = expression(paste("N deposition in 2000 (kg ", ha^{-1}," ", yr^{-1}, ")"))) +
   geom_hline(yintercept = 0, lty = 2) 
@@
   geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
-  geom_line() +
+  geom_line(lwd = 1.5) +
   labs(title = "(a) Colonization", 
@@
   geom_ribbon(aes(ymin = low, ymax = high), fill = "black", col = "white", alpha = 0.2) +
-  geom_line() +
+  geom_line(lwd = 1.5) +
   labs(title = "(b) Local survival", 
@@
 Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for oligotrophic and eutrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
-In only `r mean((sites$NTOT * 10 + 10) < 12.5)`% of the sites the deposition rate was below 12.5 kg N ha$^{-1}$ yr$^{-1}$ where the replacement of eutrophic with oligotrophic species is likely.
+In `r 100 * mean((sites$NTOT * 10 + 10) < 12.5)`% of the sites the deposition rate was below 12.5 kg N ha$^{-1}$ yr$^{-1}$ where the replacement of eutrophic with oligotrophic species is likely.
 

### 2018-08-09 R/Data_processing.R

@@ -52,10 +52,14 @@ sites <- tibble(aID_STAO = as.integer(sel)) %>%
                         NTOT_2000 = NTOT2000_Wiesen, NTOT_2010 = NTOT2010_Wiesen, NTOT_2015 = NTOT2015_Wiesen), copy = TRUE)
 
 # Standardize covariates
+sites$ElevationR <- sites$Elevation
 sites$Elevation <- (sites$Elevation - 1000) / 500
+sites$TemperatureR <- sites$Temperature 
 sites$Temperature <- (sites$Temperature - 50) / 10
+sites$PrecipitationR <- sites$Precipitation
 sites$Precipitation <- (sites$Precipitation - 1000) / 200
 sites$NTOT <- (sites$NTOT_2010 - 10) / 10
+sites$InclinationR <- sites$Inclination
 sites$Inclination <- (sites$Inclination - 10) / 10
 sites$pH <- sites$pH - 6
 

### 2018-08-09 R/Settings.R

@@ -39,7 +39,7 @@ load("RData/turnover_double.RData")
 opts_chunk$set(echo = FALSE, hide = TRUE, cache = TRUE, warning = FALSE, message = FALSE,
                fig.asp = .4, fig.width = 8, out.width = "100%")
 
-# Stting for withing text juncs
+# Setting for withing text juncs
 inline_hook <- function(x) {
   if (is.numeric(x)) {
     x <- format(x, nsmall = 2, digits = 2)

### 2018-08-09 Manuscript.Rmd

--- 2018-07-13 Manuscript.Rmd
@@
 To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
-From the same database we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change  [@Beier2012].
+The average $\pm$ SD mean annual temperature at our sites was `r mean(sites$TemperatureR / 10)` $\pm$ `r sd(sites$TemperatureR / 10)` $^\circ$C. 
+ From the same database we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change  [@Beier2012].The average annual precipitation at our sites was `r mean(sites$PrecipitationR)` $\pm$ `r sd(sites$PrecipitationR)` mm.
  Further, we estimated atmospheric N deposition for each site using a pragmatic approach  that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models [@RihmAchermann2016].
+The average nitrogen deposition at our sites was `r mean(sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ in 2000 and `r mean(sites$NTOT_2015)` $\pm$ `r sd(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$ in 2015.
 Finally, we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed [@Strebel2015].
-
+The average inclination at our sites was `r mean(sites$InclinationR)` $\pm$ `r sd(sites$InclinationR)`$^\circ$.
 ## Statistical analyses {-}
@@
 The same was also done for all species that were not observed during the second survey if they were observed or not observed (then Occ = 0) during the third survey.
-The variable 'occ' was then used as dependent variable in the logistic-LMM.
+The variable 'occ' was then used as dependent variable in the logistic-GLMM.
 As predictor variables the model contained the N deposition of the site, the indicator value for nutrients of the species and the interaction of these two variables.
 Additionally, species-ID and site-ID were included as random effects.
-The same logistic LMM was also used to investigate local survival probabilities.
+The same logistic-GLMM was also used to investigate local survival probabilities.
 In that case, however, we selected all species that were recorded during the first or second survey and the variable 'occ' then indicated whether or not the species was also observed during the next survey.

### 2018-09-13 Appendix_A.Rmd
Tidy up everything. First version of Material and methods. Spell check. Description of data in Readme.
Änderungen von Vali übernommen.

@@
 
-# Methods {-}
+# Introduction {-}
 ```{r functions}
@@
 ```
-Temporal turnover is caused by species that are lost between surveys (or overseen in the second survey) and gained between surveys (or overseen during the first survey).
-It is likely that different drivers such as Nitrogen deposition or climate warming would cause differnt species to colonize or disappear from sites.
-Also methodological issus such as imperfect detection is likely to result in species with particular characteristics that are overseen in one survey and thus result apparend species gains or losses.
-We here borrowed ideas from functional ecology in order to try to disentangle the potential drivers of species gains and losses: We compared the characteristics of the species that newly colonized or the species that disappeared from a site with the same amount of species that were randomly selected from all the species that were observed at the site (i.e.
+Temporal turnover is caused by species that are lost between surveys (or are overseen in the second survey) and/or are gained between surveys (or overseen during the first survey).
+It is likely that different drivers such as Nitrogen deposition or climate warming would cause different species to colonize or disappear from sites.
+Also methodological issues such as imperfect detection are likely to result in species with particular characteristics that are overseen in one survey and thus result in apparent species gains or losses.
+We here borrowed ideas from functional ecology in order to try to disentangle the potential drivers of species gains and losses: We compared the characteristics of the species that newly colonized or that disappeared from a site with the same amount of species that were randomly selected from all the species that were observed at the site (i.e.
 the species assemblage).
-Instead of the functional traits that are traditinally used in functional ecology such as species height, we used the Elleberg values that we though could best reflect the main drivers that are likely to cause shifts in mountain hay plant communities.
-This were the Ellenberg values for temperature (is likely to respond to climate warming), humidity (is likely to respond to changes in precipitation), nutrients (is likely to change in response to Nitrogen deposition and land-use intensity) and light (is likely to change in response to land-use intensity).
+Instead of the functional traits that are traditionally used in functional ecology such as species height, we used the Ellenberg values that we assumed could best reflect the main drivers that are likely to cause shifts in mountain hay plant communities.
+This were the Ellenberg values for temperature (that are likely to respond to climate warming), humidity (likely to respond to changes in precipitation), nutrients (likely to change in response to Nitrogen deposition and land-use intensity) and light (likely to change in response to land-use intensity).
 
@@
 
-```{r Tplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{temperature} of lost (A) and newly appearing (B) species compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e.
+```{r Tplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{temperature} of lost (A) and newly appearing (B) species, compared to the same number of species that were randomly selected from all species recorded at the sampling site (i.e.
 random communities).
@@
 
-```{r Fplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{humidity} of lost (A) and newly appearing (B) species compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e.
+```{r Fplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{humidity} of lost (A) and newly appearing (B) species, compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e.
 random communities).
@@
 
-```{r Nplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{nutrients} of lost (A) and newly appearing (B) species compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e.
+```{r Nplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{nutrients} of lost (A) and newly appearing (B) species, compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e.
 random communities).
@@
 
-```{r Lplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{light} of lost (A) and newly appearing (B) species compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e.
+```{r Lplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{light} of lost (A) and newly appearing (B) species, compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e.
 random communities).

### 2018-09-13 Manuscript.Rmd
Änderungen von Vali übernommen.

@@
 abstract: >
-  Nitrogen (N) deposition is a major threat to biodiversity of many habitats.
-The recent introduction of cleaner technologies in Switzerland has let to reductions in the emissions of nitrogen oxides, with affiliated decrease in N deposition.
-We considered different drivers of plant community change (i.e.
-N deposition, climate warming, land-use change) in Swiss mountain hay meadows using data from the Swiss biodiversity monitoring.
-For this purpose, we compared indicator values of species that disappeared or colonized a site (turnover) with the indicator values of randomly chosen species from the same site.
-While oligotrophic plant species were more likely to colonize compared to random expectation, we found only weak shifts in plant community structure.
-Particularly, the average nutrient value of plant communities remained stable over time (2003–2017).
-Despite this, we found that the nutrient values of colonization species showed largest deviations from random expectation suggesting that N deposition or other factors that change the nutrient content of soils were important drivers of the species composition change over the last 15 years in Swiss mountain hay meadows.
-In addition, we observed an overall replacement of species with lower indicator values for temperature by species with higher values.
+  Nitrogen (N) deposition is a major threat to biodiversity in many habitats.
+The recent introduction of cleaner technologies in Switzerland has led to a reduction in the emissions of nitrogen oxides, with an according decrease in N deposition.
+We examined different drivers of plant community change, i.e.
+N deposition, climate warming, and land-use change, in Swiss mountain hay meadows, using data from the Swiss biodiversity monitoring.
+We compared indicator values of species that disappeared from or colonized a site (species turnover) with the indicator values of randomly chosen species from the same site.
+While oligotrophic plant species were more likely to colonize, compared to random expectation, we found only weak shifts in plant community structure.
+In particular, the average nutrient value of plant communities remained stable over time (2003‚Äì2017).
+However, we found that the nutrient values of colonizing species showed the largest deviations from random expectation, suggesting that N deposition or other factors that change the nutrient content of soils were important drivers of the species composition change over the last 15 years in Swiss mountain hay meadows.
+In addition, we observed an overall replacement of species with lower indicator values for temperature with species with higher values.
 Apparently, the community effects of the replacement of eutrophic species with oligotrophic species was outweighed by climate warming.
@@
 # Introduction {-}
-Nitrogen (N) deposition is the entry of reactive nitrogen compounds into soil, water and vegetation input from the atmosphere to the biosphere.
-Since nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low nitrogen availability [@Vitousek1997], the addition of nitrogen is likely to alter these communities.
-Indeed, N deposition together with land-use and climate change is one of the major threats to biodiversity [@Sala2000; @Bobbink2010; @Murphy2013].
+Nitrogen (N) deposition is the entry of reactive nitrogen compounds into soil, water, and vegetation, input from the atmosphere to the biosphere.
+Since nitrogen is an essential plant nutrient and many species-rich communities are adapted to conditions of low nitrogen availability [@Vitousek1997], the addition of nitrogen is likely to change these communities.
+Indeed, together with land-use and climate change, N deposition is one of the major threats to biodiversity [@Sala2000; @Bobbink2010; @Murphy2013].
 While there is strong evidence for the reduction in diversity of species-rich grasslands due to increased N deposition [@Stevens2006; @Dupre2010; @Maskell2010; @Wesche2012], mountain grasslands have received less attention [@Humbert2016].
 
-Since the late 1980s measures to reduce atmospheric pollution have successfully diminished emissions of nitrogen oxides, with affiliated decrease in N deposition in many parts of Europe [@Torseth2012; @Fowler2007].
+In many parts of Europe, measures to reduce atmospheric pollution have successfully reduced emissions of nitrogen oxides since the late 1980s, with an according decrease in N deposition [@Torseth2012; @Fowler2007].
 However, there are a number of factors that may prevent the recovery of plant communities that suffered from increased N deposition.
-First, N deposition is still high at many sites, since, in contrast to nitrogen oxides, the recent decrease of ammonia emissions was only minor.
+First, N deposition is still high at many sites, since in contrast to nitrogen oxides, there was on a minor recent decrease of ammonia emissions.
 Thus, even if N deposition is reduced, large areas might still be above the critical load [@Slootweeg2015; @RihmAchermann2016].
-Second, there is a potential that communities reach an alternative stable state after decades of increased N deposition and these species are unlikely to disappear subsequent to reduced N deposition [@Stevens2016].
-Therefore, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of existant plant communities.
+Second, there is a possibility that communities reach an alternative stable state after decades of increased N deposition and that the respective plant species are unlikely to disappear even if N deposition is reduced [@Stevens2016].
+Therefore, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of plant communities.
 
-Recovery of existant plant communities from high N deposition would imply that the state of communities measured at different points in time is improving over time (i.e.
+Recovery of existing plant communities after high N deposition would imply that the state of communities measured at different points in time is improving over time (i.e.
 improving biodiversity endpoints sensu @Rowe2017).
-Species richness, a biodiversity endpoint that can be assessed relatively simple and communicated easily, is often negatively related to nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
-Other metrics that are potentially more useful to reflect favourable changes can be derived from the traits of the species in a community.
+Species richness, a biodiversity endpoint that can be assessed and communicated relatively easily, is often negatively related to nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
+Other metrics that are potentially more useful to reflect favorable changes can be derived from the traits of the species in a community.
 In Europe, environmental preference of plants has often been expressed using indicator values assigned to each plant species [@Ellenberg1992; @Landolt2010].
-Examples of such metrics would be the number of oligotrophic species or the average indicator value of the species in a community [@Roth2013; @Rowe2017].
+Examples of such metrics would be the number of oligotrophic species, or the average indicator value of the species in a community [@Roth2013; @Rowe2017].
 However, the lack of a temporal trend in such biodiversity endpoints -- particularly in species richness -- does not necessarily mean that species composition remains unchanged.
 This is because immigration and extinction might be equally frequent and may cancel each other out [@Hillebrand2018].
-Thus a useful approach to understanding biodiversity change is through estimates of species turnover reflecting both colonization and local extinction [@Hillebrand2018], especially if colonization and local extinction is compared to random expectation [@Chase2011].
+Thus, a useful approach to understanding biodiversity change is through estimates of species turnover reflecting both colonization and local extinction [@Hillebrand2018], especially if colonization and local extinction is compared to random expectation [@Chase2011].
 
@@
 This applies in particular to the meadows in the alpine region, where meadows with a high plant diversity are also of agronomical importance [@Leiber2006].
-In mountain hay meadows, the spatial variation of species richness in vascular plants is negative correlated with N deposition [@Roth2013], suggesting that mountain grasslands suffered from increased N deposition.
-However, between 1990 and 2010 NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% [@MaasGrennfelt2016].
-What may have led to a partial recovery of plant communities.
+In mountain hay meadows, the spatial variation of species richness in vascular plants is negatively correlated with N deposition [@Roth2013], suggesting that mountain grasslands are negatively affected by increased N deposition.
+However, between 1990 and 2010, NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% [@MaasGrennfelt2016].
+Potentially, this could have led to a partial recovery of plant communities.
 
-In addition to N deposition, Swiss mountain ecosystems are threatened also by other global change drivers.
+In addition to N deposition, Swiss mountain ecosystems are threatened also by other drivers of global change.
 In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35$^{\circ}$C per decade, which is about 1.6 times the northern hemispheric warming rate [@Ceppi2012].
-This climate warming is likely to directly affect or interact with N deposition to drive plant community changes [@Humbert2016].
-Indeed, in an earlier study we found that at the landscape scale plant communities responded to climate warming even within a comparably short time period [@Roth2014].
-Particularly the shift of plant communities at mountain summits are the result of recent climate warming and the aothors assume an interacting with airborne N deposition [@Steinbauer2018].
+This climate warming is likely to to interact with N deposition to drive plant community changes [@Humbert2016].
+Indeed, in an earlier study we found that, at the landscape scale, plant communities responded to climate warming even within a relatively short time period [@Roth2014].
+[@Steinbauer2018] say that particularly the shift of plant communities at mountain summits are the result of recent climate warming, and they assume an interaction with airborne N deposition.
 Furthermore, traditional management regimes are currently changing, which also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
-Management regimes of accessible mountainous areas are frequently intensified, while poorly accessible mountainous areas are being abandoned [@Tasser2002; @Strebel2015].
+Management regimes of easily accessible mountainous areas are often being intensified, while poorly accessible mountainous areas are abandoned [@Tasser2002; @Strebel2015].
 
-Here, we used data from the Swiss biodiversity monitoring [@Weber2004] to address the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, N deposition or land-use intensity change over the last 15 years, (2) was species turnover correlated with the average temperature, precipitation, N deposition or inclination (we expect steep areas to be less intensively managed), and (3) did species that newly colonized or disappeared (local extinction) from sites differ from random expectation according to their indicator values for temperature, humidity, nutrients or light.
+Here, we used data from the Swiss biodiversity monitoring [@Weber2004] to address the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, N deposition, or land-use intensity change over the last 15 years?
+(2) Was species turnover correlated with the average temperature, precipitation, N deposition, or inclination (we expect steep areas to be less intensively managed)?
+(3) Did species that newly colonized or disappeared from (local extinction) sites differ from random expectation, according to their indicator values for temperature, humidity, nutrients, or light?
 
@@
 ## Monitoring data and community measures {-}
-We analysed the presence/absence of vascular plants sampled within the scope of Switzerland's Biodiversity Monitoring (BDM) programme that was launched in 2001 to monitor Switzerland's biodiversity and to comply with the Convention on Biological Diversity of Rio de Janeiro [@Weber2004].
-The sampling sites were circles with a size of 10 m$^2$ and data collection was carried out by qualified botanists who visited each sampling site twice within the same season.
-During each visit all the vascular plant species detected on the plot were recorded.
+We analyzed the presence/absence of vascular plants sampled in the Swiss Biodiversity Monitoring (BDM) program that was launched in 2001 to monitor Switzerland's biodiversity and to comply with the Convention on Biological Diversity of Rio de Janeiro [@Weber2004].
+The sampling sites were circles with a size of 10 m$^2$, and data collection was carried out by qualified botanists who visited each sampling site twice within the same season.
+During each visit, all vascular plant species detected on the plot were recorded.
 For details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
 
-After the sampling of the plant data the botanists also assigned a habitat type to each sampling site according to the classification system developed for Switzerland [@Delarze2008].
+After the sampling of the plant data, the botanists also assigned a habitat type to each sampling site according to the classification system developed for Switzerland [@Delarze2008].
 We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all sampling sites in mountain hay meadows (EUNIS E2.3).
-We analysed the data from 2003 to 2017.
-During that study period each sampling site was surveyed once per five-year period: the first period lasted from 2003 to 2007, the second from 2008 to 2012 and the third from 2013 to 2017.
+We analyzed the data from 2003 to 2017.
+During that study period, each sampling site was surveyed once per five-year period: the first period lasted from 2003 to 2007, the second from 2008 to 2012, and the third from 2013 to 2017.
 These selection criteria yielded a sample of `r nrow(sites)` sites.
 
-For each survey (a survey consists of two  visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for (2) temperature, (3) humidity, (4) nutrients and (5) light.
+For each survey (that consisted of two visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for (2) temperature, (3) humidity, (4) nutrients, and (5) light.
 The indicator values were obtained from the recalibrated indicator values for the Swiss Flora [@Landolt2010].
 In addition to these five biodiversity endpoints that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
-species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points to describe the community change between two points in time.
+species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points, to describe the community change between two points in time.
 
@@
 ## Environmental gradients {-}
-We expected different global change patterns to simultaneously cause temporal change in mountain hay communities.
-To better disentangle the importance of these mechanisms we ordered the sites along four main environmental gradients.
-First, we expected communities to response to climate warming with this response being dependent on the position of the site along the temperature gradient [@Roth2014].
+We expected different drivers of global change to cause temporal change in mountain hay communities.
+To better disentangle the importance of these mechanisms, we ordered the sites along four main environmental gradients.
+First, we expected communities to response to climate warming [@Roth2014].
 To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
@@
  From the same database we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change [@Beier2012].The average annual precipitation at our sites was `r mean(sites$PrecipitationR)` $\pm$ `r sd(sites$PrecipitationR)` mm.
- Further, we estimated atmospheric N deposition for each site using a pragmatic approach  that combines monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models [@RihmAchermann2016].
+ Further, we estimated atmospheric N deposition for each site using a pragmatic approach that combined monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models [@RihmAchermann2016].
 The average nitrogen deposition at our sites was `r mean(sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ in 2000 and `r mean(sites$NTOT_2015)` $\pm$ `r sd(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$ in 2015.
-Finally, we used inclination as proxy for land-use intensity because we expected that steeper sites are likely to be less intensively managed [@Strebel2015].
+Finally, we used inclination as a proxy for land-use intensity, because we assumed that steeper sites are likely to be less intensively managed [@Strebel2015].
 The average inclination at our sites was `r mean(sites$InclinationR)` $\pm$ `r sd(sites$InclinationR)`$^\circ$.
 ## Statistical analyses {-}
-To estimate the linear trend over time for each of the five biodiversity endpoints we applied linear mixed models (LMM) with normal distribution except for species-richness with Poisson distribution.
-We specified site specific trends with the assumption that the between-site differences in intercepts and slopes can be described with normal distributions [i.e.
+To estimate the linear trend over time for each of the five biodiversity endpoints, we applied linear mixed models (LMM) with normal distribution except for species-richness with Poisson distribution.
+We specified site-specific trends with the assumption that the between-site differences in intercepts and slopes can be described with normal distributions [i.e.
 a random intercept random slope model, @Gelman2006].
@@
 
-To infer whether species turnover was changing along the gradient, we used a Binomial-GLMM with the proportion of species that differed between two surveys as dependent variable and the site gradients and period (first/second vs.
-second/third surveys) as predictors and site-ID as random effect.
+To infer whether species turnover was changing along the gradient, we used a binomial generalized linear mixed model (GLMM) with the proportion of species that differed between two surveys as dependent variable and the site gradients and period (first/second vs.
+second/third surveys) as predictors, and with site-ID as random effect.
 Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
@@
 We then calculated the community mean (CM) of the indicator value for all species that colonized the site during the three surveys (i.e.
-not recorded during first survey and recorded during second or not recorded during second and recorded during third).
+not recorded during the first survey and recorded during the second, or not recorded during the second and recorded during the third survey).
 We then randomly selected the same number of species from the total community and also calculated the community mean of the value for these species (random-CM).
 We repeated the random selection of species 1000 times.
-Then we calculated the differences of the CM minus the average of the random-CMs to obtain a standardized difference (standardized-CM) of how different the colonizing species were from random expectation.
-For example, a difference < 0 would suggest that the indicator values of colonizing species were lower than might be expected from random colonisation from the species-pool for this site.
-We applied this method for both colonizing and disappearing species and for the indicator values for temperature, humidity, nutrients and light (see Appendix A).
-We then tested whether this standardized difference is changing along the corresponding gradient (inferred from independet datasets, see Environmental gradients section) using linear models.
+We then calculated the differences of the CM minus the average of the random-CMs to obtain a standardized difference (standardized-CM) of how different the colonizing species were from random expectation.
+For example, a difference < 0 would suggest that the indicator values of colonizing species were lower than might be expected from random colonization from the species-pool for this site.
+We applied this method for both colonizing and disappearing species and for the indicator values for temperature, humidity, nutrients, and light (see Appendix A).
+We then tested whether this standardized difference was changing along the corresponding gradient (inferred from independent datasets, see Environmental gradients section) using linear models.
 Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
 
-We used logistic-GLMM to analyse whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
-To analyse the colonization probability of species, we selected all species that were not observed during the first survey.
-For each of these species we asked if they were observed (then Occ = 1) or not observed (then Occ = 0) during the second survey.
-The same was also done for all species that were not observed during the second survey if they were observed or not observed (then Occ = 0) during the third survey.
-The variable 'occ' was then used as dependent variable in the logistic-GLMM.
-As predictor variables the model contained the N deposition of the site, the indicator value for nutrients of the species and the interaction of these two variables.
+We used logistic GLMM to analyze whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
+To analyze the colonization probability of species, we selected all species that were not observed during the first survey.
+For each of these species we asked if they were observed (Occ = 1) or not observed (Occ = 0) during the second survey.
+The same was also done for all species that were not observed during the second survey if they were observed or not observed during the third survey.
+The variable 'occ' was then used as dependent variable in the logistic GLMM.
+As predictor variables, the model contained the N deposition of the site, the indicator value for nutrients of the species, and the interaction of these two variables.
 Additionally, species-ID and site-ID were included as random effects.
-The same logistic-GLMM was also used to investigate local survival probabilities.
-In that case, however, we selected all species that were recorded during the first or second survey and the variable 'occ' then indicated whether or not the species was also observed during the next survey.
+The same logistic GLMM was also used to investigate local survival probabilities.
+In that case, however, we selected all species that were recorded during the first or second survey, and the variable 'occ' then indicated whether or not the species was also observed during the next survey.
 Model parameters were estimated using an approximate Bayesian approach using the R-Package *arm* [@arm2018].
 
-To estimate the effect of N deposition on total species richness at a given point in time, we described the plant species richness at the sites using a generalized linear model with Poisson distribution and the logarithm as link function.
+To estimate the effect of N deposition on total species richness at a given point in time, we described the plant species richness at the sites using a generalized linear model (GLM) with Poisson distribution and the logarithm as link function.
 As predictors we used the four environmental gradients as described above.
@@
 ```{r communitytrendscalc}
-res <- tibble(Measures = c("Species-richness", "Temperature value", "Humidity value", "Nutrients value", "Light value"),
+res <- tibble(Measures = c("Species richness", "Temperature value", "Humidity value", "Nutrients value", "Light value"),
               'Period 1' = NA,
@@
 
-```{r communitytrendstab, echo=FALSE}
+```{r communitytrendstab}
 options(knitr.kable.NA = '') 
@@
 The temporal trends are given as change per 10 years and were estimated from linear mixed models with normal distribution (except for species richness with Poisson distribution and a log-link function).
-The measure of precision for the temporal trend is given as the 5\\% and 95\\% quantiles of the marginal posterior distribution of the linear trend.
+The measure of precision for the temporal trend is given as the 5\\% and 95\\% quantiles of the marginal posterior distribution of the linear trend (90\\% credible intervals).
 The column 'Prob.
@@
 
-Species richness and four measures of plant community structure according to Landolt indicator values (i.e.
-biodiversity endpoints) suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods the averages of species-richness and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods and the estimated trends were rather small.
-Except for average indicator value for temperature, the 90% credible-interval of the temporal trend contained zero.
+Species richness and the four measures of plant community structure according to Landolt indicator values (i.e.
+biodiversity endpoints) suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods, the averages of species richness and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods, and the estimated trends were rather small.
 The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the indicator value for temperature (probability of increase: `r
@@
 for Trend")` (`r res %>% filter(Measures == "Nutrients value") %>% pull("Prob.
-for Trend")`gives the probability for an increase).
+for Trend")` gives the probability for an increase).
 
@@
 The slopes along the gradients (estimate) are given as the change per 10 years of the logit-probability of species that differed between two surveys.
-Estimates and the 5\\% and 95\\% quantiles of the marginal posterior distribution obtained from a Binomial-GLMM.")
+Estimates and 5\\% and 95\\% quantiles of the marginal posterior distribution obtained from a binomial GLMM.")
 ```
@@
 The average $\pm$ SD percentage of species that differed between replicated surveys was `r mean(100 * turnover_double$turnover)` $\pm$ `r sd(100 * turnover_double$turnover)`%.
-This pseud-turnover was lower than the observed temporal turnover: the average percentage of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`% and the percentage of species that differ between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
-Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% Credible interval of the change in turnover estimated from the Binomial generalized linear mixed model: `r CrI`).
-Along the N deposition gradient variation in species turnover was largest, with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
+This turnover between replicated surveys was lower than the observed temporal turnover: the average percentage of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`%, and the percentage of species that differed between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
+Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% credible interval of the change in turnover estimated from the binomial GLMM: `r CrI`).
+Along the N deposition gradient, variation in species turnover was largest, with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
 The other three gradients were less conclusive to explain the variation in species turnover among sites.
@@
 ```{r difffromrandomtab}
-kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Indicator value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average indicator value of species that (a) disappeared from site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
-Shown are the results from linear model with the difference between disappeared/colonized species and random species as dependent variable and the site-measure (gradient) as predictor variable.") %>%
+kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Indicator value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average indicator value of species that (a) disappeared from a site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
+Shown are the results from linear models, with the difference between disappeared/colonized species and random species as dependent variable and the site-measure (gradient) as predictor variable.") %>%
   kable_styling() %>% 
@@
 
-High species turnover at a site is the result of species that disappeared from the site and species that newly colonized the site.
+High species turnover at a site is the result of species that disappeared from the site and/or of species that newly colonized the site.
 To better understand the factors that drive these changes we were particularly interested whether the species that disappeared or colonized the sites differed in indicator values compared to what would be expected if the same number of species randomly disappeared or colonized the sites (i.e.
-random disappearance and random colonization) and whether there is a spatial variation along the environmental gradients.
+random disappearance and random colonization), and whether there is spatial variation along the environmental gradients.
 It seems that the indicator values of newly colonizing species differed more from random colonization than the indicator values of disappearing species (Table \@ref(tab:difffromrandomtab)).
-For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with relatively low N deposition of 10 kg ha$^{-1}$ yr$^{-1}$ the newly colonizing species had in average a lower indicator value for nutrients than species under random colonization (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but this differences between colonizing species and random colonization decreased with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
-Thus, at high N deposition colonizing species did not differ from random species (see Figure 3b in Appendix A).
+For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with relatively low N deposition of 10 kg ha$^{-1}$ yr$^{-1}$, the newly colonizing species had on average a lower indicator value for nutrients than species under random colonization (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but the differences between colonizing species and random colonization decreased with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
+Thus, at high N deposition, colonizing species did not differ from random species (see Figure 3b in Appendix A).
 
-While colonizing species had higher temperature values compared to what we would expect under random colonization, the differences between colonizing species and random colonization was about four times smaller than the difference in indicator value for nutrients between colonizing species and random species.
-Nevertheless, the variation in indicator value for temperature is revealed to be important to explain the total species turnover.
-This is because disappearing species tend to have lower temperature values than random species as well as colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
-This was not the case for the indicator value for nutrients: species with lower nutrients values tended to be more likely to disappear from as well as to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
-See also Appendix A where we present detailed results for the comparison between colonizing or disappearing species with randomly selected species.
+While colonizing species had higher temperature values compared to what we would expect under random colonization, the differences between colonizing species and random colonization was about four times smaller than the difference in the indicator value for nutrients between colonizing species and random species.
+Nevertheless, the variation in the indicator value for temperature seems to be important for explaining the total species turnover.
+This is because disappearing species tend to have lower temperature values than random species, and colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
+This was not the case for the indicator value for nutrients: species with lower nutrients values tended to be more likely to disappear from, and to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
+See also Appendix A, where we present detailed results for the comparison between colonizing or disappearing species with randomly selected species.
 
 ## Potential effects of reduction in Nitrogen emmissions {-}
-```{r Ndepchange, fig.cap="(a) The nitrogen (N) deposition in 2000 and (b) the change in N deposition between 2000 and 2015 along the N deposition gradient of the study sites in 2000."}
+```{r Ndepchange, fig.cap="(a) The nitrogen (N) deposition in 2000 and (b) the change in N deposition between 2000 and 2015, along the N deposition gradient of the study sites used in 2000."}
 p1 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = NTOT_2000)) +
@@
 \@ref(fig:Ndepchange)a).
-In 2000 only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a N deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, all situated above 1000m.
-Between 2000 and 2015 the N deposition decreased in average by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$ with slightly higher net decreases at lower elevation (Fig.
+In 2000, only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a N deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, all situated above 1000 m.
+Between 2000 and 2015, the N deposition decreased on average by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$, with slightly higher net decreases at lower elevation (Fig.
 \@ref(fig:Ndepchange)b).
 
-```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic species (indicator value for nutrients $<3$; red line) and eutrophic species (indicator value for nutrients $>3$) species along the N deposition gradient.
-Given are means and 95\\%-Credible Intervals from logistic linear mixed models.
-The vertical lines indicate the deposition rate with equal colonization or survival probabilities for oligotrophic and eutrophic species with the solid line indicating the median and the dashed lines the 5\\% and 95\\% quantiles of the marginal posterior distribution."}
+```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic species (indicator value for nutrients $<3$; red line) and of eutrophic species (indicator value for nutrients $>3$) species along the N deposition gradient.
+Given are means and 95\\%-credible intervals from logistic linear mixed models.
+The vertical lines indicate the deposition rate, with equal colonization or survival probabilities for oligotrophic and eutrophic species, the solid line indicating the median, and the dashed lines the 5\\% and 95\\% quantiles of the marginal posterior distribution."}
 X <- 1:30
@@
 Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for oligotrophic and eutrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
-In `r 100 * mean((sites$NTOT * 10 + 10) < 12.5)`% of the sites the deposition rate was below 12.5 kg N ha$^{-1}$ yr$^{-1}$ where the replacement of eutrophic with oligotrophic species is likely.
+In `r 100 * mean((sites$NTOT * 10 + 10) < 12.5)`% of the sites, the deposition rate was below 12.5 kg N ha$^{-1}$ yr$^{-1}$ at which the replacement of eutrophic with oligotrophic species is likely according to Fig.
+\@ref(fig:cl).
 
-While we could not detect a consistent decrease in average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at the few sites with low N deposition rate seems to affect the spatial pattern of oligotrophic species richness: sites with low N deposition are likely to become more species rich over time.
-This likely results in steeper slope of the negative relationship between N deposition and oligotrophic species richness, when comparing this relationshep with a spatial approach as in @Roth2013.
-Indeed, if we apply such an approach at different points in time to infer the effects of N deposition on the spatial variation of oligotrophic species richness, the resulting effect size (i.e.
-the slope) became more negative over time (Fig.
+While we could not detect a consistent decrease in the average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at the few sites with low N deposition rate seems to affect the spatial pattern of oligotrophic species richness: sites with low N deposition are likely to become more species-rich over time.
+This likely results in a steeper slope of the negative relationship between N deposition and oligotrophic species richness, when comparing this relationship with a spatial approach as in @Roth2013.
+Indeed, if we apply such an approach at different points in time to infer the effects of N deposition on the spatial variation of oligotrophic species richness, the resulting effect size (i.e., the slope) became more negative over time (Fig.
 \@ref(fig:figconsequences)).
 
-```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of N deposition on oligotrophic species richness estimated from applying the Poisson-GLM with species richness as dependent variable and N deposition plus other site covariates (elevation, precipitation, inclination, mean indicator values for humidity and light) as predictors using only the surveys from one five-year interval.
-Note that within every five-year interval all plots were sampled once."}
+```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of N deposition on oligotrophic species richness estimated from applying the Poisson GLM, with species richness as dependent variable and N deposition plus other site covariates (elevation, precipitation, inclination, mean indicator values for humidity and light) as predictors, using only the surveys from one five-year interval.
+Note that within every five-year interval, all plots were sampled once."}
 # Add site covariates to 'surv'
@@
 biodiversity endpoints sensu @Rowe2017) during the same time period (Table \@ref(tab:communitytrendstab)).
-While the slight increase in average temperature indicator values suggests that plant communities adopted to increasing temperatures, the constant average nutrients value suggests that the decrease in N deposition did not yet affect plant communities.
-However, this apparent stability in community composition  was accompanied by marked temporal turnover in species identities.
+While the slight increase in average temperature indicator values suggests that plant communities adopted to increasing temperatures, the constant average nutrient value suggests that the decrease in N deposition did not yet affect plant communities.
+However, this apparent stability in community composition was accompanied by a marked temporal turnover in species identities.
 It seems unlikely that this temporal turnover can entirely be explained by methodological issues such as overlooked species.
@@
 Second, spatial variation of turnover showed patterns that can hardly be explained by methodological issues.
-For instance species turnover varied along the N deposition gradient with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
+For instance, species turnover varied along the N deposition gradient, with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
 Taken together, our results add to the increasing evidence that contemporary communities are often relatively stable at community level, but this apparent stability is often accompanied by marked turnover of species [@Vellend2013; @Dornelas2014; @Hillebrand2018].
@@
 We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site (Appendix A).
-We found that the nutrient values of colonizing species showed largest deviations from random expectation (Table  \@ref(tab:difffromrandomtab)) suggesting that N deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
+We found that the nutrient values of colonizing species showed the largest deviations from random expectation (Table  \@ref(tab:difffromrandomtab)), suggesting that N deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
 
-Nitrogen deposition at our plots decreased by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ on average between 2000 and 2015.
+Nitrogen deposition at our plots decreased, on average, by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ between 2000 and 2015.
 This is only about one tenth of the decrease in England, where N deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
@@
 Additionally, other anthropogenic pressures such as climate change might have overruled effects of N depositon on community composition.
-Particularly, we found that the species disappearing from the sites tend to have below average indicator values for temperature, while species that newly colonize sites have above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
+In particular, we found that the species disappearing from the sites tended to have below average indicator values for temperature, while species that newly colonize sites had above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
 Thus, the effect of disappearing and the effect of colonizing species on the community mean for temperature is additive resulting in increasing average temperature values (Tab.
@@
 This is in contrast to how N deposition is affecting disappearing and colonization of species: It seems that both species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
-Thus, the effect of disappearing and colonizing  species on the average community value for nutrients partially cancel each other out.
+Thus, the effects of disappearing and colonizing species on the average community value for nutrients partially cancel each other out.
 Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at higher N deposition rates.
-This suggests that mountain hay meadow communities can reach alternative stable states with eutrophic species that are unlikely to disappear even if N deposition is reduced [@Stevens2016].
+This suggests that mountain hay meadow communities can reach alternative stable states, with eutrophic species that are unlikely to disappear even if N deposition is reduced [@Stevens2016].
 Taken together, these reasons might explain why the composition of mountain hay meadow communities responded stronger to climate warming than to nitrogen reduction, although the reduction in nitrogen resulted in above average colonization of oligotrophic species.
@@
 Observational studies along a gradient of N deposition often conclude how the spatial variation in species richness is related to N deposition [@Stevens2010gradient; @Roth2013; chapter 4 in @deVries2015].
-They assume that the spatial variation in species richness (or other metrics of community composition) arose because of unequal species loss of different areas over time resulting from elevated N deposition chronically experienced by some areas.
-Although there is evidence supporting the pertinency of such a 'space for time substitution' for detecting the effects of N deposition on plant diversity [@Stevens2010], this approach can not replace studies that relate temporal patterns in species composition with N deposition [@deSchrijver2011].
-There are only a limited number of studies directly relating temporal trends of plant species diversity to varying amounts of N deposition in existant communities [@Clark2008; @Storkey2015; @Stevens2016].
-In a recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the N deposition gradient at which species richness of oligotrophic species richness start to decrease with increasing N deposition [@Roth2017].
-Using this spatial variation in species richness and N deposition we estimated a critical load for mountain hay meadows of 13.1 kg ha$^{-1}$ yr$^{-1}$.
-In the current study we estimated the rate of N deposition at which local survival probability or colonization probability was equal for oligotrophic and eutrophic species.
+They assume that the spatial variation in species richness (or other metrics of community composition) arose because of unequal species loss of different areas over time, resulting from elevated N deposition chronically experienced by some areas.
+Although there is evidence supporting the pertinency of such a 'space for time substitution' for detecting the effects of N deposition on plant diversity [@Stevens2010], this approach cannot replace studies that relate temporal patterns in species composition with N deposition [@deSchrijver2011].
+There are only a limited number of studies directly relating temporal trends of plant species diversity to varying amounts of N deposition in existing communities [@Clark2008; @Storkey2015; @Stevens2016].
+In a recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the N deposition gradient at which species richness of oligotrophic species richness starts to decrease with increasing N deposition [@Roth2017].
+Using this spatial variation in species richness and N deposition, we estimated a critical load for mountain hay meadows of 13.1 kg ha$^{-1}$ yr$^{-1}$.
+In the current study, we estimated the rate of N deposition at which local survival probability or colonization probability was equal for oligotrophic and eutrophic species.
 This rate could also be interpreted as a critical load.
-With this temporal approach we obtained very similar results as the spatial approach.
+With this temporal approach we obtained very similar results as with the spatial approach.
 Our results are thus validating the space for time approach, at least for Swiss mountain hay meadows.
@@
 When we applied the same model as in Roth et al.
-(2017) for different study periods to infer how oligotrophic species richness is decreasing along the N deposition gradient, the relationship showed to vary between study periods.
+(2017) for different study periods, to infer how oligotrophic species richness is decreasing along the N deposition gradient, the relationship appeared to vary between study periods.
 The decrease gradually became steeper (more negative) over time, except for the first two study periods.
 Our first interpretation was that the N deposition effect became stronger over time.
-This was against our prediction that the effect of N deposition should become weaker over time since N deposition is decreasing during the study period.
+This was against our prediction that the effect of N deposition should become weaker over time, since N deposition is decreasing during the study period.
 Then we realized that species turnover was highest at low N deposition sites (Tab.
 \@ref(tab:turnovertab)).
-At low N deposition rates colonizing species have below average indicator values for nutrients.
-It seems that the decrease in N deposition resulted in oligotrophic species replacing eutrophic species especially at sites with low N deposition.
-This explains why the decline in oligotrophic species richness inferred from spatial patterns of species richness and N deposition is becoming steeper over time.
-This may be interpreted as evidence that plant communities are recovering at least at low deposition sites rather than inferring that the negative N deposition effects have become stronger over time.
+At low N deposition rates, colonizing species have below average indicator values for nutrients.
+It seems that the decrease in N deposition resulted in oligotrophic species replacing eutrophic species particularly at sites with low N deposition.
+This seems to explain why the decline in oligotrophic species richness inferred from spatial patterns of species richness and N deposition is becoming steeper over time.
+And this may be interpreted as evidence that plant communities are recovering at least at low deposition sites, and that the negative N deposition effects have not become stronger over time.
 
 # Conclusions {-}
-  Comparing the indicator values of colonizing and disappearing species with random expectation we found that oligotrophic species are currently more likely to colonize mountain hay meadows than eutrophic species, which might be the result of decreasing atmospheric N deposition recently observed.
+  Comparing the indicator values of colonizing and disappearing species with random expectation, we found that oligotrophic species are currently more likely to colonize mountain hay meadows than eutrophic species, which might be the result of the recently observed decrease in atmospheric N deposition.
 However, our results also indicate that re-transferring mountain hay meadows to oligotrophic species communities after high N deposition might take much longer than transferring an oligotrophic into an eutrophic community.
-This is because eutrophic species have high local survival probabilities even after N deposition decreases again.
-However, these results are hidden if only the temporal trends of species composition are analysed instead of species turnover, which can better reveal ongoing processes in this case.
-Our study thus adds to the understanding of contemporary biodiversity change where some key findings are still missing [@Magurran2018], and it supports the notion of @Hillebrand2018 that consicering species turnover will generate far more reliable view of the biotic response to changing environments than solely tracking community composition.
+This is because eutrophic species have high local survival probabilities, even after N deposition decreases again.
+However, these results would be hidden if only the temporal trends of species composition were analyzed, instead of species turnover.
+Our study thus adds to the understanding of contemporary biodiversity change, where some key information is still missing [@Magurran2018], and it supports the notion of @Hillebrand2018 that considering species turnover will generate a far more reliable view of the biotic response to changing environments than solely tracking community composition.
 

### 2018-09-13 Manuscript.Rmd
Kleiner Anpassungen im Text.

@@
 While oligotrophic plant species were more likely to colonize, compared to random expectation, we found only weak shifts in plant community structure.
-In particular, the average nutrient value of plant communities remained stable over time (2003‚Äì2017).
+In particular, the average nutrient value of plant communities remained stable over time (2003-2017).
 However, we found that the nutrient values of colonizing species showed the largest deviations from random expectation, suggesting that N deposition or other factors that change the nutrient content of soils were important drivers of the species composition change over the last 15 years in Swiss mountain hay meadows.
@@
 However, there are a number of factors that may prevent the recovery of plant communities that suffered from increased N deposition.
-First, N deposition is still high at many sites, since in contrast to nitrogen oxides, there was on a minor recent decrease of ammonia emissions.
+First, N deposition is still high at many sites, since in contrast to nitrogen oxides, there was only a minor recent decrease of ammonia emissions.
 Thus, even if N deposition is reduced, large areas might still be above the critical load [@Slootweeg2015; @RihmAchermann2016].
@@
 Indeed, in an earlier study we found that, at the landscape scale, plant communities responded to climate warming even within a relatively short time period [@Roth2014].
-[@Steinbauer2018] say that particularly the shift of plant communities at mountain summits are the result of recent climate warming, and they assume an interaction with airborne N deposition.
+@Steinbauer2018 say that particularly the shift of plant communities at mountain summits are the result of recent climate warming, and they assume an interaction with airborne N deposition.
 Furthermore, traditional management regimes are currently changing, which also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
@@
 The average inclination at our sites was `r mean(sites$InclinationR)` $\pm$ `r sd(sites$InclinationR)`$^\circ$.
+
 ## Statistical analyses {-}
@@
 biodiversity endpoints) suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods, the averages of species richness and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods, and the estimated trends were rather small.
-The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the indicator value for temperature (probability of increase: `r
-res %>% filter(Measures == "Temperature value") %>% pull("Prob.
+The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the indicator value for temperature (probability of increase: `r res %>% filter(Measures == "Temperature value") %>% pull("Prob.
 for Trend")`), followed by the community mean of the indicator value for light (probability of decrease: `r 1 - res %>% filter(Measures == "Light value") %>% pull("Prob.
 for Trend")`).
-A linear temporal change was least likely for the species richness (probability of increase: `r res %>% filter(Measures == "Species-richness") %>% pull("Prob.
+A linear temporal change was least likely for the species richness (probability of increase: `r res %>% filter(Measures == "Species richness") %>% pull("Prob.
 for Trend")`).

### 2018-10-19 Appendix_A.Rmd
Revise Manuscript according to reviewer comments. Most of comments including. Stuff for the discussion still missing.

@@
 Instead of the functional traits that are traditionally used in functional ecology such as species height, we used the Ellenberg values that we assumed could best reflect the main drivers that are likely to cause shifts in mountain hay plant communities.
-This were the Ellenberg values for temperature (that are likely to respond to climate warming), humidity (likely to respond to changes in precipitation), nutrients (likely to change in response to Nitrogen deposition and land-use intensity) and light (likely to change in response to land-use intensity).
+This were the Ellenberg values for temperature (that are likely to respond to climate warming), soil moisture (likely to respond to changes in precipitation), nutrients (likely to change in response to Nitrogen deposition and land-use intensity) and light (likely to change in response to land-use intensity).
 
@@
 
-```{r Fplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{humidity} of lost (A) and newly appearing (B) species, compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e.
+```{r Fplot, fig.cap="\\textit{Average Ellenberg value for \\textbf{soil moisture} of lost (A) and newly appearing (B) species, compared to the same number of species that were randomly selected from all species recorded at the sample site (i.e.
 random communities).
@@
   scale_y_continuous(breaks = tylab, labels = tylab, limits = c(2, 4)) +
-  labs(x = "Annual mean precipitation", y = "Ellenberg humidity value\n(community mean)",
+  labs(x = "Annual mean precipitation", y = "Ellenberg soil moisture value\n(community mean)",
        title = "(A) Species losses") 
@@
   scale_y_continuous(breaks = tylab, labels = tylab, limits = c(2, 4)) +
-  labs(x = "Annual mean precipitation", y = "Ellenberg humidity value\n(community mean)",
+  labs(x = "Annual mean precipitation", y = "Ellenberg soil moisture value\n(community mean)",
        title = "(B) Species gains") 

### 2018-10-19 Manuscript.Rmd
Revise Manuscript according to reviewer comments. Most of comments including. Stuff for the discussion still missing.

@@
   Nitrogen (N) deposition is a major threat to biodiversity in many habitats.
-The recent introduction of cleaner technologies in Switzerland has led to a reduction in the emissions of nitrogen oxides, with an according decrease in N deposition.
+The recent introduction of cleaner technologies in Switzerland has led to a reduction in the emissions of nitrogen oxides, with a consequent decrease in N deposition.
 We examined different drivers of plant community change, i.e.
-N deposition, climate warming, and land-use change, in Swiss mountain hay meadows, using data from the Swiss biodiversity monitoring.
+N deposition, climate warming, and land-use change, in Swiss mountain hay meadows, using data from the Swiss biodiversity monitoring program.
 We compared indicator values of species that disappeared from or colonized a site (species turnover) with the indicator values of randomly chosen species from the same site.
-While oligotrophic plant species were more likely to colonize, compared to random expectation, we found only weak shifts in plant community structure.
+While oligotrophic plant species were more likely to colonize, compared to random expectation, we found only weak shifts in plant community composition.
 In particular, the average nutrient value of plant communities remained stable over time (2003-2017).
@@
 First, N deposition is still high at many sites, since in contrast to nitrogen oxides, there was only a minor recent decrease of ammonia emissions.
-Thus, even if N deposition is reduced, large areas might still be above the critical load [@Slootweeg2015; @RihmAchermann2016].
+Thus, even if N deposition is reduced, large areas might still be above the critical threshold above which harmful effects on plant diversity do occur [i.e.
+above the critical load for N deposition; @Bobbink2010; @Slootweeg2015; @RihmAchermann2016].
 Second, there is a possibility that communities reach an alternative stable state after decades of increased N deposition and that the respective plant species are unlikely to disappear even if N deposition is reduced [@Stevens2016].
@@
 This is because immigration and extinction might be equally frequent and may cancel each other out [@Hillebrand2018].
-Thus, a useful approach to understanding biodiversity change is through estimates of species turnover reflecting both colonization and local extinction [@Hillebrand2018], especially if colonization and local extinction is compared to random expectation [@Chase2011].
+Thus, a useful approach to understanding biodiversity change is through estimates of species turnover reflecting both colonization and local extinction [@Hillebrand2018], especially if colonization and local extinction are compared to random expectation [@Chase2011].
 
@@
 
-In addition to N deposition, Swiss mountain ecosystems are threatened also by other drivers of global change.
+In addition to N deposition, Swiss mountain ecosystems are also threatened by other drivers of global change.
 In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35$^{\circ}$C per decade, which is about 1.6 times the northern hemispheric warming rate [@Ceppi2012].
-This climate warming is likely to to interact with N deposition to drive plant community changes [@Humbert2016].
+This climate warming is likely to interact with N deposition to drive plant community changes [@Humbert2016].
 Indeed, in an earlier study we found that, at the landscape scale, plant communities responded to climate warming even within a relatively short time period [@Roth2014].
-@Steinbauer2018 say that particularly the shift of plant communities at mountain summits are the result of recent climate warming, and they assume an interaction with airborne N deposition.
+@Steinbauer2018 suggest that particularly the shift of plant communities at mountain summits is the result of recent climate warming, and they assume an interaction with airborne N deposition.
 Furthermore, traditional management regimes are currently changing, which also has major impacts on plant communities in mountainous regions of Europe [@Niedrist2009; @Homburger2012].
@@
 
-Here, we used data from the Swiss biodiversity monitoring [@Weber2004] to address the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, N deposition, or land-use intensity change over the last 15 years?
+Here, we used data from the Swiss biodiversity monitoring program [@Weber2004] to address the following questions: (1) Did biodiversity endpoints that are likely to reflect temperature, precipitation, N deposition, or land-use intensity change over the last 15 years?
 (2) Was species turnover correlated with the average temperature, precipitation, N deposition, or inclination (we expect steep areas to be less intensively managed)?
-(3) Did species that newly colonized or disappeared from (local extinction) sites differ from random expectation, according to their indicator values for temperature, humidity, nutrients, or light?
+(3) Did species that newly colonized or disappeared from (local extinction) sites differ from random expectation, according to their indicator values for temperature, soil moisture, nutrients, or light?
 
@@
 During that study period, each sampling site was surveyed once per five-year period: the first period lasted from 2003 to 2007, the second from 2008 to 2012, and the third from 2013 to 2017.
-These selection criteria yielded a sample of `r nrow(sites)` sites.
+These selection criteria yielded `r nrow(sites)` sites that had been sampled three times during 2003-2017.
+The distribution of sites accross Switzerland is given in Fig.
+\@ref(fig:studysite).
 
-For each survey (that consisted of two visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for (2) temperature, (3) humidity, (4) nutrients, and (5) light.
-The indicator values were obtained from the recalibrated indicator values for the Swiss Flora [@Landolt2010].
-In addition to these five biodiversity endpoints that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
+```{r studysite, fig.cap="Distribution of the 129 study sites accross Switzerland. Background data source: Swisstopo, Federal Office of Topography.", fig.asp=1}
+include_graphics("Geodata/studysite.pdf")
+```
+
+For each survey (that consisted of two visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for (2) temperature, (3) soil moisture, (4) nutrients, and (5) light.
+The indicator values were obtained from the recalibrated Ellenberg indicator values for the Swiss Flora [@Landolt2010].
+Ellenberg’s indicator values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables [@Ellenberg1992].
+In addition to the five biodiversity endpoints that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
 species exchange ratio sensu @Hillebrand2018) as the proportion of species that differ between two time points, to describe the community change between two points in time.
@@
 To better disentangle the importance of these mechanisms, we ordered the sites along four main environmental gradients.
-First, we expected communities to response to climate warming [@Roth2014].
+First, we expected communities to respond to climate warming [@Roth2014].
 To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
 The average $\pm$ SD mean annual temperature at our sites was `r mean(sites$TemperatureR / 10)` $\pm$ `r sd(sites$TemperatureR / 10)` $^\circ$C. 
- From the same database we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change [@Beier2012].The average annual precipitation at our sites was `r mean(sites$PrecipitationR)` $\pm$ `r sd(sites$PrecipitationR)` mm.
+The monthly-mean surface air temperature for Switzerland shows a linear increase of 1.29 $^\circ$C per 100 years between 1864 and 2016 with the warmest three years of the entire period measured in 2011, 2014 and 2015 [@Begert2018].
+Also from the WordClim database we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change [@Beier2012].The average annual precipitation at our sites was `r mean(sites$PrecipitationR)` $\pm$ `r sd(sites$PrecipitationR)` mm.
  Further, we estimated atmospheric N deposition for each site using a pragmatic approach that combined monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models [@RihmAchermann2016].
@@
 For example, a difference < 0 would suggest that the indicator values of colonizing species were lower than might be expected from random colonization from the species-pool for this site.
-We applied this method for both colonizing and disappearing species and for the indicator values for temperature, humidity, nutrients, and light (see Appendix A).
+We applied this method for both colonizing and disappearing species and for the indicator values for temperature, soil moisture, nutrients, and light (see Appendix A).
 We then tested whether this standardized difference was changing along the corresponding gradient (inferred from independent datasets, see Environmental gradients section) using linear models.
@@
 
-We used logistic GLMM to analyze whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
+We used logistic GLMM (i.e.
+a generalized linear mixed model with binomial distribution) to analyze whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
 To analyze the colonization probability of species, we selected all species that were not observed during the first survey.
 For each of these species we asked if they were observed (Occ = 1) or not observed (Occ = 0) during the second survey.
-The same was also done for all species that were not observed during the second survey if they were observed or not observed during the third survey.
+The same was also done for all species that were not observed during the second survey whether they were observed or not observed during the third survey.
 The variable 'occ' was then used as dependent variable in the logistic GLMM.
-As predictor variables, the model contained the N deposition of the site, the indicator value for nutrients of the species, and the interaction of these two variables.
+As predictor variables we used the N deposition of the site, the average indicator value for nutrients, and the interaction of these two variables.
 Additionally, species-ID and site-ID were included as random effects.
@@
 
-Fully executable R code and a description to produce this manuscript are provided at https://github.com/TobiasRoth/NDep-Trend.
+## Data accessibility and reproducibility of results {-}
+Data and R Markdown documents (Manuscript.Rmd and Appendix_A.Rmd) to fully reproduce this manuscript including figures and tables are provided at https://github.com/TobiasRoth/NDep-Trend.
+An R Markdown document is written in markdown (plain text format) and contains chunks of embedded R code to produce the figures and tables [@Xie2018].
+Raw data for analyses are provided in the folder 'RData' and the folder 'R' contains the R-Script that was used to export the data from the BDM database.
+The folder "Settings" contains a list of all the R packages (including version number) that were in the workspace when the manuscript was rendered.
+Note, however, that the files on the github repository might have slightly changed since submitting the mansuscript.
+The version of all files will be archived on Zenodo upon acceptance of the manuscript.
 
 # Results {-}
+
+## Plant communities {-}
+In total, `r n_distinct(pl$aID_SP)` plant species were recorded on the `r nrow(sites)` plots.
+Including the data of all three visits, `r mean(surv$SR)` $\pm$ `r sd(surv$SR)` (average $\pm$ SD) species were observed per plot.
+The lowest number of species recorded during a survey was `r min(surv$SR)` species and the highest number was `r max(surv$SR)` species.
+The average indicator value for temperature accross all surveys was `r mean(surv$T)` $\pm$ `r sd(surv$T)` ranging from `r min(surv$T)` to `r max(surv$T)`.
+The average indicator value for soil moisture was `r mean(surv$F)` $\pm$ `r sd(surv$F)` ranging from `r min(surv$F)` to `r max(surv$F)`.
+The average indicator value for nutrients was `r mean(surv$N)` $\pm$ `r sd(surv$N)` ranging from `r min(surv$N)` to `r max(surv$N)`.
+And the average indicator value for light was `r mean(surv$L)` $\pm$ `r sd(surv$L)` ranging from `r min(surv$L)` to `r max(surv$L)`.
+
 ## Temporal change in community structures {-}
 ```{r communitytrendscalc}
-res <- tibble(Measures = c("Species richness", "Temperature value", "Humidity value", "Nutrients value", "Light value"),
+res <- tibble(Measures = c("Species richness", "Temperature value", "soil moisture value", "Nutrients value", "Light value"),
               'Period 1' = NA,
@@
 Species richness and the four measures of plant community structure according to Landolt indicator values (i.e.
-biodiversity endpoints) suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods, the averages of species richness and the average indicator values for temperature, humidity, nutrients and light did not vary much among the three sampling periods, and the estimated trends were rather small.
+biodiversity endpoints) suggested that plant communities in mountain hay meadows were rather stable between 2003 and 2017 and did not show a clear increase or decrease over time (Table \@ref(tab:communitytrendstab)): for each of the three 5-year survey periods, the averages of species richness and the average indicator values for temperature, soil moisture, nutrients and light did not vary much among the three sampling periods, and the estimated trends were rather small.
 The results from the linear mixed models suggest that a linear temporal change was most likely for the community mean of the indicator value for temperature (probability of increase: `r res %>% filter(Measures == "Temperature value") %>% pull("Prob.
@@
 
-The average $\pm$ SD percentage of species that differed between replicated surveys was `r mean(100 * turnover_double$turnover)` $\pm$ `r sd(100 * turnover_double$turnover)`%.
+The average $\pm$ SD percentage of species that differed between replicated surveys (i.e.
+different botanists surveyed the sites) was `r mean(100 * turnover_double$turnover)` $\pm$ `r sd(100 * turnover_double$turnover)`%.
 This turnover between replicated surveys was lower than the observed temporal turnover: the average percentage of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`%, and the percentage of species that differed between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
@@
 Along the N deposition gradient, variation in species turnover was largest, with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
-The other three gradients were less conclusive to explain the variation in species turnover among sites.
+The other three gradients were less conclusive in explaining the variation in species turnover among sites.
 
@@
 res <- getres("Temperature", "Annual mean temperature", "lossT.RData")
-res <- rbind(res,  z <- getres("Humidity", "Annual mean precipitation", "lossF.RData"))
+res <- rbind(res,  z <- getres("Soil moisture", "Annual mean precipitation", "lossF.RData"))
 res <- rbind(res,  getres("Nutrients", "Nitrogen deposition", "lossN.RData"))
@@
 res <- rbind(res, getres("Temperature", "Annual mean temperature", "gainT.RData"))
-res <- rbind(res,  z <- getres("Humidity", "Annual mean precipitation", "gainF.RData"))
+res <- rbind(res,  z <- getres("Soil moisture", "Annual mean precipitation", "gainF.RData"))
 res <- rbind(res,  getres("Nutrients", "Nitrogen deposition", "gainN.RData"))
@@
   kable_styling() %>% 
-  group_rows("(a) Plants that disappeard from a site", 1, 4, bold = F, italic = T) %>%
+  group_rows("(a) Plants that disappeared from a site", 1, 4, bold = F, italic = T) %>%
   group_rows("(b) Plants that newly colonized a site", 5, 8, bold = F, italic = T) %>%
@@
 
-While colonizing species had higher temperature values compared to what we would expect under random colonization, the differences between colonizing species and random colonization was about four times smaller than the difference in the indicator value for nutrients between colonizing species and random species.
+While colonizing species had higher temperature values compared to what we would expect under random colonization, the effect size (i.e.
+the absolute value of the estimate for the difference from random expectation) for temperature was about four times smaller than the effect size for nutrients.
 Nevertheless, the variation in the indicator value for temperature seems to be important for explaining the total species turnover.
-This is because disappearing species tend to have lower temperature values than random species, and colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value with species with higher temperature values.
+This is because disappearing species tend to have lower temperature values than random species, and colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value by species with higher temperature values.
 This was not the case for the indicator value for nutrients: species with lower nutrients values tended to be more likely to disappear from, and to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
@@
 
-While we could not detect a consistent decrease in the average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the higher colonization rate of species with low nutrient value at the few sites with low N deposition rate seems to affect the spatial pattern of oligotrophic species richness: sites with low N deposition are likely to become more species-rich over time.
-This likely results in a steeper slope of the negative relationship between N deposition and oligotrophic species richness, when comparing this relationship with a spatial approach as in @Roth2013.
+While we could not detect a consistent decrease in the average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the nutrient value of colonizing species was below average (i.e.
+a strongly negative effect size) at low deposition rate and the difference to random expectation became smaller along  the N deposition gradient (second last line in Table \@ref(tab:difffromrandomtab)).
+This hihger colonization rate of species with low nutrient value at sites with low N deposition rate might have affected the spatial pattern of oligotrophic species richness: sites with low N deposition were likely to become more species-rich over time.
+This likely resulted in a steeper slope of the negative relationship between N deposition and oligotrophic species richness, when comparing this relationship with a spatial approach as in @Roth2013.
 Indeed, if we apply such an approach at different points in time to infer the effects of N deposition on the spatial variation of oligotrophic species richness, the resulting effect size (i.e., the slope) became more negative over time (Fig.
@@
 
-```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of N deposition on oligotrophic species richness estimated from applying the Poisson GLM, with species richness as dependent variable and N deposition plus other site covariates (elevation, precipitation, inclination, mean indicator values for humidity and light) as predictors, using only the surveys from one five-year interval.
-Note that within every five-year interval, all plots were sampled once."}
+```{r figconsequences, out.width = '50%', fig.asp=0.8, fig.width=4, fig.cap="Effect size of N deposition on oligotrophic species richness estimated from applying the Poisson GLM, with species richness as dependent variable and N deposition plus other site covariates (elevation, precipitation, inclination, mean indicator values for soil moisture and light) as predictors, using only the surveys from one five-year interval.
+Note that within every five-year interval, all plots were sampled once.
+Effect sizes are given as averages (points) and  5\\% and 95\\% quantiles (dashed lines) of the marginal posterior distribution.
+The dashed blue lines gives the linear regression with the effect size as dependent variable and the five-year interval as predictor variable.
+The blue area gives the  5\\% and 95\\% quantiles of the marginal posterior distribution of the regression line."}
 # Add site covariates to 'surv'
@@
 
-# Get estimates from the marginal the posterior distribution of the effect of N deposition
+# Get estimates from the marginal posterior distribution of the effect of N deposition
 getestim <- function(x) {
@@
 
+# Make regression with effect size as dependent variable
+possim <- lm(mean ~ period, data = res) %>% 
+  arm::sim(nsim)
+tmp <- map_dfc(1:nsim, function(x) possim@coef[x, "(Intercept)"] + 
+             possim@coef[x, "period"] * seq(2003, 2013, 0.1)) 
+ressmooth <- tibble(mean = apply(tmp, 1, quantile, probs = 0.5),
+                    low = apply(tmp, 1, quantile, probs = 0.05),
+                    up = apply(tmp, 1, quantile, probs = 0.95),
+                    period = seq(2003, 2013, 0.1))
+
 ggplot(res, aes(x = period, y = mean)) +
-  geom_errorbar(aes(ymin = low, ymax = up), width = 0.2, lty = 2, col = "grey65") +
+  geom_ribbon(data = ressmooth, fill = "lightskyblue1", aes(ymin = low, ymax = up)) +
+  geom_line(data = ressmooth, col = "steelblue3", lwd = 1, lty = 2) +
+  geom_errorbar(aes(ymin = low, ymax = up), width = 0.2, lwd = 0.3, lty = 1, col = "black") +
   geom_point() +
   geom_line() + 
-  geom_smooth(method = lm) +
-  geom_abline(intercept = 0, slope = 0, lty = 3) +
   labs(x = "five-year interval", y = "Effect of N deposition") + 
@@
 We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site (Appendix A).
-We found that the nutrient values of colonizing species showed the largest deviations from random expectation (Table  \@ref(tab:difffromrandomtab)), suggesting that N deposition or other factors that change the nutrient content of soils were the most important drivers of the change in species composition over the last 15 years in Swiss mountain hay meadows.
+We found that the nutrient values of colonizing species showed the largest deviations from random expectation (Table  \@ref(tab:difffromrandomtab)), suggesting that N deposition or other factors that change the nutrient content of soils were drivers of the change in species composition over the last 15 years in our sample.
 
@@
 \@ref(tab:communitytrendstab)).
-This is in contrast to how N deposition is affecting disappearing and colonization of species: It seems that both species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
+This is in contrast to how N deposition is affecting disappearance and colonization of species: It seems that both species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
 Thus, the effects of disappearing and colonizing species on the average community value for nutrients partially cancel each other out.
@@
 However, these results would be hidden if only the temporal trends of species composition were analyzed, instead of species turnover.
-Our study thus adds to the understanding of contemporary biodiversity change, where some key information is still missing [@Magurran2018], and it supports the notion of @Hillebrand2018 that considering species turnover will generate a far more reliable view of the biotic response to changing environments than solely tracking community composition.
+Our study adds to the understanding of contemporary biodiversity change [@Magurran2018], and it supports the notion of @Hillebrand2018 that considering species turnover will generate a far more reliable view of the biotic response to changing environments than solely tracking community composition.
 
 # Acknowledgements {-}
-We thank the dedicated and qualified botanists who conducted fieldwork for the Swiss Biodiversity Monitoring BDM.
+We thank the dedicated and qualified botanists who conducted fieldwork for the Swiss Biodiversity monitoring program.
 The Swiss Federal Office for the Environment (FOEN) kindly provided biodiversity monitoring data and topographic data.
-This work was supported by the FOEN, the Swiss National Science Foundation (grant no.
-31003A_156294), the Swiss Association Pro Petite Camargue Alsacienne, the Fondation de bienfaisance Jeanne Lovioz, and the MAVA Foundation.
 

### 2018-10-19 R/Data_processing.R

@@ -14,6 +14,8 @@ rm(list=ls(all=TRUE))
 # Libraries
 library(tidyverse)
 library(simba)
+library(RColorBrewer)
+library(BDM)
 
 # Connection to data base
 db <- src_sqlite(path = "~/Documents/Dropbox/DB_BDM.db", create = FALSE)
@@ -97,7 +99,9 @@ pl <- tbl(db, "PL") %>%
   as.tibble() %>% 
   filter(!is.na(match(aID_KD, surv$aID_KD))) %>% 
   left_join(tbl(db, "KD_Z9"), copy = TRUE) %>% 
-  transmute(aID_KD = aID_KD, aID_STAO = aID_STAO, aID_SP = aID_SP, yearP = yearP, yearPl = yearPl, 
+  left_join(tbl(db, "Arten"), copy = TRUE) %>% 
+  transmute(aID_KD = aID_KD, aID_STAO = aID_STAO, aID_SP = aID_SP, 
+            Species = paste(Gattung, Art), yearP = yearP, yearPl = yearPl, 
             Visit = as.integer(floor((yearP - 1998) / 5)),
             Occ = as.integer(1)) %>% 
   left_join(tbl(db, "Traits_PL") %>% dplyr::select(aID_SP, T, F, N, L), copy = TRUE) 
@@ -173,6 +177,34 @@ survdat <- rbind(
   d %>% filter(Occ2 == 1) %>% 
     transmute(aID_STAO = aID_STAO, aID_SP =aID_SP,
               T = T, F = F, N = N, L = L, Occ = Occ3, Period = 2))
+
+#------------------------------------------------------------------------------------------------------
+# Make map with location of study plots
+#------------------------------------------------------------------------------------------------------
+# Hintergrunddaten
+load("Geodata/ch.RData")
+load("Geodata/gadm.RData")
+load("Geodata/seen.RData")
+
+## Farbeinstellugen  
+seecol <- brewer.pal(8, name = "Paired")[1]
+sitecol <- brewer.pal(8, name = "Paired")[4]
+sitecolEPT <- brewer.pal(8, name = "Paired")[8]
+
+pdf("Geodata/studysite.pdf", width = 6, height = 4)
+par(mar = c(0,0,2,0))
+plot(NA, xlim = c(490000, 840000), ylim = c(60000, 300000), type = "n", axes = FALSE)
+plot(ch, add =TRUE)
+plot(gadm, add = TRUE)
[…]

### 2018-10-23 Manuscript.Rmd
Revise text of MS.2

@@
 However, there are a number of factors that may prevent the recovery of plant communities that suffered from increased N deposition.
-First, N deposition is still high at many sites, since in contrast to nitrogen oxides, there was only a minor recent decrease of ammonia emissions.
+Among others, N deposition is still high at many sites, since in contrast to nitrogen oxides, there was only a minor recent decrease of ammonia emissions.
 Thus, even if N deposition is reduced, large areas might still be above the critical threshold above which harmful effects on plant diversity do occur [i.e.
 above the critical load for N deposition; @Bobbink2010; @Slootweeg2015; @RihmAchermann2016].
-Second, there is a possibility that communities reach an alternative stable state after decades of increased N deposition and that the respective plant species are unlikely to disappear even if N deposition is reduced [@Stevens2016].
+Furthermore, there is a possibility that communities reach an alternative stable state after decades of increased N deposition and that the respective plant species are unlikely to disappear even if N deposition is reduced [@Stevens2016].
+Or, if oligotrophic species desappeared from the entire landscape, dispersal limitation may prevent oligotrophic species from recolonizing sites [@Dirnbock2004].
 Therefore, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of plant communities.
@@
 
-For each survey (that consisted of two visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: (1) The number of recorded species (species richness), the community mean of indicator values of recorded species for (2) temperature, (3) soil moisture, (4) nutrients, and (5) light.
-The indicator values were obtained from the recalibrated Ellenberg indicator values for the Swiss Flora [@Landolt2010].
-Ellenberg’s indicator values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables [@Ellenberg1992].
+For each survey (that consisted of two visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: We used the number of recorded species (species richness) that can be easily related to many conservation targets [@Rowe2017].
+Additionally, we calculated the community mean of the Landolt indicator values of recorded species.Similar to the Ellenberg’s indicator values [@Ellenberg1974], the Landolt values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables.
+The Landolt indicator values were developed for the specific situation in Switzerland, published the first time in @Landolt1977 and recalibrarted in @Landolt2010.
+Their predictive power was tested in different studies [e.g.
+@Scherrer2011].We analysed the indicator values for temperature (1: high elevation species; 5: low elevation species), soil moisture (1: species that grow in soils with low moisture; 5: species that grow in water-saturated soils), nutrients (in particular nitrogen, but also phosphorus; 1: species that grow in nutrient-poor soils; 5: species that grow under nutrient-rich conditions) and light (1: species that grow in shade; 5: species that predominantly occur in bright places).
+
 In addition to the five biodiversity endpoints that describe the state of plant communities for each site at a given time point, we also estimated the temporal turnover (i.e.
@@
 
-Nitrogen deposition at our plots decreased, on average, by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ between 2000 and 2015.
+In Swiss mountain hay meadows the average N deposition was still rather high with an average of `r mean(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$), which is at the upper limit of the suggested critical load [@Roth2017].
+Furthermore, Nitrogen deposition only weakly decreased by about `r mean(sites$NTOT_2015 - sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ between 2000 and 2015.
 This is only about one tenth of the decrease in England, where N deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
-This comparatively low change in N deposition in Swiss mountain hay meadows might explain why we observed no change in average nutrient value of communities.
+The still comparatively high N deposition rate and the rather low decrease in N deposition likely explain why we observed no change in average nutrient value of communities.
 Additionally, other anthropogenic pressures such as climate change might have overruled effects of N depositon on community composition.
@@
 Taken together, these reasons might explain why the composition of mountain hay meadow communities responded stronger to climate warming than to nitrogen reduction, although the reduction in nitrogen resulted in above average colonization of oligotrophic species.
+
+DISCUSSION ABOUT TEMPORAL TURNOVER.
+- Stimmt mit vielen Studien überein, dass temporal turnover sehr gross ist.
+- Unser Vergleich mit Doppelaufnahmen zeigt aber auch, dass man vorsichtig mit der Interpretation sein muss und dass auch standardisierte Monitoringprogramme ein relative hohe Pseudo-Turnover zeigen, die von unterschiedenen zwischen zwei Aufnahmen herrühren.
+- Trotzdem finden wir Muster, die kaum auf Methodische Gründe Zurückzuführen sind. Zum Beispiel ist Turnover bei tiefer N Dep am höchsten. Aber warum ist das so?
+- Ev auch noch den Anteil der Perennial diskutieren --> Turnover ist erstaunlich hoch.
 

### 2018-10-23 Manuscript.Rmd
Revise text of MS.

@@
 
-DISCUSSION ABOUT TEMPORAL TURNOVER.
-- Stimmt mit vielen Studien überein, dass temporal turnover sehr gross ist.
-- Unser Vergleich mit Doppelaufnahmen zeigt aber auch, dass man vorsichtig mit der Interpretation sein muss und dass auch standardisierte Monitoringprogramme ein relative hohe Pseudo-Turnover zeigen, die von unterschiedenen zwischen zwei Aufnahmen herrühren.
-- Trotzdem finden wir Muster, die kaum auf Methodische Gründe Zurückzuführen sind. Zum Beispiel ist Turnover bei tiefer N Dep am höchsten. Aber warum ist das so?
-- Ev auch noch den Anteil der Perennial diskutieren --> Turnover ist erstaunlich hoch.
-
 Observational studies along a gradient of N deposition often conclude how the spatial variation in species richness is related to N deposition [@Stevens2010gradient; @Roth2013; chapter 4 in @deVries2015].

### 2018-10-23 Manuscript.Rmd
Small changes in Introduction.

--- 2018-10-23 Manuscript.Rmd
@@
 
-DISCUSSION ABOUT TEMPORAL TURNOVER.
-- Stimmt mit vielen Studien überein, dass temporal turnover sehr gross ist.
+Our results conform to the results of recent reviews on biodiversity change suggesting that local-scale species communities are undergoing biodiversity change but not systematic loss [@Dornelas2014].
+However, our comparision with replicated surveys also warn that an important portion of the observed turnover could be caused by pseudo-turnover (i.e.
+the proportion of species that differed between two surveys that were conducted by two different surveyors during the same year on the same site).
+Given that the BDM program invested significantly in developing and testing reproducable methods and quality control [@Plattner2004], the pseudo-turnover was still quite high.
+A potential explanation is that there are usually few days between the replicated surveys and in few cases the situation might have changeed profoundly for example if the meadow was cut.
+Furthermore, for difficult to identify species one of the botanist may only note the species pair and not identify the plant on the species level.
+
+The BDM programme contains relatively clearly described methods and The BDM therefore invests signifi- cantly in developing and testing appropriate methods.
+
 - Unser Vergleich mit Doppelaufnahmen zeigt aber auch, dass man vorsichtig mit der Interpretation sein muss und dass auch standardisierte Monitoringprogramme ein relative hohe Pseudo-Turnover zeigen, die von unterschiedenen zwischen zwei Aufnahmen herrühren.

### 2018-10-23 R/Data_processing.R
Proportion of short-lived species added to surveys.

@@ -86,7 +86,8 @@ surv <- surv %>% left_join(
       T = mean(T, na.rm = TRUE), 
       F = mean(F, na.rm = TRUE),
       N = mean(N, na.rm = TRUE),
-      L = mean(L, na.rm = TRUE)), copy = TRUE)
+      L = mean(L, na.rm = TRUE),
+      KL = mean(KL, na.rm = TRUE)), copy = TRUE)
 
 # Standardize years
 surv$yr <- (surv$yearPl - 2010) / 10
@@ -110,10 +111,10 @@ pl <- tbl(db, "PL") %>%
 # Add data for turnover calculation to 'sites'
 #------------------------------------------------------------------------------------------------------
 for(i in 1:nrow(sites)) {
-  tt <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit <= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", listin = TRUE, listout = TRUE)
+  tt <- simba::sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit <= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", listin = TRUE, listout = TRUE)
   sites[i, "nochange1"] <- tt$a
   sites[i, "change1"] <- tt$b + tt$c
-  tt <- sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit >= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", listin = TRUE, listout = TRUE)
+  tt <- simba::sim(pl %>% filter(aID_STAO == sites$aID_STAO[i] & Visit >= 2) %>% dplyr::select(aID_KD, aID_SP, Occ), method = "cocogaston", listin = TRUE, listout = TRUE)
   sites[i, "nochange2"] <- tt$a
   sites[i, "change2"] <- tt$b + tt$c
 }

### 2018-10-23 R/Run_models.R
Species richness added as predictor in turn-over model.

@@ -62,7 +62,9 @@ tt <- rbind(
     NTOT = NTOT,
     Inclination = Inclination, 
     period = 2))
-mod <- stan_glmer(cbind(change, nochange) ~ period + Temperature + Precipitation + NTOT + Inclination + (1|aID_STAO), 
+
+tt$SR <- ((tt$change + tt$nochange) - 50) / 10
+mod <- stan_glmer(cbind(change, nochange) ~ period + SR + Temperature + Precipitation + NTOT + Inclination + (1|aID_STAO), 
                 data = tt, family = binomial, cores = ncores)
 save(mod, file = "Modelfit/turnover-trend.RData")
 

### 2018-10-24 Manuscript.Rmd
Discuss rather large pseudo-turnover.

--- 2018-10-23 Manuscript.Rmd
@@
 
-For each survey (that consisted of two visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: We used the number of recorded species (species richness) that can be easily related to many conservation targets [@Rowe2017].
+Before the analyses we removed all records that were not identified on species level.
+For each survey (that consisted of two visits per season) we then calculated the following biodiversity endpoints [@Rowe2017]: We used the number of recorded species (species richness) that can be easily related to many conservation targets [@Rowe2017].
 Additionally, we calculated the community mean of the Landolt indicator values of recorded species.Similar to the Ellenberg’s indicator values [@Ellenberg1974], the Landolt values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables.
@@
 The lowest number of species recorded during a survey was `r min(surv$SR)` species and the highest number was `r max(surv$SR)` species.
+In average `r mean(100 * surv$KL)` $\pm$ `r sd(100 * surv$KL)`% of the recorded species were annual species with one plot reaching up to `r mean(100 * surv$KL)`% annual species.
 The average indicator value for temperature accross all surveys was `r mean(surv$T)` $\pm$ `r sd(surv$T)` ranging from `r min(surv$T)` to `r max(surv$T)`.
@@
 res <- tibble()
-tparams <- c("Temperature", "Precipitation", "NTOT", "Inclination")
+tparams <- c("period", "SR", "Temperature", "Precipitation", "NTOT", "Inclination")
 res <- tibble(
-  Gradient = c("Mean annual temperature", "Mean annual precipitation", "Nitrogen deposition", "Inclination"),
+  Predictors = c("Period", "Number of species", "Mean annual temperature", "Mean annual precipitation", "Nitrogen deposition", "Inclination"),
   Estimate = fixef(mod)[tparams],
@@
   `95%` = posterior_interval(mod, pars = tparams)[,2])
-
-# 90% CrI for period effect
-tmp <-  posterior_interval(mod, pars = "period")
-CrI <- paste0(round(tmp[1,1], 2), " - ", round(tmp[1,2], 2))
 ```
@@
 options(knitr.kable.NA = '')
-kable(res, align = c("l", rep("r", 3)), digits = 2,  booktabs = T, col.names = c("Gradient", "Estimate", "5%", "95%"), caption = "Change of species turnover along the four gradients.
-The slopes along the gradients (estimate) are given as the change per 10 years of the logit-probability of species that differed between two surveys.
-Estimates and 5\\% and 95\\% quantiles of the marginal posterior distribution obtained from a binomial GLMM.")
+kable(res, align = c("l", rep("r", 3)), digits = 2,  booktabs = T, col.names = c("Predictors", "Estimate", "5%", "95%"), caption = "Change of species turnover along the four gradients when differences between the two periods (first period: turnover between first and second survey; second period: change in turnover between second and third surveys) and species richness effects are acounted for.
+Estimates for the period effect (change in turnover from first to second survey), the species richness effect (change in turnover per 10 species) and along the four gradients (slopes) with the corresponding 5\\% and 95\\% quantiles of the marginal posterior distribution were obtained from a binomial GLMM.")
 ```
@@
 This turnover between replicated surveys was lower than the observed temporal turnover: the average percentage of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`%, and the percentage of species that differed between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%.
-Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% credible interval of the change in turnover estimated from the binomial GLMM: `r CrI`).
-Along the N deposition gradient, variation in species turnover was largest, with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
-The other three gradients were less conclusive in explaining the variation in species turnover among sites.
+Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (period effect in Table \@ref(tab:turnovertab)).
+Species richness was a good predictor of species turnover: species rich sites were subject to higher turnover than sites with lower species richness (Table \@ref(tab:turnovertab)).
+The four gradients (temperature, precipitation, Nitrogen deposition and inclination) were less conclusive in explaining the variation in species turnover among sites.
 
@@
 
+Our results conform to the results of recent reviews on biodiversity change suggesting that local-scale species communities are often undergoing profound changes but not systematic loss [@Dornelas2014].
+However, our comparision with replicated surveys warn that an important portion of the observed turnover of species might be due to pseudo-turnover (i.e.
+species difference between two surveys that were conducted by two surveyors during the same year on the same site).
+Given that the BDM program had invested significantly in developing reproducable methods and has continiously invested in quality control [@Plattner2004], the recorded pseudo-turnover was quite high.
+A potential explanation is that the replicated surveys are not conducted during the same days and in few cases the situation might have changeed profoundly for example because the meadow was cut between the surveys.
+Furthermore, one of the botanist may not have identified difficult to identify species on species-level, and because we analysed only records that were identified on species level this increased pseudo-turnover even tough both botanists detected the species.
+
+- Trotzdem finden wir Muster, die kaum auf Methodische Gründe Zurückzuführen sind. Zum Beispiel ist Turnover bei tiefer N Dep am höchsten. Aber warum ist das so?
+- Ev auch noch den Anteil der Perennial diskutieren --> Turnover ist erstaunlich hoch.
+
 Observational studies along a gradient of N deposition often conclude how the spatial variation in species richness is related to N deposition [@Stevens2010gradient; @Roth2013; chapter 4 in @deVries2015].

### 2018-10-24 Manuscript.Rmd
Ergänzungen in Diskussion und Methoden.

@@
 The sampling sites were circles with a size of 10 m$^2$, and data collection was carried out by qualified botanists who visited each sampling site twice within the same season.
-During each visit, all vascular plant species detected on the plot were recorded.
+During each visit, all vascular plant species detected on the plot were recorded except for young plants that have not yet eveloped at least the first pair of leaves after the cotyledons.
 For details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
@@
 To better disentangle the importance of these mechanisms, we ordered the sites along four main environmental gradients.
-First, we expected communities to respond to climate warming [@Roth2014].
-To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
+First, we expected communities to respond to climate warming [@Roth2014]. To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
 The average $\pm$ SD mean annual temperature at our sites was `r mean(sites$TemperatureR / 10)` $\pm$ `r sd(sites$TemperatureR / 10)` $^\circ$C. 
@@
 The average nitrogen deposition at our sites was `r mean(sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ in 2000 and `r mean(sites$NTOT_2015)` $\pm$ `r sd(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$ in 2015.
+We assume that N deposition is a surrogate for N availability in the soil because we found that the spatial variation in oligotrophic species richness is clearly linked to N deposition [@Roth2013; @Roth2017].
+Unfortunately, we do not have any soil measurements to support this assumption.
+It seems that the total N as well as the soil carbon content down to 20 cm depth in the extensively used grassland sites of the Swiss soil monitoring network (NABO) were mostly stable over the last 20 years (R.
+Meuli, personal communication).
 Finally, we used inclination as a proxy for land-use intensity, because we assumed that steeper sites are likely to be less intensively managed [@Strebel2015].
@@
 Our results conform to the results of recent reviews on biodiversity change suggesting that local-scale species communities are often undergoing profound changes but not systematic loss [@Dornelas2014].
-However, our comparision with replicated surveys warn that an important portion of the observed turnover of species might be due to pseudo-turnover (i.e.
+However, our comparision with replicated surveys from the same year warn that an important portion of the observed turnover of species might be due to pseudo-turnover (i.e.
 species difference between two surveys that were conducted by two surveyors during the same year on the same site).
-Given that the BDM program had invested significantly in developing reproducable methods and has continiously invested in quality control [@Plattner2004], the recorded pseudo-turnover was quite high.
+Given that the BDM program had invested significantly in developing reproducable methods and has continously invested in quality control [@Plattner2004], the recorded pseudo-turnover was quite high.
 A potential explanation is that the replicated surveys are not conducted during the same days and in few cases the situation might have changeed profoundly for example because the meadow was cut between the surveys.
-Furthermore, one of the botanist may not have identified difficult to identify species on species-level, and because we analysed only records that were identified on species level this increased pseudo-turnover even tough both botanists detected the species.
-
-- Trotzdem finden wir Muster, die kaum auf Methodische Gründe Zurückzuführen sind. Zum Beispiel ist Turnover bei tiefer N Dep am höchsten. Aber warum ist das so?
-- Ev auch noch den Anteil der Perennial diskutieren --> Turnover ist erstaunlich hoch.
+Furthermore, one of the botanist may not have identified difficult to identify species on species-level (i.e.
+differences in botanical skills), and--because we analysed only records that were identified on species level--it increased pseudo-turnover even tough both botanists detected the species.
+This must be taken into account when evaluating the presented results.
+For example, we find that sites with high species richness have higher turnover than sites with a low species richness.
+This seems biologically plausible, since the average species coverage in species-rich sites must be lower than in species-poor sites and species with low coverage probably have a higher turnover.
+At the same time, however, the result could also be simply due to pseudo-turnover, since species identification in species-rich sites is probably more difficult than in species-poor sites.
+However, for the presented results that suggest differences in colonization or local survival in relation to the species indicator values we can hardly imagine how this could be caused by methodological issues.
+We are thus confident that most of the results presented in this study are largely unbiased.
 
@@
 ```{r}
-# The MS was build with the following R sesstion:
+# The MS was build with the following R session:
 writeLines(capture.output(sessionInfo()), "Settings/sessionInfo.txt")

### 2018-10-25 Manuscript.Rmd
Bemerkungen von Gutachter 2 eingebaut.

@@
 Management regimes of easily accessible mountainous areas are often being intensified, while poorly accessible mountainous areas are abandoned [@Tasser2002; @Strebel2015].
+Note that fires--an important driver of biodiversity in other grassland communities [@Ratajczak2014]-- hardly happen in Central European mountain hay meadows.
 
@@
 This is only about one tenth of the decrease in England, where N deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
-The still comparatively high N deposition rate and the rather low decrease in N deposition likely explain why we observed no change in average nutrient value of communities.
+The still comparatively high N deposition rate and the rather low decrease in N deposition combined with the fact that most of the species are perennials likely explain why we observed no change in average nutrient value of communities.
 Additionally, other anthropogenic pressures such as climate change might have overruled effects of N depositon on community composition.
@@
 There are only a limited number of studies directly relating temporal trends of plant species diversity to varying amounts of N deposition in existing communities [@Clark2008; @Storkey2015; @Stevens2016].
-In a recent study, we used the mountain hay meadow data from a single survey only and estimated the critical load along the N deposition gradient at which species richness of oligotrophic species richness starts to decrease with increasing N deposition [@Roth2017].
-Using this spatial variation in species richness and N deposition, we estimated a critical load for mountain hay meadows of 13.1 kg ha$^{-1}$ yr$^{-1}$.
+In a recent study, we used the mountain hay meadow data from a single survey only and estimated the empirical critical load along the N deposition gradient at which species richness of oligotrophic species richness starts to decrease with increasing N deposition [@Roth2017].
+Using this spatial variation in species richness and N deposition, @Roth2017 estimated a critical load for mountain hay meadows of 13.1 kg ha$^{-1}$ yr$^{-1}$.
 In the current study, we estimated the rate of N deposition at which local survival probability or colonization probability was equal for oligotrophic and eutrophic species.
-This rate could also be interpreted as a critical load.
-With this temporal approach we obtained very similar results as with the spatial approach.
-Our results are thus validating the space for time approach, at least for Swiss mountain hay meadows.
+Using a temporal approach in the present study we obtained very similar results as @Roth2017 using a the spatial approach.
+These results are thus validating the space for time approach, at least for Swiss mountain hay meadows.
 
@@
 \@ref(fig:figconsequences) also show that the results of spatial comparisons must be interpreted carefully.
-When we applied the same model as in Roth et al.
+When we applied inferred the spatial variation in oligotrhophic species richness with the same covariates as in Roth et al.
 (2017) for different study periods, to infer how oligotrophic species richness is decreasing along the N deposition gradient, the relationship appeared to vary between study periods.
@@
   Comparing the indicator values of colonizing and disappearing species with random expectation, we found that oligotrophic species are currently more likely to colonize mountain hay meadows than eutrophic species, which might be the result of the recently observed decrease in atmospheric N deposition.
-However, our results also indicate that re-transferring mountain hay meadows to oligotrophic species communities after high N deposition might take much longer than transferring an oligotrophic into an eutrophic community.
+However, our results also indicate that the recovering of mountain hay meadows after high N deposition might take much longer than transferring species-rich mountain hay meadows to species-poor communities with a large proportion of eutrophic species.
 This is because eutrophic species have high local survival probabilities, even after N deposition decreases again.
-However, these results would be hidden if only the temporal trends of species composition were analyzed, instead of species turnover.
 Our study adds to the understanding of contemporary biodiversity change [@Magurran2018], and it supports the notion of @Hillebrand2018 that considering species turnover will generate a far more reliable view of the biotic response to changing environments than solely tracking community composition.

### 2018-10-26 Manuscript.Rmd
Final edits in the MS before sending it to VA.

@@
 For each survey (that consisted of two visits per season) we then calculated the following biodiversity endpoints [@Rowe2017]: We used the number of recorded species (species richness) that can be easily related to many conservation targets [@Rowe2017].
-Additionally, we calculated the community mean of the Landolt indicator values of recorded species.Similar to the Ellenberg’s indicator values [@Ellenberg1974], the Landolt values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables.
+Additionally, we calculated the community mean of the Landolt indicator values of recorded species.
+Similar to the Ellenberg’s indicator values [@Ellenberg1974], the Landolt values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables.
 The Landolt indicator values were developed for the specific situation in Switzerland, published the first time in @Landolt1977 and recalibrarted in @Landolt2010.
 Their predictive power was tested in different studies [e.g.
-@Scherrer2011].We analysed the indicator values for temperature (1: high elevation species; 5: low elevation species), soil moisture (1: species that grow in soils with low moisture; 5: species that grow in water-saturated soils), nutrients (in particular nitrogen, but also phosphorus; 1: species that grow in nutrient-poor soils; 5: species that grow under nutrient-rich conditions) and light (1: species that grow in shade; 5: species that predominantly occur in bright places).
+@Scherrer2011].
+We analysed the indicator values for temperature (1: high elevation species; 5: low elevation species), soil moisture (1: species that grow in soils with low moisture; 5: species that grow in water-saturated soils), nutrients (referring in particular to nitrogen, but also to phosphorus; 1: species that grow in nutrient-poor soils; 5: species that grow under nutrient-rich conditions) and light (1: species that grow in shade; 5: species that predominantly occur in bright places).
 
@@
 
-To infer whether species turnover was changing along the gradient, we used a binomial generalized linear mixed model (GLMM) with the proportion of species that differed between two surveys as dependent variable and the site gradients and period (first/second vs.
-second/third surveys) as predictors, and with site-ID as random effect.
+To infer whether species turnover was changing along the gradient, we used a binomial generalized linear mixed model (GLMM) with the proportion of species that differed between two surveys as dependent variable and the site gradients, the period (first/second vs.
+second/third surveys) and the number of recorded species as predictors, and with site-ID as random effect.
 Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
@@
 
-We used logistic GLMM (i.e.
-a generalized linear mixed model with binomial distribution) to analyze whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
-To analyze the colonization probability of species, we selected all species that were not observed during the first survey.
-For each of these species we asked if they were observed (Occ = 1) or not observed (Occ = 0) during the second survey.
-The same was also done for all species that were not observed during the second survey whether they were observed or not observed during the third survey.
-The variable 'occ' was then used as dependent variable in the logistic GLMM.
+We used logistic GLMMs to analyze whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
+For the analysis of the colonisation probability we examined for all species whether they were also observed in the subsequent survey (1st vs 2nd, 2nd vs 3rd).
+The variable ‘occ’ (observed = 1; not observed = 0) was then used as dependent variable in the logistic GLMM.
 As predictor variables we used the N deposition of the site, the average indicator value for nutrients, and the interaction of these two variables.
@@
 The lowest number of species recorded during a survey was `r min(surv$SR)` species and the highest number was `r max(surv$SR)` species.
-In average `r mean(100 * surv$KL)` $\pm$ `r sd(100 * surv$KL)`% of the recorded species were annual species with one plot reaching up to `r mean(100 * surv$KL)`% annual species.
+In average `r mean(100 * surv$KL)` $\pm$ `r sd(100 * surv$KL)`% of the recorded species were annual species with one plot reaching up to `r max(100 * surv$KL)`% annual species.
 The average indicator value for temperature accross all surveys was `r mean(surv$T)` $\pm$ `r sd(surv$T)` ranging from `r min(surv$T)` to `r max(surv$T)`.

### 2018-10-26 Manuscript.Rmd
Spell check.

@@
 Furthermore, there is a possibility that communities reach an alternative stable state after decades of increased N deposition and that the respective plant species are unlikely to disappear even if N deposition is reduced [@Stevens2016].
-Or, if oligotrophic species desappeared from the entire landscape, dispersal limitation may prevent oligotrophic species from recolonizing sites [@Dirnbock2004].
+Or, if oligotrophic species disappeared from the entire landscape, dispersal limitation may prevent oligotrophic species from recolonizing sites [@Dirnbock2004].
 Therefore, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of plant communities.
@@
 Species richness, a biodiversity endpoint that can be assessed and communicated relatively easily, is often negatively related to nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
-Other metrics that are potentially more useful to reflect favorable changes can be derived from the traits of the species in a community.
+Other metrics that are potentially more useful to reflect favourable changes can be derived from the traits of the species in a community.
 In Europe, environmental preference of plants has often been expressed using indicator values assigned to each plant species [@Ellenberg1992; @Landolt2010].
@@
 ## Monitoring data and community measures {-}
-We analyzed the presence/absence of vascular plants sampled in the Swiss Biodiversity Monitoring (BDM) program that was launched in 2001 to monitor Switzerland's biodiversity and to comply with the Convention on Biological Diversity of Rio de Janeiro [@Weber2004].
+We analysed the presence/absence of vascular plants sampled in the Swiss Biodiversity Monitoring (BDM) program that was launched in 2001 to monitor Switzerland's biodiversity and to comply with the Convention on Biological Diversity of Rio de Janeiro [@Weber2004].
 The sampling sites were circles with a size of 10 m$^2$, and data collection was carried out by qualified botanists who visited each sampling site twice within the same season.
-During each visit, all vascular plant species detected on the plot were recorded except for young plants that have not yet eveloped at least the first pair of leaves after the cotyledons.
+During each visit, all vascular plant species detected on the plot were recorded except for young plants that have not yet developed at least the first pair of leaves after the cotyledons.
 For details on the field methods see @Plattner2004, @Roth2013 and @Roth2017.
@@
 We matched the habitat types of the Swiss classification system with the categories from the EUNIS system [level-3 classification; @Davies2004] and selected all sampling sites in mountain hay meadows (EUNIS E2.3).
-We analyzed the data from 2003 to 2017.
+We analysed the data from 2003 to 2017.
 During that study period, each sampling site was surveyed once per five-year period: the first period lasted from 2003 to 2007, the second from 2008 to 2012, and the third from 2013 to 2017.
 These selection criteria yielded `r nrow(sites)` sites that had been sampled three times during 2003-2017.
-The distribution of sites accross Switzerland is given in Fig.
+The distribution of sites across Switzerland is given in Fig.
 \@ref(fig:studysite).
@@
 Similar to the Ellenberg’s indicator values [@Ellenberg1974], the Landolt values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables.
-The Landolt indicator values were developed for the specific situation in Switzerland, published the first time in @Landolt1977 and recalibrarted in @Landolt2010.
+The Landolt indicator values were developed for the specific situation in Switzerland, published the first time in @Landolt1977 and recalibrated in @Landolt2010.
 Their predictive power was tested in different studies [e.g.
@@
 
-We used logistic GLMMs to analyze whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
+We used logistic GLMMs to analyse whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
 For the analysis of the colonisation probability we examined for all species whether they were also observed in the subsequent survey (1st vs 2nd, 2nd vs 3rd).
@@
 The folder "Settings" contains a list of all the R packages (including version number) that were in the workspace when the manuscript was rendered.
-Note, however, that the files on the github repository might have slightly changed since submitting the mansuscript.
+Note, however, that the files on the github repository might have slightly changed since submitting the manuscript.
 The version of all files will be archived on Zenodo upon acceptance of the manuscript.
@@
 In average `r mean(100 * surv$KL)` $\pm$ `r sd(100 * surv$KL)`% of the recorded species were annual species with one plot reaching up to `r max(100 * surv$KL)`% annual species.
-The average indicator value for temperature accross all surveys was `r mean(surv$T)` $\pm$ `r sd(surv$T)` ranging from `r min(surv$T)` to `r max(surv$T)`.
+The average indicator value for temperature across all surveys was `r mean(surv$T)` $\pm$ `r sd(surv$T)` ranging from `r min(surv$T)` to `r max(surv$T)`.
 The average indicator value for soil moisture was `r mean(surv$F)` $\pm$ `r sd(surv$F)` ranging from `r min(surv$F)` to `r max(surv$F)`.
@@
 a strongly negative effect size) at low deposition rate and the difference to random expectation became smaller along  the N deposition gradient (second last line in Table \@ref(tab:difffromrandomtab)).
-This hihger colonization rate of species with low nutrient value at sites with low N deposition rate might have affected the spatial pattern of oligotrophic species richness: sites with low N deposition were likely to become more species-rich over time.
+This higher colonization rate of species with low nutrient value at sites with low N deposition rate might have affected the spatial pattern of oligotrophic species richness: sites with low N deposition were likely to become more species-rich over time.
 This likely resulted in a steeper slope of the negative relationship between N deposition and oligotrophic species richness, when comparing this relationship with a spatial approach as in @Roth2013.
@@
 The still comparatively high N deposition rate and the rather low decrease in N deposition combined with the fact that most of the species are perennials likely explain why we observed no change in average nutrient value of communities.
-Additionally, other anthropogenic pressures such as climate change might have overruled effects of N depositon on community composition.
+Additionally, other anthropogenic pressures such as climate change might have overruled effects of N deposition on community composition.
 In particular, we found that the species disappearing from the sites tended to have below average indicator values for temperature, while species that newly colonize sites had above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
@@
 Our results conform to the results of recent reviews on biodiversity change suggesting that local-scale species communities are often undergoing profound changes but not systematic loss [@Dornelas2014].
-However, our comparision with replicated surveys from the same year warn that an important portion of the observed turnover of species might be due to pseudo-turnover (i.e.
+However, our comparison with replicated surveys from the same year warn that an important portion of the observed turnover of species might be due to pseudo-turnover (i.e.
 species difference between two surveys that were conducted by two surveyors during the same year on the same site).
-Given that the BDM program had invested significantly in developing reproducable methods and has continously invested in quality control [@Plattner2004], the recorded pseudo-turnover was quite high.
-A potential explanation is that the replicated surveys are not conducted during the same days and in few cases the situation might have changeed profoundly for example because the meadow was cut between the surveys.
+Given that the BDM program had invested significantly in developing reproducible methods and has continuously invested in quality control [@Plattner2004], the recorded pseudo-turnover was quite high.
+A potential explanation is that the replicated surveys are not conducted during the same days and in few cases the situation might have changed profoundly for example because the meadow was cut between the surveys.
 Furthermore, one of the botanist may not have identified difficult to identify species on species-level (i.e.
@@
 \@ref(fig:figconsequences) also show that the results of spatial comparisons must be interpreted carefully.
-When we applied inferred the spatial variation in oligotrhophic species richness with the same covariates as in Roth et al.
+When we applied inferred the spatial variation in oligotrophic species richness with the same covariates as in Roth et al.
 (2017) for different study periods, to infer how oligotrophic species richness is decreasing along the N deposition gradient, the relationship appeared to vary between study periods.

### 2018-10-26 Manuscript.Rmd
Small edits in Data accessibiility.
Small additional edits in data accessibility chapter.

@@
 The folder "Settings" contains a list of all the R packages (including version number) that were in the workspace when the manuscript was rendered.
-Note, however, that the files on the github repository might have slightly changed since submitting the manuscript.
-The version of all files will be archived on Zenodo upon acceptance of the manuscript.
+
+The v1 release of the GitHub repository is the version  that corresponds to the initial submission  (https://github.com/TobiasRoth/NDep-Trend/releases/tag/v1), v2 is the version of the repository that corresponds to the revised version of the mansucript.The final version of the repository will be archived on Zenodo upon acceptance of the manuscript.
 

### 2018-12-12 Appendix_A.Rmd
Korrekturen der Koauthoren übernommen.

@@
 Instead of the functional traits that are traditionally used in functional ecology such as species height, we used the Ellenberg values that we assumed could best reflect the main drivers that are likely to cause shifts in mountain hay plant communities.
-This were the Ellenberg values for temperature (that are likely to respond to climate warming), soil moisture (likely to respond to changes in precipitation), nutrients (likely to change in response to Nitrogen deposition and land-use intensity) and light (likely to change in response to land-use intensity).
+These were the Ellenberg values for temperature (that are likely to respond to climate warming), soil moisture (likely to respond to changes in precipitation), nutrients (likely to change in response to Nitrogen deposition and land-use intensity) and light (likely to change in response to land-use intensity).
 

### 2018-12-12 Manuscript.Rmd
Korrekturen der Koauthoren übernommen.

--- 2018-10-26 Manuscript.Rmd
@@
 In particular, the average nutrient value of plant communities remained stable over time (2003-2017).
-However, we found that the nutrient values of colonizing species showed the largest deviations from random expectation, suggesting that N deposition or other factors that change the nutrient content of soils were important drivers of the species composition change over the last 15 years in Swiss mountain hay meadows.
+We found the largest deviations from random expectation in the nutrient values of colonizing species, suggesting that N deposition or other factors that change the nutrient content of soils were important drivers of the species composition change over the last 15 years in Swiss mountain hay meadows.
 In addition, we observed an overall replacement of species with lower indicator values for temperature with species with higher values.
 Apparently, the community effects of the replacement of eutrophic species with oligotrophic species was outweighed by climate warming.
-Our results add to the increasing evidence that inferring species turnover will generate a far more reliable understanding of the biotic response to changing environments than solely tracking average community composition.
+Our results add to the increasing evidence that measures of community composition such as species richness or average indicator values may be relatively stable, but that this apparent stability is often accompanied by a marked turnover of species.
 
@@
 However, there are a number of factors that may prevent the recovery of plant communities that suffered from increased N deposition.
-Among others, N deposition is still high at many sites, since in contrast to nitrogen oxides, there was only a minor recent decrease of ammonia emissions.
+Among others, N deposition is still high at many sites, since in contrast to nitrogen oxides, ammonia emissions decreased only to a small degree.
 Thus, even if N deposition is reduced, large areas might still be above the critical threshold above which harmful effects on plant diversity do occur [i.e.
@@
 Furthermore, there is a possibility that communities reach an alternative stable state after decades of increased N deposition and that the respective plant species are unlikely to disappear even if N deposition is reduced [@Stevens2016].
-Or, if oligotrophic species disappeared from the entire landscape, dispersal limitation may prevent oligotrophic species from recolonizing sites [@Dirnbock2004].
-Therefore, it is an open question whether and how fast the reduction in N deposition rates will lead to the recovery of plant communities.
+Alternatively, if oligotrophic species disappeared from the entire landscape, dispersal limitation may prevent oligotrophic species from recolonizing sites [@Dirnbock2004].
+It is therefore an open question whether and how fast the reduction in N deposition rates will lead to the recovery of plant communities.
 
@@
 improving biodiversity endpoints sensu @Rowe2017).
-Species richness, a biodiversity endpoint that can be assessed and communicated relatively easily, is often negatively related to nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
+Species richness, a biodiversity endpoint that can be relatively easily assessed and communicated, is often negatively related to nitrogen deposition [@Maskell2010; @Field2014; @Rowe2017].
 Other metrics that are potentially more useful to reflect favourable changes can be derived from the traits of the species in a community.
@@
 With extensive cultivation, permanent grassland has a very high biodiversity.
-This applies in particular to the meadows in the alpine region, where meadows with a high plant diversity are also of agronomical importance [@Leiber2006].
-In mountain hay meadows, the spatial variation of species richness in vascular plants is negatively correlated with N deposition [@Roth2013], suggesting that mountain grasslands are negatively affected by increased N deposition.
-However, between 1990 and 2010, NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% [@MaasGrennfelt2016].
+This applies in particular to the meadows in the alpine region, where meadows with high plant diversity are also of agronomical importance [@Leiber2006].
+In mountain hay meadows, the spatial variation of species richness in vascular plants has been shown to be negatively correlated with N deposition [@Roth2013], suggesting that mountain grasslands are negatively affected by increased N deposition.
+However, between 1990 and 2010, NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% and considerable emission reductions also occured in neighboring countries [@MaasGrennfelt2016].
 Potentially, this could have led to a partial recovery of plant communities.
@@
 In Switzerland, temperatures increased from 1959 to 2008 at all altitudes, with an average warming rate of 0.35$^{\circ}$C per decade, which is about 1.6 times the northern hemispheric warming rate [@Ceppi2012].
-This climate warming is likely to interact with N deposition to drive plant community changes [@Humbert2016].
-Indeed, in an earlier study we found that, at the landscape scale, plant communities responded to climate warming even within a relatively short time period [@Roth2014].
+Climate warming is likely to interact with N deposition in driving plant community changes [@Humbert2016].
+Indeed, in an earlier study we found that, at the landscape scale, plant communities responded to climate warming within a relatively short period of time [@Roth2014].
 @Steinbauer2018 suggest that particularly the shift of plant communities at mountain summits is the result of recent climate warming, and they assume an interaction with airborne N deposition.
@@
 Management regimes of easily accessible mountainous areas are often being intensified, while poorly accessible mountainous areas are abandoned [@Tasser2002; @Strebel2015].
-Note that fires--an important driver of biodiversity in other grassland communities [@Ratajczak2014]-- hardly happen in Central European mountain hay meadows.
+Note that fires--an important driver of biodiversity in other grassland communities [@Ratajczak2014]--hardly occur in Central European mountain hay meadows.
 
@@
 (2) Was species turnover correlated with the average temperature, precipitation, N deposition, or inclination (we expect steep areas to be less intensively managed)?
-(3) Did species that newly colonized or disappeared from (local extinction) sites differ from random expectation, according to their indicator values for temperature, soil moisture, nutrients, or light?
+(3) Did species that newly colonized or disappeared from local sites differ from random expectation, according to their indicator values for temperature, soil moisture, nutrients, or light?
 
@@
 
-Before the analyses we removed all records that were not identified on species level.
-For each survey (that consisted of two visits per season) we then calculated the following biodiversity endpoints [@Rowe2017]: We used the number of recorded species (species richness) that can be easily related to many conservation targets [@Rowe2017].
+Before the analyses we removed all records that were not identified on the species level.
+For each survey (that consisted of two visits per season) we then calculated the following biodiversity endpoints: We used the number of recorded species (species richness) that can be easily related to many conservation targets [@Rowe2017].
 Additionally, we calculated the community mean of the Landolt indicator values of recorded species.
-Similar to the Ellenberg’s indicator values [@Ellenberg1974], the Landolt values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables.
+Similar to the Ellenberg indicator values [@Ellenberg1974], the Landolt values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables.
 The Landolt indicator values were developed for the specific situation in Switzerland, published the first time in @Landolt1977 and recalibrated in @Landolt2010.
@@
 
-To test data quality, independent replicate surveys are routinely performed within the BDM program by botanists who were not involved in the regular BDM surveys.
-The regular surveyors did not know if and which sites were replicated [@Plattner2004].
-We used the data from `r nrow(turnover_double)` such replicated surveys that were available to calculate the pseudo-turnover, which is the proportion of species that differed between two surveys that were conducted by two different surveyors during the same year on the same site.
+To test data quality in the BDM program, independent replicate surveys were routinely performed by botanists who were not involved in the regular BDM surveys.
+The regular surveyors did not know whether and in which sites their surveys were replicated [@Plattner2004].
+We used the data from `r nrow(turnover_double)` replicated surveys to calculate the pseudo-turnover, which is the proportion of species that differed between two surveys that were conducted by two different surveyors during the same year on the same site.
 
@@
 The monthly-mean surface air temperature for Switzerland shows a linear increase of 1.29 $^\circ$C per 100 years between 1864 and 2016 with the warmest three years of the entire period measured in 2011, 2014 and 2015 [@Begert2018].
-Also from the WordClim database we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change [@Beier2012].The average annual precipitation at our sites was `r mean(sites$PrecipitationR)` $\pm$ `r sd(sites$PrecipitationR)` mm.
+Also from the WordClim database, we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change [@Beier2012].The average annual precipitation at our sites was `r mean(sites$PrecipitationR)` $\pm$ `r sd(sites$PrecipitationR)` mm.
 Further, we estimated atmospheric N deposition for each site using a pragmatic approach that combined monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models [@RihmAchermann2016].
 The average nitrogen deposition at our sites was `r mean(sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ in 2000 and `r mean(sites$NTOT_2015)` $\pm$ `r sd(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$ in 2015.
-We assume that N deposition is a surrogate for N availability in the soil because we found that the spatial variation in oligotrophic species richness is clearly linked to N deposition [@Roth2013; @Roth2017].
-Unfortunately, we do not have any soil measurements to support this assumption.
-It seems that the total N as well as the soil carbon content down to 20 cm depth in the extensively used grassland sites of the Swiss soil monitoring network (NABO) were mostly stable over the last 20 years (R.
+We assume that N deposition is a surrogate for N availability in the soil because we found that the spatial variation in oligotrophic species richness is clearly linked to N deposition [@Roth2013; @Roth2017]; unfortunately, we do not have soil measurements to test this assumption.
+Apparently, the total N as well as the soil carbon content down to 20 cm depth were mostly stable over the last 20 years in the extensively used grassland sites of the Swiss soil monitoring network (NABO; R.
 Meuli, personal communication).
@@
 a random intercept random slope model, @Gelman2006].
-Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
+Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016; @Muth2018].
 
-To infer whether species turnover was changing along the gradient, we used a binomial generalized linear mixed model (GLMM) with the proportion of species that differed between two surveys as dependent variable and the site gradients, the period (first/second vs.
+To infer whether species turnover was changing along the gradient, we used a binomial generalized linear mixed model (GLMM) with the proportion of species that differed between two surveys as dependent variable, with the site gradients, the period (first/second vs.
 second/third surveys) and the number of recorded species as predictors, and with site-ID as random effect.
-Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
+Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016; @Muth2018].
 
@@
 We then tested whether this standardized difference was changing along the corresponding gradient (inferred from independent datasets, see Environmental gradients section) using linear models.
-Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016, @Muth2018].
+Model parameters were estimated in a Bayesian framework using the R-Package *rstanarm* [@Stan2016; @Muth2018].
 
@@
 For the analysis of the colonisation probability we examined for all species whether they were also observed in the subsequent survey (1st vs 2nd, 2nd vs 3rd).
-The variable ‘occ’ (observed = 1; not observed = 0) was then used as dependent variable in the logistic GLMM.
+The variable ‚Äòocc‚Äô (observed = 1; not observed = 0) was then used as dependent variable in the logistic GLMM.
 As predictor variables we used the N deposition of the site, the average indicator value for nutrients, and the interaction of these two variables.
@@
 The folder "Settings" contains a list of all the R packages (including version number) that were in the workspace when the manuscript was rendered.
-Note, however, that the files on the github repository might have slightly changed since submitting the manuscript.
-The version of all files will be archived on Zenodo upon acceptance of the manuscript.
+
+The v1 release of the GitHub repository is the version that corresponds to the initial submission (https://github.com/TobiasRoth/NDep-Trend/releases/tag/v1), v2 is the version of the repository that corresponds to the revised version of the mansucript.
+The final version of the repository will be archived on Zenodo upon acceptance of the manuscript.
 
@@
 The lowest number of species recorded during a survey was `r min(surv$SR)` species and the highest number was `r max(surv$SR)` species.
-In average `r mean(100 * surv$KL)` $\pm$ `r sd(100 * surv$KL)`% of the recorded species were annual species with one plot reaching up to `r max(100 * surv$KL)`% annual species.
-The average indicator value for temperature across all surveys was `r mean(surv$T)` $\pm$ `r sd(surv$T)` ranging from `r min(surv$T)` to `r max(surv$T)`.
-The average indicator value for soil moisture was `r mean(surv$F)` $\pm$ `r sd(surv$F)` ranging from `r min(surv$F)` to `r max(surv$F)`.
-The average indicator value for nutrients was `r mean(surv$N)` $\pm$ `r sd(surv$N)` ranging from `r min(surv$N)` to `r max(surv$N)`.
-And the average indicator value for light was `r mean(surv$L)` $\pm$ `r sd(surv$L)` ranging from `r min(surv$L)` to `r max(surv$L)`.
+On average, `r mean(100 * surv$KL)` $\pm$ `r sd(100 * surv$KL)`% of the recorded species were annual species with one plot reaching up to `r max(100 * surv$KL)`% annual species.
+The average indicator value for temperature across all surveys was `r mean(surv$T)` $\pm$ `r sd(surv$T)`, ranging from `r min(surv$T)` to `r max(surv$T)`.
+The average indicator value for soil moisture was `r mean(surv$F)` $\pm$ `r sd(surv$F)`, ranging from `r min(surv$F)` to `r max(surv$F)`.
+The average indicator value for nutrients was `r mean(surv$N)` $\pm$ `r sd(surv$N)`, ranging from `r min(surv$N)` to `r max(surv$N)`.
+And the average indicator value for light was `r mean(surv$L)` $\pm$ `r sd(surv$L)`, ranging from `r min(surv$L)` to `r max(surv$L)`.
 
@@
 ```{r communitytrendscalc}
-res <- tibble(Measures = c("Species richness", "Temperature value", "soil moisture value", "Nutrients value", "Light value"),
+res <- tibble(Measures = c("Species richness", "Temperature value", "soil moisture value", "Nutrient value", "Light value"),
               'Period 1' = NA,
@@
 for Trend")`).
-The chance that the community mean of the nutrient value decreased between 2003 and 2017 was `r 1- res %>% filter(Measures == "Nutrients value") %>% pull("Prob.
-for Trend")` (`r res %>% filter(Measures == "Nutrients value") %>% pull("Prob.
-for Trend")` gives the probability for an increase).
+The chance that the community mean of the nutrient value decreased between 2003 and 2017 was `r 1- res %>% filter(Measures == "Nutrient value") %>% pull("Prob.
+for Trend")` (`r res %>% filter(Measures == "Nutrient value") %>% pull("Prob.
+for Trend")` the probability for an increase was KEINE AHNUNG WAS ICH HIER EINFÜGEN MUSS...).
 
@@
 This is because disappearing species tend to have lower temperature values than random species, and colonizing species tend to have higher temperature values than random species; both processes lead to an overall replacement of species with lower temperature value by species with higher temperature values.
-This was not the case for the indicator value for nutrients: species with lower nutrients values tended to be more likely to disappear from, and to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
+This was not the case for the indicator value for nutrients: species with lower nutrient values tended to be more likely to disappear from, and to colonize sites compared to random species (Table \@ref(tab:difffromrandomtab)).
 See also Appendix A, where we present detailed results for the comparison between colonizing or disappearing species with randomly selected species.
@@
 
-```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic species (indicator value for nutrients $<3$; red line) and of eutrophic species (indicator value for nutrients $>3$) species along the N deposition gradient.
+```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic species (indicator value for nutrients $<3$; red line) and of eutrophic species (indicator value for nutrients $>3$) along the N deposition gradient.
 Given are means and 95\\%-credible intervals from logistic linear mixed models.
@@
 
-While we could not detect a consistent decrease in the average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the nutrient value of colonizing species was below average (i.e.
-a strongly negative effect size) at low deposition rate and the difference to random expectation became smaller along  the N deposition gradient (second last line in Table \@ref(tab:difffromrandomtab)).
+While we did not observe a consistent decrease in the average indicator values for nutrients (Table \@ref(tab:communitytrendstab)), the nutrient value of colonizing species was below average (i.e.
+a strongly negative effect size) at low deposition rate, and the difference to random expectation became smaller along  the N deposition gradient (second last line in Table \@ref(tab:difffromrandomtab)).
 This higher colonization rate of species with low nutrient value at sites with low N deposition rate might have affected the spatial pattern of oligotrophic species richness: sites with low N deposition were likely to become more species-rich over time.
@@
 Note that within every five-year interval, all plots were sampled once.
-Effect sizes are given as averages (points) and  5\\% and 95\\% quantiles (dashed lines) of the marginal posterior distribution.
-The dashed blue lines gives the linear regression with the effect size as dependent variable and the five-year interval as predictor variable.
+Effect sizes are given as averages (points) and  5\\% and 95\\% quantiles (lines) of the marginal posterior distribution.
+The dashed blue line gives the linear regression with the effect size as dependent variable and the five-year interval as predictor variable.
 The blue area gives the  5\\% and 95\\% quantiles of the marginal posterior distribution of the regression line."}
@@
 Although N deposition declined between 2000 and 2015 (Fig.
-\@ref(fig:Ndepchange)), we could detect only weak shifts in plant community structure (i.e.
+\@ref(fig:Ndepchange)), we observed only weak shifts in plant community structure (i.e.
 biodiversity endpoints sensu @Rowe2017) during the same time period (Table \@ref(tab:communitytrendstab)).
@@
 For instance, species turnover varied along the N deposition gradient, with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)).
-Taken together, our results add to the increasing evidence that contemporary communities are often relatively stable at community level, but this apparent stability is often accompanied by marked turnover of species [@Vellend2013; @Dornelas2014; @Hillebrand2018].
+Taken together, our results add to the increasing evidence that contemporary plant communities may be relatively stable regarding average community composition, but that this apparent stability is often accompanied by a marked turnover of species.
+[@Vellend2013; @Dornelas2014; @Hillebrand2018].
 
@@
 Such factors select for species with specific characteristics.
-Community assembly theories thus suggest that the factors that drive the composition of species in a community can be inferred from comparing the characteristics of the species in the community with random expectation [@Chase2011].
-We adopted this idea and compared the indicator values of species that disappeared or colonized a site with the indicator values from randomly chosen species from the same site (Appendix A).
-We found that the nutrient values of colonizing species showed the largest deviations from random expectation (Table  \@ref(tab:difffromrandomtab)), suggesting that N deposition or other factors that change the nutrient content of soils were drivers of the change in species composition over the last 15 years in our sample.
+Community assembly theory thus suggests that the factors driving the composition of species in a community can be inferred from comparing the characteristics of the species in the community with random expectation [@Chase2011].
+We adopted this idea and compared the indicator values of species that disappeared or colonized a site with indicator values from randomly chosen species from the same site (Appendix A).
+We found that the nutrient values of colonizing species showed the largest deviations from random expectation (Table  \@ref(tab:difffromrandomtab)), suggesting that in our sample, N deposition or other factors changing the nutrient content of soils were drivers of the change in species composition over the last 15 years.
 
-In Swiss mountain hay meadows the average N deposition was still rather high with an average of `r mean(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$), which is at the upper limit of the suggested critical load [@Roth2017].
+In Swiss mountain hay meadows, the average N deposition was still rather high with an average of `r mean(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$), which is at the upper limit of the suggested critical load [@Roth2017].
 Furthermore, Nitrogen deposition only weakly decreased by about `r mean(sites$NTOT_2015 - sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ between 2000 and 2015.
 This is only about one tenth of the decrease in England, where N deposition decreased by 24 kg ha$^{-1}$ yr$^{-1}$ from 1996 to 2011 [@Storkey2015].
-The still comparatively high N deposition rate and the rather low decrease in N deposition combined with the fact that most of the species are perennials likely explain why we observed no change in average nutrient value of communities.
-Additionally, other anthropogenic pressures such as climate change might have overruled effects of N deposition on community composition.
-In particular, we found that the species disappearing from the sites tended to have below average indicator values for temperature, while species that newly colonize sites had above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
-Thus, the effect of disappearing and the effect of colonizing species on the community mean for temperature is additive resulting in increasing average temperature values (Tab.
+The still comparatively high N deposition rate and the rather low decrease in N deposition, combined with the fact that most of the species are perennials, likely explain why we observed no change in average nutrient value of communities.
+Additionally, other anthropogenic pressures such as climate change might have outweighed effects of N deposition on community composition.
+In particular, we found that species disappearing from the sites tended to have below average indicator values for temperature, while species that newly colonize sites had above average indicator values for temperature (Table \@ref(tab:difffromrandomtab)).
+Thus, the effect of disappearing and the effect of colonizing species on the community mean for temperature is additive, resulting in increasing average temperature values (Tab.
 \@ref(tab:communitytrendstab)).
-This is in contrast to how N deposition is affecting disappearance and colonization of species: It seems that both species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
+This is in contrast to how N deposition is affecting disappearance and colonization of species: It seems that both the species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
 Thus, the effects of disappearing and colonizing species on the average community value for nutrients partially cancel each other out.
@@
 
-Our results conform to the results of recent reviews on biodiversity change suggesting that local-scale species communities are often undergoing profound changes but not systematic loss [@Dornelas2014].
+Our results conform to the patterns described in recent reviews on biodiversity change, suggesting that local-scale species communities are often undergoing profound changes, but do not necessarily show a systematic loss of species numbers [@Dornelas2014].
 However, our comparison with replicated surveys from the same year warn that an important portion of the observed turnover of species might be due to pseudo-turnover (i.e.
-species difference between two surveys that were conducted by two surveyors during the same year on the same site).
-Given that the BDM program had invested significantly in developing reproducible methods and has continuously invested in quality control [@Plattner2004], the recorded pseudo-turnover was quite high.
-A potential explanation is that the replicated surveys are not conducted during the same days and in few cases the situation might have changed profoundly for example because the meadow was cut between the surveys.
-Furthermore, one of the botanist may not have identified difficult to identify species on species-level (i.e.
-differences in botanical skills), and--because we analysed only records that were identified on species level--it increased pseudo-turnover even tough both botanists detected the species.
-This must be taken into account when evaluating the presented results.
-For example, we find that sites with high species richness have higher turnover than sites with a low species richness.
-This seems biologically plausible, since the average species coverage in species-rich sites must be lower than in species-poor sites and species with low coverage probably have a higher turnover.
+species difference between two surveys that were conducted during the same year on the same site, but by two surveyors).
+Given that the BDM program has included major efforts in developing reproducible methods and has continuously invested in quality control [@Plattner2004], the recorded pseudo-turnover was quite high.
+A potential explanation is that the replicated surveys were not conducted during the same days, and at some sites the situation might have changed profoundly, for example because a meadow was cut between the surveys.
+Furthermore, one of two botanists at a given site may not have identified some difficult species to the species-level, and--because we analysed only records that were identified to the species level--this might have increased pseudo-turnover even though both botanists actually detected the species.
+Such problems must be taken into account when evaluating the presented results.
+For example, we found that sites with high species richness had higher species turnover than sites with low species richness.
+This seems biologically plausible, since the average species coverage in species-rich sites must be lower than in species-poor sites, and species with low coverage probably have a higher turnover.
 At the same time, however, the result could also be simply due to pseudo-turnover, since species identification in species-rich sites is probably more difficult than in species-poor sites.
-However, for the presented results that suggest differences in colonization or local survival in relation to the species indicator values we can hardly imagine how this could be caused by methodological issues.
-We are thus confident that most of the results presented in this study are largely unbiased.
+However, for the presented result suggesting differences in colonization or local survival in relation to the species indicator values, we can hardly imagine how this could be caused by methodological issues.
 
 Observational studies along a gradient of N deposition often conclude how the spatial variation in species richness is related to N deposition [@Stevens2010gradient; @Roth2013; chapter 4 in @deVries2015].
-They assume that the spatial variation in species richness (or other metrics of community composition) arose because of unequal species loss of different areas over time, resulting from elevated N deposition chronically experienced by some areas.
+Such studies assume that the spatial variation in species richness (or other metrics of community composition) arose because of unequal species loss of different areas over time, resulting from elevated N deposition chronically experienced in some areas.
 Although there is evidence supporting the pertinency of such a 'space for time substitution' for detecting the effects of N deposition on plant diversity [@Stevens2010], this approach cannot replace studies that relate temporal patterns in species composition with N deposition [@deSchrijver2011].
 There are only a limited number of studies directly relating temporal trends of plant species diversity to varying amounts of N deposition in existing communities [@Clark2008; @Storkey2015; @Stevens2016].
-In a recent study, we used the mountain hay meadow data from a single survey only and estimated the empirical critical load along the N deposition gradient at which species richness of oligotrophic species richness starts to decrease with increasing N deposition [@Roth2017].
+In an earlier study, we used the mountain hay meadow data from a single survey and estimated the empirical critical load along the N deposition gradient at which species richness of oligotrophic species richness starts to decrease with increasing N deposition [@Roth2017].
 Using this spatial variation in species richness and N deposition, @Roth2017 estimated a critical load for mountain hay meadows of 13.1 kg ha$^{-1}$ yr$^{-1}$.
 In the current study, we estimated the rate of N deposition at which local survival probability or colonization probability was equal for oligotrophic and eutrophic species.
-Using a temporal approach in the present study we obtained very similar results as @Roth2017 using a the spatial approach.
-These results are thus validating the space for time approach, at least for Swiss mountain hay meadows.
+Using a temporal approach in the present study, we obtained very similar results as @Roth2017 using a the spatial approach.
+Our results may thus be taken to validate the space for time approach, at least for Swiss mountain hay meadows.
 
-However, the results of Fig.
-\@ref(fig:figconsequences) also show that the results of spatial comparisons must be interpreted carefully.
-When we applied inferred the spatial variation in oligotrophic species richness with the same covariates as in Roth et al.
-(2017) for different study periods, to infer how oligotrophic species richness is decreasing along the N deposition gradient, the relationship appeared to vary between study periods.
+However, Fig.
+\@ref(fig:figconsequences) also shows that the results of spatial comparisons must be interpreted carefully.
+When we investigated the spatial variation in oligotrophic species richness with the same covariates as in Roth et al.
+(2017) for different study periods, to infer how oligotrophic species richness was decreasing along the N deposition gradient, the relationship appeared to vary between study periods.
 The decrease gradually became steeper (more negative) over time, except for the first two study periods.
 Our first interpretation was that the N deposition effect became stronger over time.
-This was against our prediction that the effect of N deposition should become weaker over time, since N deposition is decreasing during the study period.
+This was against our prediction that the effect of N deposition should become weaker over time, since N deposition was decreasing during the study period.
 Then we realized that species turnover was highest at low N deposition sites (Tab.
@@
   Comparing the indicator values of colonizing and disappearing species with random expectation, we found that oligotrophic species are currently more likely to colonize mountain hay meadows than eutrophic species, which might be the result of the recently observed decrease in atmospheric N deposition.
-However, our results also indicate that the recovering of mountain hay meadows after high N deposition might take much longer than transferring species-rich mountain hay meadows to species-poor communities with a large proportion of eutrophic species.
+However, our results also indicate that the recovering of mountain hay meadows from high N deposition might take much longer than transferring species-rich mountain hay meadows to species-poor communities with a large proportion of eutrophic species.
 This is because eutrophic species have high local survival probabilities, even after N deposition decreases again.

### 2018-12-12 R/Data_processing.R

@@ -124,7 +124,7 @@ for(i in 1:nrow(sites)) {
 #------------------------------------------------------------------------------------------------------
 # Select sites with two surveys per year
 turnover_double <- tbl(db, "KD_Z9") %>% 
-  filter(Aufnahmetyp == "Doppelaufnahme_Z9_Pflanzen") %>% 
+  filter(Aufnahmetyp == "Doppelaufnahme_Pflanzen_Z9") %>% 
   filter(yearPl >= 2003 & yearPl <= 2017) %>% 
   as.tibble() %>% 
   filter(!is.na(match(aID_STAO, sites$aID_STAO))) %>% 
@@ -194,7 +194,7 @@ sitecolEPT <- brewer.pal(8, name = "Paired")[8]
 
 pdf("Geodata/studysite.pdf", width = 6, height = 4)
 par(mar = c(0,0,2,0))
-plot(NA, xlim = c(490000, 840000), ylim = c(60000, 300000), type = "n", axes = FALSE)
+plot(NA, xlim = c(490000, 840000), ylim = c(60000, 300000), type = "n", axes = FALSE, asp = 1)
 plot(ch, add =TRUE)
 plot(gadm, add = TRUE)
 plot(seen[1:13,], add = TRUE, col = seecol, border = seecol, lwd = 0.01)

### 2018-12-13 Appendix_A.Rmd
Typos corrected.

--- 2018-10-19 Appendix_A.Rmd
@@
 Instead of the functional traits that are traditionally used in functional ecology such as species height, we used the Ellenberg values that we assumed could best reflect the main drivers that are likely to cause shifts in mountain hay plant communities.
-This were the Ellenberg values for temperature (that are likely to respond to climate warming), soil moisture (likely to respond to changes in precipitation), nutrients (likely to change in response to Nitrogen deposition and land-use intensity) and light (likely to change in response to land-use intensity).
+These were indicator values for temperature (that are likely to respond to climate warming), soil moisture (likely to respond to changes in precipitation), nutrients (likely to change in response to Nitrogen deposition and land-use intensity) and light (likely to change in response to land-use intensity).
 

### 2018-12-13 Manuscript.Rmd
Typos corrected.

@@
 Among others, N deposition is still high at many sites, since in contrast to nitrogen oxides, ammonia emissions decreased only to a small degree.
-Thus, even if N deposition is reduced, large areas might still be above the critical threshold above which harmful effects on plant diversity do occur [i.e.
-above the critical load for N deposition; @Bobbink2010; @Slootweeg2015; @RihmAchermann2016].
+Thus, even if N deposition is reduced, large areas might still be above the critical threshold above which harmful effects on plant diversity do occur [@Bobbink2010; @Slootweeg2015; @RihmAchermann2016].
 Furthermore, there is a possibility that communities reach an alternative stable state after decades of increased N deposition and that the respective plant species are unlikely to disappear even if N deposition is reduced [@Stevens2016].
@@
 For the analysis of the colonisation probability we examined for all species whether they were also observed in the subsequent survey (1st vs 2nd, 2nd vs 3rd).
-The variable ‚Äòocc‚Äô (observed = 1; not observed = 0) was then used as dependent variable in the logistic GLMM.
+The variable `occ` (observed = 1; not observed = 0) was then used as dependent variable in the logistic GLMM.
 As predictor variables we used the N deposition of the site, the average indicator value for nutrients, and the interaction of these two variables.
@@
 In that case, however, we selected all species that were recorded during the first or second survey, and the variable 'occ' then indicated whether or not the species was also observed during the next survey.
-Model parameters were estimated using an approximate Bayesian approach using the R-Package *arm* [@arm2018].
+Model parameters were estimated in a Bayesian framework using the R-Package *arm* [@arm2018].
 
@@
 As predictors we used the four environmental gradients as described above.
-Model parameters were estimated using an approximate Bayesian approach using the R-Package *arm* [@arm2018].
+Model parameters were estimated in a Bayesian framework using the R-Package *arm* [@arm2018].
 
@@
 Raw data for analyses are provided in the folder 'RData' and the folder 'R' contains the R-Script that was used to export the data from the BDM database.
-The folder "Settings" contains a list of all the R packages (including version number) that were in the workspace when the manuscript was rendered.
+The folder 'Settings' contains a list of all the R packages (including version number) that were in the workspace when the manuscript was rendered.
 
@@
 The final version of the repository will be archived on Zenodo upon acceptance of the manuscript.
+
+## Declaration of reporting decisions {-}
+This paper presents a selection of analyses with results that appeared most promising or interesting to us. Our study should therefore be understood as being exploratory and descriptive.
 
@@
 ```{r communitytrendscalc}
-res <- tibble(Measures = c("Species richness", "Temperature value", "soil moisture value", "Nutrient value", "Light value"),
+res <- tibble(Measures = c("Species richness", "Temperature value", "Soil moisture value", "Nutrient value", "Light value"),
               'Period 1' = NA,
@@
 The chance that the community mean of the nutrient value decreased between 2003 and 2017 was `r 1- res %>% filter(Measures == "Nutrient value") %>% pull("Prob.
-for Trend")` (`r res %>% filter(Measures == "Nutrient value") %>% pull("Prob.
-for Trend")` the probability for an increase was KEINE AHNUNG WAS ICH HIER EINFÜGEN MUSS...).
+for Trend")`.
 
@@
 It seems that the indicator values of newly colonizing species differed more from random colonization than the indicator values of disappearing species (Table \@ref(tab:difffromrandomtab)).
-For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with relatively low N deposition of 10 kg ha$^{-1}$ yr$^{-1}$, the newly colonizing species had on average a lower indicator value for nutrients than species under random colonization (column "Difference from random" in Table \@ref(tab:difffromrandomtab)), but the differences between colonizing species and random colonization decreased with increasing N deposition (column "Change along gradient" in Table \@ref(tab:difffromrandomtab)).
+For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with relatively low N deposition of 10 kg ha$^{-1}$ yr$^{-1}$, the newly colonizing species had on average a lower indicator value for nutrients than species under random colonization (column 'Difference from random' in Table \@ref(tab:difffromrandomtab)), but the differences between colonizing species and random colonization decreased with increasing N deposition (column 'Change along gradient' in Table \@ref(tab:difffromrandomtab)).
 Thus, at high N deposition, colonizing species did not differ from random species (see Figure 3b in Appendix A).
@@
 In 2000, only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a N deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, all situated above 1000 m.
-Between 2000 and 2015, the N deposition decreased on average by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$, with slightly higher net decreases at lower elevation (Fig.
+Between 2000 and 2015, the N deposition decreased on average by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$, with slightly higher net decreases at sites with previously high N deposition (Fig.
 \@ref(fig:Ndepchange)b).

### 2018-12-13 Manuscript.Rmd
Credibility intervals mit compatibility interval ersetzt.

@@
 Apparently, the community effects of the replacement of eutrophic species with oligotrophic species was outweighed by climate warming.
-Our results add to the increasing evidence that measures of community composition such as species richness or average indicator values may be relatively stable, but that this apparent stability is often accompanied by a marked turnover of species.
+Our results add to the increasing evidence that plant communities in changing environments may be relatively stable regarding average community composition, but that this apparent stability is often accompanied by a marked turnover of species..
 
@@
 
+As a summary of the result of a parameter we give the 5% and 95% quantiles of the marginal posterior distribution, which we interpreted as a compatibility interval showing effect sizes most compatible with the data, under the model and prior distribution used to compute the interval [@Amrhein2018].
+
 ## Data accessibility and reproducibility of results {-}
@@
 The temporal trends are given as change per 10 years and were estimated from linear mixed models with normal distribution (except for species richness with Poisson distribution and a log-link function).
-The measure of precision for the temporal trend is given as the 5\\% and 95\\% quantiles of the marginal posterior distribution of the linear trend (90\\% credible intervals).
+The measure of precision for the temporal trend is given as the 5\\% and 95\\% quantiles of the marginal posterior distribution of the linear trend (90\\% compatibility interval).
 The column 'Prob.
@@
   tt[1, "5%"] <- tmp[1] 
-  tt[1, "90%"] <- tmp[2] 
+  tt[1, "95%"] <- tmp[2] 
   tt[1, "Estimate_Slope"] <- coef(mod)[2]
@@
   tt[1, "5%_Slope"] <- tmp[1] 
-  tt[1, "90%_Slope"] <- tmp[2] 
+  tt[1, "95%_Slope"] <- tmp[2] 
   tt
@@
 ```{r difffromrandomtab}
-kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Indicator value", "Gradient", "Estimate", "5%", "90%", "Estimate", "5%", "90%"), caption = "Difference in the average indicator value of species that (a) disappeared from a site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
+kable(res, align = c("l", "l", rep("c", 6)), digits = 3,  booktabs = T, col.names = c("Indicator value", "Gradient", "Estimate", "5%", "95%", "Estimate", "5%", "95%"), caption = "Difference in the average indicator value of species that (a) disappeared from a site or (b) newly colonized a site compared to the same number of species that were randomly selected from all species recorded at a site.
 Shown are the results from linear models, with the difference between disappeared/colonized species and random species as dependent variable and the site-measure (gradient) as predictor variable.") %>%
@@
 ```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic species (indicator value for nutrients $<3$; red line) and of eutrophic species (indicator value for nutrients $>3$) along the N deposition gradient.
-Given are means and 95\\%-credible intervals from logistic linear mixed models.
+Given are means and 95\\% compatibility intervals from logistic linear mixed models.
 The vertical lines indicate the deposition rate, with equal colonization or survival probabilities for oligotrophic and eutrophic species, the solid line indicating the median, and the dashed lines the 5\\% and 95\\% quantiles of the marginal posterior distribution."}
@@
   tibble(mean = map_dbl(tt, mean),
-         low = map_dbl(tt, quantile, probs = 0.025),
-         high = map_dbl(tt, quantile, probs = 0.975),
+         low = map_dbl(tt, quantile, probs = 0.05),
+         high = map_dbl(tt, quantile, probs = 0.95),
          X = X,

### 2018-12-16 Manuscript.Rmd
Edits von Vali und Lukas einbezogen. Kleine weitere Anpassungen vorgenommen.

@@
 Apparently, the community effects of the replacement of eutrophic species with oligotrophic species was outweighed by climate warming.
-Our results add to the increasing evidence that plant communities in changing environments may be relatively stable regarding average community composition, but that this apparent stability is often accompanied by a marked turnover of species..
+Our results add to the increasing evidence that plant communities in changing environments may be relatively stable regarding average species richness or average indicator values, but that this apparent stability is often accompanied by a marked turnover of species.
 
@@
 Examples of such metrics would be the number of oligotrophic species, or the average indicator value of the species in a community [@Roth2013; @Rowe2017].
-However, the lack of a temporal trend in such biodiversity endpoints -- particularly in species richness -- does not necessarily mean that species composition remains unchanged.
+However, the lack of a temporal trend in such biodiversity endpoints - particularly in species richness - does not necessarily mean that species composition remains unchanged.
 This is because immigration and extinction might be equally frequent and may cancel each other out [@Hillebrand2018].
@@
 In mountain hay meadows, the spatial variation of species richness in vascular plants has been shown to be negatively correlated with N deposition [@Roth2013], suggesting that mountain grasslands are negatively affected by increased N deposition.
-However, between 1990 and 2010, NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14% and considerable emission reductions also occured in neighboring countries [@MaasGrennfelt2016].
+However, between 1990 and 2010, NO$_x$ emissions in Switzerland decreased by 46% and NH$_3$ emissions by 14%, and considerable emission reductions also occurred in neighboring countries [@MaasGrennfelt2016].
 Potentially, this could have led to a partial recovery of plant communities.
@@
 Management regimes of easily accessible mountainous areas are often being intensified, while poorly accessible mountainous areas are abandoned [@Tasser2002; @Strebel2015].
-Note that fires--an important driver of biodiversity in other grassland communities [@Ratajczak2014]--hardly occur in Central European mountain hay meadows.
+Note that fires - an important driver of biodiversity in other grassland communities [@Ratajczak2014] - hardly occur in Central European mountain hay meadows.
 
@@
 To better disentangle the importance of these mechanisms, we ordered the sites along four main environmental gradients.
-First, we expected communities to respond to climate warming [@Roth2014]. To describe the temperature gradient, we used the mean annual temperature per site from the WordClim database [@Fick2017].
+First, we expected communities to respond to climate warming [@Roth2014].
+To describe the temperature gradient, we used the mean annual temperature per site based on data from the WorldClim database [@Fick2017].
 The average $\pm$ SD mean annual temperature at our sites was `r mean(sites$TemperatureR / 10)` $\pm$ `r sd(sites$TemperatureR / 10)` $^\circ$C. 
 The monthly-mean surface air temperature for Switzerland shows a linear increase of 1.29 $^\circ$C per 100 years between 1864 and 2016 with the warmest three years of the entire period measured in 2011, 2014 and 2015 [@Begert2018].
-Also from the WordClim database, we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change [@Beier2012].The average annual precipitation at our sites was `r mean(sites$PrecipitationR)` $\pm$ `r sd(sites$PrecipitationR)` mm.
+Also based on data from the WorldClim database, we used the annual precipitation per site, another key driver for plants that is likely to be affected by climate change [@Beier2012].
+The average annual precipitation at our sites was `r mean(sites$PrecipitationR)` $\pm$ `r sd(sites$PrecipitationR)` mm.
 Further, we estimated atmospheric N deposition for each site using a pragmatic approach that combined monitoring data, spatial interpolation methods, emission inventories, statistical dispersion models and inferential deposition models [@RihmAchermann2016].
@@
 
-To infer whether species that colonized or disappeared from a site had particular indicator values that differed from the other species at that site, we produced for each site a list with all species that were recorded during the three surveys (total community).
+To infer whether species that colonized a site or disappeared from a site had particular indicator values that differed from the other species at that site, we produced for each site a list with all species that were recorded during the three surveys (total community).
 We then calculated the community mean (CM) of the indicator value for all species that colonized the site during the three surveys (i.e.
@@
 We used logistic GLMMs to analyse whether the colonization probability or local survival probability was changing along the Nitrogen deposition gradient and whether this change depended on the species indicator value for nutrients.
-For the analysis of the colonisation probability we examined for all species whether they were also observed in the subsequent survey (1st vs 2nd, 2nd vs 3rd).
-The variable `occ` (observed = 1; not observed = 0) was then used as dependent variable in the logistic GLMM.
+For the analysis of the colonisation probability we examined for all species that were not observed at the first or second survey whether they were observed in the subsequent survey (1st vs 2nd, 2nd vs 3rd).
+Whether they were  observed in the subsequent survey was used as dependent variable in the logistic GLMM.
 As predictor variables we used the N deposition of the site, the average indicator value for nutrients, and the interaction of these two variables.
@@
 The same logistic GLMM was also used to investigate local survival probabilities.
-In that case, however, we selected all species that were recorded during the first or second survey, and the variable 'occ' then indicated whether or not the species was also observed during the next survey.
-Model parameters were estimated in a Bayesian framework using the R-Package *arm* [@arm2018].
+In that case, however, we selected all species that were recorded during the first or second survey, and inferred whether or not the species was observed during the subsequent survey (1st vs 2nd, 2nd vs 3rd).
+Model parameters were estimated in a Bayesian framework using the R-package *arm* [@arm2018].
 
@@
 As predictors we used the four environmental gradients as described above.
-Model parameters were estimated in a Bayesian framework using the R-Package *arm* [@arm2018].
+Model parameters were estimated in a Bayesian framework using the R-package *arm* [@arm2018].
 
-As a summary of the result of a parameter we give the 5% and 95% quantiles of the marginal posterior distribution, which we interpreted as a compatibility interval showing effect sizes most compatible with the data, under the model and prior distribution used to compute the interval [@Amrhein2018].
+As parameter estimates we give the 5% and 95% quantiles of the marginal posterior distribution, which we interpreted as a 90% compatibility interval showing effect sizes most compatible with the data, under the model and prior distribution used to compute the interval [@Amrhein2018].
+Smoothing was done using the 'loess' function of the R-library 'stats' [@R2018] with default settings.
 
@@
 High species turnover at a site is the result of species that disappeared from the site and/or of species that newly colonized the site.
-To better understand the factors that drive these changes we were particularly interested whether the species that disappeared or colonized the sites differed in indicator values compared to what would be expected if the same number of species randomly disappeared or colonized the sites (i.e.
+To better understand the factors driving these changes we were particularly interested in whether the species that disappeared or colonized the sites differed in indicator values compared to what would be expected if the same number of species randomly disappeared or colonized the sites (i.e.
 random disappearance and random colonization), and whether there is spatial variation along the environmental gradients.
@@
 ## Potential effects of reduction in Nitrogen emmissions {-}
-```{r Ndepchange, fig.cap="(a) The nitrogen (N) deposition in 2000 and (b) the change in N deposition between 2000 and 2015, along the N deposition gradient of the study sites used in 2000."}
+```{r Ndepchange, fig.cap="(a) The nitrogen (N) deposition in 2000 and (b) the change in N deposition between 2000 and 2015, along the N deposition gradient of the study sites used in 2000.
+The blue lines indicate the loess curve and the grey areas indicate the corresponding 90\\% compatibility intervals."}
 p1 <- ggplot(sites, aes(x = Elevation * 500 + 1000, y = NTOT_2000)) +
   geom_point(cex = 0.8) +
-  geom_smooth(method = loess) +
+  geom_smooth(method = loess, level = 0.9) +
   geom_abline(intercept = 0, slope = 0, lty = 3) +
@@
   geom_point(cex = 0.8) +
-  geom_smooth(method = loess)  +
+  geom_smooth(method = loess, level = 0.9)  +
   labs(title = "(b) Temporal change (2000 to 2015)", 
@@
 \@ref(fig:Ndepchange)a).
-In 2000, only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a N deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, all situated above 1000 m.
+In 2000, only `r 100 * mean(sites$NTOT_2000 < 10)`% of sites had a N deposition rate of less than 10 kg N ha$^{-1}$ yr$^{-1}$, all of which situated above 1000 m.
 Between 2000 and 2015, the N deposition decreased on average by `r mean(sites$NTOT_2015 - sites$NTOT_2000)` $\pm$ `r sd(sites$NTOT_2015 - sites$NTOT_2000)` kg N ha$^{-1}$ yr$^{-1}$, with slightly higher net decreases at sites with previously high N deposition (Fig.
@@
 ```{r cl, fig.cap="Colonization (a) and local survival (b) of oligotrophic species (indicator value for nutrients $<3$; red line) and of eutrophic species (indicator value for nutrients $>3$) along the N deposition gradient.
-Given are means and 95\\% compatibility intervals from logistic linear mixed models.
+Given are means and 90\\% compatibility intervals from logistic linear mixed models.
 The vertical lines indicate the deposition rate, with equal colonization or survival probabilities for oligotrophic and eutrophic species, the solid line indicating the median, and the dashed lines the 5\\% and 95\\% quantiles of the marginal posterior distribution."}
@@
 Local survival probability was the same for oligotrophic and euthrophic species at a deposition rate of `r CL_surv` kg N ha$^{-1}$ yr$^{-1}$; colonization probability was the same for oligotrophic and eutrophic species at a deposition rate of `r CL_col` kg N ha$^{-1}$ yr$^{-1}$.
-In `r 100 * mean((sites$NTOT * 10 + 10) < 12.5)`% of the sites, the deposition rate was below 12.5 kg N ha$^{-1}$ yr$^{-1}$ at which the replacement of eutrophic with oligotrophic species is likely according to Fig.
+In `r 100 * mean((sites$NTOT * 10 + 10) < 12.5)`% of the sites, the deposition rate was below 12.5 kg N ha$^{-1}$ yr$^{-1}$ at which the replacement of eutrophic with oligotrophic species is likely, according to Fig.
 \@ref(fig:cl).
@@
   geom_line() + 
-  labs(x = "five-year interval", y = "Effect of N deposition") + 
+  labs(x = "Five-year interval", y = "Effect of N deposition") + 
   scale_x_continuous(breaks = seq(2004, 2012, 2), labels = paste0(seq(2004, 2012, 2), "-\n", seq(2008, 2016, 2)))
@@
 
-In Swiss mountain hay meadows, the average N deposition was still rather high with an average of `r mean(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$), which is at the upper limit of the suggested critical load [@Roth2017].
+In Swiss mountain hay meadows, the average N deposition was still rather high with an average of `r mean(sites$NTOT_2015)` kg ha$^{-1}$ yr$^{-1}$, which is at the upper limit of the suggested critical load [@Roth2017].
 Furthermore, Nitrogen deposition only weakly decreased by about `r mean(sites$NTOT_2015 - sites$NTOT_2000)` kg ha$^{-1}$ yr$^{-1}$ between 2000 and 2015.
@@
 \@ref(tab:communitytrendstab)).
-This is in contrast to how N deposition is affecting disappearance and colonization of species: It seems that both the species disappearing from the sites as well as species colonizing the sites tend to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
-Thus, the effects of disappearing and colonizing species on the average community value for nutrients partially cancel each other out.
-Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species have much reduced local survival at higher N deposition rates.
+This was in contrast to how N deposition is affecting disappearance and colonization of species: It seems that both the species disappearing from the sites as well as species colonizing the sites tended to have below average indicator values for nutrients (Table \@ref(tab:difffromrandomtab)).
+Thus, the effects of disappearing and colonizing species on the average community value for nutrients partially cancelled each other out.
+Furthermore, eutrophic species had rather high local survival across the entire deposition gradient, while oligotrophic species had much reduced local survival at higher N deposition rates.
 This suggests that mountain hay meadow communities can reach alternative stable states, with eutrophic species that are unlikely to disappear even if N deposition is reduced [@Stevens2016].
-Taken together, these reasons might explain why the composition of mountain hay meadow communities responded stronger to climate warming than to nitrogen reduction, although the reduction in nitrogen resulted in above average colonization of oligotrophic species.
+Taken together, these factors might explain why the composition of mountain hay meadow communities responded stronger to climate warming than to nitrogen reduction, although the reduction in nitrogen resulted in above average colonization of oligotrophic species.
 
@@
 Given that the BDM program has included major efforts in developing reproducible methods and has continuously invested in quality control [@Plattner2004], the recorded pseudo-turnover was quite high.
-A potential explanation is that the replicated surveys were not conducted during the same days, and at some sites the situation might have changed profoundly, for example because a meadow was cut between the surveys.
-Furthermore, one of two botanists at a given site may not have identified some difficult species to the species-level, and--because we analysed only records that were identified to the species level--this might have increased pseudo-turnover even though both botanists actually detected the species.
+A potential explanation is that species that are difficult to identify were only identified by one botanist at species-level, while the other identified them at genus level.
+Although both botanists discovered the species this might have increased the pseudo-turnover because we only analysed records that were identified at species-level.
+Furthermore, replicated surveys are not conducted during the same days and in few cases the situation might have changed profoundly for example because the meadow was cut between the surveys.
 Such problems must be taken into account when evaluating the presented results.
@@
 This seems biologically plausible, since the average species coverage in species-rich sites must be lower than in species-poor sites, and species with low coverage probably have a higher turnover.
-At the same time, however, the result could also be simply due to pseudo-turnover, since species identification in species-rich sites is probably more difficult than in species-poor sites.
+At the same time, however, the result could also simply be due to pseudo-turnover, since species identification in species-rich sites is probably more difficult than in species-poor sites.
 However, for the presented result suggesting differences in colonization or local survival in relation to the species indicator values, we can hardly imagine how this could be caused by methodological issues.

### 2018-12-18 Manuscript.Rmd
Small edits according to remarks from PeerJ.

@@
 bibliography: Bibliography.bib
+csl: Settings/Refstyles.csl
 output: 
@@
 For colonizing species, we found the largest differences from random colonization in the indicator value for nutrients: at sites with relatively low N deposition of 10 kg ha$^{-1}$ yr$^{-1}$, the newly colonizing species had on average a lower indicator value for nutrients than species under random colonization (column 'Difference from random' in Table \@ref(tab:difffromrandomtab)), but the differences between colonizing species and random colonization decreased with increasing N deposition (column 'Change along gradient' in Table \@ref(tab:difffromrandomtab)).
-Thus, at high N deposition, colonizing species did not differ from random species (see Figure 3b in Appendix A).
+Thus, at high N deposition, colonizing species did not differ from random species (see Appendix A).
 
